### 2018-02-19 Merow2014_Appendix.Rmd
working through appendixes from Merow et al 2014

---
title: "Merow 2014 Appendix"
author: "Tim Szewczyk"
date: "2/16/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pkgs <- c("tidyverse", "magrittr", "fields", "IPMpack")
suppressMessages(invisible(lapply(pkgs, library, character.only=TRUE)))
theme_set(theme_bw())
```

```{r App_A}
d <- read.csv("data/Appendix_A_Simple_data.csv")
# Appendix A from Merow 2014
########
## Fit regressions
########
  p <- NULL
  # survival logistic regression
  surv.reg <- glm(surv ~ size, family="binomial", data=d)
  p$surv.int <- coef(surv.reg)[1]
  p$surv.slope <- coef(surv.reg)[2]

  # growth linear regression
  growth.reg <- gls(sizeNext ~ size + I(size^2), weights=varExp(), 
                    na.action=na.omit, data=d)
  p$growth.int <- coef(growth.reg)[1]
  p$growth.slope <- coef(growth.reg)[2]
  p$growth.sqrd <- coef(growth.reg)[3]
  p$growth.sigma2 <- summary(growth.reg)$sigma^2
  p$growth.sigma2.exp <- as.numeric(growth.reg$modelStruct$varStruct)

  # flower logistic regression
  flower.reg <- glm(fec.flower ~ size, family="binomial", data=d)
  p$flower.int <- coef(flower.reg)[1]
  p$flower.slope <- coef(flower.reg)[2]

  # seed poisson regression
  seed.reg <- glm(fec.seed ~ size, family="poisson", data=d[d$fec.flower==1,])
  p$seed.int <- coef(seed.reg)[1]
  p$seed.slope <- coef(seed.reg)[2]

  # recruit size distribution
  p$recruit.size.mean <- mean(d$sizeNext[is.na(d$size)])
  p$recruit.size.sd <- sd(d$sizeNext[is.na(d$size)])

  # establishment probability
  p$establishment.prob <- sum(is.na(d$size))/sum(d$fec.seed, na.rm=TRUE)

  par(mfrow=c(2,2), mar=c(4,4,2,1))
  xx <- data.frame(size=seq(0, 8, by=.01))
  plot(surv ~ size, data=d, col=rgb(0,0,0,0.5), 
       xlab="Size (t)", ylab="Survival to t+1")
  lines(xx$size, predict(surv.reg, xx, type="response"), col="red")
  plot(sizeNext ~ size, data=d, col=rgb(0,0,0,0.5), 
       xlab="Size (t)", ylab="Size (t+1)")
  lines(xx$size, predict(growth.reg, xx), col="red")
  plot(fec.seed ~ size, data=d, col=rgb(0,0,0,0.5), 
       xlab="Size (t)", ylab="Seed number")
  lines(xx$size, predict(seed.reg, xx, type="response"), col="red")
  lines(xx$size, predict(flower.reg, xx, type="response")*
          max(d$fec.seed, na.rm=TRUE), col="blue")
  hist(d$sizeNext[is.na(d$size)], xlab="Recruit size", main="", freq=FALSE)
  lines(xx$size, dnorm(xx$size, p$recruit.size.mean, p$recruit.size.sd),
        col="red")

########
## Life history functions
########
  # survival probability
  s.x <- function(x, p) {
    u <- exp(p$surv.int + p$surv.slope*x)
    return(u / (1+u))
  }

  # growth
  g.yx <- function(xp, x, p) {
    dnorm(xp, p$growth.int + p$growth.slope*x + p$growth.sqrd*x^2, 
          sqrt(p$growth.sigma2 * exp(2*p$growth.sigma2.exp*x)))
  }

  # reproduction
  p.flower.x <- function(x, p) {
    u <- exp(p$flower.int + p$flower.slope*x)
    return(u/(1+u))
  }
  f.yx <- function(xp, x, p) {
    p.flower.x(x, p)*
      p$establishment.prob*
      dnorm(xp, p$recruit.size.mean, p$recruit.size.sd)*
      exp(p$seed.int + p$seed.slope*x)
  }

########
## Kernel creation
########
  min.size <- .9*min(c(d$size, d$sizeNext), na.rm=TRUE)
  max.size <- 1.1*max(c(d$size, d$sizeNext), na.rm=TRUE)
  n <- 100  # ncells in matrix
  b <- min.size + c(0:n)*(max.size - min.size)/n  # boundary points
  y <- 0.5*(b[1:n] + b[2:(n+1)])  # mesh points
  h <- y[2] - y[1]  # step size

  G.mx <- h*outer(y, y, g.yx, p=p)
  S.v <- s.x(y, p)
  F.mx <- h*outer(y, y, f.yx, p=p)
  P.mx <- G.mx
  # Fix eviction of individuals below/above size limits
  for(i in 1:(n/2)) {
    G.mx[1,i] <- G.mx[1,i] + 1 - sum(G.mx[,i])
    P.mx[,i] <- G.mx[,i]*S.v[i]
  }
  for(i in (n/2+1):n) {
    G.mx[n,i] <- G.mx[n,i] + 1 - sum(G.mx[,i])
    P.mx[,i] <- G.mx[,i]*S.v[i]
  }
  K.mx <- P.mx + F.mx

########
## Summaries & plots
########
  lam <- Re(eigen(K.mx)$values[1])
  w.eigen <- Re(eigen(K.mx)$vectors[,1])
  stable.dist <- w.eigen/sum(w.eigen)
  v.eigen <- Re(eigen(t(K.mx))$vectors[,1])
  repro.val <- v.eigen/v.eigen[1]
  v.dot.w <- sum(stable.dist*repro.val)*h
  sens <- outer(repro.val, stable.dist)/v.dot.w
  elas <- matrix(as.vector(sens)*as.vector(K.mx)/lam, nrow=n)

  par(mfrow=c(2,3), mar=c(4,5,2,2))
  image.plot(y, y, t(K.mx),  
             xlab="Size (t)", ylab="Size (t+1)", main="IPM matrix")
  abline(a=0, b=1, lty=2)
  contour(y, y, t(K.mx), add=TRUE, drawlabels=TRUE)
  image.plot(y, y, t(elas), 
             xlab="Size (t)", ylab="Size (t+1)", main="Elasticity")
  image.plot(y, y, t(sens), 
             xlab="Size (t)", ylab="Size (t+1)", main="Sensitivity")
  plot(y, stable.dist, xlab="Size", type="l", main="Stable size distribution")
  plot(y, repro.val, xlab="Size", type="l", main="Reproductive values")
  plot(y, s.x(y, p), xlab="Size", ylab="Pr(surv)", main="Eviction")
  lines(y, apply(P.mx, 2, sum), col="red")
```

```{r App_B}
data("dataIPMpackSuccisa")
Sp1 <- dataIPMpackSuccisa

par(mfrow=c(3,2), mar=c(4,4,2,1)) 
plot(Sp1$size, Sp1$surv, col=rgb(0,0,0,0.5), 
     xlab="Size (t)", ylab="Survival to t+1")
plot(Sp1$size, Sp1$sizeNext, col=rgb(0,0,0,0.5), 
     xlab="Size (t)", ylab="Size (t+1)") 
plot(Sp1$size, Sp1$fec1Bolt, col=rgb(0,0,0,0.5), 
     xlab="Size (t)", ylab="Flowering probability")
plot(Sp1$size, Sp1$fec2Stem, col=rgb(0,0,0,0.5), 
     xlab="Size (t)", ylab="Stem Number")
plot(Sp1$size, Sp1$fec3Head, col=rgb(0,0,0,0.5), 
     xlab="Size (t)", ylab="Flower Head Number") 
hist(Sp1$sizeNext[is.na(Sp1$size)], main="", xlab="Recruit Size")

# size ranges
x <- seq(0, 10, length.out=1001)
x0 <- data.frame(size=x, size2=x*x)
minSize <- min(Sp1$size, na.rm=TRUE) - 1
maxSize <- max(Sp1$size, na.rm=TRUE) + 2

########
## survival and growth kernels
########
  par(mfrow=c(1,3))
  survModelComp(dataf=Sp1, makePlot=TRUE, legendPos="bottomright")
  so <- makeSurvObj(Sp1, surv ~ size)
  growthModelComp(dataf=Sp1, makePlot=TRUE, legendPos="bottomright")
  go <- makeGrowthObj(Sp1, sizeNext ~ size + size2)
  P.mx <- makeIPMPmatrix(survObj=so, growObj=go, correction="constant",
                            minSize=minSize, maxSize=maxSize)

  image.plot(P.mx@meshpoints, P.mx@meshpoints, t(P.mx), 
             main="P: Survival & growth", xlab="Size (t)", ylab="Size (t+1)")
  abline(a=0, b=1, lty=2)
  diagnosticsPmatrix(P.mx, growObj=go, survObj=so, correction="constant")

########
## fecundity kernel
########
  # flowering
  fo1 <- makeFecObj(Sp1, Formula=fec1Bolt ~ 1, Family="binomial")
  fo2 <- makeFecObj(Sp1, Formula=fec1Bolt ~ size, Family="binomial")
  fo3 <- makeFecObj(Sp1, Formula=fec1Bolt ~ size + size2, Family="binomial")
  par(mfrow=c(1,1))
  fs <- order(Sp1$size)
  fs.fec <- Sp1$fec1Bolt[fs]
  fs.size <- Sp1$size[fs]
  pfz <- tapply(fs.size, as.numeric(cut(fs.size, 21)), mean, na.rm=TRUE)
  ps <- tapply(fs.fec, as.numeric(cut(fs.size, 21)), mean, na.rm=TRUE)
  plot(as.numeric(pfz), as.numeric(ps), ylim=c(0,1), 
       xlab="Size", ylab="Proportion flowering")
  lines(x, predict(fo1@fitFec[[1]], newdata=x0, type="response"), col="red")
  lines(x, predict(fo2@fitFec[[1]], newdata=x0, type="response"), col="green")
  lines(x, predict(fo3@fitFec[[1]], newdata=x0, type="response"), col="blue")
  legend("topleft", col=2:4, lty=1, xjust=1, bg="white",
         legend=sprintf("%s: %s = %.1f", c("1", "size", "size + size2"), 
                        c("AIC"), c(AIC(fo1@fitFec[[1]]), AIC(fo2@fitFec[[1]]),
                                    AIC(fo3@fitFec[[1]]))))
  # stem production
  fo1 <- makeFecObj(Sp1, Formula=c(fec1Bolt~size+size2, fec2Stem~1), 
                    Family=c("binomial", "poisson"), Transform=c("none", "-1"))
  fo2 <- makeFecObj(Sp1, Formula=c(fec1Bolt~size+size2, fec2Stem~size), 
                    Family=c("binomial", "poisson"), Transform=c("none", "-1"))
  fo3 <- makeFecObj(Sp1, Formula=c(fec1Bolt~size+size2, fec2Stem~size+size2), 
                    Family=c("binomial", "poisson"), Transform=c("none", "-1"))
  fs <- order(Sp1$size)
  fs.fec <- Sp1$fec2Stem[fs]
  fs.size <- Sp1$size[fs]
  pfz <- tapply(fs.size, as.numeric(cut(fs.size, 21)), mean, na.rm=TRUE)
  ps <- tapply(fs.fec, as.numeric(cut(fs.size, 21)), mean, na.rm=TRUE)
  plot(as.numeric(pfz), as.numeric(ps), 
       xlab="Size", ylab="Number of stems")
  lines(x, predict(fo1@fitFec[[2]], newdata=x0, type="response")+1, col="red")
  lines(x, predict(fo2@fitFec[[2]], newdata=x0, type="response")+1, col="green")
  lines(x, predict(fo3@fitFec[[2]], newdata=x0, type="response")+1, col="blue")
  legend("topleft", col=2:4, lty=1, xjust=1, bg="white",
         legend=sprintf("%s: %s = %.1f", c("1", "size", "size + size2"), 
                        c("AIC"), c(AIC(fo1@fitFec[[2]]), AIC(fo2@fitFec[[2]]),
                                    AIC(fo3@fitFec[[2]]))))

  # inflorescence count
  fo1 <- makeFecObj(Sp1, Formula=c(fec1Bolt~size+size2, fec2Stem~1, 
                                   fec3Head~1), 
                    Family=c("binomial", "poisson", "poisson"), 
                    Transform=c("none", "-1", "none"))
  fo2 <- makeFecObj(Sp1, Formula=c(fec1Bolt~size+size2, fec2Stem~1, 
                                   fec3Head~size), 
                    Family=c("binomial", "poisson", "poisson"), 
                    Transform=c("none", "-1", "none"))
  fo3 <- makeFecObj(Sp1, Formula=c(fec1Bolt~size+size2, fec2Stem~1, 
                                   fec3Head~size+size2), 
                    Family=c("binomial", "poisson", "poisson"), 
                    Transform=c("none", "-1", "none"))
  fs <- order(Sp1$size)
  fs.fec <- Sp1$fec3Head[fs]
  fs.size <- Sp1$size[fs]
  pfz <- tapply(fs.size, as.numeric(cut(fs.size, 21)), mean, na.rm=TRUE)
  ps <- tapply(fs.fec, as.numeric(cut(fs.size, 21)), mean, na.rm=TRUE)
  plot(as.numeric(pfz), as.numeric(ps), 
       xlab="Size", ylab="Number of stems")
  lines(x, predict(fo1@fitFec[[3]], newdata=x0, type="response"), col="red")
  lines(x, predict(fo2@fitFec[[3]], newdata=x0, type="response"), col="green")
  lines(x, predict(fo3@fitFec[[3]], newdata=x0, type="response"), col="blue")
  legend("topleft", col=2:4, lty=1, xjust=1, bg="white",
         legend=sprintf("%s: %s = %.1f", c("1", "size", "size + size2"), 
                        c("AIC"), c(AIC(fo1@fitFec[[3]]), AIC(fo2@fitFec[[3]]),
                                    AIC(fo3@fitFec[[3]]))))

  # final fecundity kernel
  fo <- makeFecObj(Sp1, Formula=c(fec1Bolt~size+size2, fec2Stem~1, 
                                  fec3Head~size), 
                   Family=c("binomial", "poisson", "poisson"), 
                   Transform=c("none", "-1", "none"),
                   fecConstants=data.frame(seedsPerHead=50,
                                           seedlingEstablishmentRate=0.02))
  F.mx <- makeIPMFmatrix(fecObj=fo, minSize=minSize, maxSize=maxSize)

  image.plot(F.mx@meshpoints, F.mx@meshpoints, t(F.mx), 
             main="F: Sexual reproduction", xlab="Size (t)", ylab="Size (t+1)")

  sens.out <- sensParams(growObj=go, survObj=so, fecObj=fo, 
                         nBigMatrix=50, minSize=minSize, maxSize=maxSize)
  barplot(sens.out$elas, las=2, cex.names=0.5)
```

```{r App_C}
########
## set up
########
  data("dataIPMpackHypericumCov"); help("dataIPMpackHypericumCov")
  d <- dataIPMpackHypericumCov
  d1 <- subset(d, !is.na(d$size) | d$ontogenyNext==1)
  d1 <- subset(d1, d1$year %in% c(1994, 1996))
  fec2 <- 13.78  # nSeeds per fruit
  fec3 <- 0.001336  # pr(germinate_t)
  fec4 <- 0.14  # surv(seedling_t_to_t+1)
  goSB <- 0.08234528
  staySB <- 0.671
  d1 <- d1[,c("year", "size", "surv", "sizeNext", "fec0", "fec1")]

  # assign surviving individuals to 'continuous' stage
  d1$stageNext <- d1$stage <- "continuous"
  d1$stage[is.na(d1$size)] <- NA
  d1$stageNext[d1$surv==0] <- "dead"

  # structure for transition between discrete (SB) & continuous (stem) stages
  d1$number <- 1 
  seedbank <- data.frame(year="All", 
                         size=NA, surv=1, sizeNext=NA,
                         fec0=NA, fec1=NA, 
                         stage=c("seedbank", "seedbank", "continuous"),
                         stageNext=c("seedbank", "continuous", "seedbank"),
                         number=c(staySB, (1-staySB)*fec3*fec4, 1))
  d1 <- rbind(d1, seedbank)
  d1$stage <- as.factor(d1$stage)
  d1$stageNext <- as.factor(d1$stageNext)

########
## construct kernels
########
  # 1994
  d94 <- subset(d1, d1$year %in% c("1994", "All"))
  minSize <- min(d1$size, na.rm=TRUE)
  maxSize <- max(d1$size, na.rm=TRUE)
  x <- seq(minSize, maxSize, length.out=1001)
  x0 <- data.frame(size=x, size2=x^2, size3=x^3)

  # survival
  survModelComp(dataf=d94[!is.na(d$size),], 
                expVars=c(surv ~ 1, 
                          surv ~ size, 
                          surv ~ size + size2, 
                          surv ~ size + size2 + size3), 
                makePlot=TRUE, legendPos="bottomleft", mainTitle="Survival")
  so94 <- makeSurvObj(d94, surv ~ size + I(size^2))

  # growth
  growthModelComp(dataf=d94, 
                  expVars=c(sizeNext ~ 1, 
                            sizeNext ~ size, 
                            sizeNext ~ size + size2, 
                            sizeNext ~ size + size2 + size3),
                  makePlot=TRUE, legendPos="bottomleft", mainTitle="Growth")
  abline(a=0, b=1, lty=2)
  go94 <- makeGrowthObj(d94, sizeNext ~ size)

  # P matrix
  P.mx.94 <- makeIPMPmatrix(survObj=so94, growObj=go94, 
                            minSize=minSize, maxSize=maxSize, 
                            nBigMatrix=80, correction="constant")
  image.plot(P.mx.94@meshpoints, P.mx.94@meshpoints, t(P.mx.94),
             main="P.mx: Survival & growth", xlab="Size (t)", ylab="Size (t+1)")
  abline(a=0, b=1, lty=2, col="white", lwd=2)
  diagnosticsPmatrix(P.mx.94, growObj=go94, survObj=so94, correction="constant")

  # fecundity
  ## flowering probability
  fec0.0_94 <- makeFecObj(d94, Formula=fec0~1, Family="binomial")
  fec0.1_94 <- makeFecObj(d94, Formula=fec0~size, Family="binomial")
  fec0.2_94 <- makeFecObj(d94, Formula=fec0~size+size2, Family="binomial")
  fec0.3_94 <- makeFecObj(d94, Formula=fec0~size+size2+size3, Family="binomial")
  fs <- order(d94$size); fs.fec0 <- d94$fec0[fs]; fs.size <- d94$size[fs]
  pfz <- tapply(fs.size, as.numeric(cut(fs.size, 21)), mean, na.rm=TRUE)
  ps <- tapply(fs.fec0, as.numeric(cut(fs.size, 21)), mean, na.rm=TRUE)
  plot(as.numeric(pfz), as.numeric(ps), xlab="Size", ylab="Pr(flower)")
    y0 <- predict(fec0.0_94@fitFec[[1]], newdata=x0)
    lines(x, exp(y0)/(exp(y0) + 1), col=2)
    y0 <- predict(fec0.1_94@fitFec[[1]], newdata=x0)
    lines(x, exp(y0)/(exp(y0) + 1), col=3)
    y0 <- predict(fec0.2_94@fitFec[[1]], newdata=x0)
    lines(x, exp(y0)/(exp(y0) + 1), col=4)
    y0 <- predict(fec0.3_94@fitFec[[1]], newdata=x0)
    lines(x, exp(y0)/(exp(y0) + 1), col=5)
    legend("bottomright", col=2:5, lty=1, xjust=1, bg="white",
           legend=sprintf("%s: %s = %.1f", 
                          c("1","size","size+size2","size+size2+size3"), 
                          c("AIC"), 
                          c(AIC(fec0.0_94@fitFec[[1]]), 
                            AIC(fec0.1_94@fitFec[[1]]), 
                            AIC(fec0.2_94@fitFec[[1]]), 
                            AIC(fec0.3_94@fitFec[[1]]))))
  ## fruit number per plant
  fec1.0_94 <- makeFecObj(d94, Formula=fec1~1, Family="poisson")
  fec1.1_94 <- makeFecObj(d94, Formula=fec1~size, Family="poisson")
  fec1.2_94 <- makeFecObj(d94, Formula=fec1~size+size2, Family="poisson")
  fec1.3_94 <- makeFecObj(d94, Formula=fec1~size+size2+size3, Family="poisson")
  fs.fec1 <- d94$fec1[fs]
  ps <- tapply(fs.fec1, as.numeric(cut(fs.size, 21)), mean, na.rm=TRUE)
  plot(as.numeric(pfz), as.numeric(ps), xlab="Size", ylab="Fruits per plant")
    y0 <- predict(fec1.0_94@fitFec[[1]], newdata=x0)
    lines(x, exp(y0), col=2)
    y0 <- predict(fec1.1_94@fitFec[[1]], newdata=x0)
    lines(x, exp(y0), col=3)
    y0 <- predict(fec1.2_94@fitFec[[1]], newdata=x0)
    lines(x, exp(y0), col=4)
    y0 <- predict(fec1.3_94@fitFec[[1]], newdata=x0)
    lines(x, exp(y0), col=5)
    legend("topleft", col=2:5, lty=1, xjust=1, bg="white",
           legend=sprintf("%s: %s = %.1f", 
                          c("1","size","size+size2","size+size2+size3"), 
                          c("AIC"), 
                          c(AIC(fec1.0_94@fitFec[[1]]), 
                            AIC(fec1.1_94@fitFec[[1]]), 
                            AIC(fec1.2_94@fitFec[[1]]), 
                            AIC(fec1.3_94@fitFec[[1]]))))

  # F matrix
  fo94 <- makeFecObj(d94, 
                     Formula=c(fec0~size+size2+size3, fec1~size+size2+size3),
                     Family=c("binomial", "poisson"), Transform=c("none", -1),
                     meanOffspringSize=mean(d94[is.na(d94$size) & 
                                                  !is.na(d94$sizeNext), 
                                                "sizeNext"]),
                     sdOffspringSize=sd(d94[is.na(d94$size) & 
                                              !is.na(d94$sizeNext), 
                                            "sizeNext"]),
                     fecConstants=data.frame(fec2=fec2, fec3=fec3, fec4=fec4),
                     offspringSplitter=data.frame(seedbank=goSB,
                                                  continuous=(1-goSB)),
                     vitalRatesPerOffspringType=data.frame(
                       seedbank=c(1,1,1,0,0),
                       continuous=c(1,1,1,1,1),
                       row.names=paste0("fec", 0:4)))
  F.mx.94 <- makeIPMFmatrix(fecObj=fo94, minSize=minSize, maxSize=maxSize,
                            nBigMatrix=80, correction="constant")
  image.plot(F.mx.94@meshpoints, F.mx.94@meshpoints, t(F.mx.94[-1, -1]), 
             main="F.mx: Fecundity", xlab="Size (t)", ylab="Size (t-1)")

  # Stage transitions: discrete & continuous
  dto94 <- makeDiscreteTrans(d94)
  dummy94 <- as.matrix(fo94@offspringRel$coefficients[1], list(1, "seedbank"))
  dto94@meanToCont <- as.matrix(dummy94, dimnames=c(1, "seedbank"))
  dummy94 <- as.matrix(fo94@sdOffspringSize, dimnames=list(1, "seedbank"))
  dto94@sdToCont <- as.matrix(dummy94, dimnames=c(1, "seedbank"))
  dto94@discreteTrans[1,1] <- staySB + (1-staySB)*fec3*fec4
  P.mx.94 <- makeIPMPmatrix(growObj=go94, survObj=so94, discreteTrans=dto94,
                            minSize=minSize, maxSize=maxSize, nBigMatrix=80,
                            correction="constant")
  image.plot(P.mx.94@meshpoints, P.mx.94@meshpoints, t(P.mx.94[1:80, 1:80]),
             main="P.mx: Survival & growth", xlab="Size (t)", ylab="Size (t+1)")
  abline(a=0, b=1, lty=2, col="white", lwd=2)

  # Full kernel
  IPM94 <- P.mx.94 + F.mx.94
  Re(eigen(IPM94)$value[1])
  sens.94 <- sens(IPM94)
  image.plot(0:80, 0:80, t(sens.94),
             main="Sensitivity", xlab="Size (t)", ylab="Size (t+1)")
  elas.94 <- elas(IPM94)
  image.plot(0:80, 0:80, t(elas.94),
             main="Elasticity", xlab="SB & Size (t)", ylab="SB & Size (t+1)")
```

### 2018-02-19 exploration.Rmd
exploring options for simulation

---
title: "Exploration"
author: "Tim Szewczyk"
date: "2/16/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pkgs <- c("tidyverse", "magrittr", "fields",
          "maxent", "gbPopMod", "IPMpack", "ppmlasso")
suppressMessages(invisible(lapply(pkgs, library, character.only=TRUE)))
theme_set(theme_bw())
```

# Exploration of options for the simulation component

## Species description: Life history, characteristics, relevant attributes
The simulated species is fully described here. The object `sp_i` is a list containing three elements: one for each SDM approach. Each of these contains elements for the global parameters, relationships with environmental variables, and demographic matrices that are applicable or potentially applicable for that modeling strategy.
```{r sp_i}
set.seed(1)
sp_i <- list(c_i=list(),
             m1_i=list(),
             m2_i=list())
```

## Landscape
The landscape characteristics are aggregated here. It is a gridded landscape with climate and LULC data.
```{r landscape}
n.rows <- 10
n.cols <- 10
n.cell <- n.rows * n.cols
env.df <- tibble(x=rep(1:n.cols, each=n.rows),
                 y=rep(1:n.rows, times=n.cols),
                 x_y=paste(x, y, sep="_"),
                 temp_mean=rnorm(n.cell),
                 temp_min=rnorm(n.cell),
                 temp_max=rnorm(n.cell),
                 prec_mean=rnorm(n.cell),
                 prec_min=rnorm(n.cell),
                 prec_max=rnorm(n.cell)) %>%
  full_join(.[,3:9], by="x_y", suffix=c("", "_sq")) %>%
  mutate_at(10:15, function(x) x*x)
```

## Correlative: MaxEnt
```{r MaxEnt}

```

## Mechanistic: CA
```{r CA}

```

## Mechanistic: IPM
```{r IPM}

```

### 2018-02-19 exploration.Rmd
simple grid of populations; no interaction among cells; env covariates possible

@@
 knitr::opts_chunk$set(echo = TRUE)
-pkgs <- c("tidyverse", "magrittr", "fields",
-          "maxent", "gbPopMod", "IPMpack", "ppmlasso")
+pkgs <- c("fields", "sevcheck", "gbPopMod", "IPMpack", "tidyverse", "magrittr")
 suppressMessages(invisible(lapply(pkgs, library, character.only=TRUE)))
@@
 set.seed(1)
+n.cell <- 10
+n_i <- rpois(10, 200)
 sp_i <- list(c_i=list(),
              m1_i=list(),
-             m2_i=list())
+             m2_i=list(p=list(s_b=c(-4, 1.3), # b1 + b2*size
+                              s_x=c(-2, -2), # b1*temp + b2*temp^2
+                              g_b=c(0.8, 1.5, -0.1), # b1 + b2*size + b3*size^2
+                              g_sig2=c(0.01, 0.35), # N(E, s1*exp(2*s2*size))
+                              fl_b=c(-6.5, 1.9, -.2, .1), # b1 + b2*size + b3*size^2 + b4*size^3
+                              seed_b=c(2.9, 0.05), # b1 + b2*size
+                              rcrt=c(1.5, 0.4), #N(mean=rcrt1, sd=rcrt2)
+                              p_est=0.02)))
 ```
@@
 ```{r landscape}
-n.rows <- 10
-n.cols <- 10
-n.cell <- n.rows * n.cols
-env.df <- tibble(x=rep(1:n.cols, each=n.rows),
-                 y=rep(1:n.rows, times=n.cols),
-                 x_y=paste(x, y, sep="_"),
+# lc.df <- read_csv("data/landcover_5km.csv") %>% 
+#   filter(!is.na(nlcd1_mean) & !is.na(bio1_mean) & 
+#            !is.na(rdLen) & !is.na(pop_sum)) %>%
+#   select(-c(9:14))
+
+# n.cell <- n.rows * n.cols
+env.df <- tibble(
+  # x=rep(1:n.cols, each=n.rows),
+  #                y=rep(1:n.rows, times=n.cols),
+  #                x_y=paste(x, y, sep="_"),
                  temp_mean=rnorm(n.cell),
@@
                  prec_min=rnorm(n.cell),
-                 prec_max=rnorm(n.cell)) %>%
-  full_join(.[,3:9], by="x_y", suffix=c("", "_sq")) %>%
-  mutate_at(10:15, function(x) x*x)
+                 prec_max=rnorm(n.cell)) #%>%
+  # full_join(.[,3:9], by="x_y", suffix=c("", "_sq")) %>%
+  # mutate_at(10:15, function(x) x*x)
 ```
@@
 ## Mechanistic: IPM
-```{r IPM}
+```{r IPM_onePop}
+# Simulated based on Merow et al 2014 Appendix A
+p <- sp_i$m2_i$p
+d <- data.frame(size=runif(n=n_i[1], min=1, max=7))
+n <- nrow(d)
+lo <- min(d$size)*0.5
+hi <- max(d$size)*1.5
+size.mx <- cbind(1, d$size, d$size^2, d$size^3)
 
+# survival
+E_surv <- antilogit(size.mx[,1:length(p$s_b)] %*% p$s_b)
+d$surv <- rbinom(n, 1, E_surv)
+plot(d$size, E_surv, ylim=c(0,1)); points(d$surv~d$size, pch=19)
+
+# growth
+E_grow <- size.mx[,1:length(p$g_b)] %*% p$g_b
+d$sizeNext <- rnorm(n, mean=E_grow, sd=p$g_sig2[1]*exp(2*p$g_sig2[2]*d$size))
+d$sizeNext <- ifelse(d$surv, d$sizeNext, NA)
+plot(d$size, E_grow, ylim=c(lo,hi)); points(d$sizeNext~d$size, pch=19)
+
+# flowering 
+E_fl <- antilogit(size.mx[,1:length(p$fl_b)] %*% p$fl_b)
+d$fl <- rbinom(n, 1, E_fl) 
+plot(d$size, E_fl, ylim=c(0,1)); points(d$fl~d$size, pch=19)
+
+# seed production
+E_seed <- exp(d$fl * size.mx[,1:length(p$seed_b)] %*% p$seed_b)
+d$seed <- rpois(n, E_seed)
+d$seed <- ifelse(d$fl, d$seed, 0)
+d$fl <- ifelse(d$surv, d$fl, NA)
+d$seed <- ifelse(d$surv, d$seed, NA)
+seed.range <- range(d$seed, na.rm=TRUE)
+plot(d$size, E_seed, ylim=seed.range); points(d$seed~d$size, pch=19, col=1+d$fl)
+
+# recruits
+rcrt.n <- sum(d$seed, na.rm=TRUE)*p$p_est
+d[(n+1):(n+rcrt.n),] <- NA
+d$sizeNext[(n+1):(n+rcrt.n)] <- rnorm(rcrt.n, p$rcrt[1], p$rcrt[2])
+hist(d$sizeNext[(n+1):(n+rcrt.n)])
 ```
+
+```{r IPM_multPop}
+p <- sp_i$m2_i$p
+d <- vector("list", length=n.cell)
+for(i in 1:n.cell) {
+  d[[i]] <- data.frame(size=runif(n=n_i[i], min=1, max=7))
+  size.mx <- cbind(1, d[[i]]$size, d[[i]]$size^2, d[[i]]$size^3)
+
+  # survival
+  E_surv <- antilogit(size.mx[,1:length(p$s_b)] %*% p$s_b +
+                        rep(as.matrix(env.df[i,1:length(p$s_x)]) %*% p$s_x, 
+                            n_i[i]))
+  d[[i]]$surv <- rbinom(n_i[i], 1, E_surv)
+  plot(d[[i]]$size, E_surv, ylim=c(0,1)); points(surv~size, data=d[[i]], pch=19)
+
+  # growth
+  E_grow <- size.mx[,1:length(p$g_b)] %*% p$g_b
+  d[[i]]$sizeNext <- rnorm(n_i[i], mean=E_grow, 
+                           sd=p$g_sig2[1]*exp(2*p$g_sig2[2]*d[[i]]$size))
+  d[[i]]$sizeNext <- ifelse(d[[i]]$surv, d[[i]]$sizeNext, NA)
+  # plot(d[[i]]$size, E_grow, ylim=c(lo,hi))
+  # points(d[[i]]$sizeNext~d[[i]]$size, pch=19)
+
+  # flowering 
+  E_fl <- antilogit(size.mx[,1:length(p$fl_b)] %*% p$fl_b)
+  d[[i]]$fl <- rbinom(n_i[i], 1, E_fl) 
+  # plot(d[[i]]$size, E_fl, ylim=c(0,1))
+  # points(d[[i]]$fl~d[[i]]$size, pch=19)
+
+  # seed production
+  E_seed <- exp(d[[i]]$fl * size.mx[,1:length(p$seed_b)] %*% p$seed_b)
+  d[[i]]$seed <- rpois(n_i[i], E_seed)
+  d[[i]]$seed <- ifelse(d[[i]]$fl, d[[i]]$seed, 0)
+  d[[i]]$fl <- ifelse(d[[i]]$surv, d[[i]]$fl, NA)
+  d[[i]]$seed <- ifelse(d[[i]]$surv, d[[i]]$seed, NA)
+  seed.range <- range(d[[i]]$seed, na.rm=TRUE)
+  # plot(d[[i]]$size, E_seed, ylim=seed.range)
+  # points(d[[i]]$seed~d[[i]]$size, pch=19, col=1+d[[i]]$fl)
+
+  # recruits
+  rcrt.n <- sum(d[[i]]$seed, na.rm=TRUE)*p$p_est
+  rcrt.i <- (n_i[i]+1):(n_i[i]+rcrt.n)
+  d[[i]][rcrt.i,] <- NA
+  d[[i]]$sizeNext[rcrt.i] <- rnorm(rcrt.n, p$rcrt[1], p$rcrt[2])
+  # hist(d[[i]]$sizeNext[rcrt.i])
+}
+```

### 2018-02-23 Merow2014_Appendix.Rmd
Appendix E - good & bad years; Appendix F - environmental covariate

@@
 ```
+
+```{r App_E.1}
+########
+## simulate IPM matrices
+########
+  dff <- generateData()
+  gr1 <- makeGrowthObj(dff, sizeNext ~ size + size2)
+  sv1 <- makeSurvObj(dff, surv ~ size + size2)
+  fv1 <- makeFecObj(dff, Transform="log")
+  Pmat <- makeIPMPmatrix(nBigMatrix=100, minSize=-3, maxSize=14,
+                         growObj=gr1, survObj=sv1, correction="constant")
+  Fmat <- makeIPMFmatrix(nBigMatrix=100, minSize=-3, maxSize=14,
+                         fecObj=fv1, correction="constant")
+  IPM <- Pmat
+  IPM@.Data <- Pmat+Fmat  # retains Pmat meshpoints, etc
+
+  w.eigen <- Re(eigen(IPM)$vectors[,1])
+  stable.dist <- w.eigen/sum(w.eigen)
+
+########
+## generate starting population
+#########
+  # size distribution
+  size0 <- runif(1000, min(IPM@meshpoints), max(IPM@meshpoints))
+  # IPM breakpoints = (mdpts - h/2):(max(mdpts) + h/2)
+  breakpoints <- c(IPM@meshpoints - (IPM@meshpoints[2] - IPM@meshpoints[1])/2,
+                   IPM@meshpoints[length(IPM@meshpoints)] +
+                     (IPM@meshpoints[2] - IPM@meshpoints[1])/2)
+  # bin size0 based on breakpoints
+  new.dist.i <- findInterval(size0, breakpoints, all.inside=TRUE)
+  loc.sizes <- table(new.dist.i)
+  new.dist.n0 <- rep(0, length(IPM@meshpoints))
+  new.dist.n0[as.numeric(names(loc.sizes))] <- loc.sizes
+
+########
+## project population forward
+########
+  tmax <- 10
+  storage <- matrix(NA, nrow=dim(Pmat)[1], ncol=tmax)
+  new.dist.n <- new.dist.n0
+  for(k in 1:tmax) {
+    new.dist.n <- IPM %*% new.dist.n
+    storage[,k] <- new.dist.n
+  }
+
+########
+## plots
+########
+  par(mfrow=c(2,2), mar=c(4,4,2,1))
+  image.plot(Pmat@meshpoints, Pmat@meshpoints, t(Pmat), 
+             main="P", xlab="Size (t)", ylab="Size (t+1)")
+  contour(Pmat@meshpoints, Pmat@meshpoints, t(Pmat), add=TRUE)
+  image.plot(Pmat@meshpoints, Pmat@meshpoints, t(Fmat), 
+             main="F", xlab="Size (t)", ylab="")
+  contour(Pmat@meshpoints, Pmat@meshpoints, t(Fmat), add=TRUE)
+  hist(size0, col="grey", xlim=range(Pmat@meshpoints), freq=FALSE,
+       xlab="Size (t=0)", ylab="", main="")
+  for(k in 1:tmax) {
+    lines(Pmat@meshpoints, storage[,k]/sum(storage[,k]), col=rgb(0,0,k/tmax,1))
+  }
+  lines(Pmat@meshpoints, stable.dist, col="red")
+  image.plot(1:tmax, Pmat@meshpoints, t(log(storage)), 
+             xlab="time", ylab="Size")
+  abline(v=1.5+(1*c(0:tmax)))
+  Re(eigen(IPM)$values[1])
+```
+
+```{r App_E.2}
+dff <- generateData()
+
+# good year
+gr1 <- makeGrowthObj(dff, Formula=sizeNext ~ size, 
+                     regType="constantVar", Family="gaussian")
+sv1 <- makeSurvObj(dff, Formula=surv ~ size + size2)
+fv1 <- makeFecObj(dff, Formula=list(fec~size), Transform="log")
+fv1@fitFec[[1]]$coefficients <- c(-1, 0.35)
+
+# bad year
+sv1 <- coerceSurvObj(survObj=sv1, coeff=c(-1, 1, 2))
+sv2 <- coerceSurvObj(survObj=sv1, coeff=c(-3, 0, -2))
+gr2 <- coerceGrowthObj(gr1, coeff=c(-3, -3), sd=1)
+
+# kernel
+Pmat1 <- makeIPMPmatrix(nBigMatrix=50, minSize=-2, maxSize=12,
+                        growObj=gr1, survObj=sv1)
+Pmat2 <- makeIPMPmatrix(nBigMatrix=50, minSize=-2, maxSize=12,
+                        growObj=gr2, survObj=sv2)
+Fmat <- makeIPMFmatrix(nBigMatrix=50, minSize=-2, maxSize=12, fecObj=fv1)
+IPM1 <- Pmat1 + Fmat
+IPM2 <- Pmat2 + Fmat
+
+# plots
+image.plot(Pmat1@meshpoints, Pmat1@meshpoints, t(Pmat1), main="good year")
+image.plot(Pmat1@meshpoints, Pmat1@meshpoints, t(Pmat2), main="bad year")
+
+exp(stochGrowthRateSampleList(list(IPM1, IPM2), nRunIn=1000, tMax=5000))
+
+nsim <- 100
+tmax <- 100
+IPM.ls <- list(IPM1, IPM2)
+nt0 <- abs(Re(eigen(IPM1)$vector[,1])*10)
+rt <- rtGood <- rtBad <- matrix(NA, nsim, tmax)
+for(j in 1:nsim) {
+  ntGood <- ntBad <- nt <- nt0
+  for(k in 1:tmax) {
+    sv.t <- coerceSurvObj(sv1, coeff=runif(3, -3, 3))
+    gr.t <- coerceGrowthObj(gr1, coeff=runif(2, -3, 3), sd=1)
+    Pmat.t <- makeIPMPmatrix(nBigMatrix=50, minSize=-2, maxSize=12,
+                        growObj=gr.t, survObj=sv.t)
+    IPM.t <- Pmat.t + Fmat
+    nt <- IPM.t %*% nt
+    rt[j,k] <- sum(nt)
+    ntGood <- IPM1 %*% ntGood
+    rtGood[j,k] <- sum(ntGood)
+    ntBad <- IPM2 %*% ntBad
+    rtBad[j,k] <- sum(ntBad)
+  }
+}
+
+par(mfrow=c(1,1), bty="l")
+matplot(t(rt), type="l", xlab="time", ylab="N", log="y", 
+        ylim=c(range(c(rtGood, rtBad))), lty=1, col=rgb(0,0,0,0.5))
+points(rtGood[1,], type="l", lwd=2, lty=2)
+points(rtBad[1,], type="l", lwd=2, lty=2)
+```
+
+```{r App_E.3}
+dff <- generateData()
+go <- makeGrowthObj(dff, sizeNext ~ size + size2 + covariate)
+so <- makeSurvObj(dff, surv ~ size + size2)
+eo <- makeEnvObj(dff)
+fo <- makeFecObj(dff, Formula=list(fec~size))
+
+Pmat <- makeCompoundPmatrix(nEnvClass=2, envMatrix=eo, 
+                            minSize=-3, maxSize=14, nBigMatrix=100,
+                            growObj=go, survObj=so, 
+                            correction="constant")
+image.plot(1:nrow(Pmat), 1:ncol(Pmat), t(log(Pmat)), axes=F, 
+           xlab="Size (t)", ylab="Size (t+1)")
+axis(1, at=1:nrow(Pmat), lab=round(rep(Pmat@meshpoints, Pmat@nEnvClass)))
+axis(2, at=1:nrow(Pmat), lab=round(rep(Pmat@meshpoints, Pmat@nEnvClass)))
+abline(h=length(Pmat@meshpoints)*(1:Pmat@nEnvClass))
+abline(v=length(Pmat@meshpoints)*(1:Pmat@nEnvClass))
+```
+
+```{r App_F}
+########
+## load data
+########
+  load("data/ActaeaRegDat.RData")
+  d <- ActaeaRegDat
+
+########
+## regressions
+########
+  g.lm <- lm(y ~ x + K, data=d)
+  s.lm <- glm(surv ~ x, family="binomial", data=d)
+  fl.lm <- glm(flow ~ x, family="binomial", data=d)
+  fr.lm <- glm(fruits ~ x, family="poisson", data=d)
+
+  par(mfrow=c(2,2), mar=c(4,4,2,1))
+  xx <- data.frame(x=seq(2, 14, length.out=100), K=25)
+  plot(d$x, d$surv, col=rgb(0,0,0,0.1), xlab="Size (t)", ylab="Survival to t+1")
+  lines(xx$x, predict(s.lm, xx, type="response"), col="red", lwd=2)
+  plot(d$x, d$y, xlab="Size (t)", ylab="Size (t+1)")
+  lines(xx$x, predict(g.lm, xx, type="response"), col="red", lwd=2)
+  plot(d$x, d$flow, col=rgb(0,0,0,0.1), xlab="Size (t)", ylab="Pr(Flower)")
+  lines(xx$x, predict(fl.lm, xx, type="response"), col="red", lwd=2)
+  plot(d$x, d$fruits, xlab="Size (t)", ylab="Fruit number")
+  lines(xx$x, predict(fr.lm, xx, type="response"), col="red", lwd=2)
+
+  p <- rep(0, 15)
+  p <- list("s.b0"=coef(s.lm)[1], "s.bx"=coef(s.lm)[2], 
+         "g.b0"=coef(g.lm)[1], "g.bx"=coef(g.lm)[2], "g.bK"=coef(g.lm)[3],
+         "fl.b0"=coef(fl.lm)[1], "fl.bx"=coef(fl.lm)[2],
+         "fr.b0"=coef(fr.lm)[1], "fr.bx"=coef(fr.lm)[2],
+         "est.x.mn"=3.0721075, "est.x.sd"=1.448899,
+         "sd.xy"=1.155625, 
+         "est.sdlg"=0.0091550, "surv.sdlg"=0.2383389, 
+         "seed.pred"=0.195905)
+
+########
+## define functions
+########
+  # survival probability
+  sx <- function(x, p) {
+    u <- exp(p$s.b0 + p$s.bx*x)
+    return(u/(1+u))
+  }
+
+  # growth
+  gyx <- function(y, x, K, p) {
+    dnorm(y, p$g.b0 + p$g.bx*x + p$g.bK*K, p$sd.xy)
+  }
+
+  # fecundity
+  fx <- function(x, p) {
+    # flowering probability
+    u <- exp(p$fl.b0 + p$fl.bx*x)
+    # fruits per flowering individual
+    n.fr <- exp(p$fr.b0 + p$fr.bx*x)
+    n.fr[n.fr<0] <- 0
+    n.seed.per.fruit <- p$seed.pred*0.38901 + (1 - p$seed.pred)*9.61368
+    return(u/(1+u)*n.fr*n.seed.per.fruit)
+  }
+
+  # seedling survival and size distribution
+  py3 <- function(y, p) {
+    return(p$surv.sdlg * dnorm(y, p$est.x.mn, p$est.x.sd)/
+             (1-pnorm(minsize, p$est.x.mn, p$est.x.sd)))
+  }
+
+  # iterator function
+  iter <- function(Nt) {
+    Nt1 <- array(0, dim=c(2+n.size)) 
+    Nt1[1] <- Nt1[1] + sum(FF %*% Nt[3:(n.size+2)])
+    Nt1 <- Nt1 + as.vector(P %*% Nt)
+    return(Nt1)
+  }
+
+########
+## construct kernel
+########
+  # define breakpoints, meshpoints, stepsize
+  n.size <- 100
+  tol <- 1.e-8
+  minsize <- 0.9234274; maxsize <- 15.27885
+  L <- minsize; U <- maxsize; n <- n.size
+  b <- L + c(0:n)*(U-L)/n
+  y <- 0.5*(b[1:n]+b[2:(n+1)])
+  h <- y[2]-y[1]
+
+  # loop over forest succession (=approx(cells))
+  spruce <- 0
+  K.conc <- 0
+  for(k in 1:120) {
+    # spruce proportion & K concentration
+    spruce <- 1/(1+exp(-0.09*k+5))
+    K.conc <- exp(3.00932 - 0.53267*spruce)*1.722324
+    # kernel components
+      ## growth (including eviction correction)
+      G <- array(0, dim=c(n.size, n.size))
+      G <- h*outer(y, y, gyx, p=p, K=K.conc)
+      for(i in 1:(n.size/2)) G[1,i] <- G[1,i] + 1 - sum(G[,i])
+      for(i in (n.size/2+1):n.size) G[n.size,i] <- G[n.size,i] + 1 - sum(G[,i])
+      ## survival
+      S <- sx(y, p)
+      FF <- fx(y, p)*h
+      P <- array(0, dim=c(2+n.size, 2+n.size)) # [1,] = seed; [2,] = seedling
+      P[2,1] <- p$est.sdlg
+      P[3:(2+n.size),2] <- py3(y, p)
+      for(i in 3:(2+n.size)) P[i,3:(2+n.size)] <- S*G[(i-2),]
+
+    # population iterations
+    Nt <- array(1, dim=c(2+n.size))
+    Nt <- Nt/sum(Nt)
+    qmax <- 1000; lam <- 1
+    while(qmax > tol) {
+      Nt1 <- iter(Nt)
+      qmax <- sum(abs(Nt1 - lam*Nt))
+      lam <- sum(Nt1)
+      Nt <- Nt1/lam
+    }
+    stable.dist <- Nt/sum(Nt)
+    stable.lam <- lam
+    cat("Year:", k, "Proportion spruce:", round(spruce,2), " Lambda:", lam, "\n")
+  }
+```

### 2018-02-23 exploration.Rmd
env data for 5km NE; IPM generative process only

--- 2018-02-19 exploration.Rmd
@@
 knitr::opts_chunk$set(echo = TRUE)
-pkgs <- c("tidyverse", "magrittr", "fields",
-          "maxent", "gbPopMod", "IPMpack", "ppmlasso")
+pkgs <- c("fields", "sevcheck", "gbPopMod", "IPMpack", "tidyverse", "magrittr")
 suppressMessages(invisible(lapply(pkgs, library, character.only=TRUE)))
 theme_set(theme_bw())
+lc.df <- read_csv("data/landcover_5km.csv") %>% 
+  filter(!is.na(bio1_mean))
 ```
@@
 The simulated species is fully described here.
-The object `sp_i` is a list containing three elements: one for each SDM approach.
-Each of these contains elements for the global parameters, relationships with environmental variables, and demographic matrices that are applicable or potentially applicable for that modeling strategy.
+The object `sp_i` is a list containing elements for the global parameters, relationships with environmental variables, and demographic matrices that are applicable or potentially applicable for that modeling strategy.
+Parameters are abbreviated with the associated kernel component (e.g., `s_` for survival, `g_` for growth, etc) and the type of covariate (`_z` for size, `_x` for environmental covariates).
 ```{r sp_i}
-set.seed(1)
-sp_i <- list(c_i=list(),
-             m1_i=list(),
-             m2_i=list())
+set.seed(10)
+mean_abund <- 600
+useLC <- TRUE
+regPlots <- TRUE
+ipmPlots <- FALSE
+n.row <- ifelse(useLC, NA, 10)
+n.col <- ifelse(useLC, NA, 20)
+n.cell <- ifelse(useLC, n_distinct(lc.df$CellID), n.row*n.col)
+n_i <- rpois(n.cell, mean_abund)
+sp_i <- list(p=list(s_z=c(-2, 1.3), # b1 + b2*z
+                    s_x=c(-.5, -.7, .5, -.1), # b1*x1 + b2*x2 + ...
+                    g_z=c(1.5, 2, -0.1), # b1 + b2*z + b3*z^2
+                    g_x=c(1, -.5, 2, -.5), #b1*x1 + b2*x2 + ...
+                    g_sig2=c(0.01, 0.35), # N(E, s1*exp(2*s2*z))
+                    fl_z=c(-6.5, 1.9, -.2, .1), # b1 + b2*z + b3*z^2
+                    fl_x=c(0, -.01, 0), # b1*x1 + ...
+                    seed_z=c(2.9, 0.05), # b1 + b2*z
+                    seed_x=c(-1, -.1, .1, -.2),
+                    rcrt=c(1.5, 0.4), #N(mean=rcrt1, sd=rcrt2)
+                    p_est=0.02))
 ```
@@
 ```{r landscape}
-n.rows <- 10
-n.cols <- 10
-n.cell <- n.rows * n.cols
-env.df <- tibble(x=rep(1:n.cols, each=n.rows),
-                 y=rep(1:n.rows, times=n.cols),
-                 x_y=paste(x, y, sep="_"),
-                 temp_mean=rnorm(n.cell),
-                 temp_min=rnorm(n.cell),
-                 temp_max=rnorm(n.cell),
-                 prec_mean=rnorm(n.cell),
-                 prec_min=rnorm(n.cell),
-                 prec_max=rnorm(n.cell)) %>%
-  full_join(.[,3:9], by="x_y", suffix=c("", "_sq")) %>%
-  mutate_at(10:15, function(x) x*x)
+if(useLC) {
+  env.df <- tibble(temp=c(scale(lc.df$bio1_mean)),
+                   temp2=temp^2,
+                   prec=c(scale(lc.df$bio12_mean)),
+                   prec2=prec^2,
+                   x=lc.df$left,
+                   y=lc.df$top)
+} else {
+  env.df <- tibble(temp=rnorm(n.cell),
+                 temp2=temp^2,
+                 prec=rnorm(n.cell),
+                 prec2=prec^2,
+                 x=rep(1:n.col, each=n.row),
+                 y=rep(1:n.row, times=n.col)) 
+}
 ```
 
-## Correlative: MaxEnt
-```{r MaxEnt}
+---
+
+## Simulation: IPM framework
+```{r S}
+# set up storage objects
+p <- sp_i$p
+d <- vector("list", length=n.cell)
+lambdas <- rep(NA, n.cell)
+
+# set up environmental covariates
+n_sz <- length(p$s_z); n_sx <- length(p$s_x)
+n_gz <- length(p$g_z); n_gx <- length(p$g_x)
+n_flz <- length(p$fl_z); n_flx <- length(p$fl_x)
+n_seedz <- length(p$seed_z); n_seedx <- length(p$seed_x)
+X.s <- as.matrix(env.df[,1:n_sx])
+X.g <- as.matrix(env.df[,1:n_gx])
+X.fl <- as.matrix(env.df[,1:n_flx])
+X.seed <- as.matrix(env.df[,1:n_seedx])
+
+# loop through cells
+for(i in 1:n.cell) {
+  d[[i]] <- data.frame(size=runif(n=n_i[i], min=1, max=7))
+  lo <- min(d[[i]]$size)*0.1
+  hi <- max(d[[i]]$size)*2
+  size.mx <- cbind(1, d[[i]]$size, d[[i]]$size^2, d[[i]]$size^3)
+
+  # Survival
+  E_surv <- antilogit(size.mx[,1:n_sz] %*% p$s_z + c(X.s[i,] %*% p$s_x))
+  d[[i]]$surv <- rbinom(n_i[i], 1, E_surv)
+
+  # Growth
+  E_grow <- size.mx[,1:n_gz] %*% p$g_z + c(X.g[i,] %*% p$g_x)
+  d[[i]]$sizeNext <- rnorm(n_i[i], mean=E_grow, 
+                           sd=p$g_sig2[1]*exp(2*p$g_sig2[2]*d[[i]]$size))
+  d[[i]]$sizeNext <- ifelse(d[[i]]$surv, d[[i]]$sizeNext, NA)
+
+  # Flowering 
+  E_fl <- antilogit(size.mx[,1:n_flz] %*% p$fl_z + c(X.fl[i,] %*% p$fl_x))
+  d[[i]]$fl <- rbinom(n_i[i], 1, E_fl)
+
+  # Seed production
+  E_seed <- d[[i]]$fl * exp((size.mx[,1:n_seedz] %*% p$seed_z + 
+                               c(X.seed[i,] %*% p$seed_x)))
+
+  d[[i]]$seed <- rpois(n_i[i], E_seed)
+  d[[i]]$seed <- ifelse(d[[i]]$fl, d[[i]]$seed, 0)
+  # d[[i]]$fl <- ifelse(d[[i]]$surv, d[[i]]$fl, NA)
+  # d[[i]]$seed <- ifelse(d[[i]]$surv, d[[i]]$seed, NA)
+  seed.range <- range(d[[i]]$seed, na.rm=TRUE)
+
+  # recruits
+  rcrt.n <- max(sum(d[[i]]$seed, na.rm=TRUE)*p$p_est, 1)
+  rcrt.i <- (n_i[i]+1):(n_i[i]+rcrt.n)
+  d[[i]][rcrt.i,] <- NA
+  d[[i]]$sizeNext[rcrt.i] <- rnorm(rcrt.n, p$rcrt[1], p$rcrt[2])
+
+  # Kernel
+  so <- makeSurvObj(d[[i]], surv ~ size + size2)
+  go <- makeGrowthObj(d[[i]], sizeNext ~ size + size2)
+  fo <- makeFecObj(d[[i]], 
+                   Formula=c(fl~size+size2+size3, seed~size), 
+                   Family=c("binomial", "poisson"),
+                   Transform=c("none", "none"))
+  P.mx <- makeIPMPmatrix(growObj=go, survObj=so, minSize=lo, maxSize=hi,
+                         correction="constant")
+  F.mx <- makeIPMFmatrix(fecObj=fo, minSize=lo, maxSize=hi, correction="constant")
+  IPM <- P.mx + F.mx
+  lambdas[i] <- Re(eigen(IPM)$values[1])
+
+  # Plots
+  if(regPlots) {
+    par(mfrow=c(3,2), mar=c(4,4,2,1)); main.i <- paste("Cell", i)
+    z.t <- d[[i]][!is.na(d[[i]]$size), ]
+    plot(surv~size, data=d[[i]], ylim=c(0,1), pch=20, col=rgb(0,0,0,0.2),
+         main=paste("Surv:", main.i), xlab="Size (t)", ylab="P(surv to t+1)")
+      points(z.t$size, E_surv, cex=0.1, col="red")
+    plot(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,0.2), 
+         main=paste("Growth:", main.i), xlab="Size (t)", ylab="Size (t+1)")
+      points(z.t$size, E_grow, cex=0.1, col="red")
+      abline(a=0, b=1, col="gray", lty=2)
+    plot(fl~size, data=z.t, pch=20, ylim=c(0,1), col=rgb(0,0,0,0.2),
+         main=paste("P(Flower:", main.i), xlab="Size (t)", ylab="P(flower)")
+      points(z.t$size, E_fl, cex=0.1, col="red")
+    plot(seed~size, data=z.t, pch=23-z.t$fl*3, col=rgb(0,0,0,0.2), 
+         main=paste("nSeeds:", main.i), xlab="Size (t)", ylab="Number of seeds")
+      points(z.t$size, E_seed, cex=0.1, col="red")
+    plot(P.mx@meshpoints, Re(eigen(IPM)$vector[,1]), ylab="", type="l", lwd=2,
+         main=paste("Stable size distribution:", main.i), xlab="Size (t)")
+    hist(d[[i]]$sizeNext[rcrt.i], main=paste("Recruit sizes:", main.i), 
+         xlab="Recruit size (t+1)", col="grey", freq=FALSE)
+      xx <- seq(0, hi*.5, length.out=100)
+      lines(xx, dnorm(xx, p$rcrt[1], p$rcrt[2]), col="red")
+  }
+  if(ipmPlots) {
+    par(mfrow=c(2,2), mar=c(4,4,2,1)); main.i <- paste("Cell", i)
+    sens.mx <- sens(IPM); elas.mx <- elas(IPM)
+    image.plot(P.mx@meshpoints, P.mx@meshpoints, t(P.mx), 
+               main=paste("P:", main.i), xlab="Size (t)", ylab="Size (t+1)")
+    image.plot(F.mx@meshpoints, F.mx@meshpoints, t(F.mx),
+               main=paste("F:", main.i), xlab="Size (t)", ylab="Size (t+1)")
+    image.plot(1:dim(sens.mx)[1], 1:dim(sens.mx)[1], t(sens.mx),
+               main=paste("Sens:", main.i), xlab="Size (t)", ylab="Size (t+1)")
+    image.plot(1:dim(elas.mx)[1], 1:dim(elas.mx)[1], t(elas.mx),
+               main=paste("Elas:", main.i), xlab="Size (t)", ylab="Size (t+1)")
+  }
+}
+lam.df <- env.df %>% add_column(lambda=lambdas)
+ggplot(lam.df, aes(x=x, y=y, fill=log(lambdas))) + geom_tile() +
+  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
+par(mfrow=c(2,2))
+plot(lambdas ~ temp, data=lam.df, log="y")
+plot(lambdas ~ temp2, data=lam.df, log="y")
+plot(lambdas ~ prec, data=lam.df, log="y")
+plot(lambdas ~ prec2, data=lam.df, log="y")
+```
+
+----
+
+## Observed datasets
+The simulated species *S* is sampled appropriately in order to fit each SDM model. The observed datasets,*O*, may be random or intentionally biased to mimic common issues with empirical data.
+### Correlative: MaxEnt
+```{r O_Mx}
+
+```
+### Mechanistic: CA
+```{r O_CA}
+
+```
+### Mechanistic: IPM
+```{r O_IPM}
 
@@
 
-## Mechanistic: CA
-```{r CA}
+---
+
+## Fitting the models
+Each observed dataset *O* is used to predict the distribution of species *S*.
+The predicted distribution *P* is compared to the known truth *S* to assess the predictive accuracy of each modeling approach, given any limitations, uncertainty, or bias imposed in *O*.
+### Correlative: MaxEnt
+```{r P_Mx}
 
 ```
-
-## Mechanistic: IPM
-```{r IPM}
+### Mechanistic: CA
+```{r P_CA}
 
 ```
+### Mechanistic: IPM
+```{r P_IPM}
+
+```

### 2018-02-23 exploration.Rmd
calculating lambda based on assigned slopes rather than from simulated data with IPMpack

--- 2018-02-19 exploration.Rmd
@@
 The simulated species is fully described here.
-The object `sp_i` is a list containing three elements: one for each SDM approach.
-Each of these contains elements for the global parameters, relationships with environmental variables, and demographic matrices that are applicable or potentially applicable for that modeling strategy.
+The object `sp_i` is a list containing elements for the global parameters, relationships with environmental variables, and demographic matrices that are applicable or potentially applicable for that modeling strategy.
+Parameters are abbreviated with the associated kernel component (e.g., `s_` for survival, `g_` for growth, etc) and the type of covariate (`_z` for size, `_x` for environmental covariates).
 ```{r sp_i}
-set.seed(1)
-sp_i <- list(c_i=list(),
-             m1_i=list(),
-             m2_i=list())
+set.seed(20)
+mean_abund <- 5000
+useLC <- TRUE
+regPlots <- T
+ipmPlots <- F
+n.row <- ifelse(useLC, NA, 10)
+n.col <- ifelse(useLC, NA, 20)
+n.cell <- ifelse(useLC, n_distinct(lc.df$CellID), n.row*n.col)
+n_i <- rpois(n.cell, mean_abund)
+sp_i <- list(p=list(s_z=c(-3, 1.3), # b1 + b2*z
+                    s_x=c(-.5), # b1*x1 + ...
+                    g_z=c(1, 2, -0.1), # b1 + b2*z + b3*z^2
+                    g_x=c(-1), #b1*x1 + ...
+                    g_sig=1, # N(E, sig)
+                    fl_z=c(-6.5, 1.9, -.2, .1), # b1 + b2*z + b3*z^2
+                    fl_x=c(0, -.01, 0), # b1*x1 + ...
+                    seed_z=c(2.9, 0.05), # b1 + b2*z
+                    seed_x=c(-1), # b1*x1 + ...
+                    rcrt=c(1.5, 0.4), #N(mean=rcrt1, sd=rcrt2)
+                    p_est=0.02))
 ```
@@
 ```{r landscape}
-n.rows <- 10
-n.cols <- 10
-n.cell <- n.rows * n.cols
-env.df <- tibble(x=rep(1:n.cols, each=n.rows),
-                 y=rep(1:n.rows, times=n.cols),
-                 x_y=paste(x, y, sep="_"),
-                 temp_mean=rnorm(n.cell),
-                 temp_min=rnorm(n.cell),
-                 temp_max=rnorm(n.cell),
-                 prec_mean=rnorm(n.cell),
-                 prec_min=rnorm(n.cell),
-                 prec_max=rnorm(n.cell)) %>%
-  full_join(.[,3:9], by="x_y", suffix=c("", "_sq")) %>%
-  mutate_at(10:15, function(x) x*x)
+if(useLC) {
+  env.df <- tibble(temp=c(scale(lc.df$bio1_mean)),
+                   temp2=temp^2,
+                   prec=c(scale(lc.df$bio12_mean)),
+                   prec2=prec^2,
+                   pOpn=c(scale(lc.df$nlcd1_mean)),
+                   pOth=c(scale(lc.df$nlcd2_mean)),
+                   pDec=c(scale(lc.df$nlcd3_mean)),
+                   pEvg=c(scale(lc.df$nlcd4_mean)),
+                   pMxd=c(scale(lc.df$nlcd5_mean)),
+                   x=lc.df$left,
+                   y=lc.df$top)
+} else {
+  env.df <- tibble(temp=rnorm(n.cell),
+                 temp2=temp^2,
+                 prec=rnorm(n.cell),
+                 prec2=prec^2,
+                 x=rep(1:n.col, each=n.row),
+                 y=rep(1:n.row, times=n.col)) 
+}
 ```
 
-## Correlative: MaxEnt
-```{r MaxEnt}
+---
+
+## Simulation: IPM framework
+```{r S}
+########
+## set up
+########
+# set up storage objects
+p <- sp_i$p
+d <- vector("list", length=n.cell)
+lambdas <- lambdas2 <- rep(NA, n.cell)
+
+# set up environmental covariates
+n_sz <- length(p$s_z); n_sx <- length(p$s_x)
+n_gz <- length(p$g_z); n_gx <- length(p$g_x)
+n_flz <- length(p$fl_z); n_flx <- length(p$fl_x)
+n_seedz <- length(p$seed_z); n_seedx <- length(p$seed_x)
+X_s <- as.matrix(env.df[,1:n_sx])
+X_g <- as.matrix(env.df[,1:n_gx])
+X_fl <- as.matrix(env.df[,1:n_flx])
+X_seed <- as.matrix(env.df[,1:n_seedx])
+
+# life history functions
+z_pow <- function(z.vec, n_z) {
+  if(n_z==1) {
+    return(matrix(1, ncol=1, nrow=length(z.vec)))
+  } else {
+    z <- matrix(z.vec, nrow=length(z.vec), ncol=n_z-1)
+    for(i in 1:(n_z-1)) z[,i] <- z[,i]^i
+    z <- cbind(1, z)
+    return(z)
+  }
+}
+calc_surv <- function(z.v, n_sz, p, X.s) {
+  z <- z_pow(z.v, n_sz)
+  u <- exp(z %*% p$s_z + c(t(X.s) %*% p$s_x))
+  return(u / (1+u))
+}
+calc_grow <- function(z1, z.v, n_gz, p, X.g) {
+  z <- z_pow(z.v, n_gz)
+  g <- dnorm(z1, mean=z %*% p$g_z + c(t(X.g) %*% p$g_x), 
+             sd=p$g_sig)
+  return(g)
+}
+calc_flwr <- function(z.v, n_flz, p, X.fl) {
+  z <- z_pow(z.v, n_flz)
+  u <- exp(z %*% p$fl_z + c(t(X.fl) %*% p$fl_x))
+  return(u / (1+u))
+}
+calc_seed <- function(z1, z.v, n_seedz, n_flz, p, X.seed, X.fl) {
+  z <- z_pow(z.v, n_seedz)
+  calc_flwr(z.v, n_flz, p, X.fl) * p$p_est *
+      dnorm(z1, p$rcrt[1], p$rcrt[2]) *
+      exp(z %*% p$seed_z + c(t(X.seed) %*% p$seed_x))
+}
+
+########
+## loop through cells
+########
+for(i in 1:n.cell) {
+  d[[i]] <- data.frame(size=runif(n=n_i[i], min=1, max=12))
+  lo <- min(d[[i]]$size)*0.5
+  hi <- max(d[[i]]$size)*1.5
+  size.mx <- cbind(1, d[[i]]$size, d[[i]]$size^2, d[[i]]$size^3)
+
+  # generate data
+  ## survival
+  E_surv <- antilogit(size.mx[,1:n_sz] %*% p$s_z + c(X_s[i,] %*% p$s_x))
+  d[[i]]$surv <- rbinom(n_i[i], 1, E_surv)
+
+  ## growth
+  E_grow <- size.mx[,1:n_gz] %*% p$g_z + c(X_g[i,] %*% p$g_x)
+  d[[i]]$sizeNext <- rnorm(n_i[i], E_grow, p$g_sig)
+  d[[i]]$sizeNext <- ifelse(d[[i]]$surv, d[[i]]$sizeNext, NA)
+
+  ## flowering 
+  E_fl <- antilogit(size.mx[,1:n_flz] %*% p$fl_z + c(X_fl[i,] %*% p$fl_x))
+  d[[i]]$fl <- rbinom(n_i[i], 1, E_fl)
+
+  ## seed production
+  E_seed <- d[[i]]$fl * exp((size.mx[,1:n_seedz] %*% p$seed_z + 
+                               c(X_seed[i,] %*% p$seed_x)))
+  d[[i]]$seed <- rpois(n_i[i], E_seed)
+  d[[i]]$seed <- ifelse(d[[i]]$fl, d[[i]]$seed, 0)
+  seed.range <- range(d[[i]]$seed, na.rm=TRUE)
+
+  ## recruits
+  rcrt.n <- max(sum(d[[i]]$seed, na.rm=TRUE)*p$p_est, 1)
+  rcrt.i <- (n_i[i]+1):(n_i[i]+rcrt.n)
+  d[[i]][rcrt.i,] <- NA
+  d[[i]]$sizeNext[rcrt.i] <- rnorm(rcrt.n, p$rcrt[1], p$rcrt[2])
+
+  # fill IPM matrices
+  ## Manual: use assigned slopes
+  n <- 50  # ncells in matrix
+  b <- lo + c(0:n)*(hi - lo)/n  # boundary points
+  y <- 0.5*(b[1:n] + b[2:(n+1)])  # mesh points
+  h <- y[2] - y[1]  # step size
+
+  G.mx <- h*outer(y, y, calc_grow, n_gz=n_gz, p=p, X.g=X_g[i,])
+  S.v <- calc_surv(y, n_sz, p, X_s[i,])
+  F.mx <- h*outer(y, y, calc_seed, n_seedz=n_seedz, p=p, n_flz=n_flz,
+                  X.seed=X_seed[i,], X.fl=X_fl[i,])
+  P.mx <- G.mx
+  for(k in 1:(n/2)) { # Fix eviction of individuals below/above size limits
+    G.mx[1,k] <- G.mx[1,k] + 1 - sum(G.mx[,k])
+    P.mx[,k] <- G.mx[,k]*S.v[k]
+  }
+  for(k in (n/2+1):n) {
+    G.mx[n,k] <- G.mx[n,k] + 1 - sum(G.mx[,k])
+    P.mx[,k] <- G.mx[,k]*S.v[k]
+  }
+  IPM <- P.mx + F.mx
+
+  ## IPMpack: estimate slopes from d[[i]]
+  so <- makeSurvObj(d[[i]], surv ~ size)
+  go <- makeGrowthObj(d[[i]], sizeNext ~ size + size2)
+  fo <- makeFecObj(d[[i]],
+                   Formula=c(fl~size+size2+size3, seed~size),
+                   Family=c("binomial", "poisson"),
+                   Transform=c("none", "none"))
+  P.mx2 <- makeIPMPmatrix(growObj=go, survObj=so, minSize=lo, maxSize=hi,
+                         correction="constant")
+  F.mx2 <- makeIPMFmatrix(fecObj=fo, minSize=lo, maxSize=hi,
+                          correction="constant")
+  IPM2 <- P.mx2 + F.mx2
+  lambdas[i] <- Re(eigen(IPM)$values[1])
+  lambdas2[i] <- Re(eigen(IPM2)$values[1])
+
+  # Plots
+  z.t <- d[[i]][!is.na(d[[i]]$size),]; main.i <- paste("Cell", i)
+  if(regPlots) {
+    par(mfrow=c(3,2), mar=c(4,4,2,1))
+    plot(surv~size, data=d[[i]], ylim=c(0,1), pch=20, col=rgb(0,0,0,0.2),
+         main=paste("Surv:", main.i), xlab="Size (t)", ylab="P(surv to t+1)")
+      points(z.t$size, E_surv, cex=0.1, col="red")
+    plot(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,0.2), 
+         main=paste("Growth:", main.i), xlab="Size (t)", ylab="Size (t+1)")
+      points(z.t$size, E_grow, cex=0.1, col="red")
+      abline(a=0, b=1, col="blue", lty=2)
+    plot(fl~size, data=z.t, pch=20, ylim=c(0,1), col=rgb(0,0,0,0.2),
+         main=paste("P(Flower:", main.i), xlab="Size (t)", ylab="P(flower)")
+      points(z.t$size, E_fl, cex=0.1, col="red")
+    plot(seed~size, data=z.t, pch=23-z.t$fl*3, col=rgb(0,0,0,0.2), 
+         main=paste("nSeeds:", main.i), xlab="Size (t)", ylab="Number of seeds")
+      points(z.t$size, E_seed, cex=0.1, col="red")
+    w1 <- Re(eigen(IPM)$vectors[,1]); stable.dist1 <- w1/sum(w1)
+    w2 <- Re(eigen(IPM2)$vectors[,1]); stable.dist2 <- w2/sum(w2)
+    plot(y, stable.dist1, ylab="", type="l", lwd=2,
+         main=paste("Stable size distribution:", main.i), xlab="Size (t)")
+    lines(y, stable.dist2, lwd=2, lty=2)
+    # plot(y, S.v); points(y, apply(P.mx, 2, sum), col="red", type="l")
+    hist(d[[i]]$sizeNext[rcrt.i], main=paste("Recruit sizes:", main.i), 
+         xlab="Recruit size (t+1)", col="grey", freq=FALSE)
+      xx <- seq(0, hi*.5, length.out=100)
+      lines(xx, dnorm(xx, p$rcrt[1], p$rcrt[2]), col="red")
+  }
+  if(ipmPlots) {
+    par(mfrow=c(2,2), mar=c(4,4,2,1)); main.i <- paste("Cell", i)
+    # sens.mx <- sens(IPM2); elas.mx <- elas(IPM2)
+    sens.mx <- sens(IPM); elas.mx <- elas(IPM)
+    image.plot(y, y, t(P.mx), main=main.i, xlab="Size (t)", ylab="Size (t+1)")
+    points(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,0.2))
+    abline(a=0, b=1, col="white", lty=2)
+    image.plot(y, y, t(P.mx2), main=main.i, xlab="Size (t)", ylab="Size (t+1)")
+      points(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,0.2))
+      abline(a=0, b=1, col="white", lty=2)
+    image.plot(1:dim(sens.mx)[1], 1:dim(sens.mx)[1], t(sens.mx),
+               main=paste("Sens:", main.i), xlab="Size (t)", ylab="Size (t+1)")
+    image.plot(1:dim(elas.mx)[1], 1:dim(elas.mx)[1], t(elas.mx),
+               main=paste("Elas:", main.i), xlab="Size (t)", ylab="Size (t+1)")
+  }
+}
+lam.df <- env.df %>% add_column(lambda=lambdas, lambda2=lambdas2)
+ggplot(lam.df, aes(x=x, y=y, fill=log(lambdas))) + geom_tile() +
+  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
+par(mfrow=c(2,2))
+plot(lambdas ~ temp, data=lam.df, log="y")
+plot(lambdas ~ temp2, data=lam.df, log="y")
+plot(lambdas ~ prec, data=lam.df, log="y")
+plot(lambdas ~ prec2, data=lam.df, log="y")
+```
+
+----
+
+## Observed datasets
+The simulated species *S* is sampled appropriately in order to fit each SDM model. The observed datasets,*O*, may be random or intentionally biased to mimic common issues with empirical data.
+### Correlative: MaxEnt
+```{r O_Mx}
+
+```
+### Mechanistic: CA
+```{r O_CA}
+
+```
+### Mechanistic: IPM
+```{r O_IPM}
 

### 2018-02-26 exploration.Rmd
random subsampling option for testing

@@
 set.seed(20)
-mean_abund <- 5000
-useLC <- TRUE
-regPlots <- T
+subsamp <- 100
+mean_abund <- 2000
+regPlots <- F
 ipmPlots <- F
-n.row <- ifelse(useLC, NA, 10)
-n.col <- ifelse(useLC, NA, 20)
-n.cell <- ifelse(useLC, n_distinct(lc.df$CellID), n.row*n.col)
+if(is.na(subsamp)) { cells <- 1:n_distinct(lc.df$CellID)
+} else { cells <- sample(1:n_distinct(lc.df$CellID), subsamp, replace=FALSE) }
+n.cell <- length(cells)
 n_i <- rpois(n.cell, mean_abund)
-sp_i <- list(p=list(s_z=c(-3, 1.3), # b1 + b2*z
-                    s_x=c(-.5), # b1*x1 + ...
-                    g_z=c(1, 2, -0.1), # b1 + b2*z + b3*z^2
-                    g_x=c(-1), #b1*x1 + ...
+sp_i <- list(p=list(s_z=c(-4, .75), # b1 + b2*z
+                    s_x=c(-1, -.1, 1, -.1), # b1*x1 + ...
+                    g_z=c(.2, 2, -0.1), # b1 + b2*z + b3*z^2
+                    g_x=c(-1, -.1, 1, -.1), #b1*x1 + ...
                     g_sig=1, # N(E, sig)
-                    fl_z=c(-6.5, 1.9, -.2, .1), # b1 + b2*z + b3*z^2
-                    fl_x=c(0, -.01, 0), # b1*x1 + ...
-                    seed_z=c(2.9, 0.05), # b1 + b2*z
-                    seed_x=c(-1), # b1*x1 + ...
+                    fl_z=c(-7.5, 1.2, -.2, .1), # b1 + b2*z + b3*z^2
+                    fl_x=c(-1, -.01, 1, -.01), # b1*x1 + ...
+                    seed_z=c(2, 0.5, -.03), # b1 + b2*z
+                    seed_x=c(-1, -.1, 1, -.1), # b1*x1 + ...
                     rcrt=c(1.5, 0.4), #N(mean=rcrt1, sd=rcrt2)
@@
 ```{r landscape}
-if(useLC) {
 env.df <- tibble(temp=c(scale(lc.df$bio1_mean)),
@@
                  y=lc.df$top)
-} else {
-  env.df <- tibble(temp=rnorm(n.cell),
-                 temp2=temp^2,
-                 prec=rnorm(n.cell),
-                 prec2=prec^2,
-                 x=rep(1:n.col, each=n.row),
-                 y=rep(1:n.row, times=n.col)) 
-}
+env.df <- env.df[cells,]
 ```
@@
   d[[i]] <- data.frame(size=runif(n=n_i[i], min=1, max=12))
-  lo <- min(d[[i]]$size)*0.5
-  hi <- max(d[[i]]$size)*1.5
+  lo <- min(d[[i]]$size)*0.8
+  hi <- max(d[[i]]$size)*1.1
   size.mx <- cbind(1, d[[i]]$size, d[[i]]$size^2, d[[i]]$size^3)
@@
   }
+  if(i %% 100 == 0) cat("Finished cell", i, "\n")
 }
@@
 plot(lambdas ~ temp, data=lam.df, log="y")
-plot(lambdas ~ temp2, data=lam.df, log="y")
 plot(lambdas ~ prec, data=lam.df, log="y")
-plot(lambdas ~ prec2, data=lam.df, log="y")
+plot(lambdas, lambdas2)
+plot(density(log(lambdas))); abline(v=0, col="red", lty=2)
 ```

### 2018-02-26 exploration.Rmd
fixed establishment probability bug in IPMpack estimates

@@
 ```{r sp_i}
-set.seed(20)
-subsamp <- 100
+#set.seed(20)
+subsamp <- NA
 mean_abund <- 2000
@@
 sp_i <- list(p=list(s_z=c(-4, .75), # b1 + b2*z
-                    s_x=c(-1, -.1, 1, -.1), # b1*x1 + ...
+                    s_x=c(1, -.1, 1, -.1), # b1*x1 + ...
                     g_z=c(.2, 2, -0.1), # b1 + b2*z + b3*z^2
-                    g_x=c(-1, -.1, 1, -.1), #b1*x1 + ...
+                    g_x=c(1, -.1, 1, -.1), #b1*x1 + ...
                     g_sig=1, # N(E, sig)
-                    fl_z=c(-7.5, 1.2, -.2, .1), # b1 + b2*z + b3*z^2
-                    fl_x=c(-1, -.01, 1, -.01), # b1*x1 + ...
-                    seed_z=c(2, 0.5, -.03), # b1 + b2*z
-                    seed_x=c(-1, -.1, 1, -.1), # b1*x1 + ...
+                    fl_z=c(-3, .6), # b1 + b2*z
+                    fl_x=c(1, -.01, 1, -.01), # b1*x1 + ...
+                    seed_z=c(2, 0.5, -.03), # b1 + b2*z + b2*z^2
+                    seed_x=c(1, -.1, 1, -.1), # b1*x1 + ...
                     rcrt=c(1.5, 0.4), #N(mean=rcrt1, sd=rcrt2)
@@
   d[[i]]$seed <- rpois(n_i[i], E_seed)
-  d[[i]]$seed <- ifelse(d[[i]]$fl, d[[i]]$seed, 0)
+  d[[i]]$seed <- ifelse(d[[i]]$fl, d[[i]]$seed, NA)
   seed.range <- range(d[[i]]$seed, na.rm=TRUE)
@@
   ## recruits
-  rcrt.n <- max(sum(d[[i]]$seed, na.rm=TRUE)*p$p_est, 1)
+  rcrt.n <- max(sum(d[[i]]$seed, na.rm=TRUE)*p$p_est, 2)
   rcrt.i <- (n_i[i]+1):(n_i[i]+rcrt.n)
@@
   fo <- makeFecObj(d[[i]],
-                   Formula=c(fl~size+size2+size3, seed~size),
+                   Formula=c(fl~size, seed~size+size2),
                    Family=c("binomial", "poisson"),
-                   Transform=c("none", "none"))
+                   Transform=c("none", "none"),
+                   fecConstants=data.frame(seedlingEstablishmentRate=p$p_est))
   P.mx2 <- makeIPMPmatrix(growObj=go, survObj=so, minSize=lo, maxSize=hi,
@@
   IPM2 <- P.mx2 + F.mx2
+
   lambdas[i] <- Re(eigen(IPM)$values[1])
@@
     par(mfrow=c(3,2), mar=c(4,4,2,1))
+    xx <- tibble(size=seq(lo, hi, length.out=100), size2=size^2, size3=size^3)
     plot(surv~size, data=d[[i]], ylim=c(0,1), pch=20, col=rgb(0,0,0,0.2),
          main=paste("Surv:", main.i), xlab="Size (t)", ylab="P(surv to t+1)")
-      points(z.t$size, E_surv, cex=0.1, col="red")
+      points(z.t$size, E_surv, cex=0.2, col="red")
+      lines(xx$size, predict(so@fit, xx, type="response"), col="blue")
     plot(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,0.2), 
          main=paste("Growth:", main.i), xlab="Size (t)", ylab="Size (t+1)")
-      points(z.t$size, E_grow, cex=0.1, col="red")
-      abline(a=0, b=1, col="blue", lty=2)
+      points(z.t$size, E_grow, cex=0.2, col="red")
+      lines(xx$size, predict(go@fit, xx, type="response"), col="blue")
+      abline(a=0, b=1, lty=2)
     plot(fl~size, data=z.t, pch=20, ylim=c(0,1), col=rgb(0,0,0,0.2),
-         main=paste("P(Flower:", main.i), xlab="Size (t)", ylab="P(flower)")
-      points(z.t$size, E_fl, cex=0.1, col="red")
+         main=paste("P(Flower):", main.i), xlab="Size (t)", ylab="P(flower)")
+      points(z.t$size, E_fl, cex=0.2, col="red")
+      lines(xx$size, predict(fo@fitFec[[1]], xx, type="response"), col="blue")
     plot(seed~size, data=z.t, pch=23-z.t$fl*3, col=rgb(0,0,0,0.2), 
          main=paste("nSeeds:", main.i), xlab="Size (t)", ylab="Number of seeds")
-      points(z.t$size, E_seed, cex=0.1, col="red")
+      points(z.t$size, E_seed, cex=0.2, col="red")
+      lines(xx$size, predict(fo@fitFec[[2]], xx, type="response"), col="blue")
     w1 <- Re(eigen(IPM)$vectors[,1]); stable.dist1 <- w1/sum(w1)
     w2 <- Re(eigen(IPM2)$vectors[,1]); stable.dist2 <- w2/sum(w2)
-    plot(y, stable.dist1, ylab="", type="l", lwd=2,
+    yl <- range(c(stable.dist1, stable.dist2))
+    plot(y, stable.dist1, ylab="", type="l", lwd=2, col="red", ylim=yl,
          main=paste("Stable size distribution:", main.i), xlab="Size (t)")
-    lines(y, stable.dist2, lwd=2, lty=2)
+    lines(y, stable.dist2, lwd=2, lty=2, col="blue")
     # plot(y, S.v); points(y, apply(P.mx, 2, sum), col="red", type="l")
@@
          xlab="Recruit size (t+1)", col="grey", freq=FALSE)
-      xx <- seq(0, hi*.5, length.out=100)
-      lines(xx, dnorm(xx, p$rcrt[1], p$rcrt[2]), col="red")
+      xxx <- seq(0, hi*.5, length.out=100)
+      lines(xxx, dnorm(xxx, p$rcrt[1], p$rcrt[2]), col="red")
   }
@@
   scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
+ggplot(lam.df, aes(x=x, y=y, fill=log(lambdas2))) + geom_tile() +
+  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
 par(mfrow=c(2,2))
-plot(lambdas ~ temp, data=lam.df, log="y")
-plot(lambdas ~ prec, data=lam.df, log="y")
-plot(lambdas, lambdas2)
+plot(log(lambdas) ~ temp, data=lam.df); abline(h=0, col="red", lty=2)
+plot(log(lambdas) ~ prec, data=lam.df); abline(h=0, col="red", lty=2)
+plot(log(lambdas), log(lambdas2)); abline(v=0, h=0, a=0, b=1, lty=c(3,2,2))
 plot(density(log(lambdas))); abline(v=0, col="red", lty=2)

### 2018-03-01 exploration.Rmd
added seedbank for theoretical and simulation version; IPMpack still to do

--- 2018-02-26 exploration.Rmd
@@
 The object `sp_i` is a list containing elements for the global parameters, relationships with environmental variables, and demographic matrices that are applicable or potentially applicable for that modeling strategy.
-Parameters are abbreviated with the associated kernel component (e.g., `s_` for survival, `g_` for growth, etc) and the type of covariate (`_z` for size, `_x` for environmental covariates).
+Parameters are abbreviated with the associated kernel component (e.g., `s_*` for survival, `g_*` for growth, etc) and the type of covariate (`*_z` for size, `*_x` for environmental covariates).
 ```{r sp_i}
-set.seed(20)
-subsamp <- 100
-mean_abund <- 2000
+#set.seed(20)
+subsamp <- NA
+mean_abund <- 500
 regPlots <- F
@@
 n_i <- rpois(n.cell, mean_abund)
-sp_i <- list(p=list(s_z=c(-4, .75), # b1 + b2*z
-                    s_x=c(-1, -.1, 1, -.1), # b1*x1 + ...
+sp_i <- list(p=list(s_z=c(-1.75, .75), # b1 + b2*z
+                    s_x=c(1, -.1, -.5, -.1), # b1*x1 + ...
                     g_z=c(.2, 2, -0.1), # b1 + b2*z + b3*z^2
-                    g_x=c(-1, -.1, 1, -.1), #b1*x1 + ...
-                    g_sig=1, # N(E, sig)
-                    fl_z=c(-7.5, 1.2, -.2, .1), # b1 + b2*z + b3*z^2
-                    fl_x=c(-1, -.01, 1, -.01), # b1*x1 + ...
-                    seed_z=c(2, 0.5, -.03), # b1 + b2*z
-                    seed_x=c(-1, -.1, 1, -.1), # b1*x1 + ...
-                    rcrt=c(1.5, 0.4), #N(mean=rcrt1, sd=rcrt2)
-                    p_est=0.02))
+                    g_x=c(1, -.1, 1.5, -.1), #b1*x1 + ...
+                    g_sig=1, # growth ~ N(E, g_sig)
+                    fl_z=c(-1.5, .1, .1), # b1 + b2*z + b3*z^2 + b4*z^3
+                    fl_x=c(1, -.01, -1, -.01), # b1*x1 + ...
+                    seed_z=c(2, 0.5, -.03), # b1 + b2*z + b2*z^2
+                    seed_x=c(1, -.1, 1, -.1), # b1*x1 + ...
+                    rcr_z=c(1.5, 0.4), # N(mean=rcrt1, sd=rcrt2)
+                    p_est=0.02, # p(establishment)
+                    rcr_SB=0.4, # p(recruit from seedbank)
+                    rcr_dir=0.5, # p(recruit directly)
+                    s_SB=0.3 # p(survive in seedbank additional year)
+                    ))
 ```
@@
 d <- vector("list", length=n.cell)
-lambdas <- lambdas2 <- rep(NA, n.cell)
+lambdas <- lambdas2 <- B_1 <- rep(NA, n.cell)
+B <- rpois(n.cell, n_i)
 
@@
 }
-calc_surv <- function(z.v, n_sz, p, X.s) {
+calc_surv <- function(z.v, p, n_sz, X.s) {
   z <- z_pow(z.v, n_sz)
@@
 }
-calc_grow <- function(z1, z.v, n_gz, p, X.g) {
+calc_grow <- function(z1, z.v, p, n_gz, X.g) {
   z <- z_pow(z.v, n_gz)
@@
 }
-calc_flwr <- function(z.v, n_flz, p, X.fl) {
+calc_flwr <- function(z.v, p, n_flz, X.fl) {
   z <- z_pow(z.v, n_flz)
@@
 }
-calc_seed <- function(z1, z.v, n_seedz, n_flz, p, X.seed, X.fl) {
+calc_rcrDir <- function(z1, z.v, p, n_seedz, n_flz, X.seed, X.fl) {
   z <- z_pow(z.v, n_seedz)
-  calc_flwr(z.v, n_flz, p, X.fl) * p$p_est *
-      dnorm(z1, p$rcrt[1], p$rcrt[2]) *
-      exp(z %*% p$seed_z + c(t(X.seed) %*% p$seed_x))
+  calc_flwr(z.v, p, n_flz, X.fl) *
+    exp(z %*% p$seed_z + c(t(X.seed) %*% p$seed_x)) *
+    p$rcr_dir * 
+    p$p_est * 
+    dnorm(z1, p$rcr_z[1], p$rcr_z[2])
+}
+calc_rcrSB <- function(z1, p) {
+  p$s_SB * dnorm(z1, p$rcr_z[1], p$rcr_z[2])
+}
+calc_addSB <- function(z.v, p, n_seedz, n_flz, X.seed, X.fl) {
+  z <- z_pow(z.v, n_seedz)
+  calc_flwr(z.v, p, n_flz, X.fl) *
+    exp(z %*% p$seed_z + c(t(X.seed) %*% p$seed_x)) *
+    (1 - p$rcr_dir) *
+    p$s_SB
+}
+calc_staySB <- function(p) {
+  p$s_SB * (1 - p$rcr_SB)
 }
@@
   ## seed production
-  E_seed <- d[[i]]$fl * exp((size.mx[,1:n_seedz] %*% p$seed_z + 
-                               c(X_seed[i,] %*% p$seed_x)))
+  E_seed <- d[[i]]$fl * exp((size.mx[,1:n_seedz] %*% p$seed_z 
+                             + c(X_seed[i,] %*% p$seed_x)))
   d[[i]]$seed <- rpois(n_i[i], E_seed)
-  d[[i]]$seed <- ifelse(d[[i]]$fl, d[[i]]$seed, 0)
-  seed.range <- range(d[[i]]$seed, na.rm=TRUE)
+  d[[i]]$fl <- ifelse(d[[i]]$surv, d[[i]]$fl, NA)
+  d[[i]]$seed <- ifelse(d[[i]]$fl==1, d[[i]]$seed, NA)
+
+  ## seed bank
+  B_1[i] <- B[i] * p$s_SB * (1 - p$rcr_SB) + 
+    sum(d[[i]]$seed, na.rm=TRUE) * (1 - p$rcr_dir) * p$s_SB
 
   ## recruits
-  rcrt.n <- max(sum(d[[i]]$seed, na.rm=TRUE)*p$p_est, 1)
+  rcrt.n <- max((sum(d[[i]]$seed, na.rm=TRUE) * p$rcr_dir +
+                   B[i] * p$rcr_SB) * p$p_est, 
+                2) # need ≥ 2 for IPMpack estimate of rcr_z'_mn, rcr_z'_sd
   rcrt.i <- (n_i[i]+1):(n_i[i]+rcrt.n)
   d[[i]][rcrt.i,] <- NA
-  d[[i]]$sizeNext[rcrt.i] <- rnorm(rcrt.n, p$rcrt[1], p$rcrt[2])
+  d[[i]]$sizeNext[rcrt.i] <- rnorm(rcrt.n, p$rcr_z[1], p$rcr_z[2])
 
@@
   h <- y[2] - y[1]  # step size
+  ### growth and survival
+  G.mx <- h*outer(y, y, calc_grow, p=p, n_gz=n_gz, X.g=X_g[i,])
+  S.v <- calc_surv(y, p=p, n_sz=n_sz, X.s=X_s[i,])
+  ### Fix eviction of individuals below/above size limits
+  for(k in 1:(n/2)) G.mx[1,k] <- G.mx[1,k] + 1 - sum(G.mx[,k])
+  for(k in (n/2+1):n) G.mx[n,k] <- G.mx[n,k] + 1 - sum(G.mx[,k])
+  ### fill P matrix
+  P.mx <- F.mx <- matrix(0, ncol=n+1, nrow=n+1)
+  for(k in 2:(n+1)) P.mx[k,2:(n+1)] <- G.mx[k-1,]*S.v
+  P.mx[1,1] <- calc_staySB(p)
 
-  G.mx <- h*outer(y, y, calc_grow, n_gz=n_gz, p=p, X.g=X_g[i,])
-  S.v <- calc_surv(y, n_sz, p, X_s[i,])
-  F.mx <- h*outer(y, y, calc_seed, n_seedz=n_seedz, p=p, n_flz=n_flz,
+  ### fecundity
+  F.mx[2:(n+1), 2:(n+1)] <- h*outer(y, y, calc_rcrDir, p=p, 
+                                    n_seedz=n_seedz, n_flz=n_flz, 
                                     X.seed=X_seed[i,], X.fl=X_fl[i,])
-  P.mx <- G.mx
-  for(k in 1:(n/2)) { # Fix eviction of individuals below/above size limits
-    G.mx[1,k] <- G.mx[1,k] + 1 - sum(G.mx[,k])
-    P.mx[,k] <- G.mx[,k]*S.v[k]
-  }
-  for(k in (n/2+1):n) {
-    G.mx[n,k] <- G.mx[n,k] + 1 - sum(G.mx[,k])
-    P.mx[,k] <- G.mx[,k]*S.v[k]
-  }
+  F.mx[1,2:(n+1)] <- calc_addSB(y, p=p, n_seedz=n_seedz, n_flz=n_flz, 
+                                X.seed=X_seed[i,], X.fl=X_fl[i,])
+  F.mx[2:(n+1),1] <- calc_rcrSB(y, p)
+
+  ### final K
   IPM <- P.mx + F.mx
 
-  ## IPMpack: estimate slopes from d[[i]]
+  # ## IPMpack: estimate slopes from d[[i]]
+  # ### set up seedbank transitions
+  # d[[i]]$stage <- d[[i]]$stageNext <- "continuous"
+  # d[[i]]$stage[is.na(d[[i]]$size)] <- NA
+  # d[[i]]$stageNext[d[[i]]$surv==0] <- "dead"
+  # d[[i]]$number <- 1 
+  # seedbank <- data.frame(size=NA, surv=1, sizeNext=NA,
+  #                        fl=NA, seed=NA, 
+  #                        stage=c("seedbank", "seedbank", "continuous"),
+  #                        stageNext=c("seedbank", "continuous", "seedbank"),
+  #                        number=c(p$s_SB, p$rcr_SB*p$p_est, (1-p$rcr_dir)))
+  # d[[i]] <- rbind(d[[i]], seedbank)
+  # d[[i]]$stage <- as.factor(d[[i]]$stage)
+  # d[[i]]$stageNext <- as.factor(d[[i]]$stageNext)
+  ### construct kernels
   so <- makeSurvObj(d[[i]], surv ~ size)
@@
   fo <- makeFecObj(d[[i]],
-                   Formula=c(fl~size+size2+size3, seed~size),
+                   Formula=c(fl~size+size2, seed~size+size2),
                    Family=c("binomial", "poisson"),
-                   Transform=c("none", "none"))
-  P.mx2 <- makeIPMPmatrix(growObj=go, survObj=so, minSize=lo, maxSize=hi,
-                         correction="constant")
+                   Transform=c("none", "none"),
+                   # meanOffspringSize=mean(
+                   #   d[[i]]$sizeNext[is.na(d[[i]]$size) &
+                   #                     !is.na(d[[i]]$sizeNext)]),
+                   # sdOffspringSize=sd(
+                   #   d[[i]]$sizeNext[is.na(d[[i]]$size) &
+                   #                     !is.na(d[[i]]$sizeNext)]),
+                   fecConstants=data.frame(seedlingEstablishmentRate=p$p_est)#,
+                   # offspringSplitter=data.frame(seedbank=(1-p$rcr_dir),
+                   #                              continuous=p$rcr_dir)
+                   )
   F.mx2 <- makeIPMFmatrix(fecObj=fo, minSize=lo, maxSize=hi,
                           correction="constant")
+  # D.mx2 <- makeDiscreteTrans(d[[i]])
+  # dummy <- as.matrix(fo@offspringRel$coefficients[1], list(1, "seedbank"))
+  # D.mx2@meanToCont <- as.matrix(dummy, dimnames=c(1, "seedbank"))
+  # dummy <- as.matrix(fo@sdOffspringSize, dimnames=list(1, "seedbank"))
+  # D.mx2@sdToCont <- as.matrix(dummy, dimnames=c(1, "seedbank"))
+  # D.mx2@discreteTrans[1,1] <- p$s_SB 
+  # D.mx2@discreteTrans[3,1] <- 1 - sum(D.mx2@discreteTrans[1:2,1])
+  P.mx2 <- makeIPMPmatrix(growObj=go, survObj=so, #discreteTrans=D.mx2,
+                            minSize=lo, maxSize=hi, correction="constant")
   IPM2 <- P.mx2 + F.mx2
+
   lambdas[i] <- Re(eigen(IPM)$values[1])
@@
     par(mfrow=c(3,2), mar=c(4,4,2,1))
+    xx <- tibble(size=seq(lo, hi, length.out=100), size2=size^2, size3=size^3)
     plot(surv~size, data=d[[i]], ylim=c(0,1), pch=20, col=rgb(0,0,0,0.2),
          main=paste("Surv:", main.i), xlab="Size (t)", ylab="P(surv to t+1)")
-      points(z.t$size, E_surv, cex=0.1, col="red")
+      points(z.t$size, E_surv, cex=0.2, col="red")
+      lines(xx$size, predict(so@fit, xx, type="response"), col="blue")
     plot(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,0.2), 
          main=paste("Growth:", main.i), xlab="Size (t)", ylab="Size (t+1)")
-      points(z.t$size, E_grow, cex=0.1, col="red")
-      abline(a=0, b=1, col="blue", lty=2)
+      points(z.t$size, E_grow, cex=0.2, col="red")
+      lines(xx$size, predict(go@fit, xx, type="response"), col="blue")
+      abline(a=0, b=1, lty=2)
     plot(fl~size, data=z.t, pch=20, ylim=c(0,1), col=rgb(0,0,0,0.2),
-         main=paste("P(Flower:", main.i), xlab="Size (t)", ylab="P(flower)")
-      points(z.t$size, E_fl, cex=0.1, col="red")
+         main=paste("P(Flower):", main.i), xlab="Size (t)", ylab="P(flower)")
+      points(z.t$size, E_fl, cex=0.2, col="red")
+      lines(xx$size, predict(fo@fitFec[[1]], xx, type="response"), col="blue")
     plot(seed~size, data=z.t, pch=23-z.t$fl*3, col=rgb(0,0,0,0.2), 
          main=paste("nSeeds:", main.i), xlab="Size (t)", ylab="Number of seeds")
-      points(z.t$size, E_seed, cex=0.1, col="red")
+      points(z.t$size, E_seed, cex=0.2, col="red")
+      lines(xx$size, predict(fo@fitFec[[2]], xx, type="response"), col="blue")
     w1 <- Re(eigen(IPM)$vectors[,1]); stable.dist1 <- w1/sum(w1)
     w2 <- Re(eigen(IPM2)$vectors[,1]); stable.dist2 <- w2/sum(w2)
-    plot(y, stable.dist1, ylab="", type="l", lwd=2,
+    yl <- range(c(stable.dist1, stable.dist2))
+    plot(y, stable.dist1, ylab="", type="l", lwd=2, col="red", ylim=yl,
          main=paste("Stable size distribution:", main.i), xlab="Size (t)")
-    lines(y, stable.dist2, lwd=2, lty=2)
+    lines(y, stable.dist2, col="blue")
     # plot(y, S.v); points(y, apply(P.mx, 2, sum), col="red", type="l")
@@
          xlab="Recruit size (t+1)", col="grey", freq=FALSE)
-      xx <- seq(0, hi*.5, length.out=100)
-      lines(xx, dnorm(xx, p$rcrt[1], p$rcrt[2]), col="red")
+      xxx <- seq(0, hi*.5, length.out=100)
+      rcrt.z <- d[[i]]$sizeNext[is.na(d[[i]]$size) & !is.na(d[[i]]$sizeNext)]
+      lines(xxx, dnorm(xxx, p$rcr_z[1], p$rcr_z[2]), col="red", lwd=2)
+      lines(xxx, dnorm(xxx, mean(rcrt.z), sd(rcrt.z)), col="blue")
   }
@@
     par(mfrow=c(2,2), mar=c(4,4,2,1)); main.i <- paste("Cell", i)
+    if(length(y) != nrow(IPM)) {
+      x <- c(0, y)
+    } else {
+      x <- y
+    }
     # sens.mx <- sens(IPM2); elas.mx <- elas(IPM2)
     sens.mx <- sens(IPM); elas.mx <- elas(IPM)
-    image.plot(y, y, t(P.mx), main=main.i, xlab="Size (t)", ylab="Size (t+1)")
+    image.plot(x, x, t(IPM), main=main.i, xlab="Size (t)", ylab="Size (t+1)")
       points(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,0.2))
       abline(a=0, b=1, col="white", lty=2)
-    image.plot(y, y, t(P.mx2), main=main.i, xlab="Size (t)", ylab="Size (t+1)")
+    image.plot(P.mx2@meshpoints, P.mx2@meshpoints, t(IPM2), 
+               main=main.i, xlab="Size (t)", ylab="Size (t+1)")
       points(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,0.2))
       abline(a=0, b=1, col="white", lty=2)
-    image.plot(1:dim(sens.mx)[1], 1:dim(sens.mx)[1], t(sens.mx),
+    image.plot(x, x, t(sens.mx),
                main=paste("Sens:", main.i), xlab="Size (t)", ylab="Size (t+1)")
-    image.plot(1:dim(elas.mx)[1], 1:dim(elas.mx)[1], t(elas.mx),
+    image.plot(x, x, t(elas.mx),
                main=paste("Elas:", main.i), xlab="Size (t)", ylab="Size (t+1)")
@@
 ggplot(lam.df, aes(x=x, y=y, fill=log(lambdas))) + geom_tile() +
-  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
+  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", 
+                       midpoint=0, limits=log(range(c(lambdas, lambdas2))))
+ggplot(lam.df, aes(x=x, y=y, fill=log(lambdas2))) + geom_tile() +
+  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", 
+                       midpoint=0, limits=log(range(c(lambdas, lambdas2))))
 par(mfrow=c(2,2))
-plot(lambdas ~ temp, data=lam.df, log="y")
-plot(lambdas ~ prec, data=lam.df, log="y")
-plot(lambdas, lambdas2)
+plot(log(lambdas) ~ temp, data=lam.df); abline(h=0, col="red", lty=2)
+plot(log(lambdas) ~ prec, data=lam.df); abline(h=0, col="red", lty=2)
+plot(log(lambdas), log(lambdas2)); abline(v=0, h=0, a=0, b=1, lty=c(3,2,2))
 plot(density(log(lambdas))); abline(v=0, col="red", lty=2)

### 2018-03-01 exploration.Rmd
added IPMpack seed bank

@@
 #set.seed(20)
-subsamp <- NA
+subsamp <- 20
 mean_abund <- 500
-regPlots <- F
-ipmPlots <- F
+regPlots <- T
+ipmPlots <- T
 if(is.na(subsamp)) { cells <- 1:n_distinct(lc.df$CellID)
@@
                     p_est=0.02, # p(establishment)
-                    rcr_SB=0.4, # p(recruit from seedbank)
-                    rcr_dir=0.5, # p(recruit directly)
-                    s_SB=0.3 # p(survive in seedbank additional year)
+                    rcr_SB=0.01, # p(recruit from seedbank)
+                    rcr_dir=0.95, # p(recruit directly)
+                    s_SB=0.01 # p(survive in seedbank additional year)
                     ))
@@
 calc_rcrSB <- function(z1, p) {
-  p$s_SB * dnorm(z1, p$rcr_z[1], p$rcr_z[2])
+  p$s_SB * p$p_est * dnorm(z1, p$rcr_z[1], p$rcr_z[2])
 }
@@
   d[[i]] <- data.frame(size=runif(n=n_i[i], min=1, max=12))
-  lo <- min(d[[i]]$size)*0.8
-  hi <- max(d[[i]]$size)*1.1
+  lo <- min(d[[i]]$size)*0.75
+  hi <- max(d[[i]]$size)*1.25
   size.mx <- cbind(1, d[[i]]$size, d[[i]]$size^2, d[[i]]$size^3)
@@
   ## recruits
-  rcrt.n <- max((sum(d[[i]]$seed, na.rm=TRUE) * p$rcr_dir +
-                   B[i] * p$rcr_SB) * p$p_est, 
+  rcrt.n <- max((sum(d[[i]]$seed, na.rm=TRUE) * p$rcr_dir * p$p_est +
+                   B[i] * p$rcr_SB * p$p_est), 
                 2) # need ≥ 2 for IPMpack estimate of rcr_z'_mn, rcr_z'_sd
@@
   # ### set up seedbank transitions
-  # d[[i]]$stage <- d[[i]]$stageNext <- "continuous"
-  # d[[i]]$stage[is.na(d[[i]]$size)] <- NA
-  # d[[i]]$stageNext[d[[i]]$surv==0] <- "dead"
-  # d[[i]]$number <- 1 
-  # seedbank <- data.frame(size=NA, surv=1, sizeNext=NA,
-  #                        fl=NA, seed=NA, 
-  #                        stage=c("seedbank", "seedbank", "continuous"),
-  #                        stageNext=c("seedbank", "continuous", "seedbank"),
-  #                        number=c(p$s_SB, p$rcr_SB*p$p_est, (1-p$rcr_dir)))
-  # d[[i]] <- rbind(d[[i]], seedbank)
-  # d[[i]]$stage <- as.factor(d[[i]]$stage)
-  # d[[i]]$stageNext <- as.factor(d[[i]]$stageNext)
+  d[[i]]$stage <- d[[i]]$stageNext <- "continuous"
+  d[[i]]$stage[is.na(d[[i]]$size)] <- NA
+  d[[i]]$stageNext[d[[i]]$surv==0] <- "dead"
+  d[[i]]$number <- 1
+  seedbank <- data.frame(size=NA, surv=1, sizeNext=NA,
+                         fl=NA, seed=NA,
+                         stage=c("seedbank", "seedbank", "continuous"),
+                         stageNext=c("seedbank", "continuous", "seedbank"),
+                         number=c(p$s_SB * (1 - p$rcr_SB), 
+                                  p$rcr_SB * p$p_est, 
+                                  (1 - p$rcr_dir) * p$s_SB))
+  d[[i]] <- rbind(d[[i]], seedbank)
+  d[[i]]$stage <- as.factor(d[[i]]$stage)
+  d[[i]]$stageNext <- as.factor(d[[i]]$stageNext)
   ### construct kernels
@@
                    Transform=c("none", "none"),
-                   # meanOffspringSize=mean(
-                   #   d[[i]]$sizeNext[is.na(d[[i]]$size) &
-                   #                     !is.na(d[[i]]$sizeNext)]),
-                   # sdOffspringSize=sd(
-                   #   d[[i]]$sizeNext[is.na(d[[i]]$size) &
-                   #                     !is.na(d[[i]]$sizeNext)]),
-                   fecConstants=data.frame(seedlingEstablishmentRate=p$p_est)#,
-                   # offspringSplitter=data.frame(seedbank=(1-p$rcr_dir),
-                   #                              continuous=p$rcr_dir)
+                   meanOffspringSize=mean(
+                     d[[i]]$sizeNext[is.na(d[[i]]$size) &
+                                       !is.na(d[[i]]$sizeNext)]),
+                   sdOffspringSize=sd(
+                     d[[i]]$sizeNext[is.na(d[[i]]$size) &
+                                       !is.na(d[[i]]$sizeNext)]),
+                   fecConstants=data.frame(seedlingEstablishmentRate=p$p_est),
+                   offspringSplitter=data.frame(seedbank=(1-p$rcr_dir),
+                                                continuous=p$rcr_dir)
                    )
@@
                           correction="constant")
-  # D.mx2 <- makeDiscreteTrans(d[[i]])
-  # dummy <- as.matrix(fo@offspringRel$coefficients[1], list(1, "seedbank"))
-  # D.mx2@meanToCont <- as.matrix(dummy, dimnames=c(1, "seedbank"))
-  # dummy <- as.matrix(fo@sdOffspringSize, dimnames=list(1, "seedbank"))
-  # D.mx2@sdToCont <- as.matrix(dummy, dimnames=c(1, "seedbank"))
-  # D.mx2@discreteTrans[1,1] <- p$s_SB 
-  # D.mx2@discreteTrans[3,1] <- 1 - sum(D.mx2@discreteTrans[1:2,1])
-  P.mx2 <- makeIPMPmatrix(growObj=go, survObj=so, #discreteTrans=D.mx2,
+  D.mx2 <- makeDiscreteTrans(d[[i]], 
+                             continuousToDiscreteExplanatoryVariables=1)
+  dummy <- as.matrix(fo@offspringRel$coefficients[1], list(1, "seedbank"))
+  D.mx2@meanToCont <- as.matrix(dummy, dimnames=c(1, "seedbank"))
+  dummy <- as.matrix(fo@sdOffspringSize, dimnames=list(1, "seedbank"))
+  D.mx2@sdToCont <- as.matrix(dummy, dimnames=c(1, "seedbank"))
+  D.mx2@discreteTrans[1,1] <- p$s_SB * (1 - p$rcr_SB)
+  D.mx2@discreteTrans[3,1] <- 1 - sum(D.mx2@discreteTrans[1:2,1])
+  P.mx2 <- makeIPMPmatrix(growObj=go, survObj=so, discreteTrans=D.mx2,
                             minSize=lo, maxSize=hi, correction="constant")
@@
       lines(xx$size, predict(fo@fitFec[[2]], xx, type="response"), col="blue")
-    w1 <- Re(eigen(IPM)$vectors[,1]); stable.dist1 <- w1/sum(w1)
-    w2 <- Re(eigen(IPM2)$vectors[,1]); stable.dist2 <- w2/sum(w2)
+    w1 <- Re(eigen(IPM)$vectors[,1])[-1]; stable.dist1 <- w1/sum(w1)
+    w2 <- Re(eigen(IPM2)$vectors[,1])[-1]; stable.dist2 <- w2/sum(w2)
     yl <- range(c(stable.dist1, stable.dist2))
@@
     lines(y, stable.dist2, col="blue")
-    # plot(y, S.v); points(y, apply(P.mx, 2, sum), col="red", type="l")
     hist(d[[i]]$sizeNext[rcrt.i], main=paste("Recruit sizes:", main.i), 
@@
     }
-    # sens.mx <- sens(IPM2); elas.mx <- elas(IPM2)
+    z.t$z_int <- findInterval(z.t$size, y, all.inside=TRUE)
+    z.t$y <- y[z.t$z_int]
+    totSeeds <- z.t %>% group_by(y) %>% summarise(seed=sum(seed, na.rm=TRUE))
     sens.mx <- sens(IPM); elas.mx <- elas(IPM)
-    image.plot(x, x, t(IPM), main=main.i, xlab="Size (t)", ylab="Size (t+1)")
+    image.plot(x, x, t(P.mx), main=paste("P:", main.i), 
+               xlab="Size (t)", ylab="Size (t+1)")
       points(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,0.2))
       abline(a=0, b=1, col="white", lty=2)
-    image.plot(P.mx2@meshpoints, P.mx2@meshpoints, t(IPM2), 
-               main=main.i, xlab="Size (t)", ylab="Size (t+1)")
-      points(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,0.2))
-      abline(a=0, b=1, col="white", lty=2)
+    image.plot(x, x, t(F.mx), main=paste("F:", main.i), 
+               xlab="Size (t)", ylab="Size (t+1)")
+      points(totSeeds$y, totSeeds$seed/max(totSeeds$seed)*hi*.9, pch=19)
     image.plot(x, x, t(sens.mx),

### 2018-03-01 code/fn_IPM.R
moved functions to separate scripts

@@ -1,0 +1,100 @@
+# IPM vital rate functions
+# Comparison of SDM approaches
+# Tim Szewczyk
+
+# This script contains functions used for constructing an IPM matrix. All 
+# coefficients and probabilities must be estimated separately and used as 
+# inputs for these functions. These should be contained within the list `p`,
+# with a scalar or vector for each set of covariates, named as:
+#
+# p=list(s_z=c(-1.75, .75),  # b1 + b2*z + ...
+#        s_x=c(1, -.1, -.5, -.1),  # b1*x1 + ...
+#        g_z=c(.2, 2, -0.1),  # b1 + b2*z + b3*z^2 + ...
+#        g_x=c(1, -.1, 1.5, -.1),  #b1*x1 + ...
+#        g_sig=1,  # growth ~ N(E, g_sig)
+#        fl_z=c(-1.5, .1, .1),  # b1 + b2*z + b3*z^2 + b4*z^3 + ...
+#        fl_x=c(1, -.01, -1, -.01),  # b1*x1 + ...
+#        seed_z=c(2, 0.5, -.03),  # b1 + b2*z + b2*z^2 + ...
+#        seed_x=c(1, -.1, 1, -.1),  # b1*x1 + ...
+#        rcr_z=c(1.5, 0.4),  # N(mean=rcrt1, sd=rcrt2)
+#        p_est=0.02,  # p(establishment)
+#        rcr_SB=0.2,  # p(recruit from seedbank)
+#        rcr_dir=0.6,  # p(recruit directly)
+#        s_SB=0.3  # p(survive in seedbank additional year)
+# )
+#
+# The function `z_pow()` takes a vector of sizes (z) and the number of size 
+# slopes for each function, returning a design matrix with the correct number 
+# of columns where: z.mx[,1]=1; z.mx[,2]=z; z.mx[,3]=z^2; z.mx[,4]=z^3; etc
+
+
+##-- Survival
+##   P(s) ~ antilogit(size + environment)
+calc_surv <- function(z.v, p, n_sz, X.s) {
+  z <- z_pow(z.v, n_sz)
+  u <- exp(z %*% p$s_z + c(t(X.s) %*% p$s_x))
[…]

### 2018-03-01 code/fn_aux.R

@@ -1,0 +1,16 @@
+# Auxilliary functions
+# Comparison of SDM approaches
+# Tim Szewczyk
+
+
+##-- create design matrix from vector of sizes
+z_pow <- function(z.vec, n_z) {
+  if(n_z==1) {
+    return(matrix(1, ncol=1, nrow=length(z.vec)))
+  } else {
+    z <- matrix(z.vec, nrow=length(z.vec), ncol=n_z-1)
+    for(i in 1:(n_z-1)) z[,i] <- z[,i]^i
+    z <- cbind(1, z)
+    return(z)
+  }
+}

### 2018-03-01 exploration.Rmd
moved functions to separate script; cleaned code

--- 2018-02-26 exploration.Rmd
@@
 theme_set(theme_bw())
+source("code/fn_IPM.R"); source("code/fn_aux.R")
 lc.df <- read_csv("data/landcover_5km.csv") %>% 
@@
 The object `sp_i` is a list containing elements for the global parameters, relationships with environmental variables, and demographic matrices that are applicable or potentially applicable for that modeling strategy.
-Parameters are abbreviated with the associated kernel component (e.g., `s_` for survival, `g_` for growth, etc) and the type of covariate (`_z` for size, `_x` for environmental covariates).
+Parameters are abbreviated with the associated kernel component (e.g., `s_*` for survival, `g_*` for growth, etc) and the type of covariate (`*_z` for size, `*_x` for environmental covariates).
 ```{r sp_i}
-#set.seed(20)
 subsamp <- NA
-mean_abund <- 2000
+mean_abund <- 200
 regPlots <- F
 ipmPlots <- F
+n <- 100  # ncells in IPM matrix
 if(is.na(subsamp)) { cells <- 1:n_distinct(lc.df$CellID)
@@
 n_i <- rpois(n.cell, mean_abund)
-sp_i <- list(p=list(s_z=c(-4, .75), # b1 + b2*z
-                    s_x=c(1, -.1, 1, -.1), # b1*x1 + ...
+sp_i <- list(p=list(s_z=c(-1.75, .75), # b1 + b2*z
+                    s_x=c(1, -.1, -.5, -.1, .05), # b1*x1 + ...
                     g_z=c(.2, 2, -0.1), # b1 + b2*z + b3*z^2
-                    g_x=c(1, -.1, 1, -.1), #b1*x1 + ...
-                    g_sig=1, # N(E, sig)
-                    fl_z=c(-3, .6), # b1 + b2*z
-                    fl_x=c(1, -.01, 1, -.01), # b1*x1 + ...
+                    g_x=c(1, -.1, 1.5, -.1), #b1*x1 + ...
+                    g_sig=1, # growth ~ N(E, g_sig)
+                    fl_z=c(-1.5, .1, .1), # b1 + b2*z + b3*z^2 + b4*z^3
+                    fl_x=c(1, -.01, -1, -.01, .05), # b1*x1 + ...
                     seed_z=c(2, 0.5, -.03), # b1 + b2*z + b2*z^2
-                    seed_x=c(1, -.1, 1, -.1), # b1*x1 + ...
-                    rcrt=c(1.5, 0.4), #N(mean=rcrt1, sd=rcrt2)
-                    p_est=0.02))
+                    seed_x=c(1, -.1, 1, -.1, .05), # b1*x1 + ...
+                    rcr_z=c(1.5, 0.4), # N(mean=rcrt1, sd=rcrt2)
+                    p_est=0.02, # p(establishment)
+                    rcr_SB=0.2, # p(recruit from seedbank)
+                    rcr_dir=0.6, # p(recruit directly)
+                    s_SB=0.3 # p(survive in seedbank additional year)
+                    ))
 ```
@@
 p <- sp_i$p
+IPMs <- array(0, dim=c(n+1, n+1, n.cell))
 d <- vector("list", length=n.cell)
-lambdas <- lambdas2 <- rep(NA, n.cell)
+lambdas <- B_1 <- rep(NA, n.cell)
+B <- rpois(n.cell, n_i)
 
@@
 
-# life history functions
-z_pow <- function(z.vec, n_z) {
-  if(n_z==1) {
-    return(matrix(1, ncol=1, nrow=length(z.vec)))
-  } else {
-    z <- matrix(z.vec, nrow=length(z.vec), ncol=n_z-1)
-    for(i in 1:(n_z-1)) z[,i] <- z[,i]^i
-    z <- cbind(1, z)
-    return(z)
-  }
-}
-calc_surv <- function(z.v, n_sz, p, X.s) {
-  z <- z_pow(z.v, n_sz)
-  u <- exp(z %*% p$s_z + c(t(X.s) %*% p$s_x))
-  return(u / (1+u))
-}
-calc_grow <- function(z1, z.v, n_gz, p, X.g) {
-  z <- z_pow(z.v, n_gz)
-  g <- dnorm(z1, mean=z %*% p$g_z + c(t(X.g) %*% p$g_x), 
-             sd=p$g_sig)
-  return(g)
-}
-calc_flwr <- function(z.v, n_flz, p, X.fl) {
-  z <- z_pow(z.v, n_flz)
-  u <- exp(z %*% p$fl_z + c(t(X.fl) %*% p$fl_x))
-  return(u / (1+u))
-}
-calc_seed <- function(z1, z.v, n_seedz, n_flz, p, X.seed, X.fl) {
-  z <- z_pow(z.v, n_seedz)
-  calc_flwr(z.v, n_flz, p, X.fl) * p$p_est *
-      dnorm(z1, p$rcrt[1], p$rcrt[2]) *
-      exp(z %*% p$seed_z + c(t(X.seed) %*% p$seed_x))
-}
-
 ########
@@
   d[[i]] <- data.frame(size=runif(n=n_i[i], min=1, max=12))
-  lo <- min(d[[i]]$size)*0.8
-  hi <- max(d[[i]]$size)*1.1
+  lo <- min(d[[i]]$size)*0.75
+  hi <- max(d[[i]]$size)*1.25
   size.mx <- cbind(1, d[[i]]$size, d[[i]]$size^2, d[[i]]$size^3)
 
-  # generate data
+  #---- generate data
   ## survival
@@
   d[[i]]$surv <- rbinom(n_i[i], 1, E_surv)
-
   ## growth
@@
   d[[i]]$sizeNext <- ifelse(d[[i]]$surv, d[[i]]$sizeNext, NA)
-
   ## flowering 
@@
   d[[i]]$fl <- rbinom(n_i[i], 1, E_fl)
-
   ## seed production
-  E_seed <- d[[i]]$fl * exp((size.mx[,1:n_seedz] %*% p$seed_z + 
-                               c(X_seed[i,] %*% p$seed_x)))
+  E_seed <- d[[i]]$fl * 
+    exp((size.mx[,1:n_seedz] %*% p$seed_z + c(X_seed[i,] %*% p$seed_x)))
   d[[i]]$seed <- rpois(n_i[i], E_seed)
-  d[[i]]$seed <- ifelse(d[[i]]$fl, d[[i]]$seed, NA)
-  seed.range <- range(d[[i]]$seed, na.rm=TRUE)
-
+  d[[i]]$fl <- ifelse(d[[i]]$surv, d[[i]]$fl, NA)
+  d[[i]]$seed <- ifelse(d[[i]]$fl==1, d[[i]]$seed, NA)
+  ## seed bank
+  B_1[i] <- B[i] * p$s_SB * (1 - p$rcr_SB) + 
+    sum(d[[i]]$seed, na.rm=TRUE) * (1 - p$rcr_dir) * p$s_SB
   ## recruits
-  rcrt.n <- max(sum(d[[i]]$seed, na.rm=TRUE)*p$p_est, 2)
+  rcrt.n <- max((sum(d[[i]]$seed, na.rm=TRUE) * p$rcr_dir * p$p_est +
+                   B[i] * p$rcr_SB * p$p_est), 
+                2) # need ≥ 2 for IPMpack estimate of rcr_z'_mn, rcr_z'_sd
   rcrt.i <- (n_i[i]+1):(n_i[i]+rcrt.n)
   d[[i]][rcrt.i,] <- NA
-  d[[i]]$sizeNext[rcrt.i] <- rnorm(rcrt.n, p$rcrt[1], p$rcrt[2])
+  d[[i]]$sizeNext[rcrt.i] <- rnorm(rcrt.n, p$rcr_z[1], p$rcr_z[2])
 
-  # fill IPM matrices
-  ## Manual: use assigned slopes
-  n <- 50  # ncells in matrix
+  #---- fill IPM matrices
+  ## Underlying Truth: use assigned slopes
   b <- lo + c(0:n)*(hi - lo)/n  # boundary points
@@
   h <- y[2] - y[1]  # step size
+  z.i <- 2:(n+1)  # continuous stage matrix indices
+  ### growth and survival
+  G.mx <- h*outer(y, y, calc_grow, p=p, n_gz=n_gz, X.g=X_g[i,])
+  S.v <- calc_surv(y, p=p, n_sz=n_sz, X.s=X_s[i,])
+  ### Fix eviction of individuals below/above size limits
+  for(k in 1:(n/2)) G.mx[1,k] <- G.mx[1,k] + 1 - sum(G.mx[,k])
+  for(k in (n/2+1):n) G.mx[n,k] <- G.mx[n,k] + 1 - sum(G.mx[,k])
+  ### fill P, F, and K
+  P.mx <- F.mx <- matrix(0, ncol=n+1, nrow=n+1)
+  for(k in z.i) P.mx[k,z.i] <- G.mx[k-1,]*S.v
+  P.mx[1,1] <- calc_staySB(p)
+  F.mx[z.i,z.i] <- h*outer(y, y, calc_rcrDir, p=p, n_seedz=n_seedz, n_flz=n_flz, 
+                           X.seed=X_seed[i,], X.fl=X_fl[i,])
+  F.mx[1,z.i] <- calc_addSB(y, p=p, n_seedz=n_seedz, n_flz=n_flz, 
+                            X.seed=X_seed[i,], X.fl=X_fl[i,])
+  F.mx[z.i,1] <- calc_rcrSB(y, p)
+  IPMs[,,i] <- P.mx + F.mx
 
-  G.mx <- h*outer(y, y, calc_grow, n_gz=n_gz, p=p, X.g=X_g[i,])
-  S.v <- calc_surv(y, n_sz, p, X_s[i,])
-  F.mx <- h*outer(y, y, calc_seed, n_seedz=n_seedz, p=p, n_flz=n_flz,
-                  X.seed=X_seed[i,], X.fl=X_fl[i,])
-  P.mx <- G.mx
-  for(k in 1:(n/2)) { # Fix eviction of individuals below/above size limits
-    G.mx[1,k] <- G.mx[1,k] + 1 - sum(G.mx[,k])
-    P.mx[,k] <- G.mx[,k]*S.v[k]
+  #---- progress
+  if(i %% 100 == 0) cat("Finished cell", i, "\n")
 }
-  for(k in (n/2+1):n) {
-    G.mx[n,k] <- G.mx[n,k] + 1 - sum(G.mx[,k])
-    P.mx[,k] <- G.mx[,k]*S.v[k]
-  }
-  IPM <- P.mx + F.mx
-
-  ## IPMpack: estimate slopes from d[[i]]
-  so <- makeSurvObj(d[[i]], surv ~ size)
-  go <- makeGrowthObj(d[[i]], sizeNext ~ size + size2)
-  fo <- makeFecObj(d[[i]],
-                   Formula=c(fl~size, seed~size+size2),
-                   Family=c("binomial", "poisson"),
-                   Transform=c("none", "none"),
-                   fecConstants=data.frame(seedlingEstablishmentRate=p$p_est))
-  P.mx2 <- makeIPMPmatrix(growObj=go, survObj=so, minSize=lo, maxSize=hi,
-                         correction="constant")
-  F.mx2 <- makeIPMFmatrix(fecObj=fo, minSize=lo, maxSize=hi,
-                          correction="constant")
-  IPM2 <- P.mx2 + F.mx2
-
-  lambdas[i] <- Re(eigen(IPM)$values[1])
-  lambdas2[i] <- Re(eigen(IPM2)$values[1])
+lam.df <- env.df
+lam.df$lambda <- apply(IPMs, 3, function(x) Re(eigen(x)$values[1]))
+ggplot(lam.df, aes(x=x, y=y, fill=log(lambda))) + geom_tile() +
+  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
+par(mfrow=c(2,2))
+plot(log(lambda) ~ temp, data=lam.df); abline(h=0, col="red", lty=2)
+plot(log(lambda) ~ prec, data=lam.df); abline(h=0, col="red", lty=2)
+plot(log(lambda) ~ pOpn, data=lam.df); abline(h=0, col="red", lty=2)
+plot(density(log(lam.df$lambda))); abline(v=0, col="red", lty=2)
 
 # Plots
-  z.t <- d[[i]][!is.na(d[[i]]$size),]; main.i <- paste("Cell", i)
 if(regPlots) {
@@
   xx <- tibble(size=seq(lo, hi, length.out=100), size2=size^2, size3=size^3)
+  xxx <- seq(0, hi*.5, length.out=100)
+  for(i in 1:n.cell) {
+    z.t <- d[[i]][!is.na(d[[i]]$size),]; main.i <- paste("Cell", i)
+    w1 <- Re(eigen(IPMs[,,i])$vectors[,1])[-1]; stable.dist1 <- w1/sum(w1)
+    rcrt.z <- d[[i]]$sizeNext[is.na(d[[i]]$size) & !is.na(d[[i]]$sizeNext)]
     plot(surv~size, data=d[[i]], ylim=c(0,1), pch=20, col=rgb(0,0,0,0.2),
@@
       points(z.t$size, E_surv, cex=0.2, col="red")
-      lines(xx$size, predict(so@fit, xx, type="response"), col="blue")
     plot(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,0.2), 
          main=paste("Growth:", main.i), xlab="Size (t)", ylab="Size (t+1)")
-      points(z.t$size, E_grow, cex=0.2, col="red")
-      lines(xx$size, predict(go@fit, xx, type="response"), col="blue")
-      abline(a=0, b=1, lty=2)
+      points(z.t$size, E_grow, cex=0.2, col="red"); abline(a=0, b=1, lty=2)
     plot(fl~size, data=z.t, pch=20, ylim=c(0,1), col=rgb(0,0,0,0.2),
@@
       points(z.t$size, E_fl, cex=0.2, col="red")
-      lines(xx$size, predict(fo@fitFec[[1]], xx, type="response"), col="blue")
     plot(seed~size, data=z.t, pch=23-z.t$fl*3, col=rgb(0,0,0,0.2), 
@@
       points(z.t$size, E_seed, cex=0.2, col="red")
-      lines(xx$size, predict(fo@fitFec[[2]], xx, type="response"), col="blue")
-    w1 <- Re(eigen(IPM)$vectors[,1]); stable.dist1 <- w1/sum(w1)
-    w2 <- Re(eigen(IPM2)$vectors[,1]); stable.dist2 <- w2/sum(w2)
-    yl <- range(c(stable.dist1, stable.dist2))
-    plot(y, stable.dist1, ylab="", type="l", lwd=2, col="red", ylim=yl,
+    plot(y, stable.dist1, ylab="", type="l", lwd=2, col="red", 
          main=paste("Stable size distribution:", main.i), xlab="Size (t)")
-    lines(y, stable.dist2, lwd=2, lty=2, col="blue")
-    # plot(y, S.v); points(y, apply(P.mx, 2, sum), col="red", type="l")
     hist(d[[i]]$sizeNext[rcrt.i], main=paste("Recruit sizes:", main.i), 
          xlab="Recruit size (t+1)", col="grey", freq=FALSE)
-      xxx <- seq(0, hi*.5, length.out=100)
-      lines(xxx, dnorm(xxx, p$rcrt[1], p$rcrt[2]), col="red")
+      lines(xxx, dnorm(xxx, p$rcr_z[1], p$rcr_z[2]), col="red", lwd=2)
+  }
 }
 if(ipmPlots) {
-    par(mfrow=c(2,2), mar=c(4,4,2,1)); main.i <- paste("Cell", i)
-    # sens.mx <- sens(IPM2); elas.mx <- elas(IPM2)
-    sens.mx <- sens(IPM); elas.mx <- elas(IPM)
-    image.plot(y, y, t(P.mx), main=main.i, xlab="Size (t)", ylab="Size (t+1)")
-      points(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,0.2))
+  par(mfrow=c(2,2), mar=c(4,4,2,1))
+  x <- c(0, y)
+  for(i in 1:n.cell) {
+    z.t <- d[[i]][!is.na(d[[i]]$size),]; main.i <- paste("Cell", i)
+    z.t$y <- y[findInterval(z.t$size, y, all.inside=TRUE)]
+    totSeeds <- group_by(z.t, y) %>% summarise(seed=sum(seed, na.rm=TRUE))
+    image.plot(x, x, t(P.mx), main=paste("P matrix:", main.i), 
+               xlab="Size (t)", ylab="Size (t+1)")
+      points(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,1/(n_i[i]^.25)))
       abline(a=0, b=1, col="white", lty=2)
-    image.plot(y, y, t(P.mx2), main=main.i, xlab="Size (t)", ylab="Size (t+1)")
-      points(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,0.2))
-      abline(a=0, b=1, col="white", lty=2)
-    image.plot(1:dim(sens.mx)[1], 1:dim(sens.mx)[1], t(sens.mx),
-               main=paste("Sens:", main.i), xlab="Size (t)", ylab="Size (t+1)")
-    image.plot(1:dim(elas.mx)[1], 1:dim(elas.mx)[1], t(elas.mx),
-               main=paste("Elas:", main.i), xlab="Size (t)", ylab="Size (t+1)")
+    image.plot(x, x, t(F.mx), main=paste("F matrix:", main.i), 
+               xlab="Size (t)", ylab="Size (t+1)")
+      points(totSeeds$y, totSeeds$seed/max(totSeeds$seed)*hi*.9)
+    image.plot(x, x, t(sens(IPMs[,,i])), main=paste("Sensitivity:", main.i), 
+               xlab="Size (t)", ylab="Size (t+1)")
+    image.plot(x, x, t(elas(IPMs[,,i])), main=paste("Elasticity:", main.i),
+               xlab="Size (t)", ylab="Size (t+1)")
   }
-  if(i %% 100 == 0) cat("Finished cell", i, "\n")
 }
-lam.df <- env.df %>% add_column(lambda=lambdas, lambda2=lambdas2)
-ggplot(lam.df, aes(x=x, y=y, fill=log(lambdas))) + geom_tile() +
-  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
-ggplot(lam.df, aes(x=x, y=y, fill=log(lambdas2))) + geom_tile() +
-  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
-par(mfrow=c(2,2))
-plot(log(lambdas) ~ temp, data=lam.df); abline(h=0, col="red", lty=2)
-plot(log(lambdas) ~ prec, data=lam.df); abline(h=0, col="red", lty=2)
-plot(log(lambdas), log(lambdas2)); abline(v=0, h=0, a=0, b=1, lty=c(3,2,2))
-plot(density(log(lambdas))); abline(v=0, col="red", lty=2)
 ```
@@
 ```{r P_IPM}
+for(i in 1:ncell) {
+  # ## IPMpack: estimate slopes from d[[i]]
+  # ### set up seedbank transitions
+  d[[i]]$stage <- d[[i]]$stageNext <- "continuous"
+  d[[i]]$stage[is.na(d[[i]]$size)] <- NA
+  d[[i]]$stageNext[d[[i]]$surv==0] <- "dead"
+  d[[i]]$number <- 1
+  seedbank <- data.frame(size=NA, surv=1, sizeNext=NA,
+                         fl=NA, seed=NA,
+                         stage=c("seedbank", "seedbank", "continuous"),
+                         stageNext=c("seedbank", "continuous", "seedbank"),
+                         number=c(p$s_SB * (1 - p$rcr_SB), 
+                                  p$rcr_SB * p$p_est, 
+                                  (1 - p$rcr_dir) * p$s_SB))
+  d[[i]] <- rbind(d[[i]], seedbank)
+  d[[i]]$stage <- as.factor(d[[i]]$stage)
+  d[[i]]$stageNext <- as.factor(d[[i]]$stageNext)
+  ### construct kernels
+  so <- makeSurvObj(d[[i]], surv ~ size)
+  go <- makeGrowthObj(d[[i]], sizeNext ~ size + size2)
+  fo <- makeFecObj(d[[i]],
+                   Formula=c(fl~size+size2, seed~size+size2),
+                   Family=c("binomial", "poisson"),
+                   Transform=c("none", "none"),
+                   meanOffspringSize=mean(
+                     d[[i]]$sizeNext[is.na(d[[i]]$size) &
+                                       !is.na(d[[i]]$sizeNext)]),
+                   sdOffspringSize=sd(
+                     d[[i]]$sizeNext[is.na(d[[i]]$size) &
+                                       !is.na(d[[i]]$sizeNext)]),
+                   fecConstants=data.frame(seedlingEstablishmentRate=p$p_est),
+                   offspringSplitter=data.frame(seedbank=(1-p$rcr_dir),
+                                                continuous=p$rcr_dir)
+                   )
+  F.mx2 <- makeIPMFmatrix(fecObj=fo, minSize=lo, maxSize=hi,
+                          correction="constant")
+  D.mx2 <- makeDiscreteTrans(d[[i]], 
+                             continuousToDiscreteExplanatoryVariables=1)
+  dummy <- as.matrix(fo@offspringRel$coefficients[1], list(1, "seedbank"))
+  D.mx2@meanToCont <- as.matrix(dummy, dimnames=c(1, "seedbank"))
+  dummy <- as.matrix(fo@sdOffspringSize, dimnames=list(1, "seedbank"))
+  D.mx2@sdToCont <- as.matrix(dummy, dimnames=c(1, "seedbank"))
+  D.mx2@discreteTrans[1,1] <- p$s_SB * (1 - p$rcr_SB)
+  D.mx2@discreteTrans[3,1] <- 1 - sum(D.mx2@discreteTrans[1:2,1])
+  P.mx2 <- makeIPMPmatrix(growObj=go, survObj=so, discreteTrans=D.mx2,
+                            minSize=lo, maxSize=hi, correction="constant")
+  IPM2 <- P.mx2 + F.mx2
+  lambdas2[i] <- Re(eigen(IPM2)$values[1])
+
+  if(regPlots) {
+    par(mfrow=c(3,2), mar=c(4,4,2,1))
+    xx <- tibble(size=seq(lo, hi, length.out=100), size2=size^2, size3=size^3)
+    plot(surv~size, data=d[[i]], ylim=c(0,1), pch=20, col=rgb(0,0,0,0.2),
+         main=paste("Surv:", main.i), xlab="Size (t)", ylab="P(surv to t+1)")
+      points(z.t$size, E_surv, cex=0.2, col="red")
+      # lines(xx$size, predict(so@fit, xx, type="response"), col="blue")
+    plot(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,0.2), 
+         main=paste("Growth:", main.i), xlab="Size (t)", ylab="Size (t+1)")
+      points(z.t$size, E_grow, cex=0.2, col="red")
+      # lines(xx$size, predict(go@fit, xx, type="response"), col="blue")
+      abline(a=0, b=1, lty=2)
+    plot(fl~size, data=z.t, pch=20, ylim=c(0,1), col=rgb(0,0,0,0.2),
+         main=paste("P(Flower):", main.i), xlab="Size (t)", ylab="P(flower)")
+      points(z.t$size, E_fl, cex=0.2, col="red")
+      # lines(xx$size, predict(fo@fitFec[[1]], xx, type="response"), col="blue")
+    plot(seed~size, data=z.t, pch=23-z.t$fl*3, col=rgb(0,0,0,0.2), 
+         main=paste("nSeeds:", main.i), xlab="Size (t)", ylab="Number of seeds")
+      points(z.t$size, E_seed, cex=0.2, col="red")
+      # lines(xx$size, predict(fo@fitFec[[2]], xx, type="response"), col="blue")
+    w1 <- Re(eigen(IPMs[,,i])$vectors[,1])[-1]; stable.dist1 <- w1/sum(w1)
+    # w2 <- Re(eigen(IPM2)$vectors[,1])[-1]; stable.dist2 <- w2/sum(w2)
+    yl <- range(stable.dist1)#range(c(stable.dist1, stable.dist2))
+    plot(y, stable.dist1, ylab="", type="l", lwd=2, col="red", ylim=yl,
+         main=paste("Stable size distribution:", main.i), xlab="Size (t)")
+    # lines(y, stable.dist2, col="blue")
+    hist(d[[i]]$sizeNext[rcrt.i], main=paste("Recruit sizes:", main.i), 
+         xlab="Recruit size (t+1)", col="grey", freq=FALSE)
+      xxx <- seq(0, hi*.5, length.out=100)
+      rcrt.z <- d[[i]]$sizeNext[is.na(d[[i]]$size) & !is.na(d[[i]]$sizeNext)]
+      lines(xxx, dnorm(xxx, p$rcr_z[1], p$rcr_z[2]), col="red", lwd=2)
+      # lines(xxx, dnorm(xxx, mean(rcrt.z), sd(rcrt.z)), col="blue")
+  }
+}
 

### 2018-03-01 exploration.Rmd
arrays for P, F, K; separated plot loops

@@
 subsamp <- NA
-mean_abund <- 200
+mean_abund <- 2000
 regPlots <- F
@@
                     fl_z=c(-1.5, .1, .1), # b1 + b2*z + b3*z^2 + b4*z^3
-                    fl_x=c(1, -.01, -1, -.01, .05), # b1*x1 + ...
+                    fl_x=c(1, -.01, -1, -.01, .2), # b1*x1 + ...
                     seed_z=c(2, 0.5, -.03), # b1 + b2*z + b2*z^2
-                    seed_x=c(1, -.1, 1, -.1, .05), # b1*x1 + ...
+                    seed_x=c(1, -.1, 1, -.1, .2), # b1*x1 + ...
                     rcr_z=c(1.5, 0.4), # N(mean=rcrt1, sd=rcrt2)
@@
 p <- sp_i$p
-IPMs <- array(0, dim=c(n+1, n+1, n.cell))
+IPMs <- Ps <- Fs <- array(0, dim=c(n+1, n+1, n.cell))
 d <- vector("list", length=n.cell)
@@
   E_seed <- d[[i]]$fl * 
-    exp((size.mx[,1:n_seedz] %*% p$seed_z + c(X_seed[i,] %*% p$seed_x)))
+    exp(size.mx[,1:n_seedz] %*% p$seed_z + c(X_seed[i,] %*% p$seed_x))
   d[[i]]$seed <- rpois(n_i[i], E_seed)
@@
   ### fill P, F, and K
-  P.mx <- F.mx <- matrix(0, ncol=n+1, nrow=n+1)
-  for(k in z.i) P.mx[k,z.i] <- G.mx[k-1,]*S.v
-  P.mx[1,1] <- calc_staySB(p)
-  F.mx[z.i,z.i] <- h*outer(y, y, calc_rcrDir, p=p, n_seedz=n_seedz, n_flz=n_flz, 
+  for(k in z.i) Ps[k,z.i,i] <- G.mx[k-1,]*S.v
+  Ps[1,1,i] <- calc_staySB(p)
+  Fs[z.i,z.i,i] <- h*outer(y, y, calc_rcrDir, p=p, n_seedz=n_seedz,
+                             n_flz=n_flz, X.seed=X_seed[i,], X.fl=X_fl[i,])
+  Fs[1,z.i,i] <- calc_addSB(y, p=p, n_seedz=n_seedz, n_flz=n_flz, 
                             X.seed=X_seed[i,], X.fl=X_fl[i,])
-  F.mx[1,z.i] <- calc_addSB(y, p=p, n_seedz=n_seedz, n_flz=n_flz, 
-                            X.seed=X_seed[i,], X.fl=X_fl[i,])
-  F.mx[z.i,1] <- calc_rcrSB(y, p)
-  IPMs[,,i] <- P.mx + F.mx
+  Fs[z.i,1,i] <- calc_rcrSB(y, p)
+  IPMs[,,i] <- Ps[,,i] + Fs[,,i]
 
@@
   xx <- tibble(size=seq(lo, hi, length.out=100), size2=size^2, size3=size^3)
+  mx <- cbind(1, as.matrix(xx))
   xxx <- seq(0, hi*.5, length.out=100)
@@
     rcrt.z <- d[[i]]$sizeNext[is.na(d[[i]]$size) & !is.na(d[[i]]$sizeNext)]
-    plot(surv~size, data=d[[i]], ylim=c(0,1), pch=20, col=rgb(0,0,0,0.2),
+    plot(surv~size, data=z.t, ylim=c(0,1), pch=20, col=rgb(0,0,0,0.2),
          main=paste("Surv:", main.i), xlab="Size (t)", ylab="P(surv to t+1)")
-      points(z.t$size, E_surv, cex=0.2, col="red")
+      lines(mx[,2], antilogit(mx[,1:n_sz]%*%p$s_z + c(X_s[i,]%*%p$s_x)), col=2)
     plot(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,0.2), 
          main=paste("Growth:", main.i), xlab="Size (t)", ylab="Size (t+1)")
-      points(z.t$size, E_grow, cex=0.2, col="red"); abline(a=0, b=1, lty=2)
+      lines(mx[,2], mx[,1:n_gz]%*%p$g_z + c(X_g[i,]%*%p$g_x), col=2)
+      abline(a=0, b=1, lty=2)
     plot(fl~size, data=z.t, pch=20, ylim=c(0,1), col=rgb(0,0,0,0.2),
          main=paste("P(Flower):", main.i), xlab="Size (t)", ylab="P(flower)")
-      points(z.t$size, E_fl, cex=0.2, col="red")
+      lines(mx[,2], antilogit(mx[,1:n_flz]%*%p$fl_z + c(X_fl[i,]%*%p$fl_x)),
+            col=2)
     plot(seed~size, data=z.t, pch=23-z.t$fl*3, col=rgb(0,0,0,0.2), 
          main=paste("nSeeds:", main.i), xlab="Size (t)", ylab="Number of seeds")
-      points(z.t$size, E_seed, cex=0.2, col="red")
-    plot(y, stable.dist1, ylab="", type="l", lwd=2, col="red", 
+      lines(mx[,2], exp(mx[,1:n_seedz]%*%p$seed_z + c(X_seed[i,]%*%p$seed_x)),
+            col=2)
+    plot(y, stable.dist1, ylab="", type="l", col=2, 
          main=paste("Stable size distribution:", main.i), xlab="Size (t)")
-    hist(d[[i]]$sizeNext[rcrt.i], main=paste("Recruit sizes:", main.i), 
-         xlab="Recruit size (t+1)", col="grey", freq=FALSE)
+    hist(d[[i]]$sizeNext[is.na(d[[i]]$size)], col="grey", freq=FALSE, 
+         main=paste("Recruit sizes:", main.i), xlab="Recruit size (t+1)")
       lines(xxx, dnorm(xxx, p$rcr_z[1], p$rcr_z[2]), col="red", lwd=2)
@@
     totSeeds <- group_by(z.t, y) %>% summarise(seed=sum(seed, na.rm=TRUE))
-    image.plot(x, x, t(P.mx), main=paste("P matrix:", main.i), 
+    image.plot(x, x, t(Ps[,,i]), main=paste("P matrix:", main.i), 
                xlab="Size (t)", ylab="Size (t+1)")
@@
       abline(a=0, b=1, col="white", lty=2)
-    image.plot(x, x, t(F.mx), main=paste("F matrix:", main.i), 
+    image.plot(x, x, t(Fs[,,i]), main=paste("F matrix:", main.i), 
                xlab="Size (t)", ylab="Size (t+1)")

### 2018-03-02 exploration.Rmd
formatting

@@
 
-  #---- fill IPM matrices
-  ## Underlying Truth: use assigned slopes
+  #---- Truth: use assigned slopes to fill IPM matrix
   b <- lo + c(0:n)*(hi - lo)/n  # boundary points

### 2018-03-02 exploration.Rmd
finished seedbank

@@
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+```{r setup, include=F}
+knitr::opts_chunk$set(echo = T)
 pkgs <- c("fields", "sevcheck", "gbPopMod", "IPMpack", "tidyverse", "magrittr")
-suppressMessages(invisible(lapply(pkgs, library, character.only=TRUE)))
+suppressMessages(invisible(lapply(pkgs, library, character.only=T)))
 theme_set(theme_bw())
@@
 ```{r sp_i}
-subsamp <- NA
+subsamp <- 500
 mean_abund <- 2000
@@
 ipmPlots <- F
+z.rng <- c(1, 12)
 n <- 100  # ncells in IPM matrix
 if(is.na(subsamp)) { cells <- 1:n_distinct(lc.df$CellID)
-} else { cells <- sample(1:n_distinct(lc.df$CellID), subsamp, replace=FALSE) }
+} else { cells <- sample(1:n_distinct(lc.df$CellID), subsamp, replace=F) }
 n.cell <- length(cells)
@@
 sp_i <- list(p=list(s_z=c(-1.75, .75), # b1 + b2*z
-                    s_x=c(1, -.1, -.5, -.1, .05), # b1*x1 + ...
+                    s_x=c(1, -.1, -.5, -.1, .05, -2, -.4, -.4), # b1*x1 + ...
                     g_z=c(.2, 2, -0.1), # b1 + b2*z + b3*z^2
@@
                     fl_z=c(-1.5, .1, .1), # b1 + b2*z + b3*z^2 + b4*z^3
-                    fl_x=c(1, -.01, -1, -.01, .2), # b1*x1 + ...
+                    fl_x=c(1, -.01, -1, -.01, .2, .2), # b1*x1 + ...
                     seed_z=c(2, 0.5, -.03), # b1 + b2*z + b2*z^2
@@
 ########
-# set up storage objects
+# set up storage & IPM matrix objects
 p <- sp_i$p
+setup_IPM_matrix(n, z.rng, buffer=0.25, discrete=1)
 IPMs <- Ps <- Fs <- array(0, dim=c(n+1, n+1, n.cell))
 d <- vector("list", length=n.cell)
-lambdas <- B_1 <- rep(NA, n.cell)
 B <- rpois(n.cell, n_i)
+B_1 <- rep(NA, n.cell)
 
@@
 for(i in 1:n.cell) {
-  d[[i]] <- data.frame(size=runif(n=n_i[i], min=1, max=12))
-  lo <- min(d[[i]]$size)*0.75
-  hi <- max(d[[i]]$size)*1.25
+  d[[i]] <- data.frame(size=runif(n=n_i[i], min=z.rng[1], max=z.rng[2]))
   size.mx <- cbind(1, d[[i]]$size, d[[i]]$size^2, d[[i]]$size^3)
+
+  #---- Truth: use assigned slopes to fill IPM matrix
+  ### growth and survival
+  G.mx <- h*outer(y, y, calc_grow, p=p, n_gz=n_gz, X.g=X_g[i,])
+  S.v <- calc_surv(y, p=p, n_sz=n_sz, X.s=X_s[i,])
+  ### Fix eviction of individuals below/above size limits
+  for(k in 1:(n/2)) G.mx[1,k] <- G.mx[1,k] + 1 - sum(G.mx[,k])
+  for(k in (n/2+1):n) G.mx[n,k] <- G.mx[n,k] + 1 - sum(G.mx[,k])
+  ### fill P and F matrices
+  for(k in z.i) Ps[k,z.i,i] <- G.mx[k-1,]*S.v
+  Ps[1,1,i] <- calc_staySB(p)
+  Fs[z.i,z.i,i] <- h*outer(y, y, calc_rcrDir, p=p, n_seedz=n_seedz,
+                             n_flz=n_flz, X.seed=X_seed[i,], X.fl=X_fl[i,])
+  Fs[1,z.i,i] <- calc_addSB(y, p=p, n_seedz=n_seedz, n_flz=n_flz, 
+                            X.seed=X_seed[i,], X.fl=X_fl[i,])
+  Fs[z.i,1,i] <- calc_rcrSB(y, p)
 
@@
   B_1[i] <- B[i] * p$s_SB * (1 - p$rcr_SB) + 
-    sum(d[[i]]$seed, na.rm=TRUE) * (1 - p$rcr_dir) * p$s_SB
+    sum(d[[i]]$seed, na.rm=T) * (1 - p$rcr_dir) * p$s_SB
   ## recruits
-  rcrt.n <- max((sum(d[[i]]$seed, na.rm=TRUE) * p$rcr_dir * p$p_est +
-                   B[i] * p$rcr_SB * p$p_est), 
-                2) # need ≥ 2 for IPMpack estimate of rcr_z'_mn, rcr_z'_sd
+  rcrt.n <- max((sum(d[[i]]$seed, na.rm=T) * p$rcr_dir * p$p_est +
+                   B[i] * p$rcr_SB * p$p_est), 2) # ≥ 2 for IPMpack sd(rcr_z')
   rcrt.i <- (n_i[i]+1):(n_i[i]+rcrt.n)
@@
   d[[i]]$sizeNext[rcrt.i] <- rnorm(rcrt.n, p$rcr_z[1], p$rcr_z[2])
-
-  #---- Truth: use assigned slopes to fill IPM matrix
-  b <- lo + c(0:n)*(hi - lo)/n  # boundary points
-  y <- 0.5*(b[1:n] + b[2:(n+1)])  # mesh points
-  h <- y[2] - y[1]  # step size
-  z.i <- 2:(n+1)  # continuous stage matrix indices
-  ### growth and survival
-  G.mx <- h*outer(y, y, calc_grow, p=p, n_gz=n_gz, X.g=X_g[i,])
-  S.v <- calc_surv(y, p=p, n_sz=n_sz, X.s=X_s[i,])
-  ### Fix eviction of individuals below/above size limits
-  for(k in 1:(n/2)) G.mx[1,k] <- G.mx[1,k] + 1 - sum(G.mx[,k])
-  for(k in (n/2+1):n) G.mx[n,k] <- G.mx[n,k] + 1 - sum(G.mx[,k])
-  ### fill P, F, and K
-  for(k in z.i) Ps[k,z.i,i] <- G.mx[k-1,]*S.v
-  Ps[1,1,i] <- calc_staySB(p)
-  Fs[z.i,z.i,i] <- h*outer(y, y, calc_rcrDir, p=p, n_seedz=n_seedz,
-                             n_flz=n_flz, X.seed=X_seed[i,], X.fl=X_fl[i,])
-  Fs[1,z.i,i] <- calc_addSB(y, p=p, n_seedz=n_seedz, n_flz=n_flz, 
-                            X.seed=X_seed[i,], X.fl=X_fl[i,])
-  Fs[z.i,1,i] <- calc_rcrSB(y, p)
-  IPMs[,,i] <- Ps[,,i] + Fs[,,i]
-
-  #---- progress
-  if(i %% 100 == 0) cat("Finished cell", i, "\n")
 }
+IPMs <- Ps + Fs
 lam.df <- env.df
 lam.df$lambda <- apply(IPMs, 3, function(x) Re(eigen(x)$values[1]))
+
+########
+## plots
+########
 ggplot(lam.df, aes(x=x, y=y, fill=log(lambda))) + geom_tile() +
   scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
+ggplot(lam.df, aes(x=x, y=y, fill=lambda>1)) + geom_tile()
 par(mfrow=c(2,2))
@@
 plot(log(lambda) ~ prec, data=lam.df); abline(h=0, col="red", lty=2)
-plot(log(lambda) ~ pOpn, data=lam.df); abline(h=0, col="red", lty=2)
+plot(log(lambda) ~ pOth, data=lam.df); abline(h=0, col="red", lty=2)
 plot(density(log(lam.df$lambda))); abline(v=0, col="red", lty=2)
 
-# Plots
 if(regPlots) {
@@
          main=paste("Stable size distribution:", main.i), xlab="Size (t)")
-    hist(d[[i]]$sizeNext[is.na(d[[i]]$size)], col="grey", freq=FALSE, 
+    hist(d[[i]]$sizeNext[is.na(d[[i]]$size)], col="grey", freq=F, 
          main=paste("Recruit sizes:", main.i), xlab="Recruit size (t+1)")
@@
     z.t <- d[[i]][!is.na(d[[i]]$size),]; main.i <- paste("Cell", i)
-    z.t$y <- y[findInterval(z.t$size, y, all.inside=TRUE)]
-    totSeeds <- group_by(z.t, y) %>% summarise(seed=sum(seed, na.rm=TRUE))
+    z.t$y <- y[findInterval(z.t$size, y, all.inside=T)]
+    totSeeds <- group_by(z.t, y) %>% summarise(seed=sum(seed, na.rm=T))
     image.plot(x, x, t(Ps[,,i]), main=paste("P matrix:", main.i), 
@@
     hist(d[[i]]$sizeNext[rcrt.i], main=paste("Recruit sizes:", main.i), 
-         xlab="Recruit size (t+1)", col="grey", freq=FALSE)
+         xlab="Recruit size (t+1)", col="grey", freq=F)
       xxx <- seq(0, hi*.5, length.out=100)

### 2018-03-02 code/fn_aux.R
added setup_IPM_matrix()

@@ -3,6 +3,18 @@
 # Tim Szewczyk
 
 
+##-- set boundaries, meshpoints, step size for IPM matrix
+##   create objects in global environment
+setup_IPM_matrix <- function(n=100, z.rng=c(1,10), buffer=0.25, discrete=0) {
+  lo <<- z.rng[1]*(1-buffer)  # IPM matrix lower limit
+  hi <<- z.rng[2]*(1+buffer)  # IPM matrix upper limit
+  b <<- lo + c(0:n)*(hi - lo)/n  # boundary points
+  y <<- 0.5*(b[1:n] + b[2:(n+1)])  # mesh points
+  h <<- y[2] - y[1]  # step size
+  z.i <<- (1:n) + discrete  # continuous stage matrix indices
+}
+
+
 ##-- create design matrix from vector of sizes
 z_pow <- function(z.vec, n_z) {
   if(n_z==1) {

### 2018-03-02 exploration.Rmd
calculating SDD neighborhoods

@@
 lc.df <- read_csv("data/landcover_5km.csv") %>% 
-  filter(!is.na(bio1_mean))
+  filter(!is.na(bio1_mean)) %>%
+  mutate(x=as.integer(as.factor(lc.df$left)),
+         y=as.integer(factor(.$top,
+                             levels=rev(levels(factor(.$top))))),
+         x_y=paste(x, y, sep="_"))
 ```
@@
 ```{r sp_i}
-subsamp <- 500
 mean_abund <- 2000
@@
 n <- 100  # ncells in IPM matrix
-if(is.na(subsamp)) { cells <- 1:n_distinct(lc.df$CellID)
-} else { cells <- sample(1:n_distinct(lc.df$CellID), subsamp, replace=F) }
+cells <- 1:n_distinct(lc.df$CellID)
 n.cell <- length(cells)
@@
                     g_z=c(.2, 2, -0.1), # b1 + b2*z + b3*z^2
-                    g_x=c(1, -.1, 1.5, -.1), #b1*x1 + ...
+                    g_x=c(1, -.1, 1.5, -.1, .1, -2), #b1*x1 + ...
                     g_sig=1, # growth ~ N(E, g_sig)
@@
 ```{r landscape}
-env.df <- tibble(temp=c(scale(lc.df$bio1_mean)),
+env.rct <- as.tibble(expand.grid(x=1:max(lc.df$x),
+                                 y=1:max(lc.df$y))) %>%
+  mutate(x_y=paste(x, y, sep="_")) %>%
+  mutate(temp=c(scale(lc.df$bio1_mean))[match(.$x_y, lc.df$x_y)],
          temp2=temp^2,
-                 prec=c(scale(lc.df$bio12_mean)),
+         prec=c(scale(lc.df$bio12_mean))[match(.$x_y, lc.df$x_y)],
          prec2=prec^2,
-                 pOpn=c(scale(lc.df$nlcd1_mean)),
-                 pOth=c(scale(lc.df$nlcd2_mean)),
-                 pDec=c(scale(lc.df$nlcd3_mean)),
-                 pEvg=c(scale(lc.df$nlcd4_mean)),
-                 pMxd=c(scale(lc.df$nlcd5_mean)),
-                 x=lc.df$left,
-                 y=lc.df$top)
-env.df <- env.df[cells,]
+         pOpn=c(scale(lc.df$nlcd1_mean))[match(.$x_y, lc.df$x_y)],
+         pOth=c(scale(lc.df$nlcd2_mean))[match(.$x_y, lc.df$x_y)],
+         pDec=c(scale(lc.df$nlcd3_mean))[match(.$x_y, lc.df$x_y)],
+         pEvg=c(scale(lc.df$nlcd4_mean))[match(.$x_y, lc.df$x_y)],
+         pMxd=c(scale(lc.df$nlcd5_mean))[match(.$x_y, lc.df$x_y)],
+         inbd=!is.na(match(.$x_y, lc.df$x_y)),
+         id=row_number(), 
+         id.inbd=min_rank(na_if(inbd*id, 0)),
+         lat=lc.df$top[match(.$x_y, lc.df$x_y)],
+         lon=lc.df$left[match(.$x_y, lc.df$x_y)])
+env.rct[is.na(env.rct)] <- 0
+sdd.pr <- sdd_set_probs(ncell=n.cell, lc.df=env.rct, lc.col=8:12,
+                        g.p=list(sdd.max=4, sdd.rate=0.8, bird.hab=rep(1,5)))
+env.df <- filter(env.rct, inbd) %>% select(4:12, 1:3, 13:17)
+rm(env.rct)
 ```
@@
   ## recruits
-  rcrt.n <- max((sum(d[[i]]$seed, na.rm=T) * p$rcr_dir * p$p_est +
-                   B[i] * p$rcr_SB * p$p_est), 2) # ≥ 2 for IPMpack sd(rcr_z')
+  rcrt.n <- max(ceiling((sum(d[[i]]$seed, na.rm=T) * p$rcr_dir * p$p_est +
+                   B[i] * p$rcr_SB * p$p_est)), 2) # ≥ 2 for IPMpack sd(rcr_z')
   rcrt.i <- (n_i[i]+1):(n_i[i]+rcrt.n)
@@
 ########
-ggplot(lam.df, aes(x=x, y=y, fill=log(lambda))) + geom_tile() +
+ggplot(lam.df, aes(x=lon, y=lat, fill=log(lambda))) + geom_tile() +
   scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
-ggplot(lam.df, aes(x=x, y=y, fill=lambda>1)) + geom_tile()
+ggplot(lam.df, aes(x=lon, y=lat, fill=lambda>1)) + geom_tile()
 par(mfrow=c(2,2))

### 2018-03-02 exploration.Rmd
dispersal for simulated data; still to do theoretical IPM matrix

@@
   filter(!is.na(bio1_mean)) %>%
-  mutate(x=as.integer(as.factor(lc.df$left)),
+  mutate(x=as.integer(factor(.$left)),
          y=as.integer(factor(.$top,
@@
                     rcr_dir=0.6, # p(recruit directly)
-                    s_SB=0.3 # p(survive in seedbank additional year)
+                    s_SB=0.3, # p(survive in seedbank additional year)
+                    p_emig=0.3, # p(seed emigrates)
+                    sdd_max=4, # max SDD distance in cells
+                    sdd_rate=0.8 # SDD dispersal rate
                     ))
@@
          prec2=prec^2,
-         pOpn=c(scale(lc.df$nlcd1_mean))[match(.$x_y, lc.df$x_y)],
-         pOth=c(scale(lc.df$nlcd2_mean))[match(.$x_y, lc.df$x_y)],
-         pDec=c(scale(lc.df$nlcd3_mean))[match(.$x_y, lc.df$x_y)],
-         pEvg=c(scale(lc.df$nlcd4_mean))[match(.$x_y, lc.df$x_y)],
-         pMxd=c(scale(lc.df$nlcd5_mean))[match(.$x_y, lc.df$x_y)],
+         pOpn=lc.df$nlcd1_mean[match(.$x_y, lc.df$x_y)],
+         pOth=lc.df$nlcd2_mean[match(.$x_y, lc.df$x_y)],
+         pDec=lc.df$nlcd3_mean[match(.$x_y, lc.df$x_y)],
+         pEvg=lc.df$nlcd4_mean[match(.$x_y, lc.df$x_y)],
+         pMxd=lc.df$nlcd5_mean[match(.$x_y, lc.df$x_y)],
          inbd=!is.na(match(.$x_y, lc.df$x_y)),
@@
 sdd.pr <- sdd_set_probs(ncell=n.cell, lc.df=env.rct, lc.col=8:12,
-                        g.p=list(sdd.max=4, sdd.rate=0.8, bird.hab=rep(1,5)))
-env.df <- filter(env.rct, inbd) %>% select(4:12, 1:3, 13:17)
+                        g.p=list(sdd.max=sp_i$p$sdd_max, 
+                                 sdd.rate=sp_i$p$sdd_rate, 
+                                 bird.hab=rep(1,5)))
+env.df <- filter(env.rct, inbd) %>% select(4:12, 1:3, 13:17) %>%
+  mutate(pOpn=c(scale(pOpn)), pOth=c(scale(pOth)), pDec=c(scale(pDec)),
+         pEvg=c(scale(pEvg)), pMxd=c(scale(pMxd)))
 rm(env.rct)
+## sdd.pr[,,2,] has indexes based on env.df$id -- NOT env.df$id.inbd
+ggplot(env.df, aes(x=lon, y=lat, fill=id %in% sdd.pr[,,2,546])) + geom_tile()
 ```
@@
 B <- rpois(n.cell, n_i)
-B_1 <- rep(NA, n.cell)
+B_1 <- nSd <- D <- rep(NA, n.cell)
+sdd <- array(0, dim=dim(sdd.pr) + c(0,0,1,0)) # add layer for nEmig
+sdd[,,1:2,] <- sdd.pr
 
@@
   d[[i]]$seed <- ifelse(d[[i]]$fl==1, d[[i]]$seed, NA)
-  ## seed bank
+  nSd[i] <- sum(d[[i]]$seed, na.rm=T)
+  ## seed bank: native
   B_1[i] <- B[i] * p$s_SB * (1 - p$rcr_SB) + 
-    sum(d[[i]]$seed, na.rm=T) * (1 - p$rcr_dir) * p$s_SB
-  ## recruits
-  rcrt.n <- max(ceiling((sum(d[[i]]$seed, na.rm=T) * p$rcr_dir * p$p_est +
-                   B[i] * p$rcr_SB * p$p_est)), 2) # ≥ 2 for IPMpack sd(rcr_z')
-  rcrt.i <- (n_i[i]+1):(n_i[i]+rcrt.n)
-  d[[i]][rcrt.i,] <- NA
-  d[[i]]$sizeNext[rcrt.i] <- rnorm(rcrt.n, p$rcr_z[1], p$rcr_z[2])
+    nSd[i] * (1 - p$p_emig) * (1 - p$rcr_dir) * p$s_SB
+  ## recruits: native
+  rcr.n <- max(ceiling(nSd[i] * (1 - p$p_emig) * p$rcr_dir * p$p_est +
+                   B[i] * p$rcr_SB * p$p_est), 2) # ≥ 2 for IPMpack sd(rcr_z')
+  rcr.i <- (1:rcr.n) + n_i[i]
+  d[[i]][rcr.i,] <- NA
+  d[[i]]$sizeNext[rcr.i] <- rnorm(rcr.n, p$rcr_z[1], p$rcr_z[2])
+  ## emigrants
+  sdd[,,3,i] <- nSd[i] * p$p_emig * sdd[,,1,i]
+}
+for(i in 1:n.cell) {
+  id.i <- env.df$id[i]
+
+  #---- Truth: add immigrants
+
+  #---- generate data: add immigrants
+  # total immigrants
+  D[i] <- sum(sdd[,,3,][sdd[,,2,]==id.i])
+  # seed bank: immigrants
+  B_1[i] <- B_1[i] + D[i] * (1 - p$rcr_dir) * p$s_SB
+  # recruits: immigrants
+  rcr.imm <- max(ceiling(D[i] * p$rcr_dir * p$p_est), 2)
+  rcr.i <- (1:rcr.imm) + nrow(d[[i]])
+  d[[i]][rcr.i,] <- NA
+  d[[i]]$sizeNext[rcr.i] <- rnorm(rcr.imm, p$rcr_z[1], p$rcr_z[2])
 }
@@
 lam.df$lambda <- apply(IPMs, 3, function(x) Re(eigen(x)$values[1]))
+lam.df %<>% add_column(nSd=nSd, B0=B, B1=B_1, D=D)
 
@@
 ggplot(lam.df, aes(x=lon, y=lat, fill=lambda>1)) + geom_tile()
+ggplot(lam.df, aes(x=lon, y=lat, fill=nSd)) + geom_tile()
+ggplot(lam.df, aes(x=lon, y=lat, fill=B1>B0)) + geom_tile()
+ggplot(lam.df, aes(x=lon, y=lat, fill=D)) + geom_tile()
+ggplot(lam.df, aes(x=lon, y=lat, fill=nSd*p$p_emig > D)) + geom_tile()
+ggplot(lam.df, aes(x=lon, y=lat, fill=nSd*(1-p$p_emig) > D)) + geom_tile()
+ggplot(lam.df, aes(x=lon, y=lat, fill=nSd*(1-p$p_emig) < D & lambda < 1)) +
+  geom_tile()
 par(mfrow=c(2,2))
@@
   xxx <- seq(0, hi*.5, length.out=100)
-  for(i in 1:n.cell) {
+  for(i in 1:20) {
     z.t <- d[[i]][!is.na(d[[i]]$size),]; main.i <- paste("Cell", i)

### 2018-03-05 code/fn_IPM.R
updated functions for immigration

@@ -20,8 +20,11 @@
 #        p_est=0.02,  # p(establishment)
 #        rcr_SB=0.2,  # p(recruit from seedbank)
 #        rcr_dir=0.6,  # p(recruit directly)
-#        s_SB=0.3  # p(survive in seedbank additional year)
-# )
+#        s_SB=0.3,  # p(survive in seedbank additional year)
+#        p_emig=0.3,  # p(seed emigrates)
+#        sdd_max=2,  # max SDD distance in cells
+#        sdd_rate=0.8  # SDD dispersal rate
+#        )
 #
 # The function `z_pow()` takes a vector of sizes (z) and the number of size 
 # slopes for each function, returning a design matrix with the correct number 
@@ -56,7 +59,7 @@ calc_flwr <- function(z.v, p, n_flz, X.fl) {
 }
 
 
-##-- Seeds
+##-- Seed production
 ##   nSeeds ~ exp(size + environment)
 calc_seeds <- function(z.v, p, n_seeds, X.seed) {
   z <- z_pow(z.v, n_seedz)
@@ -64,12 +67,13 @@ calc_seeds <- function(z.v, p, n_seeds, X.seed) {
 }
 
 
-##-- Recruits (direct)
-##   z' ~ P(fl) * nSeeds * P(rcrDir) * P(estab) * N(mn, sd)
-calc_rcrDir <- function(z1, z.v, p, n_seedz, n_flz, X.seed, X.fl, SB=FALSE) {
+##-- Recruits (direct, native)
+##   z.ii' ~ P(fl.i) * nSeeds.i (1-p.emig) * P(rcrDir) * P(estab) * N(mn, sd)
+calc_rcrDir <- function(z1, z.v, p, n_seedz, n_flz, X.seed, X.fl) {
   calc_flwr(z.v, p, n_flz, X.fl) *
     calc_seeds(z.v, p, n_seedz, X.seed) *
-    ifelse(SB, p$rcr_dir, 1) * 
+    (1 - p$p_emig) *
+    p$rcr_dir * 
     p$p_est * 
     dnorm(z1, p$rcr_z[1], p$rcr_z[2])
 }
@@ -82,12 +86,39 @@ calc_rcrSB <- function(z1, p) {
 }
 
 
[…]

### 2018-03-05 exploration.Rmd
added theoretical immigration

@@
                     s_SB=0.3, # p(survive in seedbank additional year)
-                    p_emig=0.3, # p(seed emigrates)
-                    sdd_max=4, # max SDD distance in cells
+                    p_emig=0.999, # p(seed emigrates)
+                    sdd_max=3, # max SDD distance in cells
                     sdd_rate=0.8 # SDD dispersal rate
@@
 ## sdd.pr[,,2,] has indexes based on env.df$id -- NOT env.df$id.inbd
-ggplot(env.df, aes(x=lon, y=lat, fill=id %in% sdd.pr[,,2,546])) + geom_tile()
+# ggplot(env.df, aes(x=lon, y=lat, fill=id %in% sdd.pr[,,2,760])) + geom_tile()
 ```
@@
                              n_flz=n_flz, X.seed=X_seed[i,], X.fl=X_fl[i,])
-  Fs[1,z.i,i] <- calc_addSB(y, p=p, n_seedz=n_seedz, n_flz=n_flz, 
+  Fs[1,z.i,i] <- calc_DirSB(y, p=p, n_seedz=n_seedz, n_flz=n_flz, 
                             X.seed=X_seed[i,], X.fl=X_fl[i,])
@@
   d[[i]]$seed <- rpois(n_i[i], E_seed)
-  d[[i]]$fl <- ifelse(d[[i]]$surv, d[[i]]$fl, NA)
+  # d[[i]]$fl <- ifelse(d[[i]]$surv, d[[i]]$fl, NA)
   d[[i]]$seed <- ifelse(d[[i]]$fl==1, d[[i]]$seed, NA)
@@
 for(i in 1:n.cell) {
-  id.i <- env.df$id[i]
+  sdd.i <- env.df$id[i] # full rectangle grid id for SDD indexes
 
   #---- Truth: add immigrants
+  n.sdd <- nrow(which(sdd[,,2,]==sdd.i, arr.ind=TRUE))
+  sdd.j <- which(sdd[,,2,]==sdd.i, arr.ind=TRUE)[,3]
+  p.ij <- sdd[,,1,sdd.j][sdd[,,2,sdd.j]==sdd.i]
+  for(j in 1:n.sdd) {
+    Fs[z.i,z.i,i] <- Fs[z.i, z.i,i] + 
+      h*outer(y, y, calc_rcrImm, p=p, p.ij[j], n_seedz=n_seedz, n_flz=n_flz,
+              X.seed=X_seed[sdd.j[j],], X.fl=X_fl[sdd.j[j],])
+    Fs[1,z.i,i] <- Fs[1,z.i,i] + 
+      calc_ImmSB(y, p=p, p.ij[j], n_seedz=n_seedz, n_flz=n_flz,
+                            X.seed=X_seed[sdd.j[j],], X.fl=X_fl[sdd.j[j],])
+  }
 
@@
   # total immigrants
-  D[i] <- sum(sdd[,,3,][sdd[,,2,]==id.i])
+  D[i] <- sum(sdd[,,3,][sdd[,,2,]==sdd.i])
   # seed bank: immigrants

### 2018-03-05 exploration.Rmd
calculating p_emig from sdd.rate

@@
 ```{r sp_i}
-mean_abund <- 2000
+mean_abund <- 200
 regPlots <- F
@@
 z.rng <- c(1, 12)
-n <- 100  # ncells in IPM matrix
+n <- 50  # ncells in IPM matrix
 cells <- 1:n_distinct(lc.df$CellID)
@@
 sp_i <- list(p=list(s_z=c(-1.75, .75), # b1 + b2*z
-                    s_x=c(1, -.1, -.5, -.1, .05, -2, -.4, -.4), # b1*x1 + ...
+                    s_x=c(1, -.1, -.5, -.1, .05, -2, -.4), # b1*x1 + ...
                     g_z=c(.2, 2, -0.1), # b1 + b2*z + b3*z^2
@@
                     fl_z=c(-1.5, .1, .1), # b1 + b2*z + b3*z^2 + b4*z^3
-                    fl_x=c(1, -.01, -1, -.01, .2, .2), # b1*x1 + ...
+                    fl_x=c(1, -.01, -1, -.01, 0, .2), # b1*x1 + ...
                     seed_z=c(2, 0.5, -.03), # b1 + b2*z + b2*z^2
@@
                     rcr_dir=0.6, # p(recruit directly)
-                    s_SB=0.3, # p(survive in seedbank additional year)
-                    p_emig=0.999, # p(seed emigrates)
-                    sdd_max=3, # max SDD distance in cells
+                    s_SB=0.1, # p(survive in seedbank additional year)
+                    sdd_max=5, # max SDD distance in cells
                     sdd_rate=0.8 # SDD dispersal rate
                     ))
+sp_i$p$p_emig <- pexp(0.5, sp_i$p$sdd_rate, lower.tail=F) # p(seed emigrates)
 ```

### 2018-03-05 exploration.Rmd
vectorized sdd F.mx contributions

@@
   #---- Truth: add immigrants
-  n.sdd <- nrow(which(sdd[,,2,]==sdd.i, arr.ind=TRUE))
   sdd.j <- which(sdd[,,2,]==sdd.i, arr.ind=TRUE)[,3]
   p.ij <- sdd[,,1,sdd.j][sdd[,,2,sdd.j]==sdd.i]
-  for(j in 1:n.sdd) {
-    Fs[z.i,z.i,i] <- Fs[z.i, z.i,i] + 
-      h*outer(y, y, calc_rcrImm, p=p, p.ij[j], n_seedz=n_seedz, n_flz=n_flz,
-              X.seed=X_seed[sdd.j[j],], X.fl=X_fl[sdd.j[j],])
+  Fs[z.i,z.i,i] <- Fs[z.i,z.i,i] + 
+    Reduce(`+`, map2(sdd.j, p.ij, 
+                     ~h*outer(y, y, calc_rcrImm, p=p, .y, n_seedz=n_seedz,
+                              n_flz=n_flz, X.seed=X_seed[.x,], X.fl=X_fl[.x,])))
   Fs[1,z.i,i] <- Fs[1,z.i,i] + 
-      calc_ImmSB(y, p=p, p.ij[j], n_seedz=n_seedz, n_flz=n_flz,
-                            X.seed=X_seed[sdd.j[j],], X.fl=X_fl[sdd.j[j],])
-  }
+    Reduce(`+`, map2(sdd.j, p.ij,
+                     ~calc_ImmSB(y, p=p, .y, n_seedz=n_seedz, n_flz=n_flz,
+                                 X.seed=X_seed[.x,], X.fl=X_fl[.x,])))
 

### 2018-03-05 exploration.Rmd
credentials trigger

@@
                     s_SB=0.1, # p(survive in seedbank additional year)
-                    sdd_max=5, # max SDD distance in cells
-                    sdd_rate=0.8 # SDD dispersal rate
+                    sdd_max=3, # max SDD distance in cells
+                    sdd_rate=1.5 # SDD dispersal rate
                     ))

### 2018-03-05 exploration.Rmd
intermediate storage

@@
   p.ij <- sdd[,,1,sdd.j][sdd[,,2,sdd.j]==sdd.i]
-  Fs[z.i,z.i,i] <- Fs[z.i,z.i,i] + 
-    Reduce(`+`, map2(sdd.j, p.ij, 
+  imm.rcr <- map2(sdd.j, p.ij, 
                      ~h*outer(y, y, calc_rcrImm, p=p, .y, n_seedz=n_seedz,
-                              n_flz=n_flz, X.seed=X_seed[.x,], X.fl=X_fl[.x,])))
+                              n_flz=n_flz, X.seed=X_seed[.x,], X.fl=X_fl[.x,]))
+  Fs[z.i,z.i,i] <- Fs[z.i,z.i,i] + Reduce(`+`, imm.rcr)
+  imm.sb <- map2(sdd.j, p.ij,
+                     ~calc_ImmSB(y, p=p, .y, n_seedz=n_seedz, n_flz=n_flz,
+                                 X.seed=X_seed[.x,], X.fl=X_fl[.x,]))
   Fs[1,z.i,i] <- Fs[1,z.i,i] + 
-    Reduce(`+`, map2(sdd.j, p.ij,
-                     ~calc_ImmSB(y, p=p, .y, n_seedz=n_seedz, n_flz=n_flz,
-                                 X.seed=X_seed[.x,], X.fl=X_fl[.x,])))
+    Reduce(`+`, imm.sb)
 
@@
 ggplot(lam.df, aes(x=lon, y=lat, fill=nSd)) + geom_tile()
+ggplot(lam.df, aes(x=lon, y=lat, fill=D)) + geom_tile()
 ggplot(lam.df, aes(x=lon, y=lat, fill=B1>B0)) + geom_tile()
-ggplot(lam.df, aes(x=lon, y=lat, fill=D)) + geom_tile()
 ggplot(lam.df, aes(x=lon, y=lat, fill=nSd*p$p_emig > D)) + geom_tile()

### 2018-03-05 exploration.Rmd
credentials trigger

@@
 rm(env.rct)
-## sdd.pr[,,2,] has indexes based on env.df$id -- NOT env.df$id.inbd
+# sdd.pr[,,2,] has indexes based on env.df$id -- NOT env.df$id.inbd
 # ggplot(env.df, aes(x=lon, y=lat, fill=id %in% sdd.pr[,,2,760])) + geom_tile()

### 2018-03-05 exploration.Rmd
vectorized dispersal calculations

@@
 setup_IPM_matrix(n, z.rng, buffer=0.25, discrete=1)
-IPMs <- Ps <- Fs <- array(0, dim=c(n+1, n+1, n.cell))
+IPMs <- Ps <- Fs <- Fb <- array(0, dim=c(n+1, n+1, n.cell))
 d <- vector("list", length=n.cell)
@@
   Ps[1,1,i] <- calc_staySB(p)
-  Fs[z.i,z.i,i] <- h*outer(y, y, calc_rcrDir, p=p, n_seedz=n_seedz,
-                             n_flz=n_flz, X.seed=X_seed[i,], X.fl=X_fl[i,])
-  Fs[1,z.i,i] <- calc_DirSB(y, p=p, n_seedz=n_seedz, n_flz=n_flz, 
+  ### store seed production
+  Fb[z.i,z.i,i] <- outer(y, y, calc_rcrDir, p=p, n_seedz=n_seedz, n_flz=n_flz, 
                          X.seed=X_seed[i,], X.fl=X_fl[i,])
+  Fb[1,z.i,i] <- calc_addSB(y, p=p, n_seedz=n_seedz, n_flz=n_flz, 
+                            X.seed=X_seed[i,], X.fl=X_fl[i,])
+  Fs[z.i,z.i,i] <- h * (1 - p$p_emig) * Fb[z.i,z.i,i]
+  Fs[1,z.i,i] <- (1 - p$p_emig) * Fb[1,z.i,i]
   Fs[z.i,1,i] <- calc_rcrSB(y, p)
@@
   p.ij <- sdd[,,1,sdd.j][sdd[,,2,sdd.j]==sdd.i]
-  imm.rcr <- map2(sdd.j, p.ij, 
-                     ~h*outer(y, y, calc_rcrImm, p=p, .y, n_seedz=n_seedz,
-                              n_flz=n_flz, X.seed=X_seed[.x,], X.fl=X_fl[.x,]))
-  Fs[z.i,z.i,i] <- Fs[z.i,z.i,i] + Reduce(`+`, imm.rcr)
-  imm.sb <- map2(sdd.j, p.ij,
-                     ~calc_ImmSB(y, p=p, .y, n_seedz=n_seedz, n_flz=n_flz,
-                                 X.seed=X_seed[.x,], X.fl=X_fl[.x,]))
-  Fs[1,z.i,i] <- Fs[1,z.i,i] + 
-    Reduce(`+`, imm.sb)
+  Fs[z.i,z.i,i] <- Fs[z.i,z.i,i] + 
+    Reduce(`+`, map2(sdd.j, p$p_emig*p.ij, ~(Fb[z.i,z.i,.x] * .y)))
+  Fs[1,z.i,i] <- Fs[1,z.i,i] + Fb[1,z.i,sdd.j] %*% (p$p_emig*p.ij)
 

### 2018-03-05 code/fn_IPM.R

@@ -67,14 +67,13 @@ calc_seeds <- function(z.v, p, n_seeds, X.seed) {
 }
 
 
-##-- Recruits (direct, native)
-##   z.ii' ~ P(fl.i) * nSeeds.i (1-p.emig) * P(rcrDir) * P(estab) * N(mn, sd)
+##-- Recruits (direct)
+##   z' ~ P(fl) * nSeeds * P(rcrDir) * P(estab) * N(mn, sd)
 calc_rcrDir <- function(z1, z.v, p, n_seedz, n_flz, X.seed, X.fl) {
   calc_flwr(z.v, p, n_flz, X.fl) *
     calc_seeds(z.v, p, n_seedz, X.seed) *
-    (1 - p$p_emig) *
-    p$rcr_dir * 
-    p$p_est * 
+    p$rcr_dir *
+    p$p_est *
     dnorm(z1, p$rcr_z[1], p$rcr_z[2])
 }
 
@@ -86,39 +85,12 @@ calc_rcrSB <- function(z1, p) {
 }
 
 
-##-- Recruits (direct, immigrant)
-##   z.ij' ~ P(fl).j * nSeeds.j * p.emig * P(rcrDir) * P(estab) * N(mn, sd)
-calc_rcrImm <- function(z1, z.v, p, p_ij, n_seedz, n_flz, X.seed, X.fl) {
-  calc_flwr(z.v, p, n_flz, X.fl) *
-    calc_seeds(z.v, p, n_seedz, X.seed) *
-    p$p_emig *
-    p_ij *
-    p$rcr_dir * 
-    p$p_est * 
-    dnorm(z1, p$rcr_z[1], p$rcr_z[2])
-}
-
-
-##-- Add to seedbank (native)
-##   B(z).ii ~ P(fl).i * nSeeds.i * (1-P(rcrDir)) * P(s.SB)
-calc_DirSB <- function(z.v, p, n_seedz, n_flz, X.seed, X.fl) {
+##-- Add to seedbank
+##   B(z) ~ P(fl) * nSeeds * (1-P(rcrDir)) * P(s.SB)
+calc_addSB <- function(z.v, p, n_seedz, n_flz, X.seed, X.fl) {
   z <- z_pow(z.v, n_seedz)
   calc_flwr(z.v, p, n_flz, X.fl) *
     calc_seeds(z.v, p, n_seedz, X.seed) *
-    (1 - p$p_emig) *
[…]

### 2018-03-05 exploration.Rmd
added plot of relative seed influx vs production

@@
 lam.df$lambda <- apply(IPMs, 3, function(x) Re(eigen(x)$values[1]))
-lam.df %<>% add_column(nSd=nSd, B0=B, B1=B_1, D=D)
+lam.df %<>% add_column(nSd=nSd, B0=B, B1=B_1, D=D) %>%
+  mutate(nSdStay=nSd*(1-p$p_emig), nSdLeave=nSd*p$p_emig)
 
@@
 ggplot(lam.df, aes(x=lon, y=lat, fill=B1>B0)) + geom_tile()
+ggplot(lam.df, aes(x=lon, y=lat, fill=(D-nSdStay)/nSdStay)) + geom_tile() +
+  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
 ggplot(lam.df, aes(x=lon, y=lat, fill=nSd*p$p_emig > D)) + geom_tile()
-ggplot(lam.df, aes(x=lon, y=lat, fill=nSd*(1-p$p_emig) > D)) + geom_tile()
-ggplot(lam.df, aes(x=lon, y=lat, fill=nSd*(1-p$p_emig) < D & lambda < 1)) +
+ggplot(lam.df, aes(x=lon, y=lat, fill=nSdStay > D)) + geom_tile()
+ggplot(lam.df, aes(x=lon, y=lat, fill=nSdStay < D & lambda < 1)) +
   geom_tile()

### 2018-03-06 exploration.Rmd
confirmed operations, vectorized a bit more

--- 2018-03-05 exploration.Rmd
@@
   #---- Truth: add immigrants
-  sdd.j <- which(sdd[,,2,]==sdd.i, arr.ind=TRUE)[,3]
-  p.ij <- sdd[,,1,sdd.j][sdd[,,2,sdd.j]==sdd.i]
+  sdd.j <- which(sdd[,,2,]==sdd.i, arr.ind=TRUE)
+  p.ij <- sdd[,,1,][sdd.j]
   Fs[z.i,z.i,i] <- Fs[z.i,z.i,i] + 
-    Reduce(`+`, map2(sdd.j, p$p_emig*p.ij, ~(Fb[z.i,z.i,.x] * .y)))
-  Fs[1,z.i,i] <- Fs[1,z.i,i] + Fb[1,z.i,sdd.j] %*% (p$p_emig*p.ij)
+    h * Reduce(`+`, map2(sdd.j[,3], p$p_emig*p.ij, ~(Fb[z.i,z.i,.x] * .y)))
+  Fs[1,z.i,i] <- Fs[1,z.i,i] + Fb[1,z.i,sdd.j[,3]] %*% (p$p_emig*p.ij)
 
@@
   # total immigrants
-  D[i] <- sum(sdd[,,3,][sdd[,,2,]==sdd.i])
+  D[i] <- sum(sdd[,,3,][sdd.j])
   # seed bank: immigrants
@@
 ggplot(lam.df, aes(x=lon, y=lat, fill=B1>B0)) + geom_tile()
-ggplot(lam.df, aes(x=lon, y=lat, fill=nSd*p$p_emig > D)) + geom_tile()
-ggplot(lam.df, aes(x=lon, y=lat, fill=nSd*(1-p$p_emig) > D)) + geom_tile()
-ggplot(lam.df, aes(x=lon, y=lat, fill=nSd*(1-p$p_emig) < D & lambda < 1)) +
+ggplot(lam.df, aes(x=lon, y=lat, fill=(D-nSdStay)/nSdStay)) + geom_tile() +
+  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
+ggplot(lam.df, aes(x=lon, y=lat, fill=nSdStay > D)) + geom_tile()
+ggplot(lam.df, aes(x=lon, y=lat, fill=nSdStay < D & lambda < 1)) +
   geom_tile()

### 2018-03-06 exploration.Rmd
shifting step size multiplication

@@
   ### store seed production
-  Fb[z.i,z.i,i] <- outer(y, y, calc_rcrDir, p=p, n_seedz=n_seedz, n_flz=n_flz, 
+  Fb[z.i,z.i,i] <- h * outer(y, y, calc_rcrDir, p=p, n_seedz=n_seedz, n_flz=n_flz, 
                          X.seed=X_seed[i,], X.fl=X_fl[i,])
-  Fb[1,z.i,i] <- calc_addSB(y, p=p, n_seedz=n_seedz, n_flz=n_flz, 
+  Fb[1,z.i,i] <- h * calc_addSB(y, p=p, n_seedz=n_seedz, n_flz=n_flz, 
                             X.seed=X_seed[i,], X.fl=X_fl[i,])
-  Fs[z.i,z.i,i] <- h * (1 - p$p_emig) * Fb[z.i,z.i,i]
+  Fs[z.i,z.i,i] <- (1 - p$p_emig) * Fb[z.i,z.i,i]
   Fs[1,z.i,i] <- (1 - p$p_emig) * Fb[1,z.i,i]
@@
   Fs[z.i,z.i,i] <- Fs[z.i,z.i,i] + 
-    h * Reduce(`+`, map2(sdd.j[,3], p$p_emig*p.ij, ~(Fb[z.i,z.i,.x] * .y)))
+    Reduce(`+`, map2(sdd.j[,3], p$p_emig*p.ij, ~(Fb[z.i,z.i,.x] * .y)))
   Fs[1,z.i,i] <- Fs[1,z.i,i] + Fb[1,z.i,sdd.j[,3]] %*% (p$p_emig*p.ij)

### 2018-03-06 exploration.Rmd
beginning MaxEnt sampling
functional but terrible version of maxent

@@
 ```{r O_Mx}
-
+library(dismo)
+Mx.n <- 3000
+Mx.P <- sample((1:nrow(lam.df))[lam.df$lambda > 1], Mx.n, replace=T)
+pres <- 1:nrow(lam.df) %in% Mx.P
+Mx.out <- maxent(x=lam.df[,c(1,3,5:9)], p=pres)
+lam.df$MxPred <- 1:nrow(lam.df) %in% row.names(Mx.out@presence)
+lam.df$MxCorrect <- with(lam.df, (MxPred==1 & lambda > 1) |
+                           (MxPred==0 & lambda < 1))
+ggplot(lam.df, aes(x=lon, y=lat, fill=MxCorrect)) + geom_tile()
 ```

### 2018-03-06 exploration.Rmd
switch to make detailed plots for only a small number of random cells

@@
 mean_abund <- 200
-regPlots <- F
-ipmPlots <- F
+regPlots <- T
+ipmPlots <- T
+nPlots <- 5
 z.rng <- c(1, 12)
@@
                     rcr_dir=0.6, # p(recruit directly)
-                    s_SB=0.1, # p(survive in seedbank additional year)
-                    sdd_max=3, # max SDD distance in cells
-                    sdd_rate=1.5 # SDD dispersal rate
+                    s_SB=0.3, # p(survive in seedbank additional year)
+                    sdd_max=7, # max SDD distance in cells
+                    sdd_rate=1 # SDD dispersal rate
                     ))
@@
   ### store seed production
-  Fb[z.i,z.i,i] <- h * outer(y, y, calc_rcrDir, p=p, n_seedz=n_seedz, n_flz=n_flz, 
-                         X.seed=X_seed[i,], X.fl=X_fl[i,])
+  Fb[z.i,z.i,i] <- h * outer(y, y, calc_rcrDir, p=p, n_seedz=n_seedz, 
+                             n_flz=n_flz, X.seed=X_seed[i,], X.fl=X_fl[i,])
   Fb[1,z.i,i] <- h * calc_addSB(y, p=p, n_seedz=n_seedz, n_flz=n_flz, 
@@
 
+sample.cells <- sample.int(n.cell, nPlots)
 if(regPlots) {
@@
   xxx <- seq(0, hi*.5, length.out=100)
-  for(i in 1:20) {
+  for(i in sample.cells) {
     z.t <- d[[i]][!is.na(d[[i]]$size),]; main.i <- paste("Cell", i)
@@
   x <- c(0, y)
-  for(i in 1:n.cell) {
+  for(i in sample.cells) {
     z.t <- d[[i]][!is.na(d[[i]]$size),]; main.i <- paste("Cell", i)

### 2018-03-08 code/fn_IPM.R
moving p_est out of function to allow density dependence

@@ -68,20 +68,19 @@ calc_seeds <- function(z.v, p, n_seeds, X.seed) {
 
 
 ##-- Recruits (direct)
-##   z' ~ P(fl) * nSeeds * P(rcrDir) * P(estab) * N(mn, sd)
+##   z' ~ P(fl) * nSeeds * P(estab) * N(mn, sd)
 calc_rcrDir <- function(z1, z.v, p, n_seedz, n_flz, X.seed, X.fl) {
   calc_flwr(z.v, p, n_flz, X.fl) *
     calc_seeds(z.v, p, n_seedz, X.seed) *
     p$rcr_dir *
-    p$p_est *
     dnorm(z1, p$rcr_z[1], p$rcr_z[2])
 }
 
 
 ##-- Recruits (seedbank)
-##   z' ~ P(s.SB) * P(estab) * N(mn, sd)
+##   z' ~ P(s.SB) * N(mn, sd)
 calc_rcrSB <- function(z1, p) {
-  p$s_SB * p$p_est * dnorm(z1, p$rcr_z[1], p$rcr_z[2])
+  p$s_SB * dnorm(z1, p$rcr_z[1], p$rcr_z[2])
 }
 
 

### 2018-03-08 exploration.Rmd
projecting populations forward with density dependence

@@
 n <- 50  # ncells in IPM matrix
+tmax <- 10
 cells <- 1:n_distinct(lc.df$CellID)
@@
                     rcr_z=c(1.5, 0.4), # N(mean=rcrt1, sd=rcrt2)
-                    p_est=0.02, # p(establishment)
-                    rcr_SB=0.2, # p(recruit from seedbank)
-                    rcr_dir=0.6, # p(recruit directly)
+                    p_est=0.03, # p(establishment)
+                    NDD=T, # negative density dependence: p_est=p_est/N
+                    rcr_SB=0.5, # p(recruit from seedbank)
+                    rcr_dir=0.9, # p(recruit directly)
                     s_SB=0.3, # p(survive in seedbank additional year)
-                    sdd_max=7, # max SDD distance in cells
+                    sdd_max=4, # max SDD distance in cells
                     sdd_rate=1 # SDD dispersal rate
@@
 IPMs <- Ps <- Fs <- Fb <- array(0, dim=c(n+1, n+1, n.cell))
+Nt <- array(dim=c(n+1, n+1, n.cell, tmax))
 d <- vector("list", length=n.cell)
 B <- rpois(n.cell, n_i)
-B_1 <- nSd <- D <- rep(NA, n.cell)
+B_1 <- nSd <- D <- p_est.i <- rep(NA, n.cell)
 sdd <- array(0, dim=dim(sdd.pr) + c(0,0,1,0)) # add layer for nEmig
@@
 ########
+# local growth
 for(i in 1:n.cell) {
@@
   ### store seed production
-  Fb[z.i,z.i,i] <- h * outer(y, y, calc_rcrDir, p=p, n_seedz=n_seedz, 
-                             n_flz=n_flz, X.seed=X_seed[i,], X.fl=X_fl[i,])
-  Fb[1,z.i,i] <- h * calc_addSB(y, p=p, n_seedz=n_seedz, n_flz=n_flz, 
-                            X.seed=X_seed[i,], X.fl=X_fl[i,])
+  Fb[z.i,z.i,i] <- outer(y, y, calc_rcrDir, p=p, n_seedz=n_seedz, n_flz=n_flz, 
+                         X.seed=X_seed[i,], X.fl=X_fl[i,]) *h * p$p_est
+  Fb[1,z.i,i] <- calc_addSB(y, p=p, n_seedz=n_seedz, n_flz=n_flz, 
+                            X.seed=X_seed[i,], X.fl=X_fl[i,]) * h
   Fs[z.i,z.i,i] <- (1 - p$p_emig) * Fb[z.i,z.i,i]
   Fs[1,z.i,i] <- (1 - p$p_emig) * Fb[1,z.i,i]
-  Fs[z.i,1,i] <- calc_rcrSB(y, p)
+  Fs[z.i,1,i] <- calc_rcrSB(y, p) * p$p_est
 
@@
   ## recruits: native
-  rcr.n <- max(ceiling(nSd[i] * (1 - p$p_emig) * p$rcr_dir * p$p_est +
-                   B[i] * p$rcr_SB * p$p_est), 2) # ≥ 2 for IPMpack sd(rcr_z')
+  p_est.i[i] <- ifelse(p$NDD, sum(d[[i]]$surv==1)/nSd[i], p$p_est)
+  rcr.n <- ceiling(nSd[i] * (1 - p$p_emig) * p$rcr_dir * p_est.i[i] +
+                   B[i] * p$rcr_SB * p_est.i[i]) # ≥ 2 for IPMpack sd(rcr_z')
   rcr.i <- (1:rcr.n) + n_i[i]
@@
 }
+# dispersal
 for(i in 1:n.cell) {
@@
   # recruits: immigrants
-  rcr.imm <- max(ceiling(D[i] * p$rcr_dir * p$p_est), 2)
+  rcr.imm <- ceiling(D[i] * p$rcr_dir * p_est.i[i])
   rcr.i <- (1:rcr.imm) + nrow(d[[i]])
@@
   d[[i]]$sizeNext[rcr.i] <- rnorm(rcr.imm, p$rcr_z[1], p$rcr_z[2])
+}
+# density dependence
+if(p$NDD) {
+  lam.t <- p_est.t <- matrix(nrow=tmax-1, ncol=n.cell)
+  p_est.t[1,] <- p$p_est
+  for(i in 1:n.cell) {
+    Nt[,,i,1] <- rpois((n+1)*(n+1), n_i)
+    for(k in 1:(tmax-1)) {
+      F.t <- Fs[,,i]
+      p_est.t[k,i] <- min(195/(sum(F.t[z.i,] * Nt[z.i,,i,k])/p$p_est), p$p_est)
+      F.t[z.i,] <- F.t[z.i,]/p$p_est * p_est.t[k,i]
+      IPM.t <- Ps[,,i] + F.t
+      Nt[,,i,k+1] <- IPM.t %*% Nt[,,i,k]
+      lam.t[k,i] <- Re(eigen(IPM.t)$values[1])
+    }
+  }
 }
@@
 lam.df %<>% add_column(nSd=nSd, B0=B, B1=B_1, D=D) %>%
-  mutate(nSdStay=nSd*(1-p$p_emig), nSdLeave=nSd*p$p_emig)
+  mutate(nSdStay=nSd*(1-p$p_emig), nSdLeave=nSd*p$p_emig) %>%
+  add_column(Ntmax=apply(Nt[,,,tmax],3,sum), lam.tmax=lam.t[tmax-1,])
 
@@
   scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
+ggplot(lam.df, aes(x=lon, y=lat, fill=log(lam.tmax))) + geom_tile() +
+  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
 ggplot(lam.df, aes(x=lon, y=lat, fill=lambda>1)) + geom_tile()
+ggplot(lam.df, aes(x=lon, y=lat, fill=lam.tmax>1)) + geom_tile()
 ggplot(lam.df, aes(x=lon, y=lat, fill=nSd)) + geom_tile()

### 2018-03-09 exploration.Rmd
rounding recruit count instead of ceiling; additional plots

@@
 ```{r sp_i}
-mean_abund <- 200
+mean_abund <- 500
 regPlots <- T
@@
 n <- 50  # ncells in IPM matrix
-tmax <- 10
+tmax <- 100
 cells <- 1:n_distinct(lc.df$CellID)
@@
 sp_i <- list(p=list(s_z=c(-1.75, .75), # b1 + b2*z
-                    s_x=c(1, -.1, -.5, -.1, .05, -2, -.4), # b1*x1 + ...
+                    s_x=c(2, -.1, -.5, -.1, .05, -2, -.4), # b1*x1 + ...
                     g_z=c(.2, 2, -0.1), # b1 + b2*z + b3*z^2
-                    g_x=c(1, -.1, 1.5, -.1, .1, -2), #b1*x1 + ...
+                    g_x=c(2, -.1, 1.5, -.1, .1, -2), #b1*x1 + ...
                     g_sig=1, # growth ~ N(E, g_sig)
-                    fl_z=c(-1.5, .1, .1), # b1 + b2*z + b3*z^2 + b4*z^3
-                    fl_x=c(1, -.01, -1, -.01, 0, .2), # b1*x1 + ...
+                    fl_z=c(-1.5, .1, .1), # b1 + b2*z + b3*z^2
+                    fl_x=c(2, -.01, -1, -.01, 0, .2), # b1*x1 + ...
                     seed_z=c(2, 0.5, -.03), # b1 + b2*z + b2*z^2
@@
                     rcr_SB=0.5, # p(recruit from seedbank)
-                    rcr_dir=0.9, # p(recruit directly)
+                    rcr_dir=0.8, # p(recruit directly)
                     s_SB=0.3, # p(survive in seedbank additional year)
@@
   p_est.i[i] <- ifelse(p$NDD, sum(d[[i]]$surv==1)/nSd[i], p$p_est)
-  rcr.n <- ceiling(nSd[i] * (1 - p$p_emig) * p$rcr_dir * p_est.i[i] +
+  rcr.n <- round(nSd[i] * (1 - p$p_emig) * p$rcr_dir * p_est.i[i] +
                    B[i] * p$rcr_SB * p_est.i[i]) # ≥ 2 for IPMpack sd(rcr_z')
+  if(rcr.n > 0) {
     rcr.i <- (1:rcr.n) + n_i[i]
@@
     d[[i]]$sizeNext[rcr.i] <- rnorm(rcr.n, p$rcr_z[1], p$rcr_z[2])
+  }
   ## emigrants
@@
   # recruits: immigrants
-  rcr.imm <- ceiling(D[i] * p$rcr_dir * p_est.i[i])
+  rcr.imm <- round(D[i] * p$rcr_dir * p_est.i[i])
+  if(rcr.imm > 0) {
     rcr.i <- (1:rcr.imm) + nrow(d[[i]])
@@
   }
+}
 # density dependence
@@
   lam.t <- p_est.t <- matrix(nrow=tmax-1, ncol=n.cell)
-  p_est.t[1,] <- p$p_est
   for(i in 1:n.cell) {
@@
       F.t <- Fs[,,i]
+      # set p_est to density dependent value if that is < p$p_est
       p_est.t[k,i] <- min(195/(sum(F.t[z.i,] * Nt[z.i,,i,k])/p$p_est), p$p_est)
+      # F.s[,,i] was already multiplied by p$p_est, so divide by that first
       F.t[z.i,] <- F.t[z.i,]/p$p_est * p_est.t[k,i]
@@
 lam.df$lambda <- apply(IPMs, 3, function(x) Re(eigen(x)$values[1]))
-lam.df %<>% add_column(nSd=nSd, B0=B, B1=B_1, D=D) %>%
+lam.df %<>% 
+  add_column(nSd=nSd, B0=B, B1=B_1, D=D, N=map_dbl(d, ~sum(!is.na(.$sizeNext))),
+             N.surv=map_dbl(d, ~sum(.$surv, na.rm=T)),
+             N.rcr=map_dbl(d, ~sum(is.na(.$size)))) %>%
   mutate(nSdStay=nSd*(1-p$p_emig), nSdLeave=nSd*p$p_emig) %>%
@@
   scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
+ggplot(lam.df, aes(x=lon, y=lat, fill=log(Ntmax))) + geom_tile() +
+  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=1)
 ggplot(lam.df, aes(x=lon, y=lat, fill=lambda>1)) + geom_tile()
 ggplot(lam.df, aes(x=lon, y=lat, fill=lam.tmax>1)) + geom_tile()
+ggplot(lam.df, aes(x=lon, y=lat, fill=Ntmax>1)) + geom_tile() 
+ggplot(lam.df, aes(x=lon, y=lat, fill=Ntmax>1 & lam.tmax<1)) + geom_tile() 
 ggplot(lam.df, aes(x=lon, y=lat, fill=nSd)) + geom_tile()
 ggplot(lam.df, aes(x=lon, y=lat, fill=D)) + geom_tile()
+ggplot(lam.df, aes(x=lon, y=lat, fill=D + nSdStay)) + geom_tile()
 ggplot(lam.df, aes(x=lon, y=lat, fill=B1>B0)) + geom_tile()
-ggplot(lam.df, aes(x=lon, y=lat, fill=(D-nSdStay)/nSdStay)) + geom_tile() +
-  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
+ggplot(lam.df, aes(x=lon, y=lat, fill=nSdStay/D)) + geom_tile() +
+  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=1)
 ggplot(lam.df, aes(x=lon, y=lat, fill=nSdStay > D)) + geom_tile()

### 2018-03-09 exploration.Rmd
made NDD consistent in simulation and theoretical

@@
   ## recruits: native
-  p_est.i[i] <- ifelse(p$NDD, sum(d[[i]]$surv==1)/nSd[i], p$p_est)
+  p_est.i[i] <- ifelse(p$NDD, min(150/nSd[i], p$p_est), p$p_est)
   rcr.n <- round(nSd[i] * (1 - p$p_emig) * p$rcr_dir * p_est.i[i] +
@@
       # set p_est to density dependent value if that is < p$p_est
-      p_est.t[k,i] <- min(195/(sum(F.t[z.i,] * Nt[z.i,,i,k])/p$p_est), p$p_est)
+      p_est.t[k,i] <- min(150/(sum(F.t[z.i,] * Nt[z.i,,i,k])/p$p_est), p$p_est)
       # F.s[,,i] was already multiplied by p$p_est, so divide by that first

### 2018-03-09 code/fn_IPM.R
moved IPM matrix set up to fn_IPM.R

@@ -31,6 +31,10 @@
 # of columns where: z.mx[,1]=1; z.mx[,2]=z; z.mx[,3]=z^2; z.mx[,4]=z^3; etc
 
 
+################################################################################
+## Vital rate functions
+########
+
 ##-- Survival
 ##   P(s) ~ antilogit(size + environment)
 calc_surv <- function(z.v, p, n_sz, X.s) {
@@ -100,5 +104,32 @@ calc_addSB <- function(z.v, p, n_seedz, n_flz, X.seed, X.fl) {
 calc_staySB <- function(p) {
   p$s_SB * (1 - p$rcr_SB)
 }
+
+
+
+
+
+################################################################################
+## IPM matrix construction
+########
+
+##-- Set boundaries, meshpoints, & step size for IPM matrix
+##   create objects in global environment
+setup_IPM_matrix <- function(n=100, z.rng=c(1,10), buffer=0.25, discrete=0) {
+  lo <<- z.rng[1]*(1-buffer)  # IPM matrix lower limit
+  hi <<- z.rng[2]*(1+buffer)  # IPM matrix upper limit
+  b <<- lo + c(0:n)*(hi - lo)/n  # boundary points
+  y <<- 0.5*(b[1:n] + b[2:(n+1)])  # mesh points
+  h <<- y[2] - y[1]  # step size
+  z.i <<- (1:n) + discrete  # continuous stage matrix indices
+}
+
+
+##-- Fill P & F matrices: local 
+##   
+
+
+
+
 
 

### 2018-03-09 code/fn_aux.R

@@ -3,18 +3,6 @@
 # Tim Szewczyk
 
 
-##-- set boundaries, meshpoints, step size for IPM matrix
-##   create objects in global environment
-setup_IPM_matrix <- function(n=100, z.rng=c(1,10), buffer=0.25, discrete=0) {
-  lo <<- z.rng[1]*(1-buffer)  # IPM matrix lower limit
-  hi <<- z.rng[2]*(1+buffer)  # IPM matrix upper limit
-  b <<- lo + c(0:n)*(hi - lo)/n  # boundary points
-  y <<- 0.5*(b[1:n] + b[2:(n+1)])  # mesh points
-  h <<- y[2] - y[1]  # step size
-  z.i <<- (1:n) + discrete  # continuous stage matrix indices
-}
-
-
 ##-- create design matrix from vector of sizes
 z_pow <- function(z.vec, n_z) {
   if(n_z==1) {

### 2018-03-10 exploration.Rmd
simulated data across time

--- 2018-03-08 exploration.Rmd
@@
 theme_set(theme_bw())
-source("code/fn_IPM.R"); source("code/fn_aux.R")
+source("code/fn_IPM.R"); source("code/fn_aux.R"); source("code/fn_sim.R")
 lc.df <- read_csv("data/landcover_5km.csv") %>% 
@@
 ```{r sp_i}
-mean_abund <- 200
-regPlots <- T
-ipmPlots <- T
+mean_abund <- 100
+regPlots <- F
+ipmPlots <- F
 nPlots <- 5
 z.rng <- c(1, 12)
-n <- 50  # ncells in IPM matrix
-tmax <- 10
+n <- 10  # ncells in IPM matrix
+tmax <- 30
 cells <- 1:n_distinct(lc.df$CellID)
@@
 sp_i <- list(p=list(s_z=c(-1.75, .75), # b1 + b2*z
-                    s_x=c(1, -.1, -.5, -.1, .05, -2, -.4), # b1*x1 + ...
+                    s_x=c(2, -.1, -2, -.1, 2, -2, -.4), # b1*x1 + ...
                     g_z=c(.2, 2, -0.1), # b1 + b2*z + b3*z^2
-                    g_x=c(1, -.1, 1.5, -.1, .1, -2), #b1*x1 + ...
+                    g_x=c(2, -.1, 2, -.1, 2, -2), #b1*x1 + ...
                     g_sig=1, # growth ~ N(E, g_sig)
-                    fl_z=c(-1.5, .1, .1), # b1 + b2*z + b3*z^2 + b4*z^3
-                    fl_x=c(1, -.01, -1, -.01, 0, .2), # b1*x1 + ...
+                    fl_z=c(-1.5, .1, .1), # b1 + b2*z + b3*z^2
+                    fl_x=c(2, -.1, -2, -.1, 1, 1), # b1*x1 + ...
                     seed_z=c(2, 0.5, -.03), # b1 + b2*z + b2*z^2
-                    seed_x=c(1, -.1, 1, -.1, .2), # b1*x1 + ...
+                    seed_x=c(1, -.1, -1, -.1, .2), # b1*x1 + ...
                     rcr_z=c(1.5, 0.4), # N(mean=rcrt1, sd=rcrt2)
@@
                     NDD=T, # negative density dependence: p_est=p_est/N
+                    NDD_n=100, # mean number of recruits if NDD
                     rcr_SB=0.5, # p(recruit from seedbank)
-                    rcr_dir=0.9, # p(recruit directly)
+                    rcr_dir=0.5, # p(recruit directly)
                     s_SB=0.3, # p(survive in seedbank additional year)
-                    sdd_max=4, # max SDD distance in cells
+                    sdd_max=5, # max SDD distance in cells
                     sdd_rate=1 # SDD dispersal rate
@@
 # sdd.pr[,,2,] has indexes based on env.df$id -- NOT env.df$id.inbd
-# ggplot(env.df, aes(x=lon, y=lat, fill=id %in% sdd.pr[,,2,760])) + geom_tile()
+# ggplot(env.df, aes(x=lon, y=lat, fill=id %in% sdd.pr[,,2,740])) + geom_tile()
 ```
@@
 p <- sp_i$p
-setup_IPM_matrix(n, z.rng, buffer=0.25, discrete=1)
+setup_IPM_matrix(n, z.rng, buffer=0.5, discrete=1)
 IPMs <- Ps <- Fs <- Fb <- array(0, dim=c(n+1, n+1, n.cell))
 Nt <- array(dim=c(n+1, n+1, n.cell, tmax))
-d <- vector("list", length=n.cell)
-B <- rpois(n.cell, n_i)
-B_1 <- nSd <- D <- p_est.i <- rep(NA, n.cell)
+d <- sdd.j <- vector("list", length=n.cell)
+B <- cbind(rpois(n.cell, n_i), matrix(0, nrow=n.cell, ncol=tmax))
+nSd <- D <- p_est.i <- matrix(nrow=n.cell, ncol=tmax)
+lam.t <- p_est.t <- matrix(nrow=n.cell, ncol=tmax-1)
 sdd <- array(0, dim=dim(sdd.pr) + c(0,0,1,0)) # add layer for nEmig
@@
 # set up environmental covariates
-n_sz <- length(p$s_z); n_sx <- length(p$s_x)
-n_gz <- length(p$g_z); n_gx <- length(p$g_x)
-n_flz <- length(p$fl_z); n_flx <- length(p$fl_x)
-n_seedz <- length(p$seed_z); n_seedx <- length(p$seed_x)
-X_s <- as.matrix(env.df[,1:n_sx])
-X_g <- as.matrix(env.df[,1:n_gx])
-X_fl <- as.matrix(env.df[,1:n_flx])
-X_seed <- as.matrix(env.df[,1:n_seedx])
+n_z <- list(s=length(p$s_z), g=length(p$g_z),
+            fl=length(p$fl_z), seed=length(p$seed_z))
+n_x <- list(s=length(p$s_x), g=length(p$g_x),
+            fl=length(p$fl_x), seed=length(p$seed_x))
+X <- list(s=as.matrix(env.df[,1:n_x$s]), g=as.matrix(env.df[,1:n_x$g]),
+          fl=as.matrix(env.df[,1:n_x$fl]), seed=as.matrix(env.df[,1:n_x$seed]))
 
@@
 ########
-# local growth
-for(i in 1:n.cell) {
-  d[[i]] <- data.frame(size=runif(n=n_i[i], min=z.rng[1], max=z.rng[2]))
-  size.mx <- cbind(1, d[[i]]$size, d[[i]]$size^2, d[[i]]$size^3)
 
 #---- Truth: use assigned slopes to fill IPM matrix
-  ### growth and survival
-  G.mx <- h*outer(y, y, calc_grow, p=p, n_gz=n_gz, X.g=X_g[i,])
-  S.v <- calc_surv(y, p=p, n_sz=n_sz, X.s=X_s[i,])
-  ### Fix eviction of individuals below/above size limits
-  for(k in 1:(n/2)) G.mx[1,k] <- G.mx[1,k] + 1 - sum(G.mx[,k])
-  for(k in (n/2+1):n) G.mx[n,k] <- G.mx[n,k] + 1 - sum(G.mx[,k])
-  ### fill P and F matrices
-  for(k in z.i) Ps[k,z.i,i] <- G.mx[k-1,]*S.v
-  Ps[1,1,i] <- calc_staySB(p)
-  ### store seed production
-  Fb[z.i,z.i,i] <- outer(y, y, calc_rcrDir, p=p, n_seedz=n_seedz, n_flz=n_flz, 
-                         X.seed=X_seed[i,], X.fl=X_fl[i,]) *h * p$p_est
-  Fb[1,z.i,i] <- calc_addSB(y, p=p, n_seedz=n_seedz, n_flz=n_flz, 
-                            X.seed=X_seed[i,], X.fl=X_fl[i,]) * h
-  Fs[z.i,z.i,i] <- (1 - p$p_emig) * Fb[z.i,z.i,i]
-  Fs[1,z.i,i] <- (1 - p$p_emig) * Fb[1,z.i,i]
-  Fs[z.i,1,i] <- calc_rcrSB(y, p) * p$p_est
-
-  #---- generate data
-  ## survival
-  E_surv <- antilogit(size.mx[,1:n_sz] %*% p$s_z + c(X_s[i,] %*% p$s_x))
-  d[[i]]$surv <- rbinom(n_i[i], 1, E_surv)
-  ## growth
-  E_grow <- size.mx[,1:n_gz] %*% p$g_z + c(X_g[i,] %*% p$g_x)
-  d[[i]]$sizeNext <- rnorm(n_i[i], E_grow, p$g_sig)
-  d[[i]]$sizeNext <- ifelse(d[[i]]$surv, d[[i]]$sizeNext, NA)
-  ## flowering 
-  E_fl <- antilogit(size.mx[,1:n_flz] %*% p$fl_z + c(X_fl[i,] %*% p$fl_x))
-  d[[i]]$fl <- rbinom(n_i[i], 1, E_fl)
-  ## seed production
-  E_seed <- d[[i]]$fl * 
-    exp(size.mx[,1:n_seedz] %*% p$seed_z + c(X_seed[i,] %*% p$seed_x))
-  d[[i]]$seed <- rpois(n_i[i], E_seed)
-  # d[[i]]$fl <- ifelse(d[[i]]$surv, d[[i]]$fl, NA)
-  d[[i]]$seed <- ifelse(d[[i]]$fl==1, d[[i]]$seed, NA)
-  nSd[i] <- sum(d[[i]]$seed, na.rm=T)
-  ## seed bank: native
-  B_1[i] <- B[i] * p$s_SB * (1 - p$rcr_SB) + 
-    nSd[i] * (1 - p$p_emig) * (1 - p$rcr_dir) * p$s_SB
-  ## recruits: native
-  p_est.i[i] <- ifelse(p$NDD, sum(d[[i]]$surv==1)/nSd[i], p$p_est)
-  rcr.n <- ceiling(nSd[i] * (1 - p$p_emig) * p$rcr_dir * p_est.i[i] +
-                   B[i] * p$rcr_SB * p_est.i[i]) # ≥ 2 for IPMpack sd(rcr_z')
-  rcr.i <- (1:rcr.n) + n_i[i]
-  d[[i]][rcr.i,] <- NA
-  d[[i]]$sizeNext[rcr.i] <- rnorm(rcr.n, p$rcr_z[1], p$rcr_z[2])
-  ## emigrants
-  sdd[,,3,i] <- nSd[i] * p$p_emig * sdd[,,1,i]
+## local growth
+for(i in 1:n.cell) {
+  Ps[,,i] <- fill_P(h, y, z.i, n, p, n_z, n_x, X$s[i,], X$g[i,])
+  Fb[,,i] <- fill_F(h, y, z.i, n, p, n_z, n_x, X$fl[i,], X$seed[i,])
+  Fs[z.i,z.i,i] <- (1-p$p_emig) * Fb[z.i,z.i,i]
+  Fs[1,z.i,i] <- (1-p$p_emig) * Fb[1,z.i,i]
+  Fs[z.i,1,i] <- Fb[z.i,1,i]
 }
-# dispersal
+## dispersal & density dependence
 for(i in 1:n.cell) {
   sdd.i <- env.df$id[i] # full rectangle grid id for SDD indexes
-
-  #---- Truth: add immigrants
-  sdd.j <- which(sdd[,,2,]==sdd.i, arr.ind=TRUE)
-  p.ij <- sdd[,,1,][sdd.j]
+  sdd.j[[i]] <- which(sdd[,,2,]==sdd.i, arr.ind=TRUE) # j cells with target i
+  p.ij <- sdd[,,1,][sdd.j[[i]]] # p(j to i | immigrant seed from j)
   Fs[z.i,z.i,i] <- Fs[z.i,z.i,i] + 
-    Reduce(`+`, map2(sdd.j[,3], p$p_emig*p.ij, ~(Fb[z.i,z.i,.x] * .y)))
-  Fs[1,z.i,i] <- Fs[1,z.i,i] + Fb[1,z.i,sdd.j[,3]] %*% (p$p_emig*p.ij)
-
-  #---- generate data: add immigrants
-  # total immigrants
-  D[i] <- sum(sdd[,,3,][sdd.j])
-  # seed bank: immigrants
-  B_1[i] <- B_1[i] + D[i] * (1 - p$rcr_dir) * p$s_SB
-  # recruits: immigrants
-  rcr.imm <- ceiling(D[i] * p$rcr_dir * p_est.i[i])
-  rcr.i <- (1:rcr.imm) + nrow(d[[i]])
-  d[[i]][rcr.i,] <- NA
-  d[[i]]$sizeNext[rcr.i] <- rnorm(rcr.imm, p$rcr_z[1], p$rcr_z[2])
-}
-# density dependence
+    Reduce(`+`, map2(sdd.j[[i]][,3], p$p_emig*p.ij, ~(Fb[z.i,z.i,.x] * .y)))
+  Fs[1,z.i,i] <- Fs[1,z.i,i] + Fb[1,z.i,sdd.j[[i]][,3]] %*% (p$p_emig*p.ij)
   if(p$NDD) {
-  lam.t <- p_est.t <- matrix(nrow=tmax-1, ncol=n.cell)
-  p_est.t[1,] <- p$p_est
-  for(i in 1:n.cell) {
-    Nt[,,i,1] <- rpois((n+1)*(n+1), n_i)
+    Nt[,,i,1] <- rpois((n+1)*(n+1), n_i[i])
+    F.i <- F.t <- Fs[,,i]/p$p_est  # F.s[,,i] was already multiplied by p$p_est
     for(k in 1:(tmax-1)) {
-      F.t <- Fs[,,i]
-      p_est.t[k,i] <- min(195/(sum(F.t[z.i,] * Nt[z.i,,i,k])/p$p_est), p$p_est)
-      F.t[z.i,] <- F.t[z.i,]/p$p_est * p_est.t[k,i]
+      p_est.t[i,k] <- min(p$NDD_n/(sum(F.i[z.i,]*Nt[z.i,,i,k])), p$p_est)
+      F.t[z.i,] <- F.i[z.i,] * p_est.t[i,k]
       IPM.t <- Ps[,,i] + F.t
       Nt[,,i,k+1] <- IPM.t %*% Nt[,,i,k]
-      lam.t[k,i] <- Re(eigen(IPM.t)$values[1])
+      lam.t[i,k] <- Re(eigen(IPM.t)$values[1])
     }
@@
 IPMs <- Ps + Fs
+
+#---- Realization: generate simulated data
+## local growth
+for(i in 1:n.cell) {
+  d[[i]] <- data.frame(yr=1, size=runif(n=n_i[i], min=z.rng[1], max=z.rng[2]))
+  E <- sim_expected(d[[i]], p, map(X, ~(.[i,])), n_z)
+  d[[i]] <- sim_realized(d[[i]], 1:nrow(d[[i]]), E, p, lo, hi)
+  nSd[i,1] <- sum(d[[i]]$seed, na.rm=T)
+  sdd[,,3,i] <- nSd[i,1] * p$p_emig * sdd[,,1,i]
+}
+## dispersal & density dependence
+for(i in 1:n.cell) {
+  D[i,1] <- sum(sdd[,,3,][sdd.j[[i]]])
+  p_est.i[i,1] <- ifelse(p$NDD, min(p$NDD_n/(nSd[i,1]+D[i,1]),p$p_est), p$p_est)
+  d[[i]] <- sim_recruits(d[[i]], p_est.i[i,1], nSd[i,1], B[i,1], D[i,1], p)
+  B[i,2] <- sim_seedbank(nSd[i,1], B[i,1], D[i,1], p)
+}
+## iterate forward
+for(k in 2:tmax) {
+  for(i in 1:n.cell) {
+    size.k <- d[[i]]$sizeNext[!is.na(d[[i]]$sizeNext) & d[[i]]$yr == k-1]
+    if(length(size.k) > 0) {
+      k.index <- (1:length(size.k)) + nrow(d[[i]])
+      d[[i]][k.index, "size"] <- size.k
+      d[[i]][k.index, "yr"] <- k
+      E <- sim_expected(d[[i]][k.index,], p, map(X, ~(.[i,])), n_z)
+      d[[i]] <- sim_realized(d[[i]], k.index, E, p, lo, hi)
+      nSd[i,k] <- sum(d[[i]]$seed[k.index], na.rm=T)
+    } else {
+      nSd[i,k] <- 0
+    }
+    sdd[,,3,i] <- nSd[i,k] * p$p_emig * sdd[,,1,i]
+  }
+  for(i in 1:n.cell) {
+    D[i,k] <- sum(sdd[,,3,][sdd.j[[i]]])
+    p_est.i[i,k] <- ifelse(p$NDD, min(p$NDD_n/(nSd[i,k]+D[i,k]),p$p_est), p$p_est)
+    d[[i]] <- sim_recruits(d[[i]], p_est.i[i,k], nSd[i,k], B[i,k], D[i,k], p)
+    B[i,k+1] <- sim_seedbank(nSd[i,k], B[i,k], D[i,k], p)
+  }
+}
+
 lam.df <- env.df
 lam.df$lambda <- apply(IPMs, 3, function(x) Re(eigen(x)$values[1]))
-lam.df %<>% add_column(nSd=nSd, B0=B, B1=B_1, D=D) %>%
-  mutate(nSdStay=nSd*(1-p$p_emig), nSdLeave=nSd*p$p_emig) %>%
+lam.df %<>% 
+  add_column(nSeed=nSd[,tmax], B0=B[,1], Btmax=B[,tmax+1], D=D[,tmax], 
+             N=map_dbl(d, ~sum(!is.na(.$sizeNext[.$yr==tmax]))),
+             N.surv=map_dbl(d, ~sum(.$surv[.$yr==tmax], na.rm=T)),
+             N.rcr=map_dbl(d, ~sum(is.na(.$size[.$yr==tmax])))) %>%
+  mutate(nSdStay=nSeed*(1-p$p_emig), nSdLeave=nSeed*p$p_emig) %>%
   add_column(Ntmax=apply(Nt[,,,tmax],3,sum), lam.tmax=lam.t[tmax-1,])
@@
   scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
+ggplot(lam.df, aes(x=lon, y=lat, fill=log(Ntmax)*(Ntmax>10))) + geom_tile() +
+  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=1)
 ggplot(lam.df, aes(x=lon, y=lat, fill=lambda>1)) + geom_tile()
 ggplot(lam.df, aes(x=lon, y=lat, fill=lam.tmax>1)) + geom_tile()
-ggplot(lam.df, aes(x=lon, y=lat, fill=nSd)) + geom_tile()
+ggplot(lam.df, aes(x=lon, y=lat, fill=Ntmax>1)) + geom_tile() 
+ggplot(lam.df, aes(x=lon, y=lat, fill=Ntmax>1 & lam.tmax<1)) + geom_tile() 
+ggplot(lam.df, aes(x=lon, y=lat, fill=log(N))) + geom_tile() +
+  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
+ggplot(lam.df, aes(x=lon, y=lat, fill=log(N.surv))) + geom_tile() +
+  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
+ggplot(lam.df, aes(x=lon, y=lat, fill=log(N.rcr))) + geom_tile() +
+  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
+ggplot(lam.df, aes(x=lon, y=lat, fill=nSeed)) + geom_tile()
 ggplot(lam.df, aes(x=lon, y=lat, fill=D)) + geom_tile()
-ggplot(lam.df, aes(x=lon, y=lat, fill=B1>B0)) + geom_tile()
-ggplot(lam.df, aes(x=lon, y=lat, fill=(D-nSdStay)/nSdStay)) + geom_tile() +
-  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
+ggplot(lam.df, aes(x=lon, y=lat, fill=Btmax)) + geom_tile()
+ggplot(lam.df, aes(x=lon, y=lat, fill=D + nSdStay)) + geom_tile()
+ggplot(lam.df, aes(x=lon, y=lat, fill=Btmax>B0)) + geom_tile()
+ggplot(lam.df, aes(x=lon, y=lat, fill=nSdStay/D)) + geom_tile() +
+  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=1)
 ggplot(lam.df, aes(x=lon, y=lat, fill=nSdStay > D)) + geom_tile()
@@
 plot(log(lambda) ~ prec, data=lam.df); abline(h=0, col="red", lty=2)
-plot(log(lambda) ~ pOth, data=lam.df); abline(h=0, col="red", lty=2)
+plot(log(lambda) ~ pOpn, data=lam.df); abline(h=0, col="red", lty=2)
 plot(density(log(lam.df$lambda))); abline(v=0, col="red", lty=2)
@@
 }
+image.plot(x, x, t(apply(Ps, 1:2, mean)), main="P matrix: mean",
+           xlab="Size (t)", ylab="Size (t+1)")
+image.plot(x, x, t(apply(Fs, 1:2, mean)), main="F matrix: mean",
+           xlab="Size (t)", ylab="Size (t+1)")
+image.plot(x, x, sens(t(apply(IPMs, 1:2, mean))), main="Sensitivity: mean",
+           xlab="Size (t)", ylab="Size (t+1)")
+image.plot(x, x, elas(t(apply(IPMs, 1:2, mean))), main="Elasticity: mean",
+           xlab="Size (t)", ylab="Size (t+1)")
 ```

### 2018-03-10 code/fn_IPM.R

@@ -65,7 +65,7 @@ calc_flwr <- function(z.v, p, n_flz, X.fl) {
 
 ##-- Seed production
 ##   nSeeds ~ exp(size + environment)
-calc_seeds <- function(z.v, p, n_seeds, X.seed) {
+calc_seeds <- function(z.v, p, n_seedz, X.seed) {
   z <- z_pow(z.v, n_seedz)
   exp(z %*% p$seed_z + c(t(X.seed) %*% p$seed_x))
 }
@@ -125,9 +125,34 @@ setup_IPM_matrix <- function(n=100, z.rng=c(1,10), buffer=0.25, discrete=0) {
 }
 
 
-##-- Fill P & F matrices: local 
+##-- Fill P matrix: local 
 ##   
+fill_P <- function(h, y, z.i, n, p, n_z, n_x, X_s, X_g) {
+  P.mx <- matrix(0, nrow=n+1, ncol=n+1)
+  # survival & growth
+  S.v <- calc_surv(y, p=p, n_sz=n_z$s, X.s=X_s)
+  G.mx <- h*outer(y, y, calc_grow, p=p, n_gz=n_z$g, X.g=X_g)
+  # correct ejections
+  for(k in 1:(n/2)) G.mx[1,k] <- G.mx[1,k] + 1 - sum(G.mx[,k])
+  for(k in (n/2+1):n) G.mx[n,k] <- G.mx[n,k] + 1 - sum(G.mx[,k])
+  # fill P matrix
+  for(k in z.i) P.mx[k,z.i] <- G.mx[k-1,]*S.v
+  P.mx[1,1] <- calc_staySB(p)
+  return(P.mx)
+}
 
+
+##-- Fill F matrix 
+##   
+fill_F <- function(h, y, z.i, n, p, n_z, n_x, X_fl, X_seed) {
+  F.mx <- matrix(0, nrow=n+1, ncol=n+1)
+  F.mx[z.i,z.i] <- outer(y, y, calc_rcrDir, p=p, n_seedz=n_z$seed, n_flz=n_z$fl, 
+                         X.seed=X_seed, X.fl=X_fl) * h * p$p_est
+  F.mx[1,z.i] <- calc_addSB(y, p=p, n_seedz=n_z$seed, n_flz=n_z$fl, 
+                            X.seed=X_seed, X.fl=X_fl) * h
+  F.mx[z.i,1] <- calc_rcrSB(y, p) * p$p_est
+  return(F.mx)
+}
 
 
 

### 2018-03-10 code/fn_sim.R

@@ -1,0 +1,58 @@
+
+
+
+
+##--- simulate expected values
+## 
+sim_expected <- function(d_i, p, X, n_z) {
+  E <- list(s=NULL, g=NULL, fl=NULL, seed=NULL)
+  size.mx <- cbind(1, d_i$size, d_i$size^2, d_i$size^3)
+  E$s <- antilogit(size.mx[,1:n_z$s] %*% p$s_z + c(X$s %*% p$s_x))
+  E$g <- size.mx[,1:n_z$g] %*% p$g_z + c(X$g %*% p$g_x)
+  E$fl <- antilogit(size.mx[,1:n_z$fl] %*% p$fl_z + c(X$fl %*% p$fl_x))
+  E$seed <- exp(size.mx[,1:n_z$seed] %*% p$seed_z + c(X$seed %*% p$seed_x))
+  return(E)
+}
+
+
+##--- simulate realized values
+##
+sim_realized <- function(d_i, k.index, E, p, lo, hi) {
+  n_t <- length(k.index)
+  d_i$surv[k.index] <- rbinom(n_t, 1, E$s)
+  d_i$sizeNext[k.index] <- rnorm(n_t, E$g, p$g_sig) * 
+    ifelse(d_i$surv[k.index], 1, NA)
+  d_i$sizeNext[k.index] <- ifelse(d_i$sizeNext[k.index] < lo, 
+                                  lo, d_i$sizeNext[k.index])
+  d_i$sizeNext[k.index] <- ifelse(d_i$sizeNext[k.index] > hi, 
+                                  hi, d_i$sizeNext[k.index])
+  d_i$fl[k.index] <- rbinom(n_t, 1, E$fl) * ifelse(d_i$surv[k.index], 1, NA)
+  d_i$seed[k.index] <- rpois(n_t, E$seed) * 
+    ifelse(d_i$surv[k.index] & d_i$fl[k.index], 1, NA)
+  return(d_i)
+}
+
+
+##--- simulate number of recruits
+##
+sim_recruits <- function(d_i, p_est_ik, nSd_ik, B_ik, D_ik, p) {
+  n <- round(p_est_ik * (nSd_ik * (1-p$p_emig) * p$rcr_dir + 
+                           B_ik * p$rcr_SB + 
+                           D_ik * p$rcr_dir)) 
+  if(n > 0) {
+    n_index <- (1:n)+nrow(d_i)
[…]

### 2018-03-10 exploration.Rmd
added extra plots, fixed plot objects that had changed

@@
 ```{r sp_i}
-mean_abund <- 100
-regPlots <- F
-ipmPlots <- F
+mean_abund <- 500
+regPlots <- T
+ipmPlots <- T
 nPlots <- 5
 z.rng <- c(1, 12)
-n <- 10  # ncells in IPM matrix
+n <- 50  # ncells in IPM matrix
 tmax <- 30
@@
 n_i <- rpois(n.cell, mean_abund)
-sp_i <- list(p=list(s_z=c(-1.75, .75), # b1 + b2*z
+sp_i <- list(p=list(s_z=c(-8, 2.1, -.09), # b1 + b2*z + b3*z^2
                     s_x=c(2, -.1, -2, -.1, 2, -2, -.4), # b1*x1 + ...
@@
                     fl_z=c(-1.5, .1, .1), # b1 + b2*z + b3*z^2
-                    fl_x=c(2, -.1, -2, -.1, 1, 1), # b1*x1 + ...
+                    fl_x=c(-2, -.1, -2, -.1, 1, 1), # b1*x1 + ...
                     seed_z=c(2, 0.5, -.03), # b1 + b2*z + b2*z^2
@@
 B <- cbind(rpois(n.cell, n_i), matrix(0, nrow=n.cell, ncol=tmax))
-nSd <- D <- p_est.i <- matrix(nrow=n.cell, ncol=tmax)
+nSd <- D <- p_est.i <- N_sim <- matrix(0, nrow=n.cell, ncol=tmax)
 lam.t <- p_est.t <- matrix(nrow=n.cell, ncol=tmax-1)
@@
 }
+N_sim[,1] <- n_i
 ## dispersal & density dependence
@@
       nSd[i,k] <- sum(d[[i]]$seed[k.index], na.rm=T)
-    } else {
-      nSd[i,k] <- 0
+      N_sim[i,k] <- length(size.k)
     }
@@
   mutate(nSdStay=nSeed*(1-p$p_emig), nSdLeave=nSeed*p$p_emig) %>%
-  add_column(Ntmax=apply(Nt[,,,tmax],3,sum), lam.tmax=lam.t[tmax-1,])
+  add_column(Ntmax=apply(Nt[,,,tmax],3,sum), lam.tmax=lam.t[,tmax-1])
 
@@
 ggplot(lam.df, aes(x=lon, y=lat, fill=lambda>1)) + geom_tile()
+ggplot(lam.df, aes(x=lon, y=lat, fill=N>1)) + geom_tile()
 ggplot(lam.df, aes(x=lon, y=lat, fill=lam.tmax>1)) + geom_tile()
@@
 ggplot(lam.df, aes(x=lon, y=lat, fill=D)) + geom_tile()
+ggplot(lam.df, aes(x=lon, y=lat, fill=D + nSdStay)) + geom_tile()
 ggplot(lam.df, aes(x=lon, y=lat, fill=Btmax)) + geom_tile()
-ggplot(lam.df, aes(x=lon, y=lat, fill=D + nSdStay)) + geom_tile()
 ggplot(lam.df, aes(x=lon, y=lat, fill=Btmax>B0)) + geom_tile()
@@
 plot(density(log(lam.df$lambda))); abline(v=0, col="red", lty=2)
+matplot(t(log(lam.t)), type="l", lty=1, col=rgb(0,0,0,0.1))
+matplot(t(log(apply(Nt, 3:4, sum))), type="l", lty=1, col=rgb(0,0,0,0.1))
+matplot(t(N_sim[,2:tmax]/N_sim[,1:(tmax-1)]), type="l", lty=1, col=rgb(0,0,0,.1))
+matplot(t(log(N_sim+1)), type="l", lty=1, col=rgb(0,0,0,0.1))
+#matplot(t(log(nSd*(1-p$p_emig)+D)), type="l", lty=1, col=rgb(0,0,0,0.1))
 
-sample.cells <- sample.int(n.cell, nPlots)
+sample.cells <- sample((1:n.cell)[lam.df$N.surv>0], nPlots)
 if(regPlots) {
@@
   for(i in sample.cells) {
-    z.t <- d[[i]][!is.na(d[[i]]$size),]; main.i <- paste("Cell", i)
+    z.t <- d[[i]][!is.na(d[[i]]$size) & d[[i]]$yr==tmax,]
+    main.i <- paste("Cell", i)
     w1 <- Re(eigen(IPMs[,,i])$vectors[,1])[-1]; stable.dist1 <- w1/sum(w1)
@@
          main=paste("Surv:", main.i), xlab="Size (t)", ylab="P(surv to t+1)")
-      lines(mx[,2], antilogit(mx[,1:n_sz]%*%p$s_z + c(X_s[i,]%*%p$s_x)), col=2)
+      lines(mx[,2], antilogit(mx[,1:n_z$s]%*%p$s_z + c(X$s[i,]%*%p$s_x)), col=2)
     plot(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,0.2), 
          main=paste("Growth:", main.i), xlab="Size (t)", ylab="Size (t+1)")
-      lines(mx[,2], mx[,1:n_gz]%*%p$g_z + c(X_g[i,]%*%p$g_x), col=2)
+      lines(mx[,2], mx[,1:n_z$g]%*%p$g_z + c(X$g[i,]%*%p$g_x), col=2)
       abline(a=0, b=1, lty=2)
@@
          main=paste("P(Flower):", main.i), xlab="Size (t)", ylab="P(flower)")
-      lines(mx[,2], antilogit(mx[,1:n_flz]%*%p$fl_z + c(X_fl[i,]%*%p$fl_x)),
+      lines(mx[,2], antilogit(mx[,1:n_z$fl]%*%p$fl_z + c(X$fl[i,]%*%p$fl_x)),
             col=2)
@@
          main=paste("nSeeds:", main.i), xlab="Size (t)", ylab="Number of seeds")
-      lines(mx[,2], exp(mx[,1:n_seedz]%*%p$seed_z + c(X_seed[i,]%*%p$seed_x)),
+      lines(mx[,2], exp(mx[,1:n_z$seed]%*%p$seed_z + c(X$seed[i,]%*%p$seed_x)),
             col=2)
@@
   for(i in sample.cells) {
-    z.t <- d[[i]][!is.na(d[[i]]$size),]; main.i <- paste("Cell", i)
+   z.t <- d[[i]][!is.na(d[[i]]$size) & d[[i]]$yr==tmax,] 
+   main.i <- paste("Cell", i)
    z.t$y <- y[findInterval(z.t$size, y, all.inside=T)]

### 2018-03-11 exploration.Rmd
beginning code for samples from S

--- 2018-03-10 exploration.Rmd
@@
 ## Observed datasets
-The simulated species *S* is sampled appropriately in order to fit each SDM model. The observed datasets,*O*, may be random or intentionally biased to mimic common issues with empirical data.
+The simulated species *S* is sampled appropriately in order to fit each SDM model.
+The observed datasets,*O*, may be random or intentionally biased to mimic common issues with empirical data.
+The realized simulated data generated above (`d` and its abundance summary `N_sim`) are the underlying true distribution.
+The samples will be drawn from occupied cells in the final time steps where the species is approximately at equilibrium.
+```{r O_setup}
+O_n <- c(Mx=500, CA=20, IPM=20) # number of cells in sample
+```
+
 ### Correlative: MaxEnt
@@
 library(dismo)
-Mx.n <- 3000
-Mx.P <- sample((1:nrow(lam.df))[lam.df$lambda > 1], Mx.n, replace=T)
+Mx.P <- sample((1:nrow(lam.df))[lam.df$lambda > 1], O_n$Mx, replace=T)
 pres <- 1:nrow(lam.df) %in% Mx.P

### 2018-03-12 exploration.Rmd
added code for fitting IPM based on sampled cells

@@
 ```{r sp_i}
-mean_abund <- 500
+mean_abund <- 50
 regPlots <- T
@@
 n <- 50  # ncells in IPM matrix
-tmax <- 30
+tmax <- 10
 cells <- 1:n_distinct(lc.df$CellID)
@@
 Nt <- array(dim=c(n+1, n+1, n.cell, tmax))
-d <- sdd.j <- vector("list", length=n.cell)
+d <- E <- sdd.j <- vector("list", length=n.cell)
 B <- cbind(rpois(n.cell, n_i), matrix(0, nrow=n.cell, ncol=tmax))
@@
   d[[i]] <- data.frame(yr=1, size=runif(n=n_i[i], min=z.rng[1], max=z.rng[2]))
-  E <- sim_expected(d[[i]], p, map(X, ~(.[i,])), n_z)
-  d[[i]] <- sim_realized(d[[i]], 1:nrow(d[[i]]), E, p, lo, hi)
+  E[[i]] <- sim_expected(d[[i]], p, map(X, ~(.[i,])), n_z)
+  d[[i]] <- sim_realized(d[[i]], 1:nrow(d[[i]]), E[[i]], p, lo, hi)
   nSd[i,1] <- sum(d[[i]]$seed, na.rm=T)
@@
       lines(mx[,2], antilogit(mx[,1:n_z$s]%*%p$s_z + c(X$s[i,]%*%p$s_x)), col=2)
-    plot(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,0.2), 
+    plot(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,0.2), ylim=c(lo,hi),
          main=paste("Growth:", main.i), xlab="Size (t)", ylab="Size (t+1)")
@@
     plot(seed~size, data=z.t, pch=23-z.t$fl*3, col=rgb(0,0,0,0.2), 
+         ylim=range(c(exp(mx[,1:n_z$seed]%*%p$seed_z + c(X$seed[i,]%*%p$seed_x)),
+                      z.t$seed), na.rm=TRUE),
          main=paste("nSeeds:", main.i), xlab="Size (t)", ylab="Number of seeds")
@@
 ```{r O_setup}
-O_n <- c(Mx=500, CA=20, IPM=20) # number of cells in sample
+O_n <- list(Mx=500, CA=20, IPM=20) # number of cells in sample
+O_yr <- list(Mx=tmax, CA=(tmax-10):tmax, IPM=tmax) # years to sample from
+P.i <- which(lam.df$N.surv > 0) # presences: surviving individuals > seedlings
 ```
@@
 library(dismo)
-Mx.P <- sample((1:nrow(lam.df))[lam.df$lambda > 1], O_n$Mx, replace=T)
+Mx.P <- sample(P.i, O_n$Mx, replace=T)
 pres <- 1:nrow(lam.df) %in% Mx.P
@@
 ```{r O_IPM}
-
+IPM.P <- sample(P.i, O_n$IPM, replace=FALSE)
+IPM.d <- vector("list", O_n$IPM)
+for(i in IPM.P) {
+  IPM.d[[i]] <- d[[i]] %>% filter(yr %in% O_yr$IPM) %>%
+    add_column(id.inbd=i) %>%
+    full_join(env.df[i,], by="id.inbd")
+}
+IPM.d <- do.call(rbind, IPM.d)
+p.ipm <- p
 ```
@@
 ```{r P_IPM}
-for(i in 1:ncell) {
-  # ## IPMpack: estimate slopes from d[[i]]
-  # ### set up seedbank transitions
-  d[[i]]$stage <- d[[i]]$stageNext <- "continuous"
-  d[[i]]$stage[is.na(d[[i]]$size)] <- NA
-  d[[i]]$stageNext[d[[i]]$surv==0] <- "dead"
-  d[[i]]$number <- 1
-  seedbank <- data.frame(size=NA, surv=1, sizeNext=NA,
-                         fl=NA, seed=NA,
-                         stage=c("seedbank", "seedbank", "continuous"),
-                         stageNext=c("seedbank", "continuous", "seedbank"),
-                         number=c(p$s_SB * (1 - p$rcr_SB), 
-                                  p$rcr_SB * p$p_est, 
-                                  (1 - p$rcr_dir) * p$s_SB))
-  d[[i]] <- rbind(d[[i]], seedbank)
-  d[[i]]$stage <- as.factor(d[[i]]$stage)
-  d[[i]]$stageNext <- as.factor(d[[i]]$stageNext)
-  ### construct kernels
-  so <- makeSurvObj(d[[i]], surv ~ size)
-  go <- makeGrowthObj(d[[i]], sizeNext ~ size + size2)
-  fo <- makeFecObj(d[[i]],
-                   Formula=c(fl~size+size2, seed~size+size2),
-                   Family=c("binomial", "poisson"),
-                   Transform=c("none", "none"),
-                   meanOffspringSize=mean(
-                     d[[i]]$sizeNext[is.na(d[[i]]$size) &
-                                       !is.na(d[[i]]$sizeNext)]),
-                   sdOffspringSize=sd(
-                     d[[i]]$sizeNext[is.na(d[[i]]$size) &
-                                       !is.na(d[[i]]$sizeNext)]),
-                   fecConstants=data.frame(seedlingEstablishmentRate=p$p_est),
-                   offspringSplitter=data.frame(seedbank=(1-p$rcr_dir),
-                                                continuous=p$rcr_dir)
-                   )
-  F.mx2 <- makeIPMFmatrix(fecObj=fo, minSize=lo, maxSize=hi,
-                          correction="constant")
-  D.mx2 <- makeDiscreteTrans(d[[i]], 
-                             continuousToDiscreteExplanatoryVariables=1)
-  dummy <- as.matrix(fo@offspringRel$coefficients[1], list(1, "seedbank"))
-  D.mx2@meanToCont <- as.matrix(dummy, dimnames=c(1, "seedbank"))
-  dummy <- as.matrix(fo@sdOffspringSize, dimnames=list(1, "seedbank"))
-  D.mx2@sdToCont <- as.matrix(dummy, dimnames=c(1, "seedbank"))
-  D.mx2@discreteTrans[1,1] <- p$s_SB * (1 - p$rcr_SB)
-  D.mx2@discreteTrans[3,1] <- 1 - sum(D.mx2@discreteTrans[1:2,1])
-  P.mx2 <- makeIPMPmatrix(growObj=go, survObj=so, discreteTrans=D.mx2,
-                            minSize=lo, maxSize=hi, correction="constant")
-  IPM2 <- P.mx2 + F.mx2
-  lambdas2[i] <- Re(eigen(IPM2)$values[1])
+# run chunk O_IPM first to generate observed dataset `IPM.d`
 
+########
+## fit models
+########
+s.m <- glm(surv ~ size + I(size^2) + temp + temp2 + prec + prec2 + pOpn + 
+               pOth + pDec, data=IPM.d, family="binomial")
+g.m <- lm(sizeNext ~ size + I(size^2) + temp + temp2 + prec + prec2 +
+              pOpn + pOth, data=IPM.d)
+fl.m <- glm(fl ~ size + I(size^2) + temp + temp2 + prec + prec2 + pOpn +
+                pOth, data=IPM.d, family="binomial")
+seed.m <- glm(seed ~ size + I(size^2) + temp + temp2 + prec + prec2 + pOpn, 
+              data=IPM.d, family="poisson")
+p.ipm$s_z <- coef(s.m)[1:length(p.ipm$s_z)]
+p.ipm$s_x <- coef(s.m)[(1:length(p.ipm$s_x)) + length(p.ipm$s_z)]
+p.ipm$g_z <- coef(g.m)[1:length(p.ipm$g_z)]
+p.ipm$g_x <- coef(g.m)[(1:length(p.ipm$g_x)) + length(p.ipm$g_z)]
+p.ipm$g_sig <- summary(g.m)$sigma
+p.ipm$fl_z <- coef(fl.m)[1:length(p.ipm$fl_z)]
+p.ipm$fl_x <- coef(fl.m)[(1:length(p.ipm$fl_x)) + length(p.ipm$fl_z)]
+p.ipm$seed_z <- coef(seed.m)[1:length(p.ipm$seed_z)]
+p.ipm$seed_x <- coef(seed.m)[(1:length(p.ipm$seed_x)) + length(p.ipm$seed_z)]
+p.ipm$rcr_z <- c(mean(IPM.d[is.na(IPM.d$size), "sizeNext"]),
+                 sd(IPM.d[is.na(IPM.d$size), "sizeNext"]))
+
+setup_IPM_matrix(n, z.rng, buffer=0.5, discrete=1)
+IPMs.f <- Ps.f <- Fs.f <- Fb.f <- array(0, dim=c(n+1, n+1, n.cell))
+Nt.f <- array(dim=c(n+1, n+1, n.cell, tmax))
+d.f <- E.f <- sdd.j <- vector("list", length=n.cell)
+B.f <- cbind(rpois(n.cell, n_i), matrix(0, nrow=n.cell, ncol=tmax))
+nSd.f <- D.f <- p_est.i.f <- N_sim.f <- matrix(0, nrow=n.cell, ncol=tmax)
+lam.t.f <- p_est.t.f <- matrix(nrow=n.cell, ncol=tmax-1)
+sdd.f <- array(0, dim=dim(sdd.pr) + c(0,0,1,0)) # add layer for nEmig
+sdd.f[,,1:2,] <- sdd.pr
+
+# set up environmental covariates
+n_z.f <- list(s=length(p.ipm$s_z), g=length(p.ipm$g_z),
+            fl=length(p.ipm$fl_z), seed=length(p.ipm$seed_z))
+n_x.f <- list(s=length(p.ipm$s_x), g=length(p.ipm$g_x),
+            fl=length(p.ipm$fl_x), seed=length(p.ipm$seed_x))
+X.f <- list(s=as.matrix(env.df[,1:n_x.f$s]), g=as.matrix(env.df[,1:n_x.f$g]),
+          fl=as.matrix(env.df[,1:n_x.f$fl]), seed=as.matrix(env.df[,1:n_x.f$seed]))
+
+########
+## loop through cells
+########
+
+#---- Truth: use assigned slopes to fill IPM matrix
+## local growth
+for(i in 1:n.cell) {
+  Ps.f[,,i] <- fill_P(h, y, z.i, n, p.ipm, n_z.f, n_x.f, X.f$s[i,], X.f$g[i,])
+  Fb.f[,,i] <- fill_F(h, y, z.i, n, p.ipm, n_z.f, n_x.f, X.f$fl[i,], X.f$seed[i,])
+  Fs.f[z.i,z.i,i] <- (1-p.ipm$p_emig) * Fb.f[z.i,z.i,i]
+  Fs.f[1,z.i,i] <- (1-p.ipm$p_emig) * Fb.f[1,z.i,i]
+  Fs.f[z.i,1,i] <- Fb.f[z.i,1,i]
+}
+## dispersal & density dependence
+for(i in 1:n.cell) {
+  sdd.i <- env.df$id[i] # full rectangle grid id for SDD indexes
+  sdd.j[[i]] <- which(sdd.f[,,2,]==sdd.i, arr.ind=TRUE) # j cells with target i
+  p.ij <- sdd.f[,,1,][sdd.j[[i]]] # p(j to i | immigrant seed from j)
+  Fs.f[z.i,z.i,i] <- Fs.f[z.i,z.i,i] + 
+    Reduce(`+`, map2(sdd.j[[i]][,3], p.ipm$p_emig*p.ij, ~(Fb.f[z.i,z.i,.x] * .y)))
+  Fs.f[1,z.i,i] <- Fs.f[1,z.i,i] + Fb.f[1,z.i,sdd.j[[i]][,3]] %*% (p.ipm$p_emig*p.ij)
+  if(p.ipm$NDD) {
+    Nt.f[,,i,1] <- rpois((n+1)*(n+1), n_i[i])
+    F.i.f <- F.t.f <- Fs.f[,,i]/p.ipm$p_est  
+    for(k in 1:(tmax-1)) {
+      p_est.t.f[i,k] <- min(p.ipm$NDD_n/(sum(F.i.f[z.i,]*Nt.f[z.i,,i,k])), p.ipm$p_est)
+      F.t.f[z.i,] <- F.i.f[z.i,] * p_est.t[i,k]
+      IPM.t.f <- Ps.f[,,i] + F.t.f
+      Nt.f[,,i,k+1] <- IPM.t.f %*% Nt.f[,,i,k]
+      lam.t.f[i,k] <- Re(eigen(IPM.t.f)$values[1])
+    }
+  }
+}
+IPMs.f <- Ps.f + Fs.f
+
+#---- Realization: generate simulated data
+## local growth
+for(i in 1:n.cell) {
+  d.f[[i]] <- data.frame(yr=1, size=runif(n=n_i[i], min=z.rng[1], max=z.rng[2]))
+  E.f[[i]] <- sim_expected(d.f[[i]], p.ipm, map(X.f, ~(.[i,])), n_z)
+  d.f[[i]] <- sim_realized(d.f[[i]], 1:nrow(d.f[[i]]), E.f[[i]], p.ipm, lo, hi)
+  nSd.f[i,1] <- sum(d.f[[i]]$seed, na.rm=T)
+  sdd.f[,,3,i] <- nSd.f[i,1] * p.ipm$p_emig * sdd.f[,,1,i]
+}
+N_sim.f[,1] <- n_i
+## dispersal & density dependence
+for(i in 1:n.cell) {
+  D.f[i,1] <- sum(sdd.f[,,3,][sdd.j[[i]]])
+  p_est.i.f[i,1] <- ifelse(p.ipm$NDD, min(p.ipm$NDD_n/(nSd.f[i,1]+D.f[i,1]), 
+                                          p.ipm$p_est), p.ipm$p_est)
+  d.f[[i]] <- sim_recruits(d.f[[i]], p_est.i[i,1], nSd.f[i,1], B.f[i,1], D.f[i,1], p.ipm)
+  B.f[i,2] <- sim_seedbank(nSd.f[i,1], B.f[i,1], D.f[i,1], p.ipm)
+}
+## iterate forward
+for(k in 2:tmax) {
+  for(i in 1:n.cell) {
+    size.k <- d.f[[i]]$sizeNext[!is.na(d.f[[i]]$sizeNext) & d.f[[i]]$yr == k-1]
+    if(length(size.k) > 0) {
+      k.index <- (1:length(size.k)) + nrow(d.f[[i]])
+      d.f[[i]][k.index, "size"] <- size.k
+      d.f[[i]][k.index, "yr"] <- k
+      E.f <- sim_expected(d.f[[i]][k.index,], p, map(X, ~(.[i,])), n_z)
+      d.f[[i]] <- sim_realized(d.f[[i]], k.index, E.f, p.ipm, lo, hi)
+      nSd.f[i,k] <- sum(d.f[[i]]$seed[k.index], na.rm=T)
+      N_sim.f[i,k] <- length(size.k)
+    }
+    sdd.f[,,3,i] <- nSd.f[i,k] * p.ipm$p_emig * sdd.f[,,1,i]
+  }
+  for(i in 1:n.cell) {
+    D.f[i,k] <- sum(sdd.f[,,3,][sdd.j[[i]]])
+    p_est.i.f[i,k] <- ifelse(p.ipm$NDD, min(p.ipm$NDD_n/(nSd.f[i,k]+D.f[i,k]),
+                                            p.ipm$p_est), p.ipm$p_est)
+    d.f[[i]] <- sim_recruits(d.f[[i]], p_est.i.f[i,k], nSd.f[i,k], B.f[i,k], D.f[i,k], p.ipm)
+    B.f[i,k+1] <- sim_seedbank(nSd.f[i,k], B.f[i,k], D.f[i,k], p.ipm)
+  }
+}
+
+lam.df <- env.df
+lam.df$lambda.f <- apply(IPMs.f, 3, function(x) Re(eigen(x)$values[1]))
+lam.df %<>% 
+  add_column(nSeed.f=nSd.f[,tmax], B0.f=B.f[,1], Btmax.f=B.f[,tmax+1], D.f=D.f[,tmax], 
+             N.f=map_dbl(d.f, ~sum(!is.na(.$sizeNext[.$yr==tmax]))),
+             N.surv.f=map_dbl(d.f, ~sum(.$surv[.$yr==tmax], na.rm=T)),
+             N.rcr.f=map_dbl(d.f, ~sum(is.na(.$size[.$yr==tmax])))) %>%
+  mutate(nSdStay.f=nSeed.f*(1-p.ipm$p_emig), nSdLeave.f=nSeed.f*p.ipm$p_emig) %>%
+  add_column(Ntmax.f=apply(Nt.f[,,,tmax],3,sum), lam.tmax.f=lam.t.f[,tmax-1])
+
+########
+## plots
+########
+ggplot(lam.df, aes(x=lon, y=lat, fill=log(lambda.f))) + geom_tile() +
+  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
+ggplot(lam.df, aes(x=lon, y=lat, fill=log(lam.tmax.f))) + geom_tile() +
+  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
+ggplot(lam.df, aes(x=lon, y=lat, fill=log(Ntmax.f)*(Ntmax.f>10))) + geom_tile() +
+  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=1)
+ggplot(lam.df, aes(x=lon, y=lat, fill=lambda.f>1)) + geom_tile()
+ggplot(lam.df, aes(x=lon, y=lat, fill=N.f>1)) + geom_tile()
+ggplot(lam.df, aes(x=lon, y=lat, fill=lam.tmax.f>1)) + geom_tile()
+ggplot(lam.df, aes(x=lon, y=lat, fill=Ntmax.f>1)) + geom_tile() 
+ggplot(lam.df, aes(x=lon, y=lat, fill=Ntmax.f>1 & lam.tmax.f<1)) + geom_tile() 
+ggplot(lam.df, aes(x=lon, y=lat, fill=log(N.f))) + geom_tile() +
+  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
+ggplot(lam.df, aes(x=lon, y=lat, fill=log(N.surv.f))) + geom_tile() +
+  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
+ggplot(lam.df, aes(x=lon, y=lat, fill=log(N.rcr.f))) + geom_tile() +
+  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
+ggplot(lam.df, aes(x=lon, y=lat, fill=nSeed.f)) + geom_tile()
+ggplot(lam.df, aes(x=lon, y=lat, fill=D.f)) + geom_tile()
+ggplot(lam.df, aes(x=lon, y=lat, fill=D.f + nSdStay.f)) + geom_tile()
+ggplot(lam.df, aes(x=lon, y=lat, fill=Btmax.f)) + geom_tile()
+ggplot(lam.df, aes(x=lon, y=lat, fill=Btmax.f>B0.f)) + geom_tile()
+ggplot(lam.df, aes(x=lon, y=lat, fill=nSdStay.f/D.f)) + geom_tile() +
+  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=1)
+ggplot(lam.df, aes(x=lon, y=lat, fill=nSdStay.f > D.f)) + geom_tile()
+ggplot(lam.df, aes(x=lon, y=lat, fill=nSdStay.f < D.f & lambda.f < 1)) +
+  geom_tile()
+par(mfrow=c(2,2))
+plot(log(lambda.f) ~ temp, data=lam.df); abline(h=0, col="red", lty=2)
+plot(log(lambda.f) ~ prec, data=lam.df); abline(h=0, col="red", lty=2)
+plot(log(lambda.f) ~ pOpn, data=lam.df); abline(h=0, col="red", lty=2)
+plot(density(log(lam.df$lambda.f))); abline(v=0, col="red", lty=2)
+matplot(t(log(lam.t.f)), type="l", lty=1, col=rgb(0,0,0,0.1))
+matplot(t(log(apply(Nt.f, 3:4, sum))), type="l", lty=1, col=rgb(0,0,0,0.1))
+matplot(t(N_sim.f[,2:tmax]/N_sim.f[,1:(tmax-1)]), type="l", lty=1, col=rgb(0,0,0,.1))
+matplot(t(log(N_sim.f+1)), type="l", lty=1, col=rgb(0,0,0,0.1))
+#matplot(t(log(nSd*(1-p.ipm$p_emig)+D)), type="l", lty=1, col=rgb(0,0,0,0.1))
+
+sample.cells <- sample((1:n.cell)[lam.df$N.surv>0], nPlots)
 if(regPlots) {
@@
   xx <- tibble(size=seq(lo, hi, length.out=100), size2=size^2, size3=size^3)
-    plot(surv~size, data=d[[i]], ylim=c(0,1), pch=20, col=rgb(0,0,0,0.2),
+  mx <- cbind(1, as.matrix(xx))
+  xxx <- seq(0, hi*.5, length.out=100)
+  for(i in sample.cells) {
+    z.t <- d[[i]][!is.na(d[[i]]$size) & d[[i]]$yr==tmax,]
+    main.i <- paste("Cell", i)
+    w1 <- Re(eigen(IPMs[,,i])$vectors[,1])[-1]; stable.dist1 <- w1/sum(w1)
+    rcrt.z <- d[[i]]$sizeNext[is.na(d[[i]]$size) & !is.na(d[[i]]$sizeNext)]
+    plot(surv~size, data=z.t, ylim=c(0,1), pch=20, col=rgb(0,0,0,0.2),
          main=paste("Surv:", main.i), xlab="Size (t)", ylab="P(surv to t+1)")
-      points(z.t$size, E_surv, cex=0.2, col="red")
-      # lines(xx$size, predict(so@fit, xx, type="response"), col="blue")
-    plot(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,0.2), 
+      lines(mx[,2], antilogit(mx[,1:n_z$s]%*%p.ipm$s_z + c(X$s[i,]%*%p.ipm$s_x)), col=2)
+    plot(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,0.2), ylim=c(lo,hi),
          main=paste("Growth:", main.i), xlab="Size (t)", ylab="Size (t+1)")
-      points(z.t$size, E_grow, cex=0.2, col="red")
-      # lines(xx$size, predict(go@fit, xx, type="response"), col="blue")
+      lines(mx[,2], mx[,1:n_z$g]%*%p.ipm$g_z + c(X$g[i,]%*%p.ipm$g_x), col=2)
       abline(a=0, b=1, lty=2)
@@
          main=paste("P(Flower):", main.i), xlab="Size (t)", ylab="P(flower)")
-      points(z.t$size, E_fl, cex=0.2, col="red")
-      # lines(xx$size, predict(fo@fitFec[[1]], xx, type="response"), col="blue")
+      lines(mx[,2], antilogit(mx[,1:n_z$fl]%*%p.ipm$fl_z + c(X$fl[i,]%*%p.ipm$fl_x)),
+            col=2)
     plot(seed~size, data=z.t, pch=23-z.t$fl*3, col=rgb(0,0,0,0.2), 
+         ylim=range(c(exp(mx[,1:n_z$seed]%*%p.ipm$seed_z + c(X$seed[i,]%*%p.ipm$seed_x)),
+                      z.t$seed), na.rm=TRUE),
          main=paste("nSeeds:", main.i), xlab="Size (t)", ylab="Number of seeds")
-      points(z.t$size, E_seed, cex=0.2, col="red")
-      # lines(xx$size, predict(fo@fitFec[[2]], xx, type="response"), col="blue")
-    w1 <- Re(eigen(IPMs[,,i])$vectors[,1])[-1]; stable.dist1 <- w1/sum(w1)
-    # w2 <- Re(eigen(IPM2)$vectors[,1])[-1]; stable.dist2 <- w2/sum(w2)
-    yl <- range(stable.dist1)#range(c(stable.dist1, stable.dist2))
-    plot(y, stable.dist1, ylab="", type="l", lwd=2, col="red", ylim=yl,
+      lines(mx[,2], exp(mx[,1:n_z$seed]%*%p.ipm$seed_z + c(X$seed[i,]%*%p.ipm$seed_x)),
+            col=2)
+    plot(y, stable.dist1, ylab="", type="l", col=2, 
          main=paste("Stable size distribution:", main.i), xlab="Size (t)")
-    # lines(y, stable.dist2, col="blue")
-    hist(d[[i]]$sizeNext[rcrt.i], main=paste("Recruit sizes:", main.i), 
-         xlab="Recruit size (t+1)", col="grey", freq=F)
-      xxx <- seq(0, hi*.5, length.out=100)
-      rcrt.z <- d[[i]]$sizeNext[is.na(d[[i]]$size) & !is.na(d[[i]]$sizeNext)]
-      lines(xxx, dnorm(xxx, p$rcr_z[1], p$rcr_z[2]), col="red", lwd=2)
-      # lines(xxx, dnorm(xxx, mean(rcrt.z), sd(rcrt.z)), col="blue")
+    hist(d[[i]]$sizeNext[is.na(d[[i]]$size)], col="grey", freq=F, 
+         main=paste("Recruit sizes:", main.i), xlab="Recruit size (t+1)")
+      lines(xxx, dnorm(xxx, p.ipm$rcr_z[1], p.ipm$rcr_z[2]), col="red", lwd=2)
   }
 }
-
+if(ipmPlots) {
+  par(mfrow=c(2,2), mar=c(4,4,2,1))
+  x <- c(0, y)
+  for(i in sample.cells) {
+   z.t <- d[[i]][!is.na(d[[i]]$size) & d[[i]]$yr==tmax,] 
+   main.i <- paste("Cell", i)
+   z.t$y <- y[findInterval(z.t$size, y, all.inside=T)]
+   totSeeds <- group_by(z.t, y) %>% summarise(seed=sum(seed, na.rm=T))
+   image.plot(x, x, t(Ps[,,i]), main=paste("P matrix:", main.i), 
+              xlab="Size (t)", ylab="Size (t+1)")
+    points(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,1/(n_i[i]^.25)))
+    abline(a=0, b=1, col="white", lty=2)
+   image.plot(x, x, t(Fs[,,i]), main=paste("F matrix:", main.i), 
+              xlab="Size (t)", ylab="Size (t+1)")
+    points(totSeeds$y, totSeeds$seed/max(totSeeds$seed)*hi*.9)
+   image.plot(x, x, t(sens(IPMs[,,i])), main=paste("Sensitivity:", main.i), 
+              xlab="Size (t)", ylab="Size (t+1)")
+   image.plot(x, x, t(elas(IPMs[,,i])), main=paste("Elasticity:", main.i),
+              xlab="Size (t)", ylab="Size (t+1)")
+  }
+}
+image.plot(x, x, t(apply(Ps.f, 1:2, mean)), main="P matrix: mean",
+           xlab="Size (t)", ylab="Size (t+1)")
+image.plot(x, x, t(apply(Fs.f, 1:2, mean)), main="F matrix: mean",
+           xlab="Size (t)", ylab="Size (t+1)")
+image.plot(x, x, sens(t(apply(IPMs.f, 1:2, mean))), main="Sensitivity: mean",
+           xlab="Size (t)", ylab="Size (t+1)")
+image.plot(x, x, elas(t(apply(IPMs.f, 1:2, mean))), main="Elasticity: mean",
+           xlab="Size (t)", ylab="Size (t+1)")
 ```

### 2018-03-12 exploration.Rmd
made functions for fitting IPM and simulation to make O -> P simpler

--- 2018-03-11 exploration.Rmd
@@
 ```{r sp_i}
-mean_abund <- 100
-regPlots <- F
-ipmPlots <- F
+mean_abund <- 50
+regPlots <- T
+ipmPlots <- T
 nPlots <- 5
 z.rng <- c(1, 12)
-n <- 10  # ncells in IPM matrix
-tmax <- 30
+n <- 50  # ncells in IPM matrix
+tmax <- 10
 cells <- 1:n_distinct(lc.df$CellID)
@@
 ########
-# set up storage & IPM matrix objects
+# set up storage objects & environmental covariates
 p <- sp_i$p
-setup_IPM_matrix(n, z.rng, buffer=0.5, discrete=1)
-IPMs <- Ps <- Fs <- Fb <- array(0, dim=c(n+1, n+1, n.cell))
-Nt <- array(dim=c(n+1, n+1, n.cell, tmax))
-d <- sdd.j <- vector("list", length=n.cell)
-B <- cbind(rpois(n.cell, n_i), matrix(0, nrow=n.cell, ncol=tmax))
-nSd <- D <- p_est.i <- matrix(nrow=n.cell, ncol=tmax)
-lam.t <- p_est.t <- matrix(nrow=n.cell, ncol=tmax-1)
 sdd <- array(0, dim=dim(sdd.pr) + c(0,0,1,0)) # add layer for nEmig
 sdd[,,1:2,] <- sdd.pr
-
-# set up environmental covariates
 n_z <- list(s=length(p$s_z), g=length(p$g_z),
@@
 ########
-## loop through cells
+## generate underlying IPM and simulated data
 ########
@@
 #---- Truth: use assigned slopes to fill IPM matrix
-## local growth
-for(i in 1:n.cell) {
-  Ps[,,i] <- fill_P(h, y, z.i, n, p, n_z, n_x, X$s[i,], X$g[i,])
-  Fb[,,i] <- fill_F(h, y, z.i, n, p, n_z, n_x, X$fl[i,], X$seed[i,])
-  Fs[z.i,z.i,i] <- (1-p$p_emig) * Fb[z.i,z.i,i]
-  Fs[1,z.i,i] <- (1-p$p_emig) * Fb[1,z.i,i]
-  Fs[z.i,1,i] <- Fb[z.i,1,i]
-}
-## dispersal & density dependence
-for(i in 1:n.cell) {
-  sdd.i <- env.df$id[i] # full rectangle grid id for SDD indexes
-  sdd.j[[i]] <- which(sdd[,,2,]==sdd.i, arr.ind=TRUE) # j cells with target i
-  p.ij <- sdd[,,1,][sdd.j[[i]]] # p(j to i | immigrant seed from j)
-  Fs[z.i,z.i,i] <- Fs[z.i,z.i,i] + 
-    Reduce(`+`, map2(sdd.j[[i]][,3], p$p_emig*p.ij, ~(Fb[z.i,z.i,.x] * .y)))
-  Fs[1,z.i,i] <- Fs[1,z.i,i] + Fb[1,z.i,sdd.j[[i]][,3]] %*% (p$p_emig*p.ij)
-  if(p$NDD) {
-    Nt[,,i,1] <- rpois((n+1)*(n+1), n_i[i])
-    F.i <- F.t <- Fs[,,i]/p$p_est  # F.s[,,i] was already multiplied by p$p_est
-    for(k in 1:(tmax-1)) {
-      p_est.t[i,k] <- min(p$NDD_n/(sum(F.i[z.i,]*Nt[z.i,,i,k])), p$p_est)
-      F.t[z.i,] <- F.i[z.i,] * p_est.t[i,k]
-      IPM.t <- Ps[,,i] + F.t
-      Nt[,,i,k+1] <- IPM.t %*% Nt[,,i,k]
-      lam.t[i,k] <- Re(eigen(IPM.t)$values[1])
-    }
-  }
-}
-IPMs <- Ps + Fs
+U <- fill_IPM_matrices(n.cell, tmax, n_i, n, z.rng, buffer=0.5, discrete=1,
+                       p, n_z, n_x, X, sdd.pr, env.df)
 
 #---- Realization: generate simulated data
-## local growth
-for(i in 1:n.cell) {
-  d[[i]] <- data.frame(yr=1, size=runif(n=n_i[i], min=z.rng[1], max=z.rng[2]))
-  E <- sim_expected(d[[i]], p, map(X, ~(.[i,])), n_z)
-  d[[i]] <- sim_realized(d[[i]], 1:nrow(d[[i]]), E, p, lo, hi)
-  nSd[i,1] <- sum(d[[i]]$seed, na.rm=T)
-  sdd[,,3,i] <- nSd[i,1] * p$p_emig * sdd[,,1,i]
-}
-## dispersal & density dependence
-for(i in 1:n.cell) {
-  D[i,1] <- sum(sdd[,,3,][sdd.j[[i]]])
-  p_est.i[i,1] <- ifelse(p$NDD, min(p$NDD_n/(nSd[i,1]+D[i,1]),p$p_est), p$p_est)
-  d[[i]] <- sim_recruits(d[[i]], p_est.i[i,1], nSd[i,1], B[i,1], D[i,1], p)
-  B[i,2] <- sim_seedbank(nSd[i,1], B[i,1], D[i,1], p)
-}
-## iterate forward
-for(k in 2:tmax) {
-  for(i in 1:n.cell) {
-    size.k <- d[[i]]$sizeNext[!is.na(d[[i]]$sizeNext) & d[[i]]$yr == k-1]
-    if(length(size.k) > 0) {
-      k.index <- (1:length(size.k)) + nrow(d[[i]])
-      d[[i]][k.index, "size"] <- size.k
-      d[[i]][k.index, "yr"] <- k
-      E <- sim_expected(d[[i]][k.index,], p, map(X, ~(.[i,])), n_z)
-      d[[i]] <- sim_realized(d[[i]], k.index, E, p, lo, hi)
-      nSd[i,k] <- sum(d[[i]]$seed[k.index], na.rm=T)
-    } else {
-      nSd[i,k] <- 0
-    }
-    sdd[,,3,i] <- nSd[i,k] * p$p_emig * sdd[,,1,i]
-  }
-  for(i in 1:n.cell) {
-    D[i,k] <- sum(sdd[,,3,][sdd.j[[i]]])
-    p_est.i[i,k] <- ifelse(p$NDD, min(p$NDD_n/(nSd[i,k]+D[i,k]),p$p_est), p$p_est)
-    d[[i]] <- sim_recruits(d[[i]], p_est.i[i,k], nSd[i,k], B[i,k], D[i,k], p)
-    B[i,k+1] <- sim_seedbank(nSd[i,k], B[i,k], D[i,k], p)
-  }
-}
+S <- simulate_data(n.cell, tmax, n_i, z.rng, U$lo, U$hi, 
+                   p, X, n_z, sdd, U$sdd.j)
 
+#---- Aggregate results
 lam.df <- env.df
-lam.df$lambda <- apply(IPMs, 3, function(x) Re(eigen(x)$values[1]))
+lam.df$lambda <- apply(U$IPMs, 3, function(x) Re(eigen(x)$values[1]))
 lam.df %<>% 
-  add_column(nSeed=nSd[,tmax], B0=B[,1], Btmax=B[,tmax+1], D=D[,tmax], 
-             N=map_dbl(d, ~sum(!is.na(.$sizeNext[.$yr==tmax]))),
-             N.surv=map_dbl(d, ~sum(.$surv[.$yr==tmax], na.rm=T)),
-             N.rcr=map_dbl(d, ~sum(is.na(.$size[.$yr==tmax])))) %>%
+  add_column(nSeed=S$nSd[,tmax], D=S$D[,tmax], 
+             B0=S$B[,1], Btmax=S$B[,tmax+1], 
+             N.S=map_dbl(S$d, ~sum(!is.na(.$sizeNext[.$yr==tmax]))),
+             Surv.S=map_dbl(S$d, ~sum(.$surv[.$yr==tmax], na.rm=T)),
+             Rcr.S=map_dbl(S$d, ~sum(is.na(.$size[.$yr==tmax])))) %>%
   mutate(nSdStay=nSeed*(1-p$p_emig), nSdLeave=nSeed*p$p_emig) %>%
-  add_column(Ntmax=apply(Nt[,,,tmax],3,sum), lam.tmax=lam.t[tmax-1,])
+  add_column(N.U=apply(U$Nt[,,,tmax],3,sum), lam.U=U$lam.t[,tmax-1])
 
@@
   scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
-ggplot(lam.df, aes(x=lon, y=lat, fill=log(lam.tmax))) + geom_tile() +
+ggplot(lam.df, aes(x=lon, y=lat, fill=log(lam.U))) + geom_tile() +
   scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
-ggplot(lam.df, aes(x=lon, y=lat, fill=log(Ntmax)*(Ntmax>10))) + geom_tile() +
+ggplot(lam.df, aes(x=lon, y=lat, fill=log(N.U)*(N.U>10))) + geom_tile() +
   scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=1)
 ggplot(lam.df, aes(x=lon, y=lat, fill=lambda>1)) + geom_tile()
-ggplot(lam.df, aes(x=lon, y=lat, fill=lam.tmax>1)) + geom_tile()
-ggplot(lam.df, aes(x=lon, y=lat, fill=Ntmax>1)) + geom_tile() 
-ggplot(lam.df, aes(x=lon, y=lat, fill=Ntmax>1 & lam.tmax<1)) + geom_tile() 
-ggplot(lam.df, aes(x=lon, y=lat, fill=log(N))) + geom_tile() +
+ggplot(lam.df, aes(x=lon, y=lat, fill=N.S>1)) + geom_tile()
+ggplot(lam.df, aes(x=lon, y=lat, fill=lam.U>1)) + geom_tile()
+ggplot(lam.df, aes(x=lon, y=lat, fill=N.U>1)) + geom_tile() 
+ggplot(lam.df, aes(x=lon, y=lat, fill=N.U>1 & lam.U<1)) + geom_tile() 
+ggplot(lam.df, aes(x=lon, y=lat, fill=N.S>1 & lambda<1)) + geom_tile()
+ggplot(lam.df, aes(x=lon, y=lat, fill=log(N.S))) + geom_tile() +
   scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
-ggplot(lam.df, aes(x=lon, y=lat, fill=log(N.surv))) + geom_tile() +
+ggplot(lam.df, aes(x=lon, y=lat, fill=log(Surv.S))) + geom_tile() +
   scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
-ggplot(lam.df, aes(x=lon, y=lat, fill=log(N.rcr))) + geom_tile() +
+ggplot(lam.df, aes(x=lon, y=lat, fill=log(Rcr.S))) + geom_tile() +
   scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
@@
 ggplot(lam.df, aes(x=lon, y=lat, fill=D)) + geom_tile()
+ggplot(lam.df, aes(x=lon, y=lat, fill=D + nSdStay)) + geom_tile()
 ggplot(lam.df, aes(x=lon, y=lat, fill=Btmax)) + geom_tile()
-ggplot(lam.df, aes(x=lon, y=lat, fill=D + nSdStay)) + geom_tile()
 ggplot(lam.df, aes(x=lon, y=lat, fill=Btmax>B0)) + geom_tile()
-ggplot(lam.df, aes(x=lon, y=lat, fill=nSdStay/D)) + geom_tile() +
-  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=1)
 ggplot(lam.df, aes(x=lon, y=lat, fill=nSdStay > D)) + geom_tile()
@@
 plot(density(log(lam.df$lambda))); abline(v=0, col="red", lty=2)
+matplot(t(log(U$lam.t)), type="l", lty=1, col=rgb(0,0,0,0.1))
+matplot(t(log(apply(U$Nt, 3:4, sum))), type="l", lty=1, col=rgb(0,0,0,0.1))
+matplot(t(S$N_sim[,2:tmax]/S$N_sim[,1:(tmax-1)]), 
+        type="l", lty=1, col=rgb(0,0,0,.1))
+matplot(t(log(S$N_sim+1)), type="l", lty=1, col=rgb(0,0,0,0.1))
+#matplot(t(log(nSd*(1-p$p_emig)+D)), type="l", lty=1, col=rgb(0,0,0,0.1))
 
-sample.cells <- sample.int(n.cell, nPlots)
+sample.cells <- sample((1:n.cell)[lam.df$Surv.S>0], nPlots)
 if(regPlots) {
   par(mfrow=c(3,2), mar=c(4,4,2,1))
-  xx <- tibble(size=seq(lo, hi, length.out=100), size2=size^2, size3=size^3)
-  mx <- cbind(1, as.matrix(xx))
-  xxx <- seq(0, hi*.5, length.out=100)
+  xx <- tibble(size=seq(U$lo, U$hi, length.out=100), size2=size^2, size3=size^3)
+  E.t <- xx %>% add_column(s=NA, g=NA, fl=NA, seed=NA)
+  xxx <- seq(0, U$hi*.5, length.out=100)
   for(i in sample.cells) {
-    z.t <- d[[i]][!is.na(d[[i]]$size),]; main.i <- paste("Cell", i)
-    w1 <- Re(eigen(IPMs[,,i])$vectors[,1])[-1]; stable.dist1 <- w1/sum(w1)
-    rcrt.z <- d[[i]]$sizeNext[is.na(d[[i]]$size) & !is.na(d[[i]]$sizeNext)]
+    z.t <- S$d[[i]][!is.na(S$d[[i]]$size) & S$d[[i]]$yr==tmax,]
+    E.t <- sim_expected(E.t, 1:100, 1:100, E.t, p, map(X, ~.[i,]), n_z)
+    main.i <- paste("Cell", i)
+    w1 <- Re(eigen(U$IPMs[,,i])$vectors[,1])[-1]; stable.dist1 <- w1/sum(w1)
     plot(surv~size, data=z.t, ylim=c(0,1), pch=20, col=rgb(0,0,0,0.2),
          main=paste("Surv:", main.i), xlab="Size (t)", ylab="P(surv to t+1)")
-      lines(mx[,2], antilogit(mx[,1:n_sz]%*%p$s_z + c(X_s[i,]%*%p$s_x)), col=2)
-    plot(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,0.2), 
+      lines(E.t$size, E.t$s, col=2)
+    plot(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,0.2), ylim=c(U$lo,U$hi),
          main=paste("Growth:", main.i), xlab="Size (t)", ylab="Size (t+1)")
-      lines(mx[,2], mx[,1:n_gz]%*%p$g_z + c(X_g[i,]%*%p$g_x), col=2)
+      lines(E.t$size, E.t$g, col=2)
       abline(a=0, b=1, lty=2)
@@
          main=paste("P(Flower):", main.i), xlab="Size (t)", ylab="P(flower)")
-      lines(mx[,2], antilogit(mx[,1:n_flz]%*%p$fl_z + c(X_fl[i,]%*%p$fl_x)),
-            col=2)
+      lines(E.t$size, E.t$fl, col=2)
     plot(seed~size, data=z.t, pch=23-z.t$fl*3, col=rgb(0,0,0,0.2), 
+         ylim=range(c(E.t$seed, z.t$seed), na.rm=TRUE),
          main=paste("nSeeds:", main.i), xlab="Size (t)", ylab="Number of seeds")
-      lines(mx[,2], exp(mx[,1:n_seedz]%*%p$seed_z + c(X_seed[i,]%*%p$seed_x)),
-            col=2)
-    plot(y, stable.dist1, ylab="", type="l", col=2, 
+      lines(E.t$size, E.t$seed, col=2)
+    plot(U$y, stable.dist1, ylab="", type="l", col=2, 
          main=paste("Stable size distribution:", main.i), xlab="Size (t)")
-    hist(d[[i]]$sizeNext[is.na(d[[i]]$size)], col="grey", freq=F, 
+    hist(with(S$d[[i]], sizeNext[is.na(size)]), col="grey", freq=F, 
          main=paste("Recruit sizes:", main.i), xlab="Recruit size (t+1)")
-      lines(xxx, dnorm(xxx, p$rcr_z[1], p$rcr_z[2]), col="red", lwd=2)
+      lines(xxx, dnorm(xxx, p$rcr_z[1], p$rcr_z[2]), col="red")
   }
@@
   par(mfrow=c(2,2), mar=c(4,4,2,1))
-  x <- c(0, y)
+  x <- c(0, U$y)
   for(i in sample.cells) {
-    z.t <- d[[i]][!is.na(d[[i]]$size),]; main.i <- paste("Cell", i)
-    z.t$y <- y[findInterval(z.t$size, y, all.inside=T)]
+   z.t <- S$d[[i]][!is.na(S$d[[i]]$size) & S$d[[i]]$yr==tmax,]
+   main.i <- paste("Cell", i)
+   z.t$y <- U$y[findInterval(z.t$size, U$y, all.inside=T)]
    totSeeds <- group_by(z.t, y) %>% summarise(seed=sum(seed, na.rm=T))
-    image.plot(x, x, t(Ps[,,i]), main=paste("P matrix:", main.i), 
+   image.plot(x, x, t(U$Ps[,,i]), main=paste("P matrix:", main.i), 
               xlab="Size (t)", ylab="Size (t+1)")
-      points(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,1/(n_i[i]^.25)))
+    points(sizeNext~size, data=z.t, col=rgb(1,1,1,1/(nrow(z.t)^.25)))
     abline(a=0, b=1, col="white", lty=2)
-    image.plot(x, x, t(Fs[,,i]), main=paste("F matrix:", main.i), 
+   image.plot(x, x, t(U$Fs[,,i]), main=paste("F matrix:", main.i), 
               xlab="Size (t)", ylab="Size (t+1)")
-      points(totSeeds$y, totSeeds$seed/max(totSeeds$seed)*hi*.9)
-    image.plot(x, x, t(sens(IPMs[,,i])), main=paste("Sensitivity:", main.i), 
+    points(totSeeds$y, totSeeds$seed/max(totSeeds$seed)*U$hi*.9, col="white")
+   image.plot(x, x, t(sens(U$IPMs[,,i])), main=paste("Sensitivity:", main.i), 
               xlab="Size (t)", ylab="Size (t+1)")
-    image.plot(x, x, t(elas(IPMs[,,i])), main=paste("Elasticity:", main.i),
+   image.plot(x, x, t(elas(U$IPMs[,,i])), main=paste("Elasticity:", main.i),
               xlab="Size (t)", ylab="Size (t+1)")
+   points(sizeNext~size, data=z.t, col=rgb(1,1,1,1/(nrow(z.t)^.25)))
   }
 }
-image.plot(x, x, t(apply(Ps, 1:2, mean)), main="P matrix: mean",
+image.plot(x, x, t(apply(U$Ps, 1:2, mean)), main="P matrix: mean",
            xlab="Size (t)", ylab="Size (t+1)")
-image.plot(x, x, t(apply(Fs, 1:2, mean)), main="F matrix: mean",
+image.plot(x, x, t(apply(U$Fs, 1:2, mean)), main="F matrix: mean",
            xlab="Size (t)", ylab="Size (t+1)")
-image.plot(x, x, sens(t(apply(IPMs, 1:2, mean))), main="Sensitivity: mean",
+image.plot(x, x, sens(t(apply(U$IPMs, 1:2, mean))), main="Sensitivity: mean",
            xlab="Size (t)", ylab="Size (t+1)")
-image.plot(x, x, elas(t(apply(IPMs, 1:2, mean))), main="Elasticity: mean",
+image.plot(x, x, elas(t(apply(U$IPMs, 1:2, mean))), main="Elasticity: mean",
            xlab="Size (t)", ylab="Size (t+1)")
@@
 ```{r O_setup}
-O_n <- c(Mx=500, CA=20, IPM=20) # number of cells in sample
+O_n <- list(Mx=500, CA=20, IPM=20) # number of cells in sample
+O_yr <- list(Mx=tmax, CA=(tmax-10):tmax, IPM=tmax) # years to sample from
+P.i <- which(lam.df$Surv.S > 0) # presences: surviving individuals > seedlings
 ```
@@
 ```{r O_IPM}
-
+IPM.P <- sample(P.i, O_n$IPM, replace=FALSE)
+IPM.d <- vector("list", O_n$IPM)
+for(i in IPM.P) {
+  IPM.d[[i]] <- S$d[[i]] %>% filter(yr %in% O_yr$IPM) %>%
+    add_column(id.inbd=i) %>%
+    full_join(env.df[i,], by="id.inbd")
+}
+IPM.d <- do.call(rbind, IPM.d)
+p.ipm <- p
 ```
@@
 ```{r P_IPM}
-for(i in 1:ncell) {
-  # ## IPMpack: estimate slopes from d[[i]]
-  # ### set up seedbank transitions
-  d[[i]]$stage <- d[[i]]$stageNext <- "continuous"
-  d[[i]]$stage[is.na(d[[i]]$size)] <- NA
-  d[[i]]$stageNext[d[[i]]$surv==0] <- "dead"
-  d[[i]]$number <- 1
-  seedbank <- data.frame(size=NA, surv=1, sizeNext=NA,
-                         fl=NA, seed=NA,
-                         stage=c("seedbank", "seedbank", "continuous"),
-                         stageNext=c("seedbank", "continuous", "seedbank"),
-                         number=c(p$s_SB * (1 - p$rcr_SB), 
-                                  p$rcr_SB * p$p_est, 
-                                  (1 - p$rcr_dir) * p$s_SB))
-  d[[i]] <- rbind(d[[i]], seedbank)
-  d[[i]]$stage <- as.factor(d[[i]]$stage)
-  d[[i]]$stageNext <- as.factor(d[[i]]$stageNext)
-  ### construct kernels
-  so <- makeSurvObj(d[[i]], surv ~ size)
-  go <- makeGrowthObj(d[[i]], sizeNext ~ size + size2)
-  fo <- makeFecObj(d[[i]],
-                   Formula=c(fl~size+size2, seed~size+size2),
-                   Family=c("binomial", "poisson"),
-                   Transform=c("none", "none"),
-                   meanOffspringSize=mean(
-                     d[[i]]$sizeNext[is.na(d[[i]]$size) &
-                                       !is.na(d[[i]]$sizeNext)]),
-                   sdOffspringSize=sd(
-                     d[[i]]$sizeNext[is.na(d[[i]]$size) &
-                                       !is.na(d[[i]]$sizeNext)]),
-                   fecConstants=data.frame(seedlingEstablishmentRate=p$p_est),
-                   offspringSplitter=data.frame(seedbank=(1-p$rcr_dir),
-                                                continuous=p$rcr_dir)
-                   )
-  F.mx2 <- makeIPMFmatrix(fecObj=fo, minSize=lo, maxSize=hi,
-                          correction="constant")
-  D.mx2 <- makeDiscreteTrans(d[[i]], 
-                             continuousToDiscreteExplanatoryVariables=1)
-  dummy <- as.matrix(fo@offspringRel$coefficients[1], list(1, "seedbank"))
-  D.mx2@meanToCont <- as.matrix(dummy, dimnames=c(1, "seedbank"))
-  dummy <- as.matrix(fo@sdOffspringSize, dimnames=list(1, "seedbank"))
-  D.mx2@sdToCont <- as.matrix(dummy, dimnames=c(1, "seedbank"))
-  D.mx2@discreteTrans[1,1] <- p$s_SB * (1 - p$rcr_SB)
-  D.mx2@discreteTrans[3,1] <- 1 - sum(D.mx2@discreteTrans[1:2,1])
-  P.mx2 <- makeIPMPmatrix(growObj=go, survObj=so, discreteTrans=D.mx2,
-                            minSize=lo, maxSize=hi, correction="constant")
-  IPM2 <- P.mx2 + F.mx2
-  lambdas2[i] <- Re(eigen(IPM2)$values[1])
+# run chunk O_IPM first to generate observed dataset `IPM.d`
 
+########
+## fit models
+########
+s.m <- glm(surv ~ size + I(size^2) + temp + temp2 + prec + prec2 + pOpn + 
+               pOth + pDec, data=IPM.d, family="binomial")
+g.m <- lm(sizeNext ~ size + I(size^2) + temp + temp2 + prec + prec2 +
+              pOpn + pOth, data=IPM.d)
+fl.m <- glm(fl ~ size + I(size^2) + temp + temp2 + prec + prec2 + pOpn +
+                pOth, data=IPM.d, family="binomial")
+seed.m <- glm(seed ~ size + I(size^2) + temp + temp2 + prec + prec2 + pOpn, 
+              data=IPM.d, family="poisson")
+p.ipm$s_z <- coef(s.m)[1:length(p.ipm$s_z)]
+p.ipm$s_x <- coef(s.m)[(1:length(p.ipm$s_x)) + length(p.ipm$s_z)]
+p.ipm$g_z <- coef(g.m)[1:length(p.ipm$g_z)]
+p.ipm$g_x <- coef(g.m)[(1:length(p.ipm$g_x)) + length(p.ipm$g_z)]
+p.ipm$g_sig <- summary(g.m)$sigma
+p.ipm$fl_z <- coef(fl.m)[1:length(p.ipm$fl_z)]
+p.ipm$fl_x <- coef(fl.m)[(1:length(p.ipm$fl_x)) + length(p.ipm$fl_z)]
+p.ipm$seed_z <- coef(seed.m)[1:length(p.ipm$seed_z)]
+p.ipm$seed_x <- coef(seed.m)[(1:length(p.ipm$seed_x)) + length(p.ipm$seed_z)]
+p.ipm$rcr_z <- c(mean(IPM.d[is.na(IPM.d$size), "sizeNext"]),
+                 sd(IPM.d[is.na(IPM.d$size), "sizeNext"]))
+
+n_z <- list(s=length(p.ipm$s_z), g=length(p.ipm$g_z),
+            fl=length(p.ipm$fl_z), seed=length(p.ipm$seed_z))
+n_x <- list(s=length(p.ipm$s_x), g=length(p.ipm$g_x),
+            fl=length(p.ipm$fl_x), seed=length(p.ipm$seed_x))
+X <- list(s=as.matrix(env.df[,1:n_x$s]), g=as.matrix(env.df[,1:n_x$g]),
+          fl=as.matrix(env.df[,1:n_x$fl]), seed=as.matrix(env.df[,1:n_x$seed]))
+
+########
+## predict distribution with fit models
+########
+#---- Truth: use assigned slopes to fill IPM matrix
+U.f <- fill_IPM_matrices(n.cell, tmax, n_i, n, z.rng, buffer=0.5, discrete=1,
+                         p.ipm, n_z, n_x, X, sdd, env.df)
+
+#---- Realization: generate simulated data
+S.f <- simulate_data(n.cell, tmax, n_i, z.rng, U.f$lo, U.f$hi, 
+                     p.ipm, X, n_z, sdd, U.f$sdd.j)
+
+#---- Aggregate results
+lam.df$lambda.f <- apply(U.f$IPMs, 3, function(x) Re(eigen(x)$values[1]))
+lam.df %<>% 
+  add_column(nSeed.f=S.f$nSd[,tmax], D.f=S.f$D[,tmax], 
+             B0.f=S.f$B[,1], Btmax.f=S.f$B[,tmax+1], 
+             N.S.f=map_dbl(S.f$d, ~sum(!is.na(.$sizeNext[.$yr==tmax]))),
+             Surv.S.f=map_dbl(S.f$d, ~sum(.$surv[.$yr==tmax], na.rm=T)),
+             Rcr.S.f=map_dbl(S.f$d, ~sum(is.na(.$size[.$yr==tmax])))) %>%
+  mutate(nSdStay.f=nSeed.f*(1-p.ipm$p_emig), nSdLeave.f=nSeed.f*p.ipm$p_emig) %>%
+  add_column(N.U.f=apply(U.f$Nt[,,,tmax],3,sum), lam.U.f=U.f$lam.t[,tmax-1])
+
+########
+## plots
+########
+ggplot(lam.df, aes(x=lon, y=lat, fill=lambda.f-lambda)) + geom_tile() +
+  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
+ggplot(lam.df, aes(x=lon, y=lat, fill=lam.U.f-lam.U)) + geom_tile() +
+  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
+ggplot(lam.df, aes(x=lon, y=lat)) + 
+  geom_tile(aes(fill=(lambda>1 & lambda.f>1) | (lambda<1 & lambda.f<1)))
+ggplot(lam.df, aes(x=lon, y=lat)) + 
+  geom_tile(aes(fill=(N.S>1 & N.S.f>1) | (N.S<1 & N.S.f<1)))
+ggplot(lam.df, aes(x=lon, y=lat)) + 
+  geom_tile(aes(fill=(lam.U>1 & lam.U.f>1) | (lam.U<1 & lam.U.f<1)))
+ggplot(lam.df, aes(x=lon, y=lat)) + 
+  geom_tile(aes(fill=(N.U>1 & N.U.f>1) | (N.U<1 & N.U.f<1)))
+ggplot(lam.df, aes(x=lon, y=lat, fill=(N.S.f-N.S)/N.S)) + geom_tile() +
+  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
+ggplot(lam.df, aes(x=lon, y=lat, fill=(Surv.S.f-Surv.S)/Surv.S)) + geom_tile() +
+  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
+ggplot(lam.df, aes(x=lon, y=lat, fill=(Rcr.S.f-Rcr.S)/Rcr.S)) + geom_tile() +
+  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
+par(mfrow=c(2,2))
+plot((lambda.f-lambda) ~ temp, data=lam.df); abline(h=0, col="red", lty=2)
+plot((lambda.f-lambda) ~ prec, data=lam.df); abline(h=0, col="red", lty=2)
+plot((lambda.f-lambda) ~ pOpn, data=lam.df); abline(h=0, col="red", lty=2)
+plot(density((lam.df$lambda.f-lam.df$lambda))); abline(v=0, col="red", lty=2)
+matplot(t(log(U.f$lam.t)), type="l", lty=1, col=rgb(0,0,0,0.1))
+matplot(t(log(apply(U.f$Nt, 3:4, sum))), type="l", lty=1, col=rgb(0,0,0,0.1))
+matplot(t(S.f$N_sim[,2:tmax]/S$N_sim[,1:(tmax-1)]), 
+        type="l", lty=1, col=rgb(0,0,0,.1))
+matplot(t(log(S.f$N_sim+1)), type="l", lty=1, col=rgb(0,0,0,0.1))
+
+sample.cells <- sample((1:n.cell)[lam.df$Surv.S>0], nPlots)
 if(regPlots) {
   par(mfrow=c(3,2), mar=c(4,4,2,1))
-    xx <- tibble(size=seq(lo, hi, length.out=100), size2=size^2, size3=size^3)
-    plot(surv~size, data=d[[i]], ylim=c(0,1), pch=20, col=rgb(0,0,0,0.2),
+  xx <- tibble(size=seq(U.f$lo, U.f$hi, length.out=100), 
+               size2=size^2, size3=size^3)
+  E.t <- xx %>% add_column(s=NA, g=NA, fl=NA, seed=NA)
+  xxx <- seq(0, U.f$hi*.5, length.out=100)
+  for(i in sample.cells) {
+    z.t <- S.f$d[[i]][!is.na(S.f$d[[i]]$size) & S.f$d[[i]]$yr==tmax,]
+    E.t <- sim_expected(E.t, 1:100, 1:100, E.t, p.ipm, map(X, ~.[i,]), n_z)
+    E.S.t <- sim_expected(E.t, 1:100, 1:100, E.t, p, map(X, ~.[i,]), n_z)
+    main.i <- paste("Cell", i)
+    w1 <- Re(eigen(U.f$IPMs[,,i])$vectors[,1])[-1]; stable.dist1 <- w1/sum(w1)
+    w.S <- Re(eigen(U$IPMs[,,i])$vectors[,1])[-1]; stable.dist.S <- w.S/sum(w.S)
+    plot(surv~size, data=z.t, ylim=c(0,1), pch=20, col=rgb(0,0,0,0.2),
          main=paste("Surv:", main.i), xlab="Size (t)", ylab="P(surv to t+1)")
-      points(z.t$size, E_surv, cex=0.2, col="red")
-      # lines(xx$size, predict(so@fit, xx, type="response"), col="blue")
-    plot(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,0.2), 
+      lines(E.t$size, E.t$s, col=2); lines(E.S.t$size, E.S.t$s, col=4)
+    plot(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,0.2), ylim=c(U$lo,U$hi),
          main=paste("Growth:", main.i), xlab="Size (t)", ylab="Size (t+1)")
-      points(z.t$size, E_grow, cex=0.2, col="red")
-      # lines(xx$size, predict(go@fit, xx, type="response"), col="blue")
+      lines(E.t$size, E.t$g, col=2); lines(E.S.t$size, E.S.t$g, col=4)
       abline(a=0, b=1, lty=2)
@@
          main=paste("P(Flower):", main.i), xlab="Size (t)", ylab="P(flower)")
-      points(z.t$size, E_fl, cex=0.2, col="red")
-      # lines(xx$size, predict(fo@fitFec[[1]], xx, type="response"), col="blue")
+      lines(E.t$size, E.t$fl, col=2); lines(E.S.t$size, E.S.t$fl, col=4)
     plot(seed~size, data=z.t, pch=23-z.t$fl*3, col=rgb(0,0,0,0.2), 
+         ylim=range(c(E.t$seed, z.t$seed), na.rm=TRUE),
          main=paste("nSeeds:", main.i), xlab="Size (t)", ylab="Number of seeds")
-      points(z.t$size, E_seed, cex=0.2, col="red")
-      # lines(xx$size, predict(fo@fitFec[[2]], xx, type="response"), col="blue")
-    w1 <- Re(eigen(IPMs[,,i])$vectors[,1])[-1]; stable.dist1 <- w1/sum(w1)
-    # w2 <- Re(eigen(IPM2)$vectors[,1])[-1]; stable.dist2 <- w2/sum(w2)
-    yl <- range(stable.dist1)#range(c(stable.dist1, stable.dist2))
-    plot(y, stable.dist1, ylab="", type="l", lwd=2, col="red", ylim=yl,
+      lines(E.t$size, E.t$seed, col=2); lines(E.S.t$size, E.S.t$seed, col=4)
+    plot(U.f$y, stable.dist1, ylab="", type="l", col=2, 
          main=paste("Stable size distribution:", main.i), xlab="Size (t)")
-    # lines(y, stable.dist2, col="blue")
-    hist(d[[i]]$sizeNext[rcrt.i], main=paste("Recruit sizes:", main.i), 
-         xlab="Recruit size (t+1)", col="grey", freq=F)
-      xxx <- seq(0, hi*.5, length.out=100)
-      rcrt.z <- d[[i]]$sizeNext[is.na(d[[i]]$size) & !is.na(d[[i]]$sizeNext)]
-      lines(xxx, dnorm(xxx, p$rcr_z[1], p$rcr_z[2]), col="red", lwd=2)
-      # lines(xxx, dnorm(xxx, mean(rcrt.z), sd(rcrt.z)), col="blue")
+      lines(U.f$y, stable.dist.S, col=4)
+    hist(with(S.f$d[[i]], sizeNext[is.na(size)]), col="grey", freq=F, 
+         main=paste("Recruit sizes:", main.i), xlab="Recruit size (t+1)")
+      lines(xxx, dnorm(xxx, p.ipm$rcr_z[1], p.ipm$rcr_z[2]), col=2)
+      lines(xxx, dnorm(xxx, p$rcr_z[1], p$rcr_z[2]), col=4)
   }
 }
-
+if(ipmPlots) {
+  par(mfrow=c(2,2), mar=c(4,4,2,1))
+  x <- c(0, U.f$y)
+  for(i in sample.cells) {
+   z.t <- S.f$d[[i]][!is.na(S.f$d[[i]]$size) & S.f$d[[i]]$yr==tmax,]
+   main.i <- paste("Cell", i)
+   z.t$y <- U.f$y[findInterval(z.t$size, U.f$y, all.inside=T)]
+   totSeeds <- group_by(z.t, y) %>% summarise(seed=sum(seed, na.rm=T))
+   image.plot(x, x, t(U.f$Ps[,,i]), main=paste("P matrix:", main.i), 
+              xlab="Size (t)", ylab="Size (t+1)")
+    points(sizeNext~size, data=z.t, col=rgb(1,1,1,1/(nrow(z.t)^.25)))
+    abline(a=0, b=1, col="white", lty=2)
+   image.plot(x, x, t(U.f$Fs[,,i]), main=paste("F matrix:", main.i), 
+              xlab="Size (t)", ylab="Size (t+1)")
+    points(totSeeds$y, totSeeds$seed/max(totSeeds$seed)*U.f$hi*.9, col="white")
+   image.plot(x, x, t(sens(U.f$IPMs[,,i])), main=paste("Sensitivity:", main.i), 
+              xlab="Size (t)", ylab="Size (t+1)")
+   image.plot(x, x, t(elas(U.f$IPMs[,,i])), main=paste("Elasticity:", main.i),
+              xlab="Size (t)", ylab="Size (t+1)")
+   points(sizeNext~size, data=z.t, col=rgb(1,1,1,1/(nrow(z.t)^.25)))
+  }
+}
+image.plot(x, x, t(apply(U.f$Ps, 1:2, mean)), main="P matrix: mean",
+           xlab="Size (t)", ylab="Size (t+1)")
+image.plot(x, x, t(apply(U.f$Fs, 1:2, mean)), main="F matrix: mean",
+           xlab="Size (t)", ylab="Size (t+1)")
+image.plot(x, x, sens(t(apply(U.f$IPMs, 1:2, mean))), main="Sensitivity: mean",
+           xlab="Size (t)", ylab="Size (t+1)")
+image.plot(x, x, elas(t(apply(U.f$IPMs, 1:2, mean))), main="Elasticity: mean",
+           xlab="Size (t)", ylab="Size (t+1)")
 ```

### 2018-03-12 code/fn_IPM.R

@@ -115,13 +115,13 @@ calc_staySB <- function(p) {
 
 ##-- Set boundaries, meshpoints, & step size for IPM matrix
 ##   create objects in global environment
-setup_IPM_matrix <- function(n=100, z.rng=c(1,10), buffer=0.25, discrete=0) {
-  lo <<- z.rng[1]*(1-buffer)  # IPM matrix lower limit
-  hi <<- z.rng[2]*(1+buffer)  # IPM matrix upper limit
-  b <<- lo + c(0:n)*(hi - lo)/n  # boundary points
-  y <<- 0.5*(b[1:n] + b[2:(n+1)])  # mesh points
-  h <<- y[2] - y[1]  # step size
-  z.i <<- (1:n) + discrete  # continuous stage matrix indices
+setup_IPM_matrix <- function(n=100, z.rng=c(1,10), buffer=0.25) {
+  lo <- z.rng[1]*(1-buffer)  # IPM matrix lower limit
+  hi <- z.rng[2]*(1+buffer)  # IPM matrix upper limit
+  b <- lo + c(0:n)*(hi - lo)/n  # boundary points
+  y <- 0.5*(b[1:n] + b[2:(n+1)])  # mesh points
+  h <- y[2] - y[1]  # step size
+  return(list(lo=lo, hi=hi, b=b, y=y, h=h))
 }
 
 
@@ -155,6 +155,56 @@ fill_F <- function(h, y, z.i, n, p, n_z, n_x, X_fl, X_seed) {
 }
 
 
+##-- Fill all IPM objects
+##
+fill_IPM_matrices <- function(n.cell, tmax, n_i, n, z.rng, buffer, discrete,
+                              p, n_z, n_x, X, sdd, env.df, verbose=FALSE) {
+  require(tidyverse)
+  
+  # storage objects
+  IPMs <- Ps <- Fs <- Fb <- array(0, dim=c(n+1, n+1, n.cell))
+  Nt <- array(dim=c(n+1, n+1, n.cell, tmax))
+  sdd.j <- vector("list", length=n.cell)
+  lam.t <- p_est.t <- matrix(nrow=n.cell, ncol=tmax-1)
+  
+  # IPM matrix setup
[…]

### 2018-03-12 code/fn_sim.R

@@ -4,31 +4,30 @@
 
 ##--- simulate expected values
 ## 
-sim_expected <- function(d_i, p, X, n_z) {
-  E <- list(s=NULL, g=NULL, fl=NULL, seed=NULL)
-  size.mx <- cbind(1, d_i$size, d_i$size^2, d_i$size^3)
-  E$s <- antilogit(size.mx[,1:n_z$s] %*% p$s_z + c(X$s %*% p$s_x))
-  E$g <- size.mx[,1:n_z$g] %*% p$g_z + c(X$g %*% p$g_x)
-  E$fl <- antilogit(size.mx[,1:n_z$fl] %*% p$fl_z + c(X$fl %*% p$fl_x))
-  E$seed <- exp(size.mx[,1:n_z$seed] %*% p$seed_z + c(X$seed %*% p$seed_x))
+sim_expected <- function(d_i, d.k, e.k, E, p, X, n_z) {
+  size.mx <- with(d_i[d.k,], cbind(1, size, size^2, size^3))
+  E$s[e.k] <- antilogit(size.mx[,1:n_z$s] %*% p$s_z + c(X$s %*% p$s_x))
+  E$g[e.k] <- size.mx[,1:n_z$g] %*% p$g_z + c(X$g %*% p$g_x)
+  E$fl[e.k] <- antilogit(size.mx[,1:n_z$fl] %*% p$fl_z + c(X$fl %*% p$fl_x))
+  E$seed[e.k] <- exp(size.mx[,1:n_z$seed] %*% p$seed_z + c(X$seed %*% p$seed_x))
   return(E)
 }
 
 
 ##--- simulate realized values
 ##
-sim_realized <- function(d_i, k.index, E, p, lo, hi) {
-  n_t <- length(k.index)
-  d_i$surv[k.index] <- rbinom(n_t, 1, E$s)
-  d_i$sizeNext[k.index] <- rnorm(n_t, E$g, p$g_sig) * 
-    ifelse(d_i$surv[k.index], 1, NA)
-  d_i$sizeNext[k.index] <- ifelse(d_i$sizeNext[k.index] < lo, 
-                                  lo, d_i$sizeNext[k.index])
-  d_i$sizeNext[k.index] <- ifelse(d_i$sizeNext[k.index] > hi, 
-                                  hi, d_i$sizeNext[k.index])
-  d_i$fl[k.index] <- rbinom(n_t, 1, E$fl) * ifelse(d_i$surv[k.index], 1, NA)
[…]

### 2018-03-12 exploration.Rmd
added summary of IPM fit vs truth

@@
            xlab="Size (t)", ylab="Size (t+1)")
+
+IPM.fit.sum <- lam.df %>% 
+  summarise(mnLamDiff=mean(lambda.f-lambda),
+            mdLamDiff=median(lambda.f-lambda),
+            sdLamDiff=sd(lambda.f-lambda),
+            pctLamCorr=mean((lambda>1 & lambda.f>1) | (lambda<1 & lambda.f<1)),
+            mnNDiff=mean(N.S.f-N.S),
+            mdNDiff=median(N.S.f-N.S),
+            sdNDiff=sd(N.S.f-N.S),
+            pctNCorr=mean((N.S>1 & N.S.f>1) | (N.S<1 & N.S.f<1)),
+            mnSurvDiff=mean(Surv.S.f-Surv.S),
+            mdSurvDiff=median(Surv.S.f-Surv.S),
+            sdSurvDiff=sd(Surv.S.f-Surv.S),
+            pctSurvCorr=mean((Surv.S>1 & Surv.S.f>1) | (Surv.S<1 & Surv.S.f<1)),
+            mnRcrDiff=mean(Rcr.S.f-Rcr.S),
+            mdRcrDiff=median(Rcr.S.f-Rcr.S),
+            sdRcrDiff=sd(Rcr.S.f-Rcr.S),
+            pctRcrCorr=mean((Rcr.S>1 & Rcr.S.f>1) | (Rcr.S<1 & Rcr.S.f<1)))
 ```

### 2018-03-12 exploration.Rmd
corrected N>0 instead of N>1

@@
             sdNDiff=sd(N.S.f-N.S),
-            pctNCorr=mean((N.S>1 & N.S.f>1) | (N.S<1 & N.S.f<1)),
+            pctNCorr=mean((N.S>0 & N.S.f>0) | (N.S<0 & N.S.f<0)),
             mnSurvDiff=mean(Surv.S.f-Surv.S),
@@
             sdSurvDiff=sd(Surv.S.f-Surv.S),
-            pctSurvCorr=mean((Surv.S>1 & Surv.S.f>1) | (Surv.S<1 & Surv.S.f<1)),
+            pctSurvCorr=mean((Surv.S>0 & Surv.S.f>0) | (Surv.S<0 & Surv.S.f<0)),
             mnRcrDiff=mean(Rcr.S.f-Rcr.S),
@@
             sdRcrDiff=sd(Rcr.S.f-Rcr.S),
-            pctRcrCorr=mean((Rcr.S>1 & Rcr.S.f>1) | (Rcr.S<1 & Rcr.S.f<1)))
+            pctRcrCorr=mean((Rcr.S>0 & Rcr.S.f>0) | (Rcr.S<0 & Rcr.S.f<0)))
 ```

### 2018-03-12 exploration.Rmd
updated plots

@@
 ```{r sp_i}
-mean_abund <- 50
+mean_abund <- 200
 regPlots <- T
@@
 n <- 50  # ncells in IPM matrix
-tmax <- 10
+tmax <- 50
 cells <- 1:n_distinct(lc.df$CellID)
@@
 ```{r O_setup}
-O_n <- list(Mx=500, CA=20, IPM=20) # number of cells in sample
+O_n <- list(Mx=500, CA=20, IPM=10) # number of cells in sample
 O_yr <- list(Mx=tmax, CA=(tmax-10):tmax, IPM=tmax) # years to sample from
@@
 ########
-ggplot(lam.df, aes(x=lon, y=lat, fill=lambda.f-lambda)) + geom_tile() +
-  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
-ggplot(lam.df, aes(x=lon, y=lat, fill=lam.U.f-lam.U)) + geom_tile() +
+ggplot(lam.df, aes(x=lon, y=lat)) + 
+  geom_tile(aes(fill=lambda.f-lambda, colour=id.inbd %in% IPM.P)) +
+  scale_colour_manual(values=c(NA, "black")) +
   scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
 ggplot(lam.df, aes(x=lon, y=lat)) + 
-  geom_tile(aes(fill=(lambda>1 & lambda.f>1) | (lambda<1 & lambda.f<1)))
+  geom_tile(aes(fill=lam.U.f-lam.U, colour=id.inbd %in% IPM.P)) +
+  scale_colour_manual(values=c(NA, "black")) +
+  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
 ggplot(lam.df, aes(x=lon, y=lat)) + 
-  geom_tile(aes(fill=(N.S>1 & N.S.f>1) | (N.S<1 & N.S.f<1)))
+  geom_tile(aes(fill=(lambda>1 & lambda.f>1) | (lambda<1 & lambda.f<1),
+                colour=id.inbd %in% IPM.P)) +
+  scale_colour_manual(values=c(NA, "black")) + 
+  scale_fill_manual(values=c("red", "dodgerblue"))
 ggplot(lam.df, aes(x=lon, y=lat)) + 
-  geom_tile(aes(fill=(lam.U>1 & lam.U.f>1) | (lam.U<1 & lam.U.f<1)))
+  geom_tile(aes(fill=(N.S>1 & N.S.f>1) | (N.S<1 & N.S.f<1),
+                colour=id.inbd %in% IPM.P)) +
+  scale_colour_manual(values=c(NA, "black")) + 
+  scale_fill_manual(values=c("red", "dodgerblue"))
 ggplot(lam.df, aes(x=lon, y=lat)) + 
-  geom_tile(aes(fill=(N.U>1 & N.U.f>1) | (N.U<1 & N.U.f<1)))
-ggplot(lam.df, aes(x=lon, y=lat, fill=(N.S.f-N.S)/N.S)) + geom_tile() +
+  geom_tile(aes(fill=(lam.U>1 & lam.U.f>1) | (lam.U<1 & lam.U.f<1),
+                colour=id.inbd %in% IPM.P)) +
+  scale_colour_manual(values=c(NA, "black")) + 
+  scale_fill_manual(values=c("red", "dodgerblue"))
+ggplot(lam.df, aes(x=lon, y=lat)) + 
+  geom_tile(aes(fill=(N.U>1 & N.U.f>1) | (N.U<1 & N.U.f<1),
+                colour=id.inbd %in% IPM.P)) +
+  scale_colour_manual(values=c(NA, "black")) + 
+  scale_fill_manual(values=c("red", "dodgerblue"))
+ggplot(lam.df, aes(x=lon, y=lat)) + 
+  geom_tile(aes(fill=(N.S.f-N.S)/N.S, colour=id.inbd %in% IPM.P)) +
+  scale_colour_manual(values=c(NA, "black")) +
   scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
-ggplot(lam.df, aes(x=lon, y=lat, fill=(Surv.S.f-Surv.S)/Surv.S)) + geom_tile() +
+ggplot(lam.df, aes(x=lon, y=lat)) + 
+  geom_tile(aes( fill=(Surv.S.f-Surv.S)/Surv.S, colour=id.inbd %in% IPM.P)) +
+  scale_colour_manual(values=c(NA, "black")) +
   scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
-ggplot(lam.df, aes(x=lon, y=lat, fill=(Rcr.S.f-Rcr.S)/Rcr.S)) + geom_tile() +
+ggplot(lam.df, aes(x=lon, y=lat)) + 
+  geom_tile(aes(fill=(Rcr.S.f-Rcr.S)/Rcr.S, colour=id.inbd %in% IPM.P)) +
+  scale_colour_manual(values=c(NA, "black")) +
   scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
@@
             sdNDiff=sd(N.S.f-N.S),
-            pctNCorr=mean((N.S>0 & N.S.f>0) | (N.S<0 & N.S.f<0)),
+            pctNCorr=mean((N.S>0 & N.S.f>0) | (N.S==0 & N.S.f==0)),
             mnSurvDiff=mean(Surv.S.f-Surv.S),
@@
             sdSurvDiff=sd(Surv.S.f-Surv.S),
-            pctSurvCorr=mean((Surv.S>0 & Surv.S.f>0) | (Surv.S<0 & Surv.S.f<0)),
+            pctSurvCorr=mean((Surv.S>0 & Surv.S.f>0) | (Surv.S==0 & Surv.S.f==0)),
             mnRcrDiff=mean(Rcr.S.f-Rcr.S),
@@
             sdRcrDiff=sd(Rcr.S.f-Rcr.S),
-            pctRcrCorr=mean((Rcr.S>0 & Rcr.S.f>0) | (Rcr.S<0 & Rcr.S.f<0)))
+            pctRcrCorr=mean((Rcr.S>0 & Rcr.S.f>0) | (Rcr.S==0 & Rcr.S.f==0)))
+str(IPM.fit.sum)
 ```

### 2018-03-12 exploration.Rmd
adding lambda estimate from simulated data

@@
 lam.df %<>% 
-  add_column(nSeed=S$nSd[,tmax], D=S$D[,tmax], 
+  add_column(lam.S=S$N_sim[,tmax]/(S$N_sim[,tmax-1]+.01),
+             nSeed=S$nSd[,tmax], D=S$D[,tmax], 
              B0=S$B[,1], Btmax=S$B[,tmax+1], 
@@
   scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
+ggplot(lam.df, aes(x=lon, y=lat, fill=log(lam.S))) + geom_tile() +
+  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
 ggplot(lam.df, aes(x=lon, y=lat, fill=log(lam.U))) + geom_tile() +
   scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
-ggplot(lam.df, aes(x=lon, y=lat, fill=log(N.U)*(N.U>10))) + geom_tile() +
-  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=1)
 ggplot(lam.df, aes(x=lon, y=lat, fill=lambda>1)) + geom_tile()
-ggplot(lam.df, aes(x=lon, y=lat, fill=N.S>1)) + geom_tile()
+ggplot(lam.df, aes(x=lon, y=lat, fill=lam.S>1)) + geom_tile()
+ggplot(lam.df, aes(x=lon, y=lat, fill=Surv.S>0)) + geom_tile()
+ggplot(lam.df, aes(x=lon, y=lat, fill=N.S>0)) + geom_tile()
 ggplot(lam.df, aes(x=lon, y=lat, fill=lam.U>1)) + geom_tile()
@@
 ggplot(lam.df, aes(x=lon, y=lat, fill=N.U>1 & lam.U<1)) + geom_tile() 
-ggplot(lam.df, aes(x=lon, y=lat, fill=N.S>1 & lambda<1)) + geom_tile()
+ggplot(lam.df, aes(x=lon, y=lat, fill=Surv.S>0 & lambda<1)) + geom_tile()
 ggplot(lam.df, aes(x=lon, y=lat, fill=log(N.S))) + geom_tile() +
@@
 ggplot(lam.df, aes(x=lon, y=lat, fill=nSdStay > D)) + geom_tile()
-ggplot(lam.df, aes(x=lon, y=lat, fill=nSdStay < D & lambda < 1)) +
-  geom_tile()
 par(mfrow=c(2,2))
@@
 ########
-#---- Truth: use assigned slopes to fill IPM matrix
+#---- Truth: use estimated slopes to fill IPM matrix
 U.f <- fill_IPM_matrices(n.cell, tmax, n_i, n, z.rng, buffer=0.5, discrete=1,
@@
 
-#---- Realization: generate simulated data
+#---- Realization: use estimated slopes to generate simulated data
 S.f <- simulate_data(n.cell, tmax, n_i, z.rng, U.f$lo, U.f$hi, 
@@
 lam.df %<>% 
-  add_column(nSeed.f=S.f$nSd[,tmax], D.f=S.f$D[,tmax], 
+  add_column(lam.S.f=S.f$N_sim[,tmax]/(S.f$N_sim[,tmax-1]+.01),
+             nSeed.f=S.f$nSd[,tmax], D.f=S.f$D[,tmax], 
              B0.f=S.f$B[,1], Btmax.f=S.f$B[,tmax+1], 
@@
 ggplot(lam.df, aes(x=lon, y=lat)) + 
+  geom_tile(aes(fill=lam.S.f-lam.S, colour=id.inbd %in% IPM.P)) +
+  scale_colour_manual(values=c(NA, "black")) +
+  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
+ggplot(lam.df, aes(x=lon, y=lat)) + 
   geom_tile(aes(fill=lam.U.f-lam.U, colour=id.inbd %in% IPM.P)) +
@@
 ggplot(lam.df, aes(x=lon, y=lat)) + 
+  geom_tile(aes(fill=lam.S>1, colour=id.inbd %in% IPM.P)) +
+  scale_colour_manual(values=c(NA, "black"))
+ggplot(lam.df, aes(x=lon, y=lat)) + 
+  geom_tile(aes(fill=lam.S.f>1, colour=id.inbd %in% IPM.P)) +
+  scale_colour_manual(values=c(NA, "black"))
+ggplot(lam.df, aes(x=lon, y=lat)) + 
   geom_tile(aes(fill=(lambda>1 & lambda.f>1) | (lambda<1 & lambda.f<1),
                 colour=id.inbd %in% IPM.P)) +
-  scale_colour_manual(values=c(NA, "black")) + 
-  scale_fill_manual(values=c("red", "dodgerblue"))
+  scale_colour_manual(values=c(NA, "black"))
+ggplot(lam.df, aes(x=lon, y=lat)) + 
+  geom_tile(aes(fill=(lam.S>1 & lam.S.f>1) | (lam.S<1 & lam.S.f<1),
+                colour=id.inbd %in% IPM.P)) +
+  scale_colour_manual(values=c(NA, "black"))
 ggplot(lam.df, aes(x=lon, y=lat)) + 
@@
                 colour=id.inbd %in% IPM.P)) +
-  scale_colour_manual(values=c(NA, "black")) + 
-  scale_fill_manual(values=c("red", "dodgerblue"))
+  scale_colour_manual(values=c(NA, "black"))
 ggplot(lam.df, aes(x=lon, y=lat)) + 
@@
             pctLamCorr=mean((lambda>1 & lambda.f>1) | (lambda<1 & lambda.f<1)),
+            mnLamSDiff=mean(lam.S.f-lam.S, na.rm=T),
+            mdLamSDiff=median(lam.S.f-lam.S, na.rm=T),
+            sdLamSDiff=sd(lam.S.f-lam.S, na.rm=T),
+            pctLamSCorr=mean((lam.S>1 & lam.S.f>1) | (lam.S<1 & lam.S.f<1)),
             mnNDiff=mean(N.S.f-N.S),

### 2018-03-13 code/fn_IPM.R
vector instead of matrix

@@ -163,9 +163,9 @@ fill_IPM_matrices <- function(n.cell, tmax, n_i, n, z.rng, buffer, discrete,
   
   # storage objects
   IPMs <- Ps <- Fs <- Fb <- array(0, dim=c(n+1, n+1, n.cell))
-  Nt <- array(dim=c(n+1, n+1, n.cell, tmax))
+  Nt <- array(dim=c(n+1, n.cell, tmax+1))
   sdd.j <- vector("list", length=n.cell)
-  lam.t <- p_est.t <- matrix(nrow=n.cell, ncol=tmax-1)
+  lam.t <- p_est.t <- matrix(nrow=n.cell, ncol=tmax)
   
   # IPM matrix setup
   L <- setup_IPM_matrix(n, z.rng, buffer)
@@ -189,13 +189,13 @@ fill_IPM_matrices <- function(n.cell, tmax, n_i, n, z.rng, buffer, discrete,
       Reduce(`+`, map2(sdd.j[[i]][,3], p$p_emig*p.ij, ~(Fb[z.i,z.i,.x] * .y)))
     Fs[1,z.i,i] <- Fs[1,z.i,i] + Fb[1,z.i,sdd.j[[i]][,3]] %*% (p$p_emig*p.ij)
     if(p$NDD) {
-      Nt[,,i,1] <- rpois((n+1)*(n+1), n_i[i])
+      Nt[,i,1] <- rpois((n+1), n_i[i]/(n+1))
       F.i <- F.t <- Fs[,,i]/p$p_est  # F.s[,,i] was already multiplied by p$p_est
-      for(k in 1:(tmax-1)) {
-        p_est.t[i,k] <- min(p$NDD_n/(sum(F.i[z.i,]*Nt[z.i,,i,k])), p$p_est)
+      for(k in 1:tmax) {
+        p_est.t[i,k] <- min(p$NDD_n/(sum(F.i[z.i,]*Nt[z.i,i,k])), p$p_est)
         F.t[z.i,] <- F.i[z.i,] * p_est.t[i,k]
         IPM.t <- Ps[,,i] + F.t
-        Nt[,,i,k+1] <- IPM.t %*% Nt[,,i,k]
+        Nt[,i,k+1] <- round(IPM.t %*% Nt[,i,k])
         lam.t[i,k] <- Re(eigen(IPM.t)$values[1])
       }
     }

### 2018-03-13 exploration.Rmd
tying NDD_n to mean_abund

@@
                     NDD=T, # negative density dependence: p_est=p_est/N
-                    NDD_n=100, # mean number of recruits if NDD
+                    NDD_n=mean_abund/2, # mean number of recruits if NDD
                     rcr_SB=0.5, # p(recruit from seedbank)

### 2018-03-13 exploration.Rmd
Nt as vector instead of matrix

@@
 #---- Truth: use assigned slopes to fill IPM matrix
-U <- fill_IPM_matrices(n.cell, tmax, n_i, n, z.rng, buffer=0.5, discrete=1,
-                       p, n_z, n_x, X, sdd.pr, env.df)
+U <- fill_IPM_matrices(n.cell, tmax, n_i, n, z.rng, buffer=0.75, discrete=1,
+                       p, n_z, n_x, X, sdd.pr, env.df, verbose=T)
 
@@
 S <- simulate_data(n.cell, tmax, n_i, z.rng, U$lo, U$hi, 
-                   p, X, n_z, sdd, U$sdd.j)
+                   p, X, n_z, sdd, U$sdd.j, verbose=T)
 
@@
   mutate(nSdStay=nSeed*(1-p$p_emig), nSdLeave=nSeed*p$p_emig) %>%
-  add_column(N.U=apply(U$Nt[,,,tmax],3,sum), lam.U=U$lam.t[,tmax-1])
+  add_column(N.U=apply(U$Nt[-1,,tmax+1],2,sum), lam.U=U$lam.t[,tmax])
 
@@
 ggplot(lam.df, aes(x=lon, y=lat, fill=lam.S>1)) + geom_tile()
+ggplot(lam.df, aes(x=lon, y=lat, fill=lam.U>1)) + geom_tile()
 ggplot(lam.df, aes(x=lon, y=lat, fill=Surv.S>0)) + geom_tile()
@@
 ggplot(lam.df, aes(x=lon, y=lat, fill=lam.U>1)) + geom_tile()
-ggplot(lam.df, aes(x=lon, y=lat, fill=N.U>1)) + geom_tile() 
-ggplot(lam.df, aes(x=lon, y=lat, fill=N.U>1 & lam.U<1)) + geom_tile() 
+ggplot(lam.df, aes(x=lon, y=lat, fill=N.U>0)) + geom_tile() 
+ggplot(lam.df, aes(x=lon, y=lat, fill=N.U>0 & lam.U<1)) + geom_tile() 
 ggplot(lam.df, aes(x=lon, y=lat, fill=Surv.S>0 & lambda<1)) + geom_tile()
+ggplot(lam.df, aes(x=lon, y=lat, fill=log(N.U))) + geom_tile() +
+  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
 ggplot(lam.df, aes(x=lon, y=lat, fill=log(N.S))) + geom_tile() +
@@
 matplot(t(log(U$lam.t)), type="l", lty=1, col=rgb(0,0,0,0.1))
-matplot(t(log(apply(U$Nt, 3:4, sum))), type="l", lty=1, col=rgb(0,0,0,0.1))
-matplot(t(S$N_sim[,2:tmax]/S$N_sim[,1:(tmax-1)]), 
+matplot(t(log(apply(U$Nt[-1,,], 2:3, sum))), type="l", lty=1, col=rgb(0,0,0,0.1))
+matplot(t(log(S$N_sim[,2:tmax]/(S$N_sim[,1:(tmax-1)]))), 
         type="l", lty=1, col=rgb(0,0,0,.1))
@@
 ```{r O_setup}
-O_n <- list(Mx=500, CA=20, IPM=10) # number of cells in sample
+O_n <- list(Mx=500, CA=20, IPM=20) # number of cells in sample
 O_yr <- list(Mx=tmax, CA=(tmax-10):tmax, IPM=tmax) # years to sample from
@@
 U.f <- fill_IPM_matrices(n.cell, tmax, n_i, n, z.rng, buffer=0.5, discrete=1,
-                         p.ipm, n_z, n_x, X, sdd, env.df)
+                         p.ipm, n_z, n_x, X, sdd, env.df, verbose=T)
 
@@
 S.f <- simulate_data(n.cell, tmax, n_i, z.rng, U.f$lo, U.f$hi, 
-                     p.ipm, X, n_z, sdd, U.f$sdd.j)
+                     p.ipm, X, n_z, sdd, U.f$sdd.j, verbose=T)
 
@@
   mutate(nSdStay.f=nSeed.f*(1-p.ipm$p_emig), nSdLeave.f=nSeed.f*p.ipm$p_emig) %>%
-  add_column(N.U.f=apply(U.f$Nt[,,,tmax],3,sum), lam.U.f=U.f$lam.t[,tmax-1])
+  add_column(N.U.f=apply(U.f$Nt[,,tmax],2,sum), lam.U.f=U.f$lam.t[,tmax-1])
 

### 2018-03-13 exploration.Rmd
copied IPM image plot axes outside if{ipmPlots}

@@
 }
+par(mfrow=c(2,2), mar=c(4,4,2,1))
+x <- c(0, U$y)
 image.plot(x, x, t(apply(U$Ps, 1:2, mean)), main="P matrix: mean",
@@
 if(ipmPlots) {
-  par(mfrow=c(2,2), mar=c(4,4,2,1))
-  x <- c(0, U.f$y)
   for(i in sample.cells) {
@@
 }
+par(mfrow=c(2,2), mar=c(4,4,2,1))
+x <- c(0, U.f$y)
 image.plot(x, x, t(apply(U.f$Ps, 1:2, mean)), main="P matrix: mean",

### 2018-03-13 exploration.Rmd
added more covariate plots

@@
 ggplot(lam.df, aes(x=lon, y=lat, fill=nSdStay > D)) + geom_tile()
-par(mfrow=c(2,2))
+par(mfrow=c(3,3))
 plot(log(lambda) ~ temp, data=lam.df); abline(h=0, col="red", lty=2)
@@
 plot(log(lambda) ~ pOpn, data=lam.df); abline(h=0, col="red", lty=2)
+plot(log(lambda) ~ pOth, data=lam.df); abline(h=0, col="red", lty=2)
+plot(log(lambda) ~ pDec, data=lam.df); abline(h=0, col="red", lty=2)
+plot(log(lambda) ~ pEvg, data=lam.df); abline(h=0, col="red", lty=2)
+plot(log(lambda) ~ pMxd, data=lam.df); abline(h=0, col="red", lty=2)
 plot(density(log(lam.df$lambda))); abline(v=0, col="red", lty=2)
-matplot(t(log(U$lam.t)), type="l", lty=1, col=rgb(0,0,0,0.1))
-matplot(t(log(apply(U$Nt[-1,,], 2:3, sum))), type="l", lty=1, col=rgb(0,0,0,0.1))
-matplot(t(log(S$N_sim[,2:tmax]/(S$N_sim[,1:(tmax-1)]))), 
-        type="l", lty=1, col=rgb(0,0,0,.1))
-matplot(t(log(S$N_sim+1)), type="l", lty=1, col=rgb(0,0,0,0.1))
+plot(density(log(lam.df$lam.U))); abline(v=0, col="red", lty=2)
+# matplot(t(log(U$lam.t)), type="l", lty=1, col=rgb(0,0,0,0.1))
+# matplot(t(log(apply(U$Nt[-1,,], 2:3, sum))), type="l", lty=1, col=rgb(0,0,0,0.1))
+# matplot(t(log(S$N_sim[,2:tmax]/(S$N_sim[,1:(tmax-1)]))), 
+#         type="l", lty=1, col=rgb(0,0,0,.1))
+# matplot(t(log(S$N_sim+1)), type="l", lty=1, col=rgb(0,0,0,0.1))
 #matplot(t(log(nSd*(1-p$p_emig)+D)), type="l", lty=1, col=rgb(0,0,0,0.1))
@@
   scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
-par(mfrow=c(2,2))
+par(mfrow=c(3,3))
 plot((lambda.f-lambda) ~ temp, data=lam.df); abline(h=0, col="red", lty=2)
@@
 plot((lambda.f-lambda) ~ pOpn, data=lam.df); abline(h=0, col="red", lty=2)
-plot(density((lam.df$lambda.f-lam.df$lambda))); abline(v=0, col="red", lty=2)
-matplot(t(log(U.f$lam.t)), type="l", lty=1, col=rgb(0,0,0,0.1))
-matplot(t(log(apply(U.f$Nt, 3:4, sum))), type="l", lty=1, col=rgb(0,0,0,0.1))
-matplot(t(S.f$N_sim[,2:tmax]/S$N_sim[,1:(tmax-1)]), 
-        type="l", lty=1, col=rgb(0,0,0,.1))
-matplot(t(log(S.f$N_sim+1)), type="l", lty=1, col=rgb(0,0,0,0.1))
+plot((lambda.f-lambda) ~ pOth, data=lam.df); abline(h=0, col="red", lty=2)
+plot((lambda.f-lambda) ~ pDec, data=lam.df); abline(h=0, col="red", lty=2)
+plot((lambda.f-lambda) ~ pEvg, data=lam.df); abline(h=0, col="red", lty=2)
+plot((lambda.f-lambda) ~ pMxd, data=lam.df); abline(h=0, col="red", lty=2)
+plot(density(lambda.f-lambda)); abline(v=0, col="red", lty=2)
+plot(density(lam.U.f-lam.U)); abline(v=0, col="red", lty=2)
+# matplot(t(log(U.f$lam.t)), type="l", lty=1, col=rgb(0,0,0,0.1))
+# matplot(t(log(apply(U.f$Nt, 3:4, sum))), type="l", lty=1, col=rgb(0,0,0,0.1))
+# matplot(t(S.f$N_sim[,2:tmax]/S$N_sim[,1:(tmax-1)]), 
+#         type="l", lty=1, col=rgb(0,0,0,.1))
+# matplot(t(log(S.f$N_sim+1)), type="l", lty=1, col=rgb(0,0,0,0.1))
 

### 2018-03-13 exploration.Rmd
redundant plot

--- 2018-03-13 exploration.Rmd
@@
 ggplot(lam.df, aes(x=lon, y=lat, fill=N.S>0)) + geom_tile()
-ggplot(lam.df, aes(x=lon, y=lat, fill=lam.U>1)) + geom_tile()
 ggplot(lam.df, aes(x=lon, y=lat, fill=N.U>0)) + geom_tile() 

### 2018-03-13 exploration.Rmd
copied

@@
 if(ipmPlots) {
+  par(mfrow=c(2,2), mar=c(4,4,2,1))
+  x <- c(0, U.f$y)
   for(i in sample.cells) {

### 2018-03-13 exploration.Rmd
needed dataframe

@@
 plot((lambda.f-lambda) ~ pMxd, data=lam.df); abline(h=0, col="red", lty=2)
-plot(density(lambda.f-lambda)); abline(v=0, col="red", lty=2)
-plot(density(lam.U.f-lam.U)); abline(v=0, col="red", lty=2)
+plot(density(lam.df$lambda.f-lam.df$lambda)); abline(v=0, col="red", lty=2)
+plot(density(lam.df$lam.U.f-lam.df$lam.U)); abline(v=0, col="red", lty=2)
 # matplot(t(log(U.f$lam.t)), type="l", lty=1, col=rgb(0,0,0,0.1))

### 2018-03-14 code/fn_IPM.R
vectorized operations

@@ -48,8 +48,7 @@ calc_surv <- function(z.v, p, n_sz, X.s) {
 ##   z' ~ N(size + environment, sd)
 calc_grow <- function(z1, z.v, p, n_gz, X.g) {
   z <- z_pow(z.v, n_gz)
-  g <- dnorm(z1, mean=z %*% p$g_z + c(t(X.g) %*% p$g_x), 
-             sd=p$g_sig)
+  g <- dnorm(z1, mean=z %*% p$g_z + c(t(X.g) %*% p$g_x), sd=p$g_sig)
   return(g)
 }
 
@@ -157,9 +156,10 @@ fill_F <- function(h, y, z.i, n, p, n_z, n_x, X_fl, X_seed) {
 
 ##-- Fill all IPM objects
 ##
-fill_IPM_matrices <- function(n.cell, tmax, n_i, n, z.rng, buffer, discrete,
+fill_IPM_matrices <- function(n.cell, tmax, n0, n, z.rng, buffer, discrete,
                               p, n_z, n_x, X, sdd, env.df, verbose=FALSE) {
   require(tidyverse)
+  i <- 1:n.cell
   
   # storage objects
   IPMs <- Ps <- Fs <- Fb <- array(0, dim=c(n+1, n+1, n.cell))
@@ -172,34 +172,38 @@ fill_IPM_matrices <- function(n.cell, tmax, n_i, n, z.rng, buffer, discrete,
   z.i <- (1:n) + discrete  # continuous stage matrix indices
 
   ## local growth
-  for(i in 1:n.cell) {
-    Ps[,,i] <- fill_P(L$h, L$y, z.i, n, p, n_z, n_x, X$s[i,], X$g[i,])
-    Fb[,,i] <- fill_F(L$h, L$y, z.i, n, p, n_z, n_x, X$fl[i,], X$seed[i,])
-    Fs[z.i,z.i,i] <- (1-p$p_emig) * Fb[z.i,z.i,i]
-    Fs[1,z.i,i] <- (1-p$p_emig) * Fb[1,z.i,i]
-    Fs[z.i,1,i] <- Fb[z.i,1,i]
-  }
+  Ps <- vapply(i, function(x) fill_P(L$h, L$y, z.i, n, p, n_z, n_x, 
+                                            X$s[x,], X$g[x,]), Ps[,,1])
[…]

### 2018-03-14 code/fn_sim.R
vectorized operations

@@ -4,44 +4,50 @@
 
 ##--- simulate expected values
 ## 
-sim_expected <- function(d_i, d.k, e.k, E, p, X, n_z) {
-  size.mx <- with(d_i[d.k,], cbind(1, size, size^2, size^3))
-  E$s[e.k] <- antilogit(size.mx[,1:n_z$s] %*% p$s_z + c(X$s %*% p$s_x))
-  E$g[e.k] <- size.mx[,1:n_z$g] %*% p$g_z + c(X$g %*% p$g_x)
-  E$fl[e.k] <- antilogit(size.mx[,1:n_z$fl] %*% p$fl_z + c(X$fl %*% p$fl_x))
-  E$seed[e.k] <- exp(size.mx[,1:n_z$seed] %*% p$seed_z + c(X$seed %*% p$seed_x))
+sim_expected <- function(k, z.i, e.k, E, p, X, n_z) {
+  if(length(z.i) > 0) {
+    z.mx <- cbind(1, z.i, z.i^2, z.i^3)
+    E$yr[e.k] <- k
+    E$size[e.k] <- z.i
+    E$s[e.k] <- antilogit(z.mx[,1:n_z$s] %*% p$s_z + c(X$s %*% p$s_x))
+    E$g[e.k] <- z.mx[,1:n_z$g] %*% p$g_z + c(X$g %*% p$g_x)
+    E$fl[e.k] <- antilogit(z.mx[,1:n_z$fl] %*% p$fl_z + c(X$fl %*% p$fl_x))
+    E$seed[e.k] <- exp(z.mx[,1:n_z$seed] %*% p$seed_z + c(X$seed %*% p$seed_x))
+  }
   return(E)
 }
 
 
 ##--- simulate realized values
 ##
-sim_realized <- function(d_i, d.k, e.k, E, p, lo, hi) {
+sim_realized <- function(k, z.i, d_i, d.k, e.k, E, p, lo, hi) {
   n_t <- length(d.k)
-  d_i$surv[d.k] <- rbinom(n_t, 1, E$s[e.k])
-  d_i$sizeNext[d.k] <- rnorm(n_t, E$g[e.k], p$g_sig) * 
-    ifelse(d_i$surv[d.k], 1, NA)
-  d_i$sizeNext[d.k] <- ifelse(d_i$sizeNext[d.k] < lo, 
-                                  lo, d_i$sizeNext[d.k])
[…]

### 2018-03-14 exploration.Rmd
one mean n0 instead of randomizing across cells

@@
 ```{r sp_i}
-mean_abund <- 200
+n0 <- 100
 regPlots <- T
@@
 n <- 50  # ncells in IPM matrix
-tmax <- 50
+tmax <- 30
 cells <- 1:n_distinct(lc.df$CellID)
 n.cell <- length(cells)
-n_i <- rpois(n.cell, mean_abund)
 sp_i <- list(p=list(s_z=c(-8, 2.1, -.09), # b1 + b2*z + b3*z^2
@@
                     NDD=T, # negative density dependence: p_est=p_est/N
-                    NDD_n=mean_abund/2, # mean number of recruits if NDD
+                    NDD_n=n0/2, # mean number of recruits if NDD
                     rcr_SB=0.5, # p(recruit from seedbank)
@@
 #---- Truth: use assigned slopes to fill IPM matrix
-U <- fill_IPM_matrices(n.cell, tmax, n_i, n, z.rng, buffer=0.75, discrete=1,
+U <- fill_IPM_matrices(n.cell, tmax, n0, n, z.rng, buffer=0.75, discrete=1,
                        p, n_z, n_x, X, sdd.pr, env.df, verbose=T)
@@
 #---- Realization: generate simulated data
-S <- simulate_data(n.cell, tmax, n_i, z.rng, U$lo, U$hi, 
-                   p, X, n_z, sdd, U$sdd.j, verbose=T)
+S <- simulate_data(n.cell, tmax, n0, z.rng, U$lo, U$hi,
+                   p, X, n_z, sdd.pr, U$sdd.j, verbose=T)
 

### 2018-03-14 exploration.Rmd
sample set number of cells for MaxEnt

@@
 library(dismo)
-Mx.P <- sample(P.i, O_n$Mx, replace=T)
+Mx.P <- sample(P.i, O_n$Mx, replace=F)
 pres <- 1:nrow(lam.df) %in% Mx.P

### 2018-03-14 exploration.Rmd
n0 instead of n_i

@@
 #---- Truth: use estimated slopes to fill IPM matrix
-U.f <- fill_IPM_matrices(n.cell, tmax, n_i, n, z.rng, buffer=0.5, discrete=1,
+U.f <- fill_IPM_matrices(n.cell, tmax, n0, n, z.rng, buffer=0.5, discrete=1,
                          p.ipm, n_z, n_x, X, sdd, env.df, verbose=T)
@@
 #---- Realization: use estimated slopes to generate simulated data
-S.f <- simulate_data(n.cell, tmax, n_i, z.rng, U.f$lo, U.f$hi, 
+S.f <- simulate_data(n.cell, tmax, n0, z.rng, U.f$lo, U.f$hi, 
                      p.ipm, X, n_z, sdd, U.f$sdd.j, verbose=T)

### 2018-03-14 exploration.Rmd
making sampled cells more clear

@@
 ggplot(lam.df, aes(x=lon, y=lat)) + 
-  geom_tile(aes(fill=lambda.f-lambda, colour=id.inbd %in% IPM.P)) +
+  geom_tile(aes(fill=lambda.f-lambda, colour=id.inbd %in% IPM.P), size=.5) +
   scale_colour_manual(values=c(NA, "black")) +
@@
 ggplot(lam.df, aes(x=lon, y=lat)) + 
-  geom_tile(aes(fill=lam.S.f-lam.S, colour=id.inbd %in% IPM.P)) +
+  geom_tile(aes(fill=lam.S.f-lam.S, colour=id.inbd %in% IPM.P), size=.5) +
   scale_colour_manual(values=c(NA, "black")) +
@@
 ggplot(lam.df, aes(x=lon, y=lat)) + 
-  geom_tile(aes(fill=lam.U.f-lam.U, colour=id.inbd %in% IPM.P)) +
+  geom_tile(aes(fill=lam.U.f-lam.U, colour=id.inbd %in% IPM.P), size=.5) +
   scale_colour_manual(values=c(NA, "black")) +
@@
 ggplot(lam.df, aes(x=lon, y=lat)) + 
-  geom_tile(aes(fill=lam.S>1, colour=id.inbd %in% IPM.P)) +
+  geom_tile(aes(fill=lam.S>1, colour=id.inbd %in% IPM.P), size=.5) +
   scale_colour_manual(values=c(NA, "black"))
 ggplot(lam.df, aes(x=lon, y=lat)) + 
-  geom_tile(aes(fill=lam.S.f>1, colour=id.inbd %in% IPM.P)) +
+  geom_tile(aes(fill=lam.S.f>1, colour=id.inbd %in% IPM.P), size=.5) +
   scale_colour_manual(values=c(NA, "black"))
@@
   geom_tile(aes(fill=(lambda>1 & lambda.f>1) | (lambda<1 & lambda.f<1),
-                colour=id.inbd %in% IPM.P)) +
+                colour=id.inbd %in% IPM.P), size=.5) +
   scale_colour_manual(values=c(NA, "black"))
@@
   geom_tile(aes(fill=(lam.S>1 & lam.S.f>1) | (lam.S<1 & lam.S.f<1),
-                colour=id.inbd %in% IPM.P)) +
+                colour=id.inbd %in% IPM.P), size=.5) +
   scale_colour_manual(values=c(NA, "black"))
@@
   geom_tile(aes(fill=(N.S>1 & N.S.f>1) | (N.S<1 & N.S.f<1),
-                colour=id.inbd %in% IPM.P)) +
+                colour=id.inbd %in% IPM.P), size=.5) +
   scale_colour_manual(values=c(NA, "black"))
@@
   geom_tile(aes(fill=(lam.U>1 & lam.U.f>1) | (lam.U<1 & lam.U.f<1),
-                colour=id.inbd %in% IPM.P)) +
+                colour=id.inbd %in% IPM.P), size=.5) +
   scale_colour_manual(values=c(NA, "black"))
@@
   geom_tile(aes(fill=(N.U>1 & N.U.f>1) | (N.U<1 & N.U.f<1),
-                colour=id.inbd %in% IPM.P)) +
+                colour=id.inbd %in% IPM.P), size=.5) +
   scale_colour_manual(values=c(NA, "black"))
 ggplot(lam.df, aes(x=lon, y=lat)) + 
-  geom_tile(aes(fill=(N.S.f-N.S)/N.S, colour=id.inbd %in% IPM.P)) +
+  geom_tile(aes(fill=(N.S.f-N.S)/N.S, colour=id.inbd %in% IPM.P), size=.5) +
   scale_colour_manual(values=c(NA, "black")) +
@@
 ggplot(lam.df, aes(x=lon, y=lat)) + 
-  geom_tile(aes( fill=(Surv.S.f-Surv.S)/Surv.S, colour=id.inbd %in% IPM.P)) +
+  geom_tile(aes(fill=(Surv.S.f-Surv.S)/Surv.S, 
+                colour=id.inbd %in% IPM.P), size=.5) +
   scale_colour_manual(values=c(NA, "black")) +
@@
 ggplot(lam.df, aes(x=lon, y=lat)) + 
-  geom_tile(aes(fill=(Rcr.S.f-Rcr.S)/Rcr.S, colour=id.inbd %in% IPM.P)) +
+  geom_tile(aes(fill=(Rcr.S.f-Rcr.S)/Rcr.S, 
+                colour=id.inbd %in% IPM.P), size=.5) +
   scale_colour_manual(values=c(NA, "black")) +

### 2018-03-14 exploration.Rmd
redundant object

@@
 p <- sp_i$p
-sdd <- array(0, dim=dim(sdd.pr) + c(0,0,1,0)) # add layer for nEmig
-sdd[,,1:2,] <- sdd.pr
 n_z <- list(s=length(p$s_z), g=length(p$g_z),

### 2018-03-14 code/fn_sim.R
fixed indexing bug

@@ -26,7 +26,7 @@ sim_realized <- function(k, z.i, d_i, d.k, e.k, E, p, lo, hi) {
     d_i$yr[d.k] <- k
     d_i$size[d.k] <- z.i
     d_i$surv[d.k] <- rbinom(n_t, 1, E$s[e.k])
-    surv.ik <- ifelse(surv.ik, 1, NA)
+    surv.ik <- ifelse(d_i$surv[d.k], 1, NA)
     d_i$sizeNext[d.k] <- rnorm(n_t, E$g[e.k], p$g_sig) * surv.ik
     d_i$sizeNext[d.k] <- pmax(d_i$sizeNext[d.k], lo)
     d_i$sizeNext[d.k] <- pmin(d_i$sizeNext[d.k], hi)
@@ -80,8 +80,8 @@ simulate_data <- function(n.cell, tmax, n0, z.rng, lo, hi,
   for(k in 1:tmax) {
     ## local growth
     if(k>1) z.k <- map(d, ~.$sizeNext[!is.na(.$sizeNext) & .$yr == k-1])
-    d.k <- lapply(i, function(x) seq_along(z.k[[x]]) + nrow(d[[x]]))
-    e.k <- lapply(i, function(x) seq_along(z.k[[x]]) + nrow(E[[x]]))
+    d.k <- lapply(i, function(x) seq_along(z.k[[x]]) + length(d[[x]]$yr))
+    e.k <- lapply(i, function(x) seq_along(z.k[[x]]) + length(E[[x]]$yr))
     E <- lapply(i, function(x) sim_expected(k, z.k[[x]], e.k[[x]], E[[x]], 
                                             p, X_map[[x]], n_z))
     d <- lapply(i, function(x) sim_realized(k, z.k[[x]], d[[x]], d.k[[x]], 

### 2018-03-14 exploration.Rmd
set up for averaging multiple simulations from same O sample; eventually will need to take multiple samples

@@
   xx <- tibble(size=seq(U$lo, U$hi, length.out=100), size2=size^2, size3=size^3)
-  E.t <- xx %>% add_column(s=NA, g=NA, fl=NA, seed=NA)
+  E.t <- xx %>% add_column(yr=NA, s=NA, g=NA, fl=NA, seed=NA)
   xxx <- seq(0, U$hi*.5, length.out=100)
   for(i in sample.cells) {
-    z.t <- S$d[[i]][!is.na(S$d[[i]]$size) & S$d[[i]]$yr==tmax,]
-    E.t <- sim_expected(E.t, 1:100, 1:100, E.t, p, map(X, ~.[i,]), n_z)
+    z.t <- data.frame(S$d[[i]]) %>% filter(!is.na(size) & yr==tmax)
+    pt.col <- rgb(0,0,0,1/(nrow(z.t)^.25))
+    E.t <- sim_expected(1, xx$size, 1:100, E.t, p, map(X, ~.[i,]), n_z)
     main.i <- paste("Cell", i)
     w1 <- Re(eigen(U$IPMs[,,i])$vectors[,1])[-1]; stable.dist1 <- w1/sum(w1)
-    plot(surv~size, data=z.t, ylim=c(0,1), pch=20, col=rgb(0,0,0,0.2),
+    plot(surv~size, data=z.t, ylim=c(0,1), pch=20, col=pt.col,
          main=paste("Surv:", main.i), xlab="Size (t)", ylab="P(surv to t+1)")
       lines(E.t$size, E.t$s, col=2)
-    plot(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,0.2), ylim=c(U$lo,U$hi),
+    plot(sizeNext~size, data=z.t, pch=20, col=pt.col, ylim=c(U$lo,U$hi),
          main=paste("Growth:", main.i), xlab="Size (t)", ylab="Size (t+1)")
@@
       abline(a=0, b=1, lty=2)
-    plot(fl~size, data=z.t, pch=20, ylim=c(0,1), col=rgb(0,0,0,0.2),
+    plot(fl~size, data=z.t, pch=20, ylim=c(0,1), col=pt.col,
          main=paste("P(Flower):", main.i), xlab="Size (t)", ylab="P(flower)")
       lines(E.t$size, E.t$fl, col=2)
-    plot(seed~size, data=z.t, pch=23-z.t$fl*3, col=rgb(0,0,0,0.2), 
+    plot(seed~size, data=z.t, pch=23-z.t$fl*3, col=pt.col, 
          ylim=range(c(E.t$seed, z.t$seed), na.rm=TRUE),
@@
   for(i in sample.cells) {
-   z.t <- S$d[[i]][!is.na(S$d[[i]]$size) & S$d[[i]]$yr==tmax,]
+   z.t <- data.frame(S$d[[i]]) %>% filter(!is.na(size) & yr==tmax)
+   pt.col <- rgb(1,1,1,1/(nrow(z.t)^.25))
    main.i <- paste("Cell", i)
    z.t$y <- U$y[findInterval(z.t$size, U$y, all.inside=T)]
-   totSeeds <- group_by(z.t, y) %>% summarise(seed=sum(seed, na.rm=T))
+   totSd <- group_by(z.t, y) %>% summarise(seed=sum(seed, na.rm=T))
    image.plot(x, x, t(U$Ps[,,i]), main=paste("P matrix:", main.i), 
               xlab="Size (t)", ylab="Size (t+1)")
-    points(sizeNext~size, data=z.t, col=rgb(1,1,1,1/(nrow(z.t)^.25)))
+    points(sizeNext~size, data=z.t, col=pt.col, pch=20)
     abline(a=0, b=1, col="white", lty=2)
@@
               xlab="Size (t)", ylab="Size (t+1)")
-    points(totSeeds$y, totSeeds$seed/max(totSeeds$seed)*U$hi*.9, col="white")
+    points(totSd$y, totSd$seed/max(totSd$seed)*U$hi*.9, col="white", pch=20)
    image.plot(x, x, t(sens(U$IPMs[,,i])), main=paste("Sensitivity:", main.i), 
@@
               xlab="Size (t)", ylab="Size (t+1)")
-   points(sizeNext~size, data=z.t, col=rgb(1,1,1,1/(nrow(z.t)^.25)))
+   points(sizeNext~size, data=z.t, col=pt.col, pch=20)
   }
@@
 x <- c(0, U$y)
+S_d <- map_dfr(S$d, data.frame) %>% sample_n(1000)
+S_d$y <- U$y[findInterval(S_d$size, U$y, all.inside=T)]
+totSd <- group_by(S_d, y) %>% summarise(seed=sum(seed, na.rm=T))
+pt.col <- rgb(1,1,1,1/(nrow(S_d)^.25))
 image.plot(x, x, t(apply(U$Ps, 1:2, mean)), main="P matrix: mean",
            xlab="Size (t)", ylab="Size (t+1)")
+points(sizeNext~size, data=S_d, col=pt.col, pch=20)
 image.plot(x, x, t(apply(U$Fs, 1:2, mean)), main="F matrix: mean",
            xlab="Size (t)", ylab="Size (t+1)")
+points(totSd$y, totSd$seed/max(totSd$seed)*U$hi*.9, col="white", pch=20)
 image.plot(x, x, sens(t(apply(U$IPMs, 1:2, mean))), main="Sensitivity: mean",
@@
            xlab="Size (t)", ylab="Size (t+1)")
+points(sizeNext~size, data=S_d, col=pt.col, pch=20)
+
 ```
@@
 P.i <- which(lam.df$Surv.S > 0) # presences: surviving individuals > seedlings
+n_sim <- 20
 ```
@@
 IPM.d <- vector("list", O_n$IPM)
+# sample individuals from each cell rather than full census -- proportion?
 for(i in IPM.P) {
-  IPM.d[[i]] <- S$d[[i]] %>% filter(yr %in% O_yr$IPM) %>%
+  IPM.d[[i]] <- data.frame(S$d[[i]]) %>% filter(yr %in% O_yr$IPM) %>%
     add_column(id.inbd=i) %>%
@@
 U.f <- fill_IPM_matrices(n.cell, tmax, n0, n, z.rng, buffer=0.5, discrete=1,
-                         p.ipm, n_z, n_x, X, sdd, env.df, verbose=T)
+                         p.ipm, n_z, n_x, X, sdd.pr, env.df, verbose=T)
 
 #---- Realization: use estimated slopes to generate simulated data
-S.f <- simulate_data(n.cell, tmax, n0, z.rng, U.f$lo, U.f$hi, 
-                     p.ipm, X, n_z, sdd, U.f$sdd.j, verbose=T)
+S.i <- vector("list", n_sim)
+for(s in 1:n_sim) {
+  S.i[[s]] <- simulate_data(n.cell, tmax, n0, z.rng, U.f$lo, U.f$hi, 
+                     p.ipm, X, n_z, sdd.pr, U.f$sdd.j, verbose=T)
+  cat("||-- Finished simulation", s, "\n||--\n")
+}
+S.sum <- list(B=map(S.i, ~.$B),
+              nSd=map(S.i, ~.$nSd),
+              D=map(S.i, ~.$D),
+              p_est=map(S.i, ~.$p_est.i),
+              N_sim=map(S.i, ~.$N_sim),
+              N_tot=map(S.i, ~map_dbl(.$d, 
+                                    ~sum(!is.na(.$sizeNext[.$yr==tmax])))),
+              N_surv=map(S.i, ~map_dbl(.$d, 
+                                    ~sum(.$surv[.$yr==tmax], na.rm=T))),
+              N_rcr=map(S.i, ~map_dbl(.$d, 
+                                    ~sum(is.na(.$size[.$yr==tmax]))))) %>% 
+  map(simplify2array)
+rm(S.i)
+
+S.f <- list(B.mn=apply(S.sum$B, c(1,2), mean), 
+            B.sd=apply(S.sum$B, c(1,2), sd), 
+            nSd.mn=apply(S.sum$nSd, c(1,2), mean), 
+            nSd.sd=apply(S.sum$nSd, c(1,2), sd), 
+            D.mn=apply(S.sum$D, c(1,2), mean), 
+            D.sd=apply(S.sum$D, c(1,2), sd), 
+            p_est.mn=apply(S.sum$p_est, c(1,2), mean), 
+            p_est.sd=apply(S.sum$p_est, c(1,2), sd), 
+            N_sim.mn=apply(S.sum$N_sim, c(1,2), mean), 
+            N_sim.sd=apply(S.sum$N_sim, c(1,2), sd),
+            N_tot.mn=apply(S.sum$N_tot, 1, mean),
+            N_tot.sd=apply(S.sum$N_tot, 1, sd),
+            N_surv.mn=apply(S.sum$N_surv, 1, mean),
+            N_surv.sd=apply(S.sum$N_surv, 1, mean),
+            N_rcr.mn=apply(S.sum$N_rcr, 1, mean),
+            N_rcr.sd=apply(S.sum$N_rcr, 1, mean))
 
@@
 lam.df %<>%
-  add_column(lam.S.f=S.f$N_sim[,tmax]/(S.f$N_sim[,tmax-1]+.01),
-             nSeed.f=S.f$nSd[,tmax], D.f=S.f$D[,tmax], 
-             B0.f=S.f$B[,1], Btmax.f=S.f$B[,tmax+1], 
-             N.S.f=map_dbl(S.f$d, ~sum(!is.na(.$sizeNext[.$yr==tmax]))),
-             Surv.S.f=map_dbl(S.f$d, ~sum(.$surv[.$yr==tmax], na.rm=T)),
-             Rcr.S.f=map_dbl(S.f$d, ~sum(is.na(.$size[.$yr==tmax])))) %>%
+  add_column(lam.S.f=rowMeans(S.f$N_sim.mn[,(tmax-3):tmax]/
+                                (S.f$N_sim.mn[,(tmax-4):(tmax-1)]+0.0001)),
+             nSeed.f=S.f$nSd.mn[,tmax], 
+             D.f=S.f$D.mn[,tmax],
+             B0.f=S.f$B.mn[,1], 
+             Btmax.f=S.f$B.mn[,tmax+1],
+             N.S.f=S.f$N_tot.mn, 
+             Surv.S.f=S.f$N_surv.mn, 
+             Rcr.S.f=S.f$N_rcr.mn) %>%
   mutate(nSdStay.f=nSeed.f*(1-p.ipm$p_emig), nSdLeave.f=nSeed.f*p.ipm$p_emig) %>%
@@
                size2=size^2, size3=size^3)
-  E.t <- xx %>% add_column(s=NA, g=NA, fl=NA, seed=NA)
+  E.t <- xx %>% add_column(yr=NA, s=NA, g=NA, fl=NA, seed=NA)
   xxx <- seq(0, U.f$hi*.5, length.out=100)
   for(i in sample.cells) {
-    z.t <- S.f$d[[i]][!is.na(S.f$d[[i]]$size) & S.f$d[[i]]$yr==tmax,]
-    E.t <- sim_expected(E.t, 1:100, 1:100, E.t, p.ipm, map(X, ~.[i,]), n_z)
-    E.S.t <- sim_expected(E.t, 1:100, 1:100, E.t, p, map(X, ~.[i,]), n_z)
+    z.t <- data.frame(S$d[[i]]) %>% filter(!is.na(size) & yr==tmax)
+    pt.col <- rgb(0,0,0,1/(nrow(z.t)^.25))
+    E.t <- sim_expected(1, xx$size, 1:100, E.t, p.ipm, map(X, ~.[i,]), n_z)
+    E.S.t <- sim_expected(1, xx$size, 1:100, E.t, p, map(X, ~.[i,]), n_z)
     main.i <- paste("Cell", i)
@@
     w.S <- Re(eigen(U$IPMs[,,i])$vectors[,1])[-1]; stable.dist.S <- w.S/sum(w.S)
-    plot(surv~size, data=z.t, ylim=c(0,1), pch=20, col=rgb(0,0,0,0.2),
+    plot(surv~size, data=z.t, ylim=c(0,1), pch=20, col=pt.col,
          main=paste("Surv:", main.i), xlab="Size (t)", ylab="P(surv to t+1)")
       lines(E.t$size, E.t$s, col=2); lines(E.S.t$size, E.S.t$s, col=4)
-    plot(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,0.2), ylim=c(U$lo,U$hi),
+    plot(sizeNext~size, data=z.t, pch=20, col=pt.col, ylim=c(U$lo,U$hi),
          main=paste("Growth:", main.i), xlab="Size (t)", ylab="Size (t+1)")
@@
       abline(a=0, b=1, lty=2)
-    plot(fl~size, data=z.t, pch=20, ylim=c(0,1), col=rgb(0,0,0,0.2),
+    plot(fl~size, data=z.t, pch=20, ylim=c(0,1), col=pt.col,
          main=paste("P(Flower):", main.i), xlab="Size (t)", ylab="P(flower)")
       lines(E.t$size, E.t$fl, col=2); lines(E.S.t$size, E.S.t$fl, col=4)
-    plot(seed~size, data=z.t, pch=23-z.t$fl*3, col=rgb(0,0,0,0.2), 
+    plot(seed~size, data=z.t, pch=23-z.t$fl*3, col=pt.col, 
          ylim=range(c(E.t$seed, z.t$seed), na.rm=TRUE),
@@
       lines(U.f$y, stable.dist.S, col=4)
-    hist(with(S.f$d[[i]], sizeNext[is.na(size)]), col="grey", freq=F, 
+    hist(with(S$d[[i]], sizeNext[is.na(size)]), col="grey", freq=F, 
          main=paste("Recruit sizes:", main.i), xlab="Recruit size (t+1)")
@@
   for(i in sample.cells) {
-   z.t <- S.f$d[[i]][!is.na(S.f$d[[i]]$size) & S.f$d[[i]]$yr==tmax,]
+   z.t <- data.frame(S$d[[i]]) %>% filter(!is.na(size) & yr==tmax)
+   pt.col <- rgb(1,1,1,1/(nrow(z.t)^.25))
    main.i <- paste("Cell", i)
    z.t$y <- U.f$y[findInterval(z.t$size, U.f$y, all.inside=T)]
-   totSeeds <- group_by(z.t, y) %>% summarise(seed=sum(seed, na.rm=T))
+   totSd <- group_by(z.t, y) %>% summarise(seed=sum(seed, na.rm=T))
    image.plot(x, x, t(U.f$Ps[,,i]), main=paste("P matrix:", main.i), 
               xlab="Size (t)", ylab="Size (t+1)")
-    points(sizeNext~size, data=z.t, col=rgb(1,1,1,1/(nrow(z.t)^.25)))
+    points(sizeNext~size, data=z.t, col=pt.col, pch=20)
     abline(a=0, b=1, col="white", lty=2)
@@
               xlab="Size (t)", ylab="Size (t+1)")
-    points(totSeeds$y, totSeeds$seed/max(totSeeds$seed)*U.f$hi*.9, col="white")
+    points(totSd$y, totSd$seed/max(totSd$seed)*U.f$hi*.9, col="white", pch=20)
    image.plot(x, x, t(sens(U.f$IPMs[,,i])), main=paste("Sensitivity:", main.i), 
@@
 x <- c(0, U.f$y)
+S_d <- map_dfr(S$d, data.frame) %>% sample_n(1000)
+S_d$y <- U.f$y[findInterval(S_d$size, U.f$y, all.inside=T)]
+totSd <- group_by(S_d, y) %>% summarise(seed=sum(seed, na.rm=T))
+pt.col <- rgb(1,1,1,1/(nrow(S_d)^.25))
 image.plot(x, x, t(apply(U.f$Ps, 1:2, mean)), main="P matrix: mean",
            xlab="Size (t)", ylab="Size (t+1)")
+  points(sizeNext~size, data=S_d, col=pt.col, pch=20)
 image.plot(x, x, t(apply(U.f$Fs, 1:2, mean)), main="F matrix: mean",
            xlab="Size (t)", ylab="Size (t+1)")
+  points(totSd$y, totSd$seed/max(totSd$seed)*U.f$hi*.9, col="white", pch=20)
 image.plot(x, x, sens(t(apply(U.f$IPMs, 1:2, mean))), main="Sensitivity: mean",
@@
            xlab="Size (t)", ylab="Size (t+1)")
+points(sizeNext~size, data=S_d, col=pt.col, pch=20)
 

### 2018-03-14 exploration.Rmd
enforce plot limits

@@
 P.i <- which(lam.df$Surv.S > 0) # presences: surviving individuals > seedlings
-n_sim <- 20
+n_sim <- 15
 ```
@@
   par(mfrow=c(3,2), mar=c(4,4,2,1))
-  xx <- tibble(size=seq(U.f$lo, U.f$hi, length.out=100), 
+  xx <- tibble(size=seq(U$lo, U$hi, length.out=100), 
                size2=size^2, size3=size^3)
   E.t <- xx %>% add_column(yr=NA, s=NA, g=NA, fl=NA, seed=NA)
-  xxx <- seq(0, U.f$hi*.5, length.out=100)
+  xxx <- seq(0, U$hi*.5, length.out=100)
   for(i in sample.cells) {
@@
     w.S <- Re(eigen(U$IPMs[,,i])$vectors[,1])[-1]; stable.dist.S <- w.S/sum(w.S)
-    plot(surv~size, data=z.t, ylim=c(0,1), pch=20, col=pt.col,
+    plot(z.t$surv~z.t$size, ylim=c(0,1), pch=20, col=pt.col, xlim=c(U$lo, U$hi),
          main=paste("Surv:", main.i), xlab="Size (t)", ylab="P(surv to t+1)")
       lines(E.t$size, E.t$s, col=2); lines(E.S.t$size, E.S.t$s, col=4)
-    plot(sizeNext~size, data=z.t, pch=20, col=pt.col, ylim=c(U$lo,U$hi),
+    plot(z.t$sizeNext~z.t$size, pch=20, col=pt.col, ylim=c(U$lo,U$hi),
+         xlim=c(U$lo, U$hi),
          main=paste("Growth:", main.i), xlab="Size (t)", ylab="Size (t+1)")
@@
       abline(a=0, b=1, lty=2)
-    plot(fl~size, data=z.t, pch=20, ylim=c(0,1), col=pt.col,
+    plot(fl~size, data=z.t, pch=20, ylim=c(0,1), col=pt.col, xlim=c(U$lo, U$hi),
          main=paste("P(Flower):", main.i), xlab="Size (t)", ylab="P(flower)")
       lines(E.t$size, E.t$fl, col=2); lines(E.S.t$size, E.S.t$fl, col=4)
-    plot(seed~size, data=z.t, pch=23-z.t$fl*3, col=pt.col, 
+    plot(seed~size, data=z.t, pch=23-z.t$fl*3, col=pt.col, xlim=c(U$lo, U$hi),
          ylim=range(c(E.t$seed, z.t$seed), na.rm=TRUE),

### 2018-03-14 exploration.Rmd
added function for summarizing IPM simulations

@@
 }
-S.sum <- list(B=map(S.i, ~.$B),
-              nSd=map(S.i, ~.$nSd),
-              D=map(S.i, ~.$D),
-              p_est=map(S.i, ~.$p_est.i),
-              N_sim=map(S.i, ~.$N_sim),
-              N_tot=map(S.i, ~map_dbl(.$d, 
-                                    ~sum(!is.na(.$sizeNext[.$yr==tmax])))),
-              N_surv=map(S.i, ~map_dbl(.$d, 
-                                    ~sum(.$surv[.$yr==tmax], na.rm=T))),
-              N_rcr=map(S.i, ~map_dbl(.$d, 
-                                    ~sum(is.na(.$size[.$yr==tmax]))))) %>% 
-  map(simplify2array)
-rm(S.i)
-
-S.f <- list(B.mn=apply(S.sum$B, c(1,2), mean), 
-            B.sd=apply(S.sum$B, c(1,2), sd), 
-            nSd.mn=apply(S.sum$nSd, c(1,2), mean), 
-            nSd.sd=apply(S.sum$nSd, c(1,2), sd), 
-            D.mn=apply(S.sum$D, c(1,2), mean), 
-            D.sd=apply(S.sum$D, c(1,2), sd), 
-            p_est.mn=apply(S.sum$p_est, c(1,2), mean), 
-            p_est.sd=apply(S.sum$p_est, c(1,2), sd), 
-            N_sim.mn=apply(S.sum$N_sim, c(1,2), mean), 
-            N_sim.sd=apply(S.sum$N_sim, c(1,2), sd),
-            N_tot.mn=apply(S.sum$N_tot, 1, mean),
-            N_tot.sd=apply(S.sum$N_tot, 1, sd),
-            N_surv.mn=apply(S.sum$N_surv, 1, mean),
-            N_surv.sd=apply(S.sum$N_surv, 1, mean),
-            N_rcr.mn=apply(S.sum$N_rcr, 1, mean),
-            N_rcr.sd=apply(S.sum$N_rcr, 1, mean))
+S.f <- summarize_IPM_simulations(S.i)
 

### 2018-03-14 code/fn_aux.R

@@ -14,3 +14,42 @@ z_pow <- function(z.vec, n_z) {
     return(z)
   }
 }
+
+
+##-- calculate means and sds from multiple IPM simulations
+summarize_IPM_simulations <- function(S.i) {
+  
+  require(tidyverse)
+  
+  S.sum <- list(B=map(S.i, ~.$B),
+                nSd=map(S.i, ~.$nSd),
+                D=map(S.i, ~.$D),
+                p_est=map(S.i, ~.$p_est.i),
+                N_sim=map(S.i, ~.$N_sim),
+                N_tot=map(S.i, ~map_dbl(.$d, 
+                                        ~sum(!is.na(.$sizeNext[.$yr==tmax])))),
+                N_surv=map(S.i, ~map_dbl(.$d, 
+                                         ~sum(.$surv[.$yr==tmax], na.rm=T))),
+                N_rcr=map(S.i, ~map_dbl(.$d, 
+                                        ~sum(is.na(.$size[.$yr==tmax]))))) %>% 
+    map(simplify2array)
+  rm(S.i)
+  
+  return(list(B.mn=apply(S.sum$B, c(1,2), mean), 
+              B.sd=apply(S.sum$B, c(1,2), sd), 
+              nSd.mn=apply(S.sum$nSd, c(1,2), mean), 
+              nSd.sd=apply(S.sum$nSd, c(1,2), sd), 
+              D.mn=apply(S.sum$D, c(1,2), mean), 
+              D.sd=apply(S.sum$D, c(1,2), sd), 
+              p_est.mn=apply(S.sum$p_est, c(1,2), mean), 
+              p_est.sd=apply(S.sum$p_est, c(1,2), sd), 
+              N_sim.mn=apply(S.sum$N_sim, c(1,2), mean), 
+              N_sim.sd=apply(S.sum$N_sim, c(1,2), sd),
+              N_tot.mn=apply(S.sum$N_tot, 1, mean),
+              N_tot.sd=apply(S.sum$N_tot, 1, sd),
+              N_surv.mn=apply(S.sum$N_surv, 1, mean),
+              N_surv.sd=apply(S.sum$N_surv, 1, mean),
+              N_rcr.mn=apply(S.sum$N_rcr, 1, mean),
+              N_rcr.sd=apply(S.sum$N_rcr, 1, mean)))
+}
+

### 2018-03-14 exploration.Rmd
adjusting landscape setup to allow for subgrids for testing

--- 2018-03-14 exploration.Rmd
@@
 source("code/fn_IPM.R"); source("code/fn_aux.R"); source("code/fn_sim.R")
+```
+
+## Landscape
+The landscape characteristics are aggregated here. It is a gridded landscape with climate and LULC data.
+```{r landscape}
+x_max <- 100
+y_max <- 100
+
 lc.df <- read_csv("data/landcover_5km.csv") %>% 
@@
          x_y=paste(x, y, sep="_"))
+env.rct <- as.tibble(expand.grid(x=1:max(lc.df$x),
+                                 y=1:max(lc.df$y))) %>%
+  mutate(x_y=paste(x, y, sep="_")) %>%
+  mutate(temp=c(scale(lc.df$bio1_mean))[match(.$x_y, lc.df$x_y)],
+         temp2=temp^2,
+         prec=c(scale(lc.df$bio12_mean))[match(.$x_y, lc.df$x_y)],
+         prec2=prec^2,
+         pOpn=lc.df$nlcd1_mean[match(.$x_y, lc.df$x_y)],
+         pOth=lc.df$nlcd2_mean[match(.$x_y, lc.df$x_y)],
+         pDec=lc.df$nlcd3_mean[match(.$x_y, lc.df$x_y)],
+         pEvg=lc.df$nlcd4_mean[match(.$x_y, lc.df$x_y)],
+         pMxd=lc.df$nlcd5_mean[match(.$x_y, lc.df$x_y)],
+         inbd=!is.na(match(.$x_y, lc.df$x_y)),
+         id=row_number(), 
+         id.inbd=min_rank(na_if(inbd*id, 0)),
+         lat=lc.df$top[match(.$x_y, lc.df$x_y)],
+         lon=lc.df$left[match(.$x_y, lc.df$x_y)])
+env.rct[is.na(env.rct)] <- 0
+env.df <- filter(env.rct, inbd) %>% select(4:12, 1:3, 13:17) %>%
+  mutate(pOpn=c(scale(pOpn)), pOth=c(scale(pOth)), pDec=c(scale(pDec)),
+         pEvg=c(scale(pEvg)), pMxd=c(scale(pMxd)))
+rm(env.rct)
 ```
@@
 z.rng <- c(1, 12)
-n <- 50  # ncells in IPM matrix
-tmax <- 30
-cells <- 1:n_distinct(lc.df$CellID)
+n <- 20  # ncells in IPM matrix
+tmax <- 10
+cells <- n_distinct(lc.df$CellID)
 n.cell <- length(cells)
@@
 sp_i$p$p_emig <- pexp(0.5, sp_i$p$sdd_rate, lower.tail=F) # p(seed emigrates)
-```
-
-## Landscape
-The landscape characteristics are aggregated here. It is a gridded landscape with climate and LULC data.
-```{r landscape}
-env.rct <- as.tibble(expand.grid(x=1:max(lc.df$x),
-                                 y=1:max(lc.df$y))) %>%
-  mutate(x_y=paste(x, y, sep="_")) %>%
-  mutate(temp=c(scale(lc.df$bio1_mean))[match(.$x_y, lc.df$x_y)],
-         temp2=temp^2,
-         prec=c(scale(lc.df$bio12_mean))[match(.$x_y, lc.df$x_y)],
-         prec2=prec^2,
-         pOpn=lc.df$nlcd1_mean[match(.$x_y, lc.df$x_y)],
-         pOth=lc.df$nlcd2_mean[match(.$x_y, lc.df$x_y)],
-         pDec=lc.df$nlcd3_mean[match(.$x_y, lc.df$x_y)],
-         pEvg=lc.df$nlcd4_mean[match(.$x_y, lc.df$x_y)],
-         pMxd=lc.df$nlcd5_mean[match(.$x_y, lc.df$x_y)],
-         inbd=!is.na(match(.$x_y, lc.df$x_y)),
-         id=row_number(), 
-         id.inbd=min_rank(na_if(inbd*id, 0)),
-         lat=lc.df$top[match(.$x_y, lc.df$x_y)],
-         lon=lc.df$left[match(.$x_y, lc.df$x_y)])
-env.rct[is.na(env.rct)] <- 0
+# sdd.pr[,,2,] has indexes based on env.df$id -- NOT env.df$id.inbd
 sdd.pr <- sdd_set_probs(ncell=n.cell, lc.df=env.rct, lc.col=8:12,
@@
                                  bird.hab=rep(1,5)))
-env.df <- filter(env.rct, inbd) %>% select(4:12, 1:3, 13:17) %>%
-  mutate(pOpn=c(scale(pOpn)), pOth=c(scale(pOth)), pDec=c(scale(pDec)),
-         pEvg=c(scale(pEvg)), pMxd=c(scale(pMxd)))
-rm(env.rct)
-# sdd.pr[,,2,] has indexes based on env.df$id -- NOT env.df$id.inbd
-# ggplot(env.df, aes(x=lon, y=lat, fill=id %in% sdd.pr[,,2,740])) + geom_tile()
 ```

### 2018-03-14 exploration.Rmd
added axis limits

@@
     plot(surv~size, data=z.t, ylim=c(0,1), pch=20, col=pt.col, 
+         xlim=c(U$lo, U$hi),
          main=paste("Surv:", main.i), xlab="Size (t)", ylab="P(surv to t+1)")
@@
     plot(sizeNext~size, data=z.t, pch=20, col=pt.col, ylim=c(U$lo,U$hi), 
+         xlim=c(U$lo, U$hi),
          main=paste("Growth:", main.i), xlab="Size (t)", ylab="Size (t+1)")
@@
       abline(a=0, b=1, lty=2)
-    plot(fl~size, data=z.t, pch=20, ylim=c(0,1), col=pt.col,
+    plot(fl~size, data=z.t, pch=20, ylim=c(0,1), col=pt.col, xlim=c(U$lo, U$hi),
          main=paste("P(Flower):", main.i), xlab="Size (t)", ylab="P(flower)")
       lines(E.t$size, E.t$fl, col=2)
-    plot(seed~size, data=z.t, pch=23-z.t$fl*3, col=pt.col, 
+    plot(seed~size, data=z.t, pch=23-z.t$fl*3, col=pt.col, xlim=c(U$lo, U$hi),
          ylim=range(c(E.t$seed, z.t$seed), na.rm=TRUE),

### 2018-03-15 code/fn_IPM.R
storing more parameters in p instead of feeding as separate arguments

@@ -126,14 +126,14 @@ setup_IPM_matrix <- function(n=100, z.rng=c(1,10), buffer=0.25) {
 
 ##-- Fill P matrix: local 
 ##   
-fill_P <- function(h, y, z.i, n, p, n_z, n_x, X_s, X_g) {
-  P.mx <- matrix(0, nrow=n+1, ncol=n+1)
+fill_P <- function(h, y, z.i, p, n_z, n_x, X_s, X_g) {
+  P.mx <- matrix(0, nrow=p$n+1, ncol=p$n+1)
   # survival & growth
   S.v <- calc_surv(y, p=p, n_sz=n_z$s, X.s=X_s)
   G.mx <- h*outer(y, y, calc_grow, p=p, n_gz=n_z$g, X.g=X_g)
   # correct ejections
-  for(k in 1:(n/2)) G.mx[1,k] <- G.mx[1,k] + 1 - sum(G.mx[,k])
-  for(k in (n/2+1):n) G.mx[n,k] <- G.mx[n,k] + 1 - sum(G.mx[,k])
+  for(k in 1:(p$n/2)) G.mx[1,k] <- G.mx[1,k] + 1 - sum(G.mx[,k])
+  for(k in (p$n/2+1):p$n) G.mx[p$n,k] <- G.mx[p$n,k] + 1 - sum(G.mx[,k])
   # fill P matrix
   for(k in z.i) P.mx[k,z.i] <- G.mx[k-1,]*S.v
   P.mx[1,1] <- calc_staySB(p)
@@ -143,8 +143,8 @@ fill_P <- function(h, y, z.i, n, p, n_z, n_x, X_s, X_g) {
 
 ##-- Fill F matrix 
 ##   
-fill_F <- function(h, y, z.i, n, p, n_z, n_x, X_fl, X_seed) {
-  F.mx <- matrix(0, nrow=n+1, ncol=n+1)
+fill_F <- function(h, y, z.i, p, n_z, n_x, X_fl, X_seed) {
+  F.mx <- matrix(0, nrow=p$n+1, ncol=p$n+1)
   F.mx[z.i,z.i] <- outer(y, y, calc_rcrDir, p=p, n_seedz=n_z$seed, n_flz=n_z$fl, 
                          X.seed=X_seed, X.fl=X_fl) * h * p$p_est
   F.mx[1,z.i] <- calc_addSB(y, p=p, n_seedz=n_z$seed, n_flz=n_z$fl, 
[…]

### 2018-03-15 code/fn_aux.R
function for building landscape

@@ -3,6 +3,48 @@
 # Tim Szewczyk
 
 
+##-- generate landscape
+build_landscape <- function(f, x_max=Inf, y_max=Inf) {
+  require(tidyverse)
+  
+  # load GIS data
+  lc.df <- read_csv(f) %>% 
+    filter(!is.na(bio1_mean)) %>%
+    mutate(x=as.integer(factor(.$left)),
+           y=as.integer(factor(.$top, levels=rev(levels(factor(.$top))))),
+           x_y=paste(x, y, sep="_"))
+  # establish grid with x=1:ncol, y=1:nrow
+  env.rct <- as.tibble(expand.grid(x=1:max(lc.df$x), y=1:max(lc.df$y))) %>%
+    mutate(x_y=paste(x, y, sep="_")) 
+  match_id <- match(env.rct$x_y, lc.df$x_y)
+  # pair environmental variables
+  env.rct <- env.rct %>%
+    mutate(temp=c(scale(lc.df$bio1_mean))[match_id],
+           temp2=temp^2,
+           prec=c(scale(lc.df$bio12_mean))[match_id],
+           prec2=prec^2,
+           pOpn=lc.df$nlcd1_mean[match_id],
+           pOth=lc.df$nlcd2_mean[match_id],
+           pDec=lc.df$nlcd3_mean[match_id],
+           pEvg=lc.df$nlcd4_mean[match_id],
+           pMxd=lc.df$nlcd5_mean[match_id],
+           inbd=!is.na(match(.$x_y, lc.df$x_y)),
+           lat=lc.df$top[match_id],
+           lon=lc.df$left[match_id],
+           rdLen=lc.df$rdLen[match_id]) %>%
+    filter(x <= x_max & y >= (max(.$y)-y_max)) %>%
+    mutate(id=row_number(), 
+           id.inbd=min_rank(na_if(inbd*id, 0)))
+  env.rct[is.na(env.rct)] <- 0
+  # subset inbound cells
+  env.in <- filter(env.rct, inbd) %>% select(4:12, 1:3, 13:18) %>%
+    mutate(pOpn=c(scale(pOpn)), pOth=c(scale(pOth)), pDec=c(scale(pDec)),
+           pEvg=c(scale(pEvg)), pMxd=c(scale(pMxd)))
+  
+  return(list(env.rct=env.rct, env.in=env.in))
+}
+
+
 ##-- create design matrix from vector of sizes
 z_pow <- function(z.vec, n_z) {
   if(n_z==1) {

### 2018-03-15 code/fn_aux.R
antilogit function to avoid needing library(sevcheck)

@@ -95,3 +95,7 @@ summarize_IPM_simulations <- function(S.i) {
               N_rcr.sd=apply(S.sum$N_rcr, 1, mean)))
 }
 
+
+##-- antilogit
+antilogit <- function (x) {
+  exp(x)/(1 + exp(x))

### 2018-03-15 code/fn_aux.R
reformatting for clarity

@@ -68,25 +68,25 @@ summarize_IPM_simulations <- function(S.i) {
                 D=map(S.i, ~.$D),
                 p_est=map(S.i, ~.$p_est.i),
                 N_sim=map(S.i, ~.$N_sim),
-                N_tot=map(S.i, ~map_dbl(.$d, 
-                                        ~sum(!is.na(.$sizeNext[.$yr==tmax])))),
-                N_surv=map(S.i, ~map_dbl(.$d, 
-                                         ~sum(.$surv[.$yr==tmax], na.rm=T))),
-                N_rcr=map(S.i, ~map_dbl(.$d, 
-                                        ~sum(is.na(.$size[.$yr==tmax]))))) %>% 
+                N_tot=map(S.i, 
+                          ~map_dbl(.$d, ~sum(!is.na(.$sizeNext[.$yr==tmax])))),
+                N_surv=map(S.i, 
+                           ~map_dbl(.$d, ~sum(.$surv[.$yr==tmax], na.rm=T))),
+                N_rcr=map(S.i, 
+                          ~map_dbl(.$d, ~sum(is.na(.$size[.$yr==tmax]))))) %>% 
     map(simplify2array)
   rm(S.i)
   
-  return(list(B.mn=apply(S.sum$B, c(1,2), mean), 
-              B.sd=apply(S.sum$B, c(1,2), sd), 
-              nSd.mn=apply(S.sum$nSd, c(1,2), mean), 
-              nSd.sd=apply(S.sum$nSd, c(1,2), sd), 
-              D.mn=apply(S.sum$D, c(1,2), mean), 
-              D.sd=apply(S.sum$D, c(1,2), sd), 
-              p_est.mn=apply(S.sum$p_est, c(1,2), mean), 
-              p_est.sd=apply(S.sum$p_est, c(1,2), sd), 
-              N_sim.mn=apply(S.sum$N_sim, c(1,2), mean), 
-              N_sim.sd=apply(S.sum$N_sim, c(1,2), sd),
+  return(list(B.mn=apply(S.sum$B, 1:2, mean), 
+              B.sd=apply(S.sum$B, 1:2, sd), 
+              nSd.mn=apply(S.sum$nSd, 1:2, mean), 
+              nSd.sd=apply(S.sum$nSd, 1:2, sd), 
+              D.mn=apply(S.sum$D, 1:2, mean), 
[…]

### 2018-03-15 code/fn_sim.R
storing more parameters in p instead of feeding separately

@@ -64,20 +64,19 @@ sim_seedbank <- function(nSd_ik, B_ik, D_ik, p) {
 
 ##--- simulate full dataset
 ##
-simulate_data <- function(n.cell, tmax, n0, z.rng, lo, hi, 
-                          p, X, n_z, sdd, sdd.j, verbose=FALSE) {
+simulate_data <- function(n.cell, lo, hi, p, X, n_z, sdd, sdd.j, verbose=F) {
   require(tidyverse)
   i <- 1:n.cell
   
   # storage objects
   E <- d <- map(i, ~list())  # much faster than using dataframes
-  B <- cbind(rpois(n.cell, n0), matrix(0, nrow=n.cell, ncol=tmax))
-  nSd <- D <- N_sim <- matrix(0, nrow=n.cell, ncol=tmax)
-  p_est.i <- matrix(p$p_est, nrow=n.cell, ncol=tmax)
-  z.k <- map(i, ~runif(n0, z.rng[1], z.rng[2]))
+  B <- cbind(rpois(n.cell, p$n0), matrix(0, nrow=n.cell, ncol=p$tmax))
+  nSd <- D <- N_sim <- matrix(0, nrow=n.cell, ncol=p$tmax)
+  p_est.i <- matrix(p$p_est, nrow=n.cell, ncol=p$tmax)
+  z.k <- map(i, ~runif(p$n0, p$z.rng[1], p$z.rng[2]))
   X_map <- lapply(i, function(x) map(X, ~.[x,]))
   
-  for(k in 1:tmax) {
+  for(k in 1:p$tmax) {
     ## local growth
     if(k>1) z.k <- map(d, ~.$sizeNext[!is.na(.$sizeNext) & .$yr == k-1])
     d.k <- lapply(i, function(x) seq_along(z.k[[x]]) + length(d[[x]]$yr))

### 2018-03-15 code/1_simulateSpecies.R
initial commit

@@ -1,0 +1,111 @@
+# 1: Simulate species
+# Comparison of SDM approaches
+# Tim Szewczyk
+
+# This script generates data for hypothetical species. The simulation process is
+# based on an IPM framework. Both the underlying IPM objects and the simulated
+# individuals are produced and stored. This serves as the perfectly known true
+# distribution from which samples are taken for fitting each SDM.
+
+########
+## Setup
+########
+# file specifications
+sp <- "sp1"
+env.f <- "data/landcover_5km.csv"  # file with environmental data
+
+# load workspace
+pkgs <- c("tidyverse", "magrittr")
+suppressMessages(invisible(lapply(pkgs, library, character.only=T)))
+source("code/fn_IPM.R"); source("code/fn_aux.R"); source("code/fn_sim.R")
+L <- build_landscape(f=env.f, 
+                     x_max=30, # ncol in landscape; Inf for full dataset
+                     y_max=30) # nrow in landscape; Inf for full dataset
+n.cell <- sum(L$env.rct$inbd)
+
+
+########
+## Set species traits
+########
+p=list(n=20,  # ncells in IPM matrix
+       tmax=10,  # time steps for NDD & simulations
+       n0=100,  # initial pop sizes
+       z.rng=c(1,12),  # initial size range
+       s_z=c(-8, 2.1, -.09),  # b1 + b2*z + b3*z^2
+       s_x=c(2, -.1, -2, -.1, 2, -2, -.4),  # b1*x1 + ...
+       g_z=c(.2, 2, -0.1),  # b1 + b2*z + b3*z^2
+       g_x=c(2, -.1, 2, -.1, 2, -2),  # b1*x1 + ...
+       g_sig=1,  # growth ~ N(E, g_sig)
+       fl_z=c(-1.5, .1, .1),  # b1 + b2*z + b3*z^2
+       fl_x=c(-2, -.1, -2, -.1, 1, 1),  # b1*x1 + ...
+       seed_z=c(2, 0.5, -.03),  # b1 + b2*z + b2*z^2
+       seed_x=c(1, -.1, -1, -.1, .2),  # b1*x1 + ...
+       rcr_z=c(1.5, 0.4),  # N(mean=rcrt1, sd=rcrt2)
+       p_est=0.03,  # p(establishment)
+       NDD=T,  # negative density dependent p_est
[…]

### 2018-03-15 code/2_sampleDistribution.R

@@ -1,0 +1,103 @@
+# 2: Sample from simulated distribution
+# Comparison of SDM approaches
+# Tim Szewczyk
+
+# This script samples from the distribution simulated in `1_simulateSpecies.R` 
+# and formats the samples for the purposes of fitting a correlative Maxent SDM,
+# a mechanistic CA-style SDM, and a mechanistic IPM-style SDM. The issues we
+# explore that are sampling-based are imposed here.
+
+########
+## Setup
+########
+# file specifications
+sp <- "sp1"
+sampling.issue <- c("none", "error", "geog", "bias")[1]
+
+# load workspace
+pkgs <- c("tidyverse", "magrittr")
+suppressMessages(invisible(lapply(pkgs, library, character.only=T)))
+source("code/fn_IPM.R"); source("code/fn_aux.R"); source("code/fn_sim.R")
+env.in <- readRDS(paste0("out/", sp, "_env_in.rds"))
+n.cell <- nrow(env.in)
+p <- readRDS(paste0("out/", sp, "_p.rds"))
+S <- readRDS(paste0("out/", sp, "_S.rds"))
+lam.df <- readRDS(paste0("out/", sp, "_lam_df.rds"))
+
+
+########
+## Set sampling details
+########
+n_samp <- 2  # number of unique samples to average across
+O_n <- list(Corr=100, Mech=20)  # number of cells in sample
+O_yr <- list(Mx=p$tmax, CA=(p$tmax-10):p$tmax, IPM=p$tmax)  # years to sample
+P.i <- which(lam.df$Surv.S > 0)  # presences: survival past recruit stage
+P.pr <- rep(1, length(P.i))  # pr(sample cell | presence)
+prop.sampled <- 0.8  # proportion of individuals sampled per sampled cell 
+geog.excl <- which(env.in$x > 10 & env.in$y > 45)
+prop.err <- 0.2
+
+
+########
+## Draw samples
+########
+# determine cells
+if(sampling.issue=="geog") {
+  P.pr[P.i %in% geog.excl] <- 0
+} else if(sampling.issue=="bias") {
+  P.pr <- P.pr * env.in$rdLen[P.i]
+}
[…]

### 2018-03-15 code/2_sampleDistribution.R
calculating size^2 and size^3 as part of the sample

@@ -22,7 +22,7 @@ env.in <- readRDS(paste0("out/", sp, "_env_in.rds"))
 n.cell <- nrow(env.in)
 p <- readRDS(paste0("out/", sp, "_p.rds"))
 S <- readRDS(paste0("out/", sp, "_S.rds"))
-lam.df <- readRDS(paste0("out/", sp, "_lam_df.rds"))
+lam.df <- readRDS(paste0("out/", sp, "_lam_df.rds")) # env + true pop values
 
 
 ########
@@ -72,6 +72,7 @@ for(s in 1:n_samp) {
   IPM.d <- vector("list", O_n$Mech)
   for(i in Mech.sample[[s]]) {
     IPM.d[[i]] <- data.frame(S$d[[i]]) %>% filter(yr %in% O_yr$IPM) %>%
+      mutate(size2=size^2, size3=size^3) %>%
       add_column(id.inbd=i) %>%
       full_join(env.in[i,], by="id.inbd")
     IPM.d[[i]] <- sample_frac(IPM.d[[i]], prop.sampled)

### 2018-03-15 code/fn_aux.R
need tmax in summarize_IPM_simulations()

@@ -59,7 +59,7 @@ z_pow <- function(z.vec, n_z) {
 
 
 ##-- calculate means and sds from multiple IPM simulations
-summarize_IPM_simulations <- function(S.i) {
+summarize_IPM_simulations <- function(S.i, tmax) {
   
   require(tidyverse)
   

### 2018-03-15 code/3_fitModels.R
initial commit

@@ -1,0 +1,125 @@
+# 3: Fit models from samples
+# Comparison of SDM approaches
+# Tim Szewczyk
+
+# This script uses the samples generated in `2_sampleDistribution.R` to fit a
+# correlative MaxEnt SDM, a mechanistic CA-style SDM, and a mechanistic 
+# IPM-style SDM. Model selection is performed for each model
+
+########
+## Setup
+########
+# file specifications
+sp <- "sp1"
+sampling.issue <- c("none", "error", "geog", "bias")[1]
+modeling.issue <- c("none")[1]
+
+# load workspace
+pkgs <- c("gbPopMod", "tidyverse", "magrittr", "MuMIn")
+suppressMessages(invisible(lapply(pkgs, library, character.only=T)))
+source("code/fn_IPM.R"); source("code/fn_aux.R"); source("code/fn_sim.R")
+p <- readRDS(paste0("out/", sp, "_p.rds"))
+S <- readRDS(paste0("out/", sp, "_S.rds"))
+U <- readRDS(paste0("out/", sp, "_U.rds"))
+sdd.pr <- readRDS(paste0("out/", sp, "_sdd.rds"))
+lam.df <- readRDS(paste0("out/", sp, "_lam_df.rds"))
+env.in <- readRDS(paste0("out/", sp, "_env_in.rds"))
+n.cell <- nrow(env.in)
+O_Mx <- readRDS(paste0("out/", sp, "_O_Mx_", sampling.issue, ".rds"))
+O_CA <- readRDS(paste0("out/", sp, "_O_CA_", sampling.issue, ".rds"))
+O_IPM <- readRDS(paste0("out/", sp, "_O_IPM_", sampling.issue, ".rds"))
+
+
+########
+## Set model details
+########
+n_sim <- 15  # number of simulations per sample (mechanistic only)
+v <- c("(Intercept)"=0, "size"=0, "size2"=0, "size3"=0, 
+       "temp"=0, "temp2"=0, "prec"=0, "prec2"=0, 
+       "pOpn"=0, "pOth"=0, "pDec"=0, "pEvg"=0, "pMxd"=0)
+m.full <- paste(names(v)[-1], collapse=" + ")
+n_z <- sum(grepl("size", names(v))) + 1  # include intercept
+n_x <- length(v) - n_z
[…]

### 2018-03-15 code/1_simulateSpecies.R
need gbPopMod::sdd_set_probs()

@@ -15,7 +15,7 @@ sp <- "sp1"
 env.f <- "data/landcover_5km.csv"  # file with environmental data
 
 # load workspace
-pkgs <- c("tidyverse", "magrittr")
+pkgs <- c("gbPopMod", "tidyverse", "magrittr")
 suppressMessages(invisible(lapply(pkgs, library, character.only=T)))
 source("code/fn_IPM.R"); source("code/fn_aux.R"); source("code/fn_sim.R")
 L <- build_landscape(f=env.f, 

### 2018-03-15 code/3_fitModels.R
summarizing output from multiple samples

@@ -109,15 +109,36 @@ for(i in 1:length(O_IPM)) {
   
   #---- Realization: use estimated slopes to generate simulated data
   S.i <- vector("list", n_sim)
+  cat("||-- Starting simulations\n||--\n")
   for(s in 1:n_sim) {
     S.i[[s]] <- simulate_data(n.cell, U$lo, U$hi, p.ipm, X, n_z, 
                               sdd.pr, U$sdd.j, verbose=T)
     cat("||-- Finished simulation", s, "\n||--\n")
   }
   S.f[[i]] <- summarize_IPM_simulations(S.i, p.ipm$tmax)
+  cat("\n  Finished dataset", i, "\n")
 }
+out <- summarize_IPM_samples(U.f, S.f)
+
+P_IPM <- lam.df %>% 
+  mutate(lambda.f=apply(out$Uf$IPM.mn, 3, function(x) Re(eigen(x)$values[1])),
+         lam.S.f=rowMeans(out$Sf$N_sim.mn[,(-3:0)+p.ipm$tmax]/
+                            (out$Sf$N_sim.mn[,(-4:-1)+p.ipm$tmax]+0.0001)),
+         nSeed.f=out$Sf$nSd.mn[,p.ipm$tmax], 
+         D.f=out$Sf$D.mn[,p.ipm$tmax],
+         B0.f=out$Sf$B.mn[,1], 
+         Btmax.f=out$Sf$B.mn[,p.ipm$tmax+1],
+         N.S.f=out$Sf$N_tot.mn, 
+         Surv.S.f=out$Sf$N_surv.mn, 
+         Rcr.S.f=out$Sf$N_rcr.mn) %>%
+  mutate(nSdStay.f=nSeed.f*(1-p.ipm$p_emig), 
+         nSdLeave.f=nSeed.f*p.ipm$p_emig,
+         N.U.f=apply(out$Uf$Nt.mn[,,p.ipm$tmax],2,sum), 
+         lam.U.f=out$Uf$lam.mn[,p.ipm$tmax-1])
 
 
+saveRDS(P_IPM, paste0("out/", sp, "_P_IPM_", sampling.issue, "_",
+                      modeling.issue, ".rds"))
 
 
 

### 2018-03-15 code/fn_aux.R
function for summarizing IPM output from multiple samples

@@ -75,7 +75,6 @@ summarize_IPM_simulations <- function(S.i, tmax) {
                 N_rcr=map(S.i, 
                           ~map_dbl(.$d, ~sum(is.na(.$size[.$yr==tmax]))))) %>% 
     map(simplify2array)
-  rm(S.i)
   
   return(list(B.mn=apply(S.sum$B, 1:2, mean), 
               B.sd=apply(S.sum$B, 1:2, sd), 
@@ -93,6 +92,42 @@ summarize_IPM_simulations <- function(S.i, tmax) {
               N_surv.sd=apply(S.sum$N_surv, 1, mean),
               N_rcr.mn=apply(S.sum$N_rcr, 1, mean),
               N_rcr.sd=apply(S.sum$N_rcr, 1, mean)))
+}
+
+
+##-- calculate means from simulation means of multiple IPM samples
+summarize_IPM_samples <- function(U.f, S.f) {
+  Ua <- list(IPM.mn=map(U.f, ~.$IPMs),
+             P.mn=map(U.f, ~.$Ps),
+             F.mn=map(U.f, ~.$Fs),
+             Nt.mn=map(U.f, ~.$Nt),
+             lam.mn=map(U.f, ~.$lam.t),
+             p_est.mn=map(U.f, ~.$p_est.t)) %>% 
+    map(simplify2array)
+  Uf <- list(IPM.mn=apply(Ua$IPM.mn, 1:3, mean),
+             P.mn=apply(Ua$P.mn, 1:3, mean),
+             F.mn=apply(Ua$F.mn, 1:3, mean),
+             Nt.mn=apply(Ua$Nt.mn, 1:3, mean),
+             lam.mn=apply(Ua$lam.mn, 1:2, mean),
+             p_est.mn=apply(Ua$p_est.mn, 1:2, mean))
+  Sa <- list(B.mn=map(S.f, ~.$B.mn),
+             nSd.mn=map(S.f, ~.$nSd.mn),
+             D.mn=map(S.f, ~.$D.mn),
+             p_est.mn=map(S.f, ~.$p_est.mn),
+             N_sim.mn=map(S.f, ~.$N_sim.mn),
+             N_tot.mn=map(S.f, ~.$N_tot.mn),
+             N_surv.mn=map(S.f, ~.$N_surv.mn),
+             N_rcr.mn=map(S.f, ~.$N_rcr.mn)) %>%
[…]

### 2018-03-16 code/1_simulateSpecies.R
clearer lam.df creation

@@ -27,9 +27,9 @@ n.cell <- sum(L$env.rct$inbd)
 ########
 ## Set species traits
 ########
-p=list(n=20,  # ncells in IPM matrix
+p=list(n=50,  # ncells in IPM matrix
        tmax=10,  # time steps for NDD & simulations
-       n0=100,  # initial pop sizes
+       n0=200,  # initial pop sizes
        z.rng=c(1,12),  # initial size range
        s_z=c(-8, 2.1, -.09),  # b1 + b2*z + b3*z^2
        s_x=c(2, -.1, -2, -.1, 2, -2, -.4),  # b1*x1 + ...
@@ -84,17 +84,20 @@ U <- fill_IPM_matrices(n.cell, buffer=0.75, discrete=1, p, n_z, n_x,
 S <- simulate_data(n.cell, U$lo, U$hi, p, X, n_z, sdd.pr, U$sdd.j, verbose=T)
 
 # Aggregate results
-lam.df <- L$env.in
-lam.df$lambda <- apply(U$IPMs, 3, function(x) Re(eigen(x)$values[1]))
-lam.df %<>% 
-  add_column(lam.S=S$N_sim[,p$tmax]/(S$N_sim[,p$tmax-1]+.01),
-             nSeed=S$nSd[,p$tmax], D=S$D[,p$tmax], 
-             B0=S$B[,1], Btmax=S$B[,p$tmax+1], 
-             N.S=map_dbl(S$d, ~sum(!is.na(.$sizeNext[.$yr==p$tmax]))),
-             Surv.S=map_dbl(S$d, ~sum(.$surv[.$yr==p$tmax], na.rm=T)),
-             Rcr.S=map_dbl(S$d, ~sum(is.na(.$size[.$yr==p$tmax])))) %>%
-  mutate(nSdStay=nSeed*(1-p$p_emig), nSdLeave=nSeed*p$p_emig) %>%
-  add_column(N.U=apply(U$Nt[-1,,p$tmax+1],2,sum), lam.U=U$lam.t[,p$tmax])
+lam.df <- L$env.in %>%
+  mutate(lambda=apply(U$IPMs, 3, function(x) Re(eigen(x)$values[1])),
+         lam.S=S$N_sim[,p$tmax]/(S$N_sim[,p$tmax-1]+.01),
+         nSeed=S$nSd[,p$tmax], 
+         D=S$D[,p$tmax], 
[…]

### 2018-03-16 code/2_sampleDistribution.R
testing options

@@ -12,7 +12,7 @@
 ########
 # file specifications
 sp <- "sp1"
-sampling.issue <- c("none", "error", "geog", "bias")[1]
+sampling.issue <- c("none", "error", "geog", "bias")[4]
 
 # load workspace
 pkgs <- c("tidyverse", "magrittr")
@@ -28,12 +28,12 @@ lam.df <- readRDS(paste0("out/", sp, "_lam_df.rds")) # env + true pop values
 ########
 ## Set sampling details
 ########
-n_samp <- 2  # number of unique samples to average across
+n_samp <- 5  # number of unique samples to average across
 O_n <- list(Corr=100, Mech=20)  # number of cells in sample
 O_yr <- list(Mx=p$tmax, CA=(p$tmax-10):p$tmax, IPM=p$tmax)  # years to sample
 P.i <- which(lam.df$Surv.S > 0)  # presences: survival past recruit stage
 P.pr <- rep(1, length(P.i))  # pr(sample cell | presence)
-prop.sampled <- 0.8  # proportion of individuals sampled per sampled cell 
+prop.sampled <- 1  # proportion of individuals sampled per sampled cell 
 geog.excl <- which(env.in$x > 10 & env.in$y > 45)
 prop.err <- 0.2
 

### 2018-03-16 code/3_fitModels.R
cleaning messy code

@@ -11,7 +11,7 @@
 ########
 # file specifications
 sp <- "sp1"
-sampling.issue <- c("none", "error", "geog", "bias")[1]
+sampling.issue <- c("none", "error", "geog", "bias")[4]
 modeling.issue <- c("none")[1]
 
 # load workspace
@@ -33,30 +33,25 @@ O_IPM <- readRDS(paste0("out/", sp, "_O_IPM_", sampling.issue, ".rds"))
 ########
 ## Set model details
 ########
-n_sim <- 15  # number of simulations per sample (mechanistic only)
-v <- c("(Intercept)"=0, "size"=0, "size2"=0, "size3"=0, 
+n_sim <- 10  # number of simulations per sample (mechanistic only)
+v <- c("(Intercept)"=0, "size"=0, "size2"=0,# "size3"=0, 
        "temp"=0, "temp2"=0, "prec"=0, "prec2"=0, 
-       "pOpn"=0, "pOth"=0, "pDec"=0, "pEvg"=0, "pMxd"=0)
+       "pOpn"=0, "pOth"=0, "pDec"=0)#, "pEvg"=0, "pMxd"=0)
 m.full <- paste(names(v)[-1], collapse=" + ")
-n_z <- sum(grepl("size", names(v))) + 1  # include intercept
-n_x <- length(v) - n_z
-n_z <- rep(list(n_z), 4)
-names(n_z) <- c("s", "g", "fl", "seed")
-n_x <- rep(list(n_x), 4)
-names(n_x) <- c("s", "g", "fl", "seed")
-X <- list(s=as.matrix(env.in[,1:n_x$s]), g=as.matrix(env.in[,1:n_x$g]),
-          fl=as.matrix(env.in[,1:n_x$fl]), seed=as.matrix(env.in[,1:n_x$seed]))
-
+n_z <- rep(list(sum(grepl("size", names(v))) + 1), 4)# include intercept
+n_x <- rep(list(length(v) - n_z[[1]]), 4)
+names(n_z) <- names(n_x) <- c("s", "g", "fl", "seed")
+X <- map(n_x, ~as.matrix(env.in[,1:.]))
 
 ########
 ## fit models
 ########
-# MaxEnt
-# CA
-# IPM
+##--- MaxEnt
+##--- CA
+##--- IPM
 S.f <- U.f <- vector("list", length(O_IPM))
 for(i in 1:length(O_IPM)) {
-  # separate data from MuMIn::dredge()
+  # separate data for MuMIn::dredge()
[…]

### 2018-03-16 code/4_analysis.R
initial commit

@@ -1,0 +1,11 @@
+# 4: Analyze fit models
+# Comparison of SDM approaches
+# Tim Szewczyk
+
+# This script aggregates and analyzes the output from `3_fitModels.R`.
+
+########
+## Setup
+########
+# file specifications
+sp <- "sp1"

### 2018-03-16 code/5_plots.R
initial commit

@@ -1,0 +1,11 @@
+# 5: Plot output
+# Comparison of SDM approaches
+# Tim Szewczyk
+
+# This script plots the output from `3_fitModels.R` and `4_analysis.R`.
+
+########
+## Setup
+########
+# file specifications
+sp <- "sp1"

### 2018-03-16 code/3_fitModels.R
fixed naming bug

@@ -85,7 +85,7 @@ for(i in 1:length(O_IPM)) {
   p.ipm$s_x <- vars.ls$s[(n_z$s+1):length(v)]
   p.ipm$g_z <- vars.ls$g[1:n_z$g]
   p.ipm$g_x <- vars.ls$g[(n_z$g+1):length(v)]
-  p.ipm$g_sig <- summary(g.opt)$sigma
+  p.ipm$g_sig <- summary(opt.m$g)$sigma
   p.ipm$fl_z <- vars.ls$fl[1:n_z$fl]
   p.ipm$fl_x <- vars.ls$fl[(n_z$fl+1):length(v)]
   p.ipm$seed_z <- vars.ls$seed[1:n_z$seed]

### 2018-03-16 code/fn_aux.R
added function for pulling SDM details from file name

@@ -135,3 +135,11 @@ summarize_IPM_samples <- function(U.f, S.f) {
 antilogit <- function (x) {
   exp(x)/(1 + exp(x))
 }
+
+
+##-- extract SDM sampling & modeling issues from file names
+extract_SDM_issues <- function(f, SDM) {
+  require(stringr)
+  issues <- str_remove(str_split(f, paste0(SDM, "_"), Inf, T)[,2], ".rds")
+  return(str_split(issues, "_"))
+}

### 2018-03-16 code/1_simulateSpecies.R
simplifying code

@@ -59,10 +59,7 @@ n_x <- list(s=length(p$s_x), # n env covariates for each vital rate
             g=length(p$g_x),
             fl=length(p$fl_x), 
             seed=length(p$seed_x))
-X <- list(s=as.matrix(L$env.in[,1:n_x$s]), # env covariates for each vital rate
-          g=as.matrix(L$env.in[,1:n_x$g]),
-          fl=as.matrix(L$env.in[,1:n_x$fl]),
-          seed=as.matrix(L$env.in[,1:n_x$seed]))
+X <- map(n_x, ~as.matrix(L$env.in[,1:.]))  # env covariates for each vital rate
 sdd.pr <- sdd_set_probs(ncell=n.cell, lc.df=L$env.rct, lc.col=8:12,
                         g.p=list(sdd.max=p$sdd_max, 
                                  sdd.rate=p$sdd_rate, 
@@ -78,10 +75,10 @@ sdd.pr <- sdd_set_probs(ncell=n.cell, lc.df=L$env.rct, lc.col=8:12,
 ########
 # Truth: use assigned slopes to fill IPM matrix
 U <- fill_IPM_matrices(n.cell, buffer=0.75, discrete=1, p, n_z, n_x, 
-                       X, sdd.pr, L$env.in$id, verbose=T)
+                       X, sdd.pr, L$env.in$id)
 
 # Realization: generate simulated data
-S <- simulate_data(n.cell, U$lo, U$hi, p, X, n_z, sdd.pr, U$sdd.j, verbose=T)
+S <- simulate_data(n.cell, U$lo, U$hi, p, X, n_z, sdd.pr, U$sdd.j)
 
 # Aggregate results
 lam.df <- L$env.in %>%

### 2018-03-16 code/4_analysis.R
aggregating IPM output

@@ -9,3 +9,18 @@
 ########
 # file specifications
 sp <- "sp1"
+
+# load workspace
+pkgs <- c("tidyverse", "magrittr", "stringr")
+suppressMessages(invisible(lapply(pkgs, library, character.only=T)))
+source("code/fn_IPM.R"); source("code/fn_aux.R"); source("code/fn_sim.R")
+lam.df <- readRDS(paste0("out/", sp, "_lam_df.rds"))
+f_IPM <- list.files("out", pattern=paste0(sp, "_P_IPM"), full.names=T)
+i_IPM <- extract_SDM_issues(f_IPM, "IPM")
+out_IPM <- map2(f_IPM, i_IPM, ~readRDS(.x) %>% 
+                  mutate(SDM="IPM", s_Iss=.y[1], m_Iss=.y[2])) %>%
+  do.call("rbind", .) %>%
+  full_join(select(lam.df, -c(10:15,17)), ., by="id.inbd")
+
+write.csv(out_IPM, paste0("out/", sp, "_IPM.csv"))
+

### 2018-03-16 code/1_simulateSpecies.R
testing landscape sizes

@@ -19,8 +19,8 @@ pkgs <- c("gbPopMod", "tidyverse", "magrittr")
 suppressMessages(invisible(lapply(pkgs, library, character.only=T)))
 source("code/fn_IPM.R"); source("code/fn_aux.R"); source("code/fn_sim.R")
 L <- build_landscape(f=env.f, 
-                     x_max=30, # ncol in landscape; Inf for full dataset
-                     y_max=30) # nrow in landscape; Inf for full dataset
+                     x_max=40, # ncol in landscape; Inf for full dataset
+                     y_max=40) # nrow in landscape; Inf for full dataset
 n.cell <- sum(L$env.rct$inbd)
 
 

### 2018-03-16 code/2_sampleDistribution.R
added noise for IPM

@@ -12,7 +12,7 @@
 ########
 # file specifications
 sp <- "sp1"
-sampling.issue <- c("none", "error", "geog", "bias")[4]
+sampling.issue <- c("none", "noise", "geog", "bias")[4]
 
 # load workspace
 pkgs <- c("tidyverse", "magrittr")
@@ -22,6 +22,7 @@ env.in <- readRDS(paste0("out/", sp, "_env_in.rds"))
 n.cell <- nrow(env.in)
 p <- readRDS(paste0("out/", sp, "_p.rds"))
 S <- readRDS(paste0("out/", sp, "_S.rds"))
+U <- readRDS(paste0("out/", sp, "_U.rds"))
 lam.df <- readRDS(paste0("out/", sp, "_lam_df.rds")) # env + true pop values
 
 
@@ -34,8 +35,14 @@ O_yr <- list(Mx=p$tmax, CA=(p$tmax-10):p$tmax, IPM=p$tmax)  # years to sample
 P.i <- which(lam.df$Surv.S > 0)  # presences: survival past recruit stage
 P.pr <- rep(1, length(P.i))  # pr(sample cell | presence)
 prop.sampled <- 1  # proportion of individuals sampled per sampled cell 
-geog.excl <- which(env.in$x > 10 & env.in$y > 45)
-prop.err <- 0.2
+geog.excl <- which(env.in$x > 20 & env.in$y > 45)
+noise <- list(Mx=0.2, # proportion of observed presences that are false
+            CA=NA,
+            IPM=list(s=0.1,  # proportion of incorrectly assessed surv
+                     g=0.2,  # sizeNext.obs = SizeNext.true + rnorm(0,SD) 
+                     fl=0.1,  # proportion of incorrectly assessed fl
+                     seed=10)  # seed.obs = seed.true + rnorm(0,SD)
+            )
 
 
 ########
@@ -81,15 +88,25 @@ for(s in 1:n_samp) {
 }
 
 # add error
-if(sampling.issue=="error") {
+if(sampling.issue=="noise") {
   # MaxEnt: substitute in some % false presences
-  n.err <- prop.err*O_n$Corr
+  n.err <- noise$Mx*O_n$Corr
   for(s in 1:n_samp) {
     Corr.sample[[s]][1:n.err] <- sample((1:n.cell)[-P.i], n.err, F)
   }
[…]

### 2018-03-16 code/5_plots.R
idea for plot

@@ -9,3 +9,8 @@
 ########
 # file specifications
 sp <- "sp1"
+
+
+# barplots showing % correct for presence/absence; fill=SDM; facet=issues
+ggplot(out_IPM, aes(x=sign(Surv.S.f) == sign(Surv.S), fill=s_Iss)) + 
+  geom_bar(position="dodge")

### 2019-01-03 code/fn_fit.R
avoiding loading dismo bc conflicts

@@ -208,17 +208,17 @@ add_misDisperse <- function(p.mod, p, sdd_max_adj=2, sdd_rate_adj=.1, ldd=5) {
 #'   columns for cell information, mean(probability of presence), and
 #'   sd(probability of presence); and [["diag"]] containing diagnostics
 fit_MxE <- function(sp, issue, samp.issue, lam.df, v) {
-  library(dismo); library(here); library(tidyverse); library(raster)
+  library(here); library(tidyverse)
   path_iss <- paste0("out/maxent/", sp, "/", issue, "/")
   # load observations
   O_Mx <- readRDS(here("vs", sp, paste0("O_Mx_", samp.issue, ".rds")))
   X.Mx <- lam.df[,names(lam.df) %in% v]
   temp.rast <- vector("list", ncol(X.Mx))
   for(vi in 1:ncol(X.Mx)) {
-    temp.rast[[vi]] <- rasterFromXYZ(cbind(lam.df[,c("lon","lat")], X.Mx[,vi]))
+    temp.rast[[vi]] <- raster::rasterFromXYZ(cbind(lam.df[,c("lon","lat")], X.Mx[,vi]))
   }
   names(temp.rast) <- names(X.Mx)
-  rast.Mx <- stack(temp.rast)
+  rast.Mx <- raster::stack(temp.rast)
 
   bias_type="none"
   MxE.f <- MxE.p <- vector("list", length(O_Mx))
@@ -228,9 +228,11 @@ fit_MxE <- function(sp, issue, samp.issue, lam.df, v) {
                 "outputformat=logistic")
   for(i in seq_along(O_Mx)) {
     if(!dir.exists(paste0(path_iss, i))) dir.create(paste0(path_iss, i), recursive=T)
-    MxE.f[[i]] <- maxent(x=rast.Mx, p=as.matrix(lam.df[O_Mx[[i]], c("lon", "lat")]),
-                        args=fit.args, path=paste0(path_iss, i))
-    MxE.p[[i]] <- mean(predict(MxE.f[[i]], rast.Mx, args="outputformat=logistic"))
+    MxE.f[[i]] <- dismo::maxent(x=rast.Mx, 
+                                p=as.matrix(lam.df[O_Mx[[i]], c("lon", "lat")]),
+                                args=fit.args, path=paste0(path_iss, i))
+    MxE.p[[i]] <- mean(dismo::predict(MxE.f[[i]], rast.Mx, 
[…]

### 2019-01-03 code/fn_fit.R
no LC categories to aggregate across

@@ -439,8 +439,9 @@ fit_CA <- function(sp, sp_i, samp.issue, mod.issue, p, env.rct, env.rct.unsc,
     N.init[X.CA$id[X.CA$inbd], p.CA$m] <- N_init
     N.init[X.CA$id[X.CA$inbd], -p.CA$m] <- round(N_init/5)
     for(s in 1:n_sim) {
-      sim.ls[[s]] <- gbPopMod::run_sim(n.grid, n.cell, p.CA, X.CA, sdd.pr, N.init,
-                                       NULL, T, (-1:0)+p.CA$tmax, 1e6)
+      sim.ls[[s]] <- gbPopMod::run_sim(n.grid, n.cell, p.CA, X.CA, sdd.pr, 
+                                       N.init, NULL, T, (-1:0)+p.CA$tmax, 
+                                       1e6, dem_out=TRUE, FALSE)
     }
     
     i_pad <- str_pad(i, 2, pad="0")

### 2019-01-03 code/1_simulateSpecies.R
rounding

@@ -41,8 +41,8 @@ n.cell <- sum(L$env.rct$inbd)
 p <- fit_PNAS_species(sp, env.f, nlcd.sp, clim_X, FALSE, max_z_pow, habitat,
                       x_min, x_max, y_min, y_max)
 p$n <- 10
-p$tmax <- 120
-p$tnonEq <- p$tmax/3
+p$tmax <- 80
+p$tnonEq <- floor(p$tmax/3)
 p$n0 <- 10
 p$prop_init <- 0.001
 p$NDD <- T

### 2019-01-03 code/1_simulateSpecies.R
if(plots)

@@ -141,51 +141,53 @@ lam.gg <- ggplot(lam.df, aes(x=lon, y=lat)) + theme_bw() +
   scale_fill_viridis(name="", option="B") +
   ggtitle(paste0(sp, ": 3km x 3km, favorable habitat"))
 
-lam.gg + geom_tile(aes(fill=log(lambda))) + 
-  labs(subtitle="log(lambda)") +
-  geom_point(data=lam.df[N_init>0,], colour="white", shape=1)
-lam.gg + geom_tile(aes(fill=log(Surv.S))) + 
-  labs(subtitle=paste("log(N): year", p$tmax)) +
-  geom_point(data=lam.df[N_init>0,], colour="white", shape=1)
-lam.gg + geom_tile(aes(fill=log(Surv.S_nonEq))) + 
-  labs(subtitle=paste("log(N): year", p$tmax/3)) +
-  geom_point(data=lam.df[N_init>0,], colour="white", shape=1)
-
-lam.gg + geom_tile(aes(fill=log(nSeed))) + 
-  labs(subtitle="log(Seed production)") +
-  geom_point(data=lam.df[N_init>0,], colour="white", shape=1)
-lam.gg + geom_tile(aes(fill=nSeed/nRepro)) + 
-  labs(subtitle="Per capita seed production") +
-  geom_point(data=lam.df[N_init>0,], colour="white", shape=1)
-lam.gg + geom_tile(aes(fill=log(round(D)))) + 
-  labs(subtitle="log(Immigrant seeds)") +
-  geom_point(data=lam.df[N_init>0,], colour="white", shape=1)
-lam.gg + geom_tile(aes(fill=log(nSdStay+round(D)))) + 
-  labs(subtitle="log(Propagule pressure)") +
-  geom_point(data=lam.df[N_init>0,], colour="white", shape=1)
-lam.gg + geom_tile(aes(fill=log(B))) + 
-  labs(subtitle="log(Seed bank)") +
-  geom_point(data=lam.df[N_init>0,], colour="white", shape=1)
-
-lam.gg + geom_tile(aes(fill=lambda>1)) +  
-  labs(subtitle="lambda > 1") +
-  scale_fill_manual("", values=c("gray30", "dodgerblue")) +
-  geom_point(data=lam.df[N_init>0,], colour="white", shape=1)
-lam.gg + geom_tile(aes(fill=Surv.S>0)) + 
-  labs(subtitle="N > 0") +
[…]

### 2019-01-03 code/3_fitModels.R
correcting variable order

@@ -38,24 +38,24 @@ clim_X_wrong <- paste0("bio10_", c(1, 5, 12, "prMay")) %>%
 issue_i <- read.csv("data/issues.csv", stringsAsFactors=F)
 vars_correct <- rep(0, length(clim_X_correct)*2+4)
 names(vars_correct) <- c("(Intercept)", "size", "size2", "size3",
-                         clim_X_correct, paste0(clim_X_correct, "_sq"))
+                         paste0(rep(clim_X_correct, each=2), c("", "_sq")))
 vars_correct <- vars_correct[-4]
 vars_wrong <- rep(0, length(clim_X_wrong)*2+4)
 names(vars_wrong) <- c("(Intercept)", "size", "size2", "size3",
-                         clim_X_wrong, paste0(clim_X_wrong, "_sq"))
+                       paste0(rep(clim_X_wrong, each=2), c("", "_sq")))
 vars_wrong <- vars_wrong[-4]
 out.dir <- paste0("out/", sp_i$Num)
 if(!dir.exists(here(out.dir))) dir.create(here(out.dir), recursive=T)
 
 
 p.c <- makeCluster(n_core_iss); registerDoSNOW(p.c)
-foreach(i=seq_along(issue_i$Issue), .packages=pkgs) %dopar% {
+foreach(i=seq_along(issue_i$Issue), .packages=c("dismo", pkgs)) %dopar% {
   # load issues
   issue <- issue_i$Issue[i]
   samp_iss <- issue_i$Sampling[i]
   mod_iss <- issue_i$Modeling[i]
   
-  # set model details
+  # set variables & model formulas
   v <- m <- list(CA=NULL, IPM=NULL)
   if(mod_iss=="wrongCov") {
     vars <- vars_wrong

### 2019-01-03 code/3_fitModels.R
not loading dismo bc conflicts

@@ -10,14 +10,14 @@
 ## Setup
 ########
 # file specifications
-sp <- c("barberry", "garlic_mustard")[2]
+sp <- c("barberry", "garlic_mustard")[1]
 overwrite <- TRUE
 n_core_iss <- 2  # number of issues to run in parallel
-n_core_obs <- 2  # number of simulations to run in parallel for each issue
+n_core_obs <- 4  # number of simulations to run in parallel for each issue
 n_sim <- 1 # number of simulations per sample (mechanistic only)
 
 # load workspace
-pkgs <- c("dismo", "gbPopMod", "tidyverse", "magrittr", "MuMIn", "here", "doSNOW")
+pkgs <- c("gbPopMod", "tidyverse", "magrittr", "MuMIn", "here", "doSNOW")
 suppressMessages(invisible(lapply(pkgs, library, character.only=T)))
 walk(dir("code", "fn", full.names=T), source)
 sp_i <- read.csv("data/species_3km.csv") %>% filter(Name==sp)

### 2019-01-03 code/3_fitModels.R
running maxent

@@ -72,17 +72,23 @@ foreach(i=seq_along(issue_i$Issue), .packages=c("dismo", pkgs)) %dopar% {
   v$IPM <- vars[c(1, v.size, v.i)]
   m$IPM <- paste(names(v$IPM)[-1], collapse=" + ")
   n_z <- rep(list(1+length(v.size)), 4)  # adds intercept
-  n_x <- rep(list(length(v.i)), 4)
-  names(n_z) <- names(n_x) <- c("s", "g", "fl", "seed")
+  names(n_z) <- c("s", "g", "fl", "seed")
+  if(is.null(p$germ_x)) {
+    n_x <- rep(list(length(v.i)), 5)
+    names(n_x) <- c(names(n_z), "germ")
+  } else {
+    n_x <- rep(list(length(v.i)), 4)
+    names(n_x) <- names(n_z)
+  }
   
   # fit MaxEnt
-  # if(issue %in% issue_i$Issue[c(1:4,8:9)]) {
-  #   P_MxE <- fit_MxE(sp_i$Num, issue, samp_iss, lam.df, v$Mx)
-  #   if(overwrite) {
-  #     saveRDS(P_MxE$diag, here(out.dir, paste0("Diag_MxE_", issue, ".rds")))
-  #     saveRDS(P_MxE$P_MxE, here(out.dir, paste0("P_MxE_", issue, ".rds")))
-  #   }
-  # }
+  if(issue %in% issue_i$Issue[c(1:4,8)]) {
+    P_MxE <- fit_MxE(sp_i$Num, issue, samp_iss, lam.df, v$Mx)
+    if(overwrite) {
+      saveRDS(P_MxE$diag, here(out.dir, paste0("Diag_MxE_", issue, ".rds")))
+      saveRDS(P_MxE$P_MxE, here(out.dir, paste0("P_MxE_", issue, ".rds")))
+    }
+  }
   # fit CA-demographic
   P_CA <- fit_CA(sp, sp_i, samp_iss, mod_iss, p, 
                  L$env_rct, L$env_rct_unscaled, lam.df, v$CA, m$CA, 
@@ -93,8 +99,8 @@ foreach(i=seq_along(issue_i$Issue), .packages=c("dismo", pkgs)) %dopar% {
     # saveRDS(P_CA$P_CAl, here(out.dir, paste0("P_CAl_", issue, ".rds")))
   }
   # fit IPM, CA-individual
[…]

### 2019-01-03 code/5_plots.R
updated issue names

@@ -8,7 +8,7 @@
 ## Setup
 ########
 # file specifications
-sp <- "sp2"
+sp <- "sp1"
 
 # load workspace
 pkgs <- c("tidyverse", "magrittr", "stringr", "here", "viridis")
@@ -18,11 +18,11 @@ theme_set(theme_bw())
 lam.df <- readRDS(here("vs", sp, "lam_df.rds"))
 out <- read.csv(here("out", sp, "out.csv"))
 out$issue <- factor(out$issue, 
-                    levels=c("none", "noise", "geogBias", "sampBias",
-                             "noSB", "underDisp", "overDisp", "clim", "lc"),
-                    labels=c("None", "Measurement error", "Geographic bias", 
-                             "Sampling bias", "No Seedbank", "Under dispersal",
-                             "Over dispersal", "Climate Only", "LC only"))
+                    levels=c("none", "noise", "sampBias", "nonEq",
+                             "noSB", "underDisp", "overDisp", "wrongCov"),
+                    labels=c("None", "Measurement error", "Sampling bias", 
+                             "Non-equilibrium", "No Seedbank", "Under dispersal",
+                             "Over dispersal", "Incorrect covariates"))
 SDM_col <- c(MxE="#3f007d", IPM="#014636", CAi="#02818a", CAd="#67a9cf")
 
 par(mfrow=c(3,3))

### 2019-01-03 code/fn_sim.R
simplifying

@@ -166,10 +166,10 @@ simulate_data <- function(n.cell, lo, hi, p, X, n_z, sdd.ji, p.ji, N_init, sp,
   }
   
   # does germination probability vary with environment?
-  if(!is.null(X$germ)) { 
-    pr_germ <- c(antilogit(X$germ %*% p$germ_x))
-  } else { 
+  if(is.null(X$germ)) { 
     pr_germ <- NULL 
+  } else { 
+    pr_germ <- c(antilogit(X$germ %*% p$germ_x))
   }
   
   # storage objects

### 2019-01-03 code/gis_processing.R
species observations for sampBias

@@ -148,6 +148,19 @@ grd.df$rd_len[rd.grd$layer] <- rd.grd$length
 
 
 
+##------ Species observation locations
+## Geolocated presences from Allen & Bradley 2016. Used as a measure of the 
+## probability of sampling in a given location.
+##------
+library(sf)
+spp_obs <- read.csv("data/AllenBradley2016/IAS_occurences_final_analysis.csv") %>%
+  st_as_sf(coords=c("LONGITUDE_DECIMAL", "LATITUDE_DECIMAL")) %>%
+  st_set_crs(4326) %>%
+  st_transform(crs=alb_CRS@projargs) %>%
+  st_crop(., extent(ENF.us))
+
+
+
 write_csv(grd.df, paste0("data/ENF_", cell_side/1000, "km.csv"))
 
 

### 2019-01-03 code/gis_processing.R
prMay

@@ -57,7 +57,9 @@ grd.wgs <- spTransform(grd, wgs_CRS)
 grd.df <- as.data.frame(coordinates(grd))
 names(grd.df) <- c("long", "lat")
 
-
+clim.rast <- raster("data/climate/CHELSA_prec_5_V1.2_land.tif")
+# run crop(), projectRaster() below
+clim.df <- zonal.stats(grd, clim.rast, mean, trace=F, plot=F)
 
 ##------ Climate data
 ## Climate raster files are stored in climate/ with no other files. Each file
@@ -66,7 +68,7 @@ names(grd.df) <- c("long", "lat")
 ## from CHELSA <http://chelsa-climate.org/downloads/>
 ##------
 # Load and extract bioclimatic variable numbers
-clim.f <- dir("data/climate")
+clim.f <- dir("data/climate", "bio10")
 clim.var <- str_split_fixed(clim.f, pattern="[[:punct:]]", 4)[,2:3] %>% 
   apply(1, str_flatten, collapse="_")
 clim.rast <- stack(paste0("data/climate/", clim.f))
@@ -148,7 +150,7 @@ grd.df$rd_len[rd.grd$layer] <- rd.grd$length
 
 
 
-write_csv(grd.df, paste0("data/ENF_", cell_side/1000, "km.csv"))
+write_csv(grd.df, paste0("data/ENF_", cell_side/1000, "km_pr_may.csv"))
 
 
 

### 2019-01-03 background/exploration.Rmd

--- 2018-03-14 exploration.Rmd
@@
 ```{r landscape}
-x_max <- 100
-y_max <- 100
+x_max <- 20
+y_max <- 20
 
@@
          inbd=!is.na(match(.$x_y, lc.df$x_y)),
-         id=row_number(), 
-         id.inbd=min_rank(na_if(inbd*id, 0)),
          lat=lc.df$top[match(.$x_y, lc.df$x_y)],
-         lon=lc.df$left[match(.$x_y, lc.df$x_y)])
+         lon=lc.df$left[match(.$x_y, lc.df$x_y)]) %>%
+  filter(y >= (max(.$y) - y_max) & x <= x_max) %>%
+  mutate(id=row_number(), 
+         id.inbd=min_rank(na_if(inbd*id, 0)))
 env.rct[is.na(env.rct)] <- 0
@@
          pEvg=c(scale(pEvg)), pMxd=c(scale(pMxd)))
-rm(env.rct)
+n.cell <- sum(env.rct$inbd)
 ```
@@
 tmax <- 10
-cells <- n_distinct(lc.df$CellID)
-n.cell <- length(cells)
 sp_i <- list(p=list(s_z=c(-8, 2.1, -.09), # b1 + b2*z + b3*z^2
@@
 points(sizeNext~size, data=S_d, col=pt.col, pch=20)
-
 ```
@@
 ----
+
+# Exploration of options for the sampling component
+
+```{r loadSpecies}
+env.in <- readRDS("out/sp1_env_in.rds")
+n.cell <- nrow(env.in)
+p <- readRDS("out/sp1_p.rds")
+sdd.pr <- readRDS("out/sp1_sdd.rds")
+U <- readRDS("out/sp1_U.rds")
+S <- readRDS("out/sp1_S.rds")
+lam.df <- readRDS("out/sp1_lam_df.rds")
+```
 
@@
 ```{r O_setup}
-O_n <- list(Mx=500, CA=20, IPM=20) # number of cells in sample
-O_yr <- list(Mx=tmax, CA=(tmax-10):tmax, IPM=tmax) # years to sample from
+O_n <- list(Corr=100, Mech=20) # number of cells in sample
+O_yr <- list(Mx=p$tmax, CA=(p$tmax-10):p$tmax, IPM=p$tmax) # years to sample from
 P.i <- which(lam.df$Surv.S > 0) # presences: surviving individuals > seedlings
+Mech.sample <- sample(P.i, O_n$Mech, replace=FALSE)
+Corr.sample <- sample(P.i, O_n$Corr, replace=FALSE)
 n_sim <- 15
@@
 ```{r O_Mx}
-library(dismo)
-Mx.P <- sample(P.i, O_n$Mx, replace=F)
-pres <- 1:nrow(lam.df) %in% Mx.P
-Mx.out <- maxent(x=lam.df[,c(1,3,5:9)], p=pres)
-lam.df$MxPred <- 1:nrow(lam.df) %in% row.names(Mx.out@presence)
-lam.df$MxCorrect <- with(lam.df, (MxPred==1 & lambda > 1) |
-                           (MxPred==0 & lambda < 1))
-ggplot(lam.df, aes(x=lon, y=lat, fill=MxCorrect)) + geom_tile()
+pres <- 1:nrow(lam.df) %in% Corr.sample
 ```
@@
 ```{r O_IPM}
-IPM.P <- sample(P.i, O_n$IPM, replace=FALSE)
-IPM.d <- vector("list", O_n$IPM)
+IPM.d <- vector("list", O_n$Mech)
 # sample individuals from each cell rather than full census -- proportion?
-for(i in IPM.P) {
+for(i in Mech.sample) {
   IPM.d[[i]] <- data.frame(S$d[[i]]) %>% filter(yr %in% O_yr$IPM) %>%
     add_column(id.inbd=i) %>%
-    full_join(env.df[i,], by="id.inbd")
+    full_join(env.in[i,], by="id.inbd")
 }
@@
 
+# Exploration of options for the fitting component
+
 ## Fitting the models
@@
 ```{r P_Mx}
-
+Mx.out <- maxent(x=lam.df[,c(1,3,5:9)], p=pres)
+lam.df$MxPred <- 1:nrow(lam.df) %in% row.names(Mx.out@presence)
+lam.df$MxCorrect <- with(lam.df, (MxPred==1 & lambda > 1) |
+                           (MxPred==0 & lambda < 1))
+ggplot(lam.df, aes(x=lon, y=lat, fill=MxCorrect)) + geom_tile()
 ```
@@
 
+########
+## comparison with truth
+########
 IPM.fit.sum <- lam.df %>% 

### 2019-01-04 code/fn_fit.R
revised LDD misspecification

@@ -363,10 +363,10 @@ fit_CA <- function(sp, sp_i, samp.issue, mod.issue, p, env.rct, env.rct.unsc,
   # impose dispersal issues
   if(mod.issue=="underDisp") {
     p.CA <- add_misDisperse(p.CA, p, sdd_max_adj=-2, sdd_rate_adj=2, 
-                            ldd=round(p$ldd/5))
+                            ldd=round(p$ldd/2))
   } else if(mod.issue=="overDisp") {
     p.CA <- add_misDisperse(p.CA, p, sdd_max_adj=2, sdd_rate_adj=.5, 
-                            ldd=p$ldd*5)
+                            ldd=p$ldd*2)
   }
   if(grepl("Disp", mod.issue)) {
     sdd.pr <- sdd_set_probs(ncell=n.cell, lc.df=env.rct.unsc, 

### 2019-01-04 code/1_simulateSpecies.R
testing quadratic terms

@@ -11,17 +11,18 @@
 ## Setup
 ########
 # file specifications
-sp <- c("barberry", "garlic_mustard")[2]
+sp <- c("barberry", "garlic_mustard")[1]
 res <- "10km" # "3km", "5km", "10km", "50km"
 overwrite <- TRUE
+plots <- TRUE
 clim_X <- paste0("bio10_", c(5, "prMay"))
 habitat <- 3
 max_z_pow <- 1
 n.cores <- 4
-x_min <- 200#675#
+x_min <- 0#200#675#
 x_max <- Inf
 y_min <- 0
-y_max <- 75#250#
+y_max <- Inf#75#250#
 
 # load workspace
 pkgs <- c("gbPopMod", "tidyverse", "magrittr", "here", "doSNOW", "foreach")
@@ -40,6 +41,11 @@ n.cell <- sum(L$env.rct$inbd)
 ########
 p <- fit_PNAS_species(sp, env.f, nlcd.sp, clim_X, FALSE, max_z_pow, habitat,
                       x_min, x_max, y_min, y_max)
+p$s_x[1] <- p$s_x[1]*50
+p$s_x[c(2,4)] <- c(-0.8, -0.5)
+p$g_x[c(2,4)] <- -0.1
+p$germ_x[2] <- p$germ_x[2]*-1
+p$germ_x[c(3,5)] <- c(-2, -0.1)
 p$n <- 10
 p$tmax <- 80
 p$tnonEq <- floor(p$tmax/3)

### 2019-12-13 ms/ms_latex/ms_sdmComp.tex
returned to latex for equations

%% This is file `elsarticle-template-1-num.tex',
%%
%% Copyright 2009 Elsevier Ltd
%%
%% This file is part of the 'Elsarticle Bundle'.
%% ---------------------------------------------
%%
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%%
%% Template article for Elsevier's document class `elsarticle'
%% with numbered style bibliographic references
%%
%% $Id: elsarticle-template-1-num.tex 149 2009-10-08 05:01:15Z rishi $
%% $URL: http://lenova.river-valley.com/svn/elsbst/trunk/elsarticle-template-1-num.tex $
%%
\documentclass[preprint,review,times,12pt]{elsarticle}

%% Use the option review to obtain double line spacing
%% \documentclass[preprint,review,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,1p,times]{elsarticle}
%% \documentclass[final,1p,times,twocolumn]{elsarticle}
%% \documentclass[final,3p,times]{elsarticle}
%% \documentclass[final,3p,times,twocolumn]{elsarticle}
%% \documentclass[final,5p,times]{elsarticle}
%% \documentclass[final,5p,times,twocolumn]{elsarticle}

\usepackage[left=1in, right=1in, top=1in, bottom=1in]{geometry}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{lineno}
\usepackage{caption}
\usepackage{natbib}

\journal{Ecological Modelling}
\setcitestyle{authoryear,round,semicolon,sort}
\bibliographystyle{apalike}
\begin{document}
\begin{frontmatter}

\title{The performance of presence-based and process-based species distribution models}

\author[NREN,CS]{Tim M. Szewczyk}
\author[CS]{Marek Petrik}
\author[NREN]{Jenica M. Allen}

\address[NREN]{Department of Natural Resources and the Environment, University of New Hampshire}
\address[CS]{Department of Computer Science, University of New Hampshire}

\begin{abstract}
Abstract here
\end{abstract}

\begin{keyword}
keywords
\end{keyword}

\end{frontmatter}
\linenumbers

\section{Introduction}
\label{S:1}
% SDMs and the prevalence of occurrenc-based methods (And)
1. General introduction about species distribution models and the historical prevalence of occurrence-based SDMs.

% Process-based SDMs as an alternative, including a specific example or two (But)
2. The rise of process-based SDMs (or at least the rise in advocacy for process-based SDMs), including some specific examples of their use, since that was the main question raised at ISEM.

% Expected benefits of process-based SDMs (And)
3. Expected benefits of process-based SDMs

% Benefits of occurrence-based, and the unknowns of process-based performance (But) 
4. Potential downsides of process-based SDMs

% Mission statement (Therefore)
We evaluate the performance of an occurrence-based SDM and three process-based SDMs for two virtual species. We investigate the ability of each SDM  to predict the true distributions under ideal conditions, as well as under various realistic data and modelling scenarios. Specifically, we ask: 1) Do process-based models outperform MaxEnt overall? 2) are process-based or occurrence-based SDMs more susceptible to particular data deficiencies? And 3) do common modelling errors negate any improved performance observed in the process-based SDMs?

\section{Methods}
% General outline of the process (?)
% Range boundary definitions
% Virtual species simulation
% Virtual ecologist simulation
% SDM methods we compared

\label{S:2}
\begin{figure}
	\centering\includegraphics[width=5in]{../../figs/process_outline.png}
	\caption{\label{fig:outline} Simulation, sampling, and prediction process. Distributions were simulated for virtual species based on environmental effects on vital rates and demographic processes.}
\end{figure}
% General outline
We evaluated the performance of four species distribution models (SDMs) in predicting the ranges of two simulated species under a variety of data quality and modelling scenarios (Fig. 1), including one occurrence-based method (MaxEnt) and three process-based methods (Integral Projection Model: IPM; individual-level cellular automata: CA\textsubscript{i}; and population-level cellular automata: CA\textsubscript{p}). IPMs predict a deterministic intrinsic growth rate $\lambda$ in each cell of the landscape based on regressions of vital rates with environmental variables and individual sizes. In contrast, CA models are simulation-based and explicitly model spatio-temporal dynamics among cells. The CA\textsubscript{i} model simulates the growth, survival, and fecundity of individuals, while the CA\textsubscript{p} model uses population averages to summarize similar processes. See additional details for the process-based models in Appendix 1.

\subsection{Virtual Species Simulation}
We simulated two virtual species in the eastern United States, based approximately on Japanese barberry (\emph{Berberis thunbergii}), an invasive shrub with bird-dispersed seeds, and garlic mustard (\emph{Alliaria petiolata}), an invasive biennial with seeds dispersed by non-volant animals and water. Both species have previously been the focus of process-based SDMs (CITE). To simulate each species, we defined relationships between environmental variables, individual size, and vital rates (Fig. \ref{fig:outline}, Appendix 1) on a gridded landscape spanning the Eastern Temperate Forest and Northern Forest ecozones within the United States ($\sim$5x5km resolution: 115,105 cells, CITE). Vital rates included annual individual growth, annual survival, flowering probability, seed production, and germination probability. For environmental covariates, we used the minimum temperature of the coldest month and the precipitation in May, following previous work (CITE PNAS, Chelsa). Based on these regressions, we calculated the intrinsic growth rate $\lambda$ in each cell following the methods used for IPMs, where each cell was treated as a population. Cells with $\lambda > 1$ were considered to be capable of containing persistent populations, constituting the first definition of the true distribution for each species.

To simulate populations, we initialized 10 random cells with 10 individuals, with individual sizes drawn from a uniform distribution constrained by the allowable range for each species. Then, we calculated the expected survival probability, growth rate, flowering probability, and seed production for each individual according to the vital rate regressions (Appendix 1). Then, we drew random values from the appropriate distributions for each vital rate for each individual, implemented short- and long-distance dispersal of seeds according to the dispersal mode of each species, and generated new individuals in each cell from germinating seeds. Populations in each cell were subject to density dependence through reduced seedling establishment (CITE) after the abundance exceeded a predetermined threshold. The populations were simulated through 300 years, at which point the distribution appeared stable, indicating that equilibrium had been reached. The cells with $N > 0$ in the final year thus constituted the second definition of the true distribution for each species.

\subsection{Sampling and Modelling Scenarios}
We evaluated four sampling scenarios: \emph{ideal} (year 300; uniform sampling probability among occupied cells), \emph{non-equilibrium} (year 100; uniform sampling probability among occupied cells), \emph{geographic bias} (year 300; sampling probability $\propto$ human and road density in occupied cells), and \emph{measurement error} (year 300; MaxEnt: 3\% of observations drawn from full landscape regardless of occupancy; process-based: error added to growth, seed production, and abundance observations). For each species, we generated 100 sets of sampled cells for each scenario. For MaxEnt, each sample consisted of 300 cells (CITE), while for the process-based SDMs, each sample consisted of 25 cells (CITE). For process-based models, a maximum of 1000 (CA\textsubscript{p}) or 100 (CA\textsubscript{i}, IPM) individuals were sampled from within each selected grid cell to reflect real-world logistical limitations (Appendix 1, https://github.com/Sz-Tim/sdmMethodComp).

In addition, we assessed the effect of four scenarios of modeling misspecification: \emph{incorrect covariates}, \emph{no seed bank} (process-based only), \emph{over-estimated dispersal} (process-based only), \emph{under-estimated dispersal} (process-based only). Each of these misspecifications was implemented using the \emph{ideal} datasets. For \emph{incorrect covariates}, the SDMs were fit with correlated but more general climate variables (mean annual temperature, annual precipitation) rather than the more specific covariates used in the generative process and for all other scenarios (minimum temperature of the coldest month, precipitation in May).

For each scenario (Fig \ref{fig:outline}), the SDMs were parameterized using the corresponding sampled datasets. To select environmental covariates for each vital rate regression for each process-based SDMs, we compared all combinations of the climatic variables and their squares with the R package \emph{MuMIn}. For the CA\textsubscript{p} and CA\textsubscript{i} models, simulations were run for 300 years to generate predicted distributions. Using each definition of the true simulated ranges (i.e., $\lambda > 1$ and $N > 0$), we calculated the True Skill Statistic (TSS) for each of the 100 predicted distributions per scenario per SDM. We calculated TSS as $sensitivity + specificity - 1$, where $sensitivity$ is the proportion of true presences that are predicted correctly, and $specificity$ is the proportion of true absences that are predicted correctly, such that TSS ranges from $-1$ (all cells predicted incorrectly) to $1$ (all cells predicted correctly).

\section{Results}
\label{S:3}
% Talk results
% 1. Across all scenarios, IPM is best for lambda, CAi/p are best for N, with MaxEnt in the middle for both
% 2. Under ideal conditions, process-based outperform MaxEnt
% 3. For non-equilibrium distributions, process-based outperform MaxEnt
% 4. Process-based are much, much more sensitive to covariate choice -- implies sensitivity to regression accuracy
% 5. Dispersal is critical for CA models
% 6. Structural errors seem to have erratic effects
% 0. Process-based are better IF you are confident of the process, but MaxEnt is more robust to error and uncertainty

% Possible paragraphs:
% Rankings:
%	Ideal & mean(all scenarios): corresponding process-based is best on average
%	lambda = IPM, MxE, CA (avgs)
%	N = CA, MxE, IPM (avgs)
%	MaxEnt somewhat better for lambda
% Effect of data issues
%	Measurement error
% Effect of modelling issues
% Differences in output

% Rankings (overall, ideal)
All models performed well in recovering the true $\lambda$-based ($\lambda > 1$) and $N$-based distributions ($N>0$) for both species across all scenarios (all TSS medians $>$ 0.77), though no single model performed best universally (Fig. \ref{fig:TSSmn}). Across scenarios applicable to all four SDMs (Table \ref{table:ranks}), the IPM best predicted the $\lambda$-based distributions (mean rank: 1.5) and the CA models best predicted the $N$-based distributions (mean rank: CA\textsubscript{i}=1.6, CA\textsubscript{p}=1.9). MaxEnt was intermediate for both range boundaries on average (mean rank: $\lambda$-based=2.7, $N$-based=2.9)

\begin{table}
	\centering
	\captionsetup{width=.75\textwidth}
	\begin{tabular}{l c c c c}
		\hline \\[-3.5ex]
		\textbf{SDM} & \textbf{Best} & \textbf{Mean} & \textbf{Median} & \textbf{Worst}\\
		\hline 
		\hline \\[-2.5ex]
		\multicolumn{5}{l}{\textsc{True range: $\lambda > 1$}} \\
		CA\textsubscript{p} & 2 & 2.9 & 2.5 & 4\\
		CA\textsubscript{i} & 2 & 2.9 & 3 & 4\\
		IPM & 1 & 1.5 & 1 & 4\\
		MaxEnt & 1 & 2.7 & 2.5 & 4\\
		[.5ex]\hline \\[-2.5ex]
		\multicolumn{5}{l}{\textsc{True range: $N > 0$}} \\
		CA\textsubscript{p} & 1 & 1.9 & 2 & 3\\
		CA\textsubscript{i} & 1 & 1.6 & 1 & 4\\
		IPM & 3 & 3.6 & 4 & 4\\
		MaxEnt & 1 & 2.9 & 3 & 4\\
		[.5ex]\hline \\[-2ex]
	\end{tabular}
	\caption{\label{table:ranks}Summary of ranks across species and all scenarios applicable to all four SDMs. For each scenario and species, SDMs were ranked based on the median TSS.}
\end{table}

% Effect of data issues
For MaxEnt, TSS decreased when fit using samples from non-equilibrium populations, driven by smaller predicted ranges as indicated by a decrease in sensitivity (Fig. SXX:Sensitivity-Ideal) and a smaller increase in specificity (Fig. SXX:Specificity-Ideal). In contrast, non-equilibrium samples had minimal effect on TSS for the CA\textsubscript{p} model, with a modest improvement for the long-lived shrub in the IPM and CA\textsubscript{i} models driven by increases in sensitivity (Fig. SXX:Sensitivity-Ideal). Measurement error and sampling bias had negligible effects on the process-based models, and small to moderate effects on the performance of MaxEnt (Fig. \ref{fig:TSSmn}, Fig. SXX:Sensitivity-Ideal, Fig. SXX:Specificity-Ideal).

\begin{figure}
	\centering\includegraphics[width=5in]{../../figs/fig_TSS_mn+CI.jpg}
	\caption{\label{fig:TSSmn} True skill statistic (TSS) mean and 95\% confidence intervals across 100 sampled datasets for each SDM and scenario, compared to true distributions defined by $\lambda > 1$ and $N > 0$. Scenarios include: no sampling or modeling issues (ideal), sampling issues (measurement error, sampling bias, non-equilibrium), and modeling issues (incorrect covariates, no seed bank, under dispersal, over dispersal).}
\end{figure}

\begin{figure}
	\centering\includegraphics[width=\linewidth]{../../figs/fig_TSSvIdeal.jpg}
	\caption{\label{fig:TSSvIdeal} Effect of scenario on median true skill statistic (TSS) relative to the '\emph{ideal}' scenario. Positive values indicate improved predictive ability, while negative values indicate decreased predictive ability.}
\end{figure}

% Effect of modelling issues
Modeling with incorrect covariates was universally problematic for model performance, though it had a greater impact on TSS in the shrub than the biennial species (Fig. \ref{fig:TSSvIdeal}). Incorrect covariates did not lead to systematic over-prediction or under-prediction, but rather a decrease in both sensitivity and specificity, particularly for the process-based models (Fig. SXX:Sensitivity-Ideal, Fig. SXX:Specificity-Ideal). Excluding the seed bank in the process-based models had minimal effect on TSS for the shrub, but variable effects on TSS for the biennial. Relative to ‘ideal’, all process-based models showed reduced sensitivity and increased specificity for the biennial, indicating a smaller predicted range. The IPM was generally more resistant to mischaracterized dispersal compared to the simulation-based CA\textsubscript{i} and CA\textsubscript{p} models. Underestimation of dispersal parameters led to decreased predicted ranges, and overestimation to increased predicted ranges relative to ‘ideal’. The effects were more extreme for the biennial than for the shrub.

% Process-based predictions (lambda, N, etc)
The process-based SDMs generate predictions for a number of biological quantities of interest. For instance, the IPM predicts $\lambda$ in each cell of the landscape. Discrepancies in predicted $\lambda$ values were biased downward in the northern portion of the study region (FIGURE 4), more strongly so for the shrub. Similarly, the CA models predict the abundance $N$ for each species. Discrepancies in predicted abundance from the CA\textsubscript{p} model showed a spatial banding of upward bias at the range margins for both species (FIGURE 4). Immediately interior to the range margin, the CA\textsubscript{p} and CA\textsubscript{i} model showed upward bias for the shrub (FIGURE 4).

\section{Discussion}
\label{S:4}

% Talk Discussion
% 1. Across all scenarios, IPM is best for lambda, CAi/p are best for N, with MaxEnt in the middle for both
% 2. Under ideal conditions, process-based outperform MaxEnt
% 3. For non-equilibrium distributions, process-based outperform MaxEnt
% 4. Process-based are much, much more sensitive to covariate choice -- implies sensitivity to regression accuracy
% 5. Dispersal is critical for CA models
% 6. Structural errors seem to have erratic effects
% 0. Process-based are better IF you are confident of the process, but MaxEnt is more robust to error and uncertainty

% Overview
This is the first paragraph summarizing some general stuff. The process-based SDMs provide a lot more information, and they do, in fact, perform a bit better on average. They are less impacted by samples from non-equilibrium populations, so that is in line with expectations. However, they are more susceptible to mismatches in the covariates than MaxEnt, and the structural errors in the models don't necessarily have predictable effects.

% Process-based > MaxEnt under ideal conditions, and especially non-equilibrium distributions
Expand on the summary, bringing in past work on non-equilibrium distributions. Since process-based models aim to describe the biological processes that ultimately produce an emergent distribution, it really doesn't matter whether the samples are from a non-equilibrium species as long as the populations contain a sufficiently broad span of ages or sizes. This points to the fact that in parameterizing process-based models, care should be taken to select populations with a mix of ages or sizes rather from a range of environmental conditions than focusing on the geographic distribution per se, in essence capturing more of the parameter space. Thus, for non-equilibrium species, a well-described process-based SDM should do better than an occurrence-based model.
.
% However, process-based are highly sensitive
%	- covariate choice: even highly correlated covariates lead to dramatically decreased ability
%	- dispersal: critical for CA models
% 	- structural errors seem to have erratic effects
However, process-based models appear to be sensitive to several parts of the modelling process. First, they are much more sensitive to covariate choice, and by generalization, to the alignment of the modelled relationship between vital rates and the environment with reality. The flexible structure of MaxEnt, perhaps along with the top-down perspective, makes it much less influenced by errors in the covariates. Even with covariates that are highly correlated with the true covariates, there is still a large decrease in predictive ability for the process-based models.

% So: Why or when to use each?
%	- MaxEnt is probably best for most species: fairly robust, already know about non-equilibrium, etc
%		- But what about climate change? Isn't everything at disequilibrium now? Shit...
%	- is it justifiable to draw conclusions about long-lived vs short-lived species? 
%		- We should have done these simulations with everything else identical but that (climatic preferences, etc)
%		- redo each with just the lifespan changed? But that will probably just make it more complicated....
% Range boundaries: What is each SDM predicting?

% Range boundaries
Range boundaries can be defined in several ways. For the true distribution of each species, we use two definitions: abundance-based (cells with $N > 0$) and $\lambda$-based (cells with $\lambda > 1$). The process-based SDMs compared here each predict a specific type of range boundary. IPMs predict $\lambda$ in each cell of the landscape, while the simulation-based CA models predict abundance. For MaxEnt, the range boundary definition is less explicit. The geo-located presences used as input imply an abundance-based range, but predicted distributions are frequently constrained by a data-based threshold (e.g., relative suitability values that encapsulate 95\% of the observed data) to exclude both erroneous data and sink populations, which implies an effort to predict a $\lambda$-based range. Generally, process-based SDMs explicitly predict one type of range or another, while occurrence-based SDMs may be rather ambiguous. There may or may not be a big difference between the two range definitions, but it will depend on the species. For example, the magnitude of the difference should vary with the number or distribution of marginal cells (lots of marginal cells mean large areas with middling occupancy probabilities), the typical annual stochasticity in the population abundances (more stochasticity means more discrepancies are likely), and dispersal limitation (high dispersal or a long introduction history means fewer areas that have not been reached, but also possibly a larger proportion of sink populations).

\section{Acknowledgments}
This project was funding through National Science Foundation award IIS-1717368.

\section{Bibliography}

\end{document}

### 2019-12-13 ms/supp/App_1.Rmd
returned to latex for equations

---
title: 'Appendix 1: Supplemental Methods'
subtitle: 'The performance of presence-based and process-based species distribution models'
author: 'Tim M. Szewczyk, Marek Petrik, Jenica M. Allen'
output:
  pdf_document: 
    number_sections: yes
    toc: yes
    toc_depth: 1
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 1
editor_options:
  chunk_output_type: inline
header-includes:
  - \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{A.\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{A.\arabic{figure}}}  \usepackage{longtable}  \usepackage{caption}
---

\beginsupplement

---

This appendix contains supplemental methods pertaining to the virtual species generation, species distribution model specifics, and scenario particulars.

---

\section{General model structure}
\subsection{Integral Projection Model overview}
To generate fully-known true distributions for the virtual species, we used the general structure of an Integral Projection Model (IPM) to calculate the intrinsic growth rate $\lambda$ in each cell of the gridded landscape, and adapted the regression-based structure of an IPM into an individual-level, simulation-based cellular automata (CA) model to produce spatiotemporally dynamic abundance distributions. The equations below apply to both models.

IPMs use the size distribution, $z$, of individuals at time $t$, along with a kernel $K(z^{\prime}, z)$, to predict the size distribution, $z^{\prime}$, at time $t+1$:
\begin{equation}
n_{t+1}(z^{\prime}) = \int_{\Omega} K(z^{\prime}, z) n_t(z) dz
\end{equation}

Here, $\Omega$ represents the range of possible sizes for the species or population. The kernel $K(z^{\prime}, z)$ is composed of a growth and survival component, $P(z, z^{\prime})$, representing the fate of individuals from time $t$ to $t+1$, and a fecundity component, $F(z, z^{\prime})$, representing new individuals added between time $t$ and $t+1$. In practice, the integral is approximated using a discretized transition matrix, and the intrinsic growth rate $\lambda$ is calculated as the first eigenvalue of the transition matrix.

The $P$ and $F$ kernels are decomposed further into more mechanistic conditional probabilities and parameters, many of which are functions of the size distribution $z$. For example:
\begin{equation}
\begin{split}
K(z^{\prime}, z) & = P(z^{\prime}, z) + F(z^{\prime}, z) \\
	& = s(z) g(z^{\prime}|z) + p_{flower}(z) f_{seeds}(z) p_{estab} f_{rcrSize}(z^{\prime})
\end{split}
\end{equation}

where $s(z)$ is the survival probability of individuals based on size, $g(z^{\prime}|z)$ is the probability density of size $z^{\prime}$ for an individual of size $z$, $p_{flower}(z)$ is the probability that an individual of size $z$ produces flowers, $f_{seeds}(z)$ is the expected number of seeds produced by an individual of size $z$ given that they flower, $p_{estab}$ is the probability that a seed germinates and establishes as a new recruit, and $f_{rcrSize}(z^{\prime})$ is the expected size distribution of new recruits at time $t+1$. On a gridded  landscape, the population in each cell could be modelled independently using the above structure, with environmental effects incorporated by allowing the environment to influence parameter values (e.g., $f_{seeds}$).

\subsection{Adding a seed bank}
This basic IPM structure is very flexible, allowing for discrete life stages, reproduction-dependent mortality, and environmental covariates. To incorporate a seed bank where seeds that do not germinate between time $t$ and $t+1$ may survive to $t+2$ or beyond, the fecundity kernel is altered, with seed bank $B$, such that:
\begin{align}
n_{t+1}(z^{\prime}) &= B_t s_{rcrB} p_{estab} f_{rcrSize}(z^{\prime}) + \int_{\Omega}[P(z^{\prime}, z)+F(z^{\prime}, z)] n_t(z) dz \\
F(z^{\prime}, z) &= p_{flower}(z)  f_{seeds}(z) s_{rcrDirect} p_{estab} f_{rcrSize}(z^{\prime}) \\
B_{t+1} &= B_t s_{survB} (1-s_{rcrB})  + p_{flower}(z)  f_{seeds}(z) (1-s_{rcrDirect}) s_{survB} n_t(z) dz
\end{align}

where $B_t$ is the number of seeds in the seed bank at time $t$, $s_{rcrB}$ is the probability a seed recruits from the seed bank, $s_{survB}$ is the probability a seed survives in the seed bank from time $t$ to $t+1$, and $s_{rcrDirect}$ is the probability a seed produced in year $t$ germinates between year $t$ and $t+1$. Two notes about the structure and definitions: 1) a seed added to the seed bank must fail to recruit in time $t$ and must also survive from $t$ to $t+1$, and 2) the probability of recruiting is best interpreted as the probability of germinating and is therefore separate from the probability of establishing. Both of these could be defined differently to combine each set of processes, though keeping them separate allows for seeds to perish.

\subsection{Adding dispersal}
The above equations assume isolated populations in each cell. However, for a typical plant species, dispersal occurs when seeds move from the cell where they are produced to a different cell. In an IPM with a seed bank, this will affect the fecundity kernel $F(z^{\prime}, z)$ and the seed bank $B$. Specifically, the number of seeds in the seed bank in cell $i$ at time $t+1$ will be the number of seeds surviving in the seed bank $B_i$ from time $t$ to $t+1$, plus the number of seeds produced in cell $i$ at time $t$ that remain in cell $i$ and do not recruit directly, plus the number of seeds entering cell $i$ from cells $j = 1, ..., J$ as immigrants in time $t$ that then fail to recruit directly:
\begin{equation}
	\begin{aligned}
		B_{i,t+1} = \enspace
		& B_{i,t} s_{survSB} (1-s_{rcrSB}) \enspace+  \\
		& \int_{\Omega} p_{flower,i}(z)  f_{seeds,i}(z) (1-p_{emig}) (1-s_{rcrDirect}) s_{survB} n_{i,t}(z) dz \enspace+  \\
		& \sum_{j=1}^{J} \int_{\Omega} [p_{flower,j}(z)  f_{seeds,j}(z) p_{emig} n_{t,j}(z) dz] p_{SDD,ji} (1-s_{rcrDirect}) s_{survB}\\
	\end{aligned}
\end{equation}
\begin{equation}
\begin{aligned}
n_{i,t+1}(z^{\prime}) = \enspace
& B_{i,t} s_{rcrSB} p_{estab_i} f_{rcrSize}(z^{\prime}) \enspace+   \\
& \int_{\Omega} [s_i(z) g_i(z^{\prime}|z) + p_{flower_i}(z)  f_{seeds_i}(z) (1-p_{emig}) s_{rcrDirect} p_{estab_i} f_{rcrSize}(z^{\prime})] n_{i,t}(z) dz \enspace +  \\
& \sum_{j=1}^{J} \int_{\Omega} [p_{flower,j}(z)  f_{seeds,j}(z) p_{emig} n_{t, j}(z) dz] p_{SDD,ji} s_{rcrDirect} p_{estab,i} f_{rcrSize}(z^{\prime})\\
\end{aligned}
\end{equation}
for each cell $i$ which is a target cell of each cell $j$ of $J$ cells, where the integral describes the seed production in each cell $j$ and $p_{SDD,ji}$ is the probability that a seed dispersed from $j$ lands in $i$. Note that, as above, seeds added to the seed bank must survive overwinter as well as fail to recruit directly.

---

\newpage
\section{Regression equations for the virtual species}
We used the above IPM structure, including a seed bank and dispersal, as the basis for the virtual species. Thus, the population in each cell $i$ has size distribution $z$, where $z^{\prime}$ is the size distribution the next year, and $\boldsymbol{z_i}$ is a matrix with columns for: 1, $z$, $z^2$, and $z^3$.

\subsection{Survival}   
Annual survival, $\boldsymbol{s_i}$, was modelled for each individual in cell $i$ as a binary outcome (0: mortality; 1: survival) following a Bernoulli distribution with probability $\boldsymbol{\psi_{si}}$ such that:
\begin{align}
\boldsymbol{s_i} & \sim Bern(\boldsymbol{\psi_{si}}) \\
logit(\boldsymbol{\psi_{si}}) & = \boldsymbol{z_i}\beta_{s} + \boldsymbol{X_i}\theta_{s}
\end{align}
where $\beta_{s}$ is a vector of covariates for size, $\boldsymbol{X_i}$ is a set of cell-level environmental covariates, and $\theta_{s}$ is a vector of responses to the environmental covariates.

\subsection{Growth} 
The size distribution, $\boldsymbol{z^{\prime}_i}$ of individuals in cell $i$ for time $t+1$ was distributed normally about the vector of expected sizes, $\boldsymbol{\mu_{gi}}$ with standard deviation $\sigma_g$, such that
\begin{align}
\boldsymbol{z^{\prime}_i} &\sim Norm(\boldsymbol{\mu_{gi}}, \sigma_g) \\
\boldsymbol{\mu_{gi}} &= \boldsymbol{z_i}\beta_{g} + \boldsymbol{X_i}\theta_{g}
\end{align}
where $\beta_{g}$ is a vector of covariates for size, $\boldsymbol{X_i}$ is a set of cell-level environmental covariates, and $\theta_{g}$ is a vector of responses to the environmental covariates.

\subsection{Flowering}  
Individual flowering, $\boldsymbol{l_i}$, was modelled for each individual in cell $i$ as a binary outcome (0: no flowers; 1: flowers) following a Bernoulli distribution with probability $\boldsymbol{\psi_{li}}$ such that:
\begin{align}
\boldsymbol{l_i} &\sim Bern(\boldsymbol{\psi_{li}}) \\
logit(\boldsymbol{\psi_{li}}) &= \boldsymbol{z_i}\beta_{l} + \boldsymbol{X_i}\theta_{l}
\end{align}
where $\beta_{l}$ is a vector of covariates for size, $\boldsymbol{X_i}$ is a set of cell-level environmental covariates, and $\theta_{l}$ is a vector of responses to the environmental covariates.

\subsection{Seeds}   
The number of seeds produced, $\boldsymbol{d_i}$ by each flowering individual in cell $i$ for time $t$ was Poisson distributed about the vector of expected seed counts, $\boldsymbol{\mu_{di}}$, such that
\begin{align}
\boldsymbol{d_i} &\sim Poisson(\boldsymbol{\mu_{di}}) \\
log(\boldsymbol{\mu_{di}}) &= \boldsymbol{z_i}\beta_{d} + \boldsymbol{X_i}\theta_{d}
\end{align}
where $\beta_{d}$ is a vector of covariates for size, $\boldsymbol{X_i}$ is a set of cell-level environmental covariates, and $\theta_{d}$ is a vector of responses to the environmental covariates. In each cell $i$, the seeds produced stay in the cell, emigrate through short distance dispersal, or perish. Seeds that survive may either enter the seed bank or recruit directly.

\subsection{Dispersal}   
The total number of immigrant seeds arriving in a cell, $D_i$ is calculated as the sum of the seeds from nearby cells that disperse from cell $j$ to cell $i$, such that:
\begin{equation}
D_i = \sum\limits_{j=1}^J \sum\limits_{n=1}^{N_j} \boldsymbol{d_j}p_{emig}p_{ji}
\end{equation}
where $J$ is the number of cells dispersing into $i$, $N_j$ is the number of individuals in cell $j$, $d_j$ is the vector of seed numbers produced by individuals in cell $j$, $p_{emig}$ is the probability that seeds produced in cell $j$ emigrate, and $p_{ji}$ is the probability that a seed emigrating from cell $j$ is dispersed to cell $i$.

\subsection{Recruits} 
The number of recruits, $n_{rcr,i}$ in cell $i$ in time $t$, may originate either from the seed bank, from seeds produced in cell $i$ in year $t$, or from immigrant seeds arriving in year $t$, such that:
\begin{align}
n_{rcr,i} &= B_i p_{rcrB}p_{est} + \sum\limits_{n=1}^{N_i} \boldsymbol{d_i}(1-p_{emig})p_{rcrDirect}p_{est} + D_i p_{rcrDirect}p_{est}\\
\boldsymbol{z^{\prime}_{rcr,i}} &\sim Norm(\mu_{rcr.z}, \sigma_{rcr.z})
\end{align}
where $B_i$ is the number of seeds in the seed bank, $p_{rcrB}$ is the probability that a seed germinates from the seed bank, $p_{est}$ is the probability that a new seedling establishes, $p_rcrDirect$ is the probability that a seed germinates in the year it was produced, $z^{\prime}_{rcr,i}$ is the size distribution of new recruits in year $t+1$, $\mu_{rcr.z}$ is the mean recruit size, and $\sigma_{rcr.z}$ is the standard deviation of recruit size.

\subsection{Seed Bank} 
Finally, the seed bank for year $t+1$ is calculated as the sum of the seeds remaining in the seed bank, the seeds produced in cell $i$ and entering into the seed bank, and the immigrant seeds entering into the seed bank, such that:
\begin{equation}
B^{\prime}_i = B_i(1-p_{rcrB})s_B + \sum\limits_{n=1}^{N_i} \boldsymbol{d_i}(1-p_{emig})(1-p_{rcrDirect})s_B + D_i(1-p_{rcrDirect})s_B
\end{equation}
where $s_B$ is the probability that a seed survives in the seed bank to the next year.

---

\newpage
\section{Species Distribution Model details}
This section of the appendix contains additional information regarding the structure of the species distribution models. The full R code is available from https://github.com/Sz-Tim/sdmMethodComp. Include more information on the CA\textsubscript{p} model and some on the CA\textsubscript{i} model.

---

\newpage
\section{Scenario details}
This section of the appendix contains additional information regarding the data and modelling scenarios. The full R code is available from https://github.com/Sz-Tim/sdmMethodComp. Include the table showing the exact implementation of each scenario.

### 2019-12-13 ms/supp/App_1.tex
returned to latex for equations

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Appendix 1: Supplemental Methods},
            pdfauthor={Tim M. Szewczyk, Marek Petrik, Jenica M. Allen},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\providecommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{Appendix 1: Supplemental Methods}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \subtitle{The performance of presence-based and process-based species distribution
models}
  \author{Tim M. Szewczyk, Marek Petrik, Jenica M. Allen}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
    \date{}
    \predate{}\postdate{}

\newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{A.\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{A.\arabic{figure}}}  \usepackage{longtable}  \usepackage{caption}

\begin{document}
\maketitle

{
\setcounter{tocdepth}{1}
\tableofcontents
}
\setcounter{table}{0}  \renewcommand{\thetable}{A.\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{A.\arabic{figure}}

\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}

This appendix contains supplemental methods pertaining to the virtual
species generation, species distribution model specifics, and scenario
particulars.

\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}

\section{General model structure}
\subsection{Integral Projection Model overview}

To generate fully-known true distributions for the virtual species, we
used the general structure of an Integral Projection Model (IPM) to
calculate the intrinsic growth rate \(\lambda\) in each cell of the
gridded landscape, and adapted the regression-based structure of an IPM
into an individual-level, simulation-based cellular automata (CA) model
to produce spatiotemporally dynamic abundance distributions. The
equations below apply to both models.

IPMs use the size distribution, \(z\), of individuals at time \(t\),
along with a kernel \(K(z^{\prime}, z)\), to predict the size
distribution, \(z^{\prime}\), at time \(t+1\): \begin{equation}
n_{t+1}(z^{\prime}) = \int_{\Omega} K(z^{\prime}, z) n_t(z) dz
\end{equation}

Here, \(\Omega\) represents the range of possible sizes for the species
or population. The kernel \(K(z^{\prime}, z)\) is composed of a growth
and survival component, \(P(z, z^{\prime})\), representing the fate of
individuals from time \(t\) to \(t+1\), and a fecundity component,
\(F(z, z^{\prime})\), representing new individuals added between time
\(t\) and \(t+1\). In practice, the integral is approximated using a
discretized transition matrix, and the intrinsic growth rate \(\lambda\)
is calculated as the first eigenvalue of the transition matrix.

The \(P\) and \(F\) kernels are decomposed further into more mechanistic
conditional probabilities and parameters, many of which are functions of
the size distribution \(z\). For example: \begin{equation}
\begin{split}
K(z^{\prime}, z) & = P(z^{\prime}, z) + F(z^{\prime}, z) \\
    & = s(z) g(z^{\prime}|z) + p_{flower}(z) f_{seeds}(z) p_{estab} f_{rcrSize}(z^{\prime})
\end{split}
\end{equation}

where \(s(z)\) is the survival probability of individuals based on size,
\(g(z^{\prime}|z)\) is the probability density of size \(z^{\prime}\)
for an individual of size \(z\), \(p_{flower}(z)\) is the probability
that an individual of size \(z\) produces flowers, \(f_{seeds}(z)\) is
the expected number of seeds produced by an individual of size \(z\)
given that they flower, \(p_{estab}\) is the probability that a seed
germinates and establishes as a new recruit, and
\(f_{rcrSize}(z^{\prime})\) is the expected size distribution of new
recruits at time \(t+1\). On a gridded landscape, the population in each
cell could be modelled independently using the above structure, with
environmental effects incorporated by allowing the environment to
influence parameter values (e.g., \(f_{seeds}\)).

\subsection{Adding a seed bank}

This basic IPM structure is very flexible, allowing for discrete life
stages, reproduction-dependent mortality, and environmental covariates.
To incorporate a seed bank where seeds that do not germinate between
time \(t\) and \(t+1\) may survive to \(t+2\) or beyond, the fecundity
kernel is altered, with seed bank \(B\), such that: \begin{align}
n_{t+1}(z^{\prime}) &= B_t s_{rcrB} p_{estab} f_{rcrSize}(z^{\prime}) + \int_{\Omega}[P(z^{\prime}, z)+F(z^{\prime}, z)] n_t(z) dz \\
F(z^{\prime}, z) &= p_{flower}(z)  f_{seeds}(z) s_{rcrDirect} p_{estab} f_{rcrSize}(z^{\prime}) \\
B_{t+1} &= B_t s_{survB} (1-s_{rcrB})  + p_{flower}(z)  f_{seeds}(z) (1-s_{rcrDirect}) s_{survB} n_t(z) dz
\end{align}

where \(B_t\) is the number of seeds in the seed bank at time \(t\),
\(s_{rcrB}\) is the probability a seed recruits from the seed bank,
\(s_{survB}\) is the probability a seed survives in the seed bank from
time \(t\) to \(t+1\), and \(s_{rcrDirect}\) is the probability a seed
produced in year \(t\) germinates between year \(t\) and \(t+1\). Two
notes about the structure and definitions: 1) a seed added to the seed
bank must fail to recruit in time \(t\) and must also survive from \(t\)
to \(t+1\), and 2) the probability of recruiting is best interpreted as
the probability of germinating and is therefore separate from the
probability of establishing. Both of these could be defined differently
to combine each set of processes, though keeping them separate allows
for seeds to perish.

\subsection{Adding dispersal}

The above equations assume isolated populations in each cell. However,
for a typical plant species, dispersal occurs when seeds move from the
cell where they are produced to a different cell. In an IPM with a seed
bank, this will affect the fecundity kernel \(F(z^{\prime}, z)\) and the
seed bank \(B\). Specifically, the number of seeds in the seed bank in
cell \(i\) at time \(t+1\) will be the number of seeds surviving in the
seed bank \(B_i\) from time \(t\) to \(t+1\), plus the number of seeds
produced in cell \(i\) at time \(t\) that remain in cell \(i\) and do
not recruit directly, plus the number of seeds entering cell \(i\) from
cells \(j = 1, ..., J\) as immigrants in time \(t\) that then fail to
recruit directly: \begin{equation}
    \begin{aligned}
        B_{i,t+1} = \enspace
        & B_{i,t} s_{survSB} (1-s_{rcrSB}) \enspace+  \\
        & \int_{\Omega} p_{flower,i}(z)  f_{seeds,i}(z) (1-p_{emig}) (1-s_{rcrDirect}) s_{survB} n_{i,t}(z) dz \enspace+  \\
        & \sum_{j=1}^{J} \int_{\Omega} [p_{flower,j}(z)  f_{seeds,j}(z) p_{emig} n_{t,j}(z) dz] p_{SDD,ji} (1-s_{rcrDirect}) s_{survB}\\
    \end{aligned}
\end{equation} \begin{equation}
\begin{aligned}
n_{i,t+1}(z^{\prime}) = \enspace
& B_{i,t} s_{rcrSB} p_{estab_i} f_{rcrSize}(z^{\prime}) \enspace+   \\
& \int_{\Omega} [s_i(z) g_i(z^{\prime}|z) + p_{flower_i}(z)  f_{seeds_i}(z) (1-p_{emig}) s_{rcrDirect} p_{estab_i} f_{rcrSize}(z^{\prime})] n_{i,t}(z) dz \enspace +  \\
& \sum_{j=1}^{J} \int_{\Omega} [p_{flower,j}(z)  f_{seeds,j}(z) p_{emig} n_{t, j}(z) dz] p_{SDD,ji} s_{rcrDirect} p_{estab,i} f_{rcrSize}(z^{\prime})\\
\end{aligned}
\end{equation} for each cell \(i\) which is a target cell of each cell
\(j\) of \(J\) cells, where the integral describes the seed production
in each cell \(j\) and \(p_{SDD,ji}\) is the probability that a seed
dispersed from \(j\) lands in \(i\). Note that, as above, seeds added to
the seed bank must survive overwinter as well as fail to recruit
directly.

\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}

\newpage
\section{Regression equations for the virtual species}

We used the above IPM structure, including a seed bank and dispersal, as
the basis for the virtual species. Thus, the population in each cell
\(i\) has size distribution \(z\), where \(z^{\prime}\) is the size
distribution the next year, and \(\boldsymbol{z_i}\) is a matrix with
columns for: 1, \(z\), \(z^2\), and \(z^3\).

\subsection{Survival}

Annual survival, \(\boldsymbol{s_i}\), was modelled for each individual
in cell \(i\) as a binary outcome (0: mortality; 1: survival) following
a Bernoulli distribution with probability \(\boldsymbol{\psi_{si}}\)
such that: \begin{align}
\boldsymbol{s_i} & \sim Bern(\boldsymbol{\psi_{si}}) \\
logit(\boldsymbol{\psi_{si}}) & = \boldsymbol{z_i}\beta_{s} + \boldsymbol{X_i}\theta_{s}
\end{align} where \(\beta_{s}\) is a vector of covariates for size,
\(\boldsymbol{X_i}\) is a set of cell-level environmental covariates,
and \(\theta_{s}\) is a vector of responses to the environmental
covariates.

\subsection{Growth}

The size distribution, \(\boldsymbol{z^{\prime}_i}\) of individuals in
cell \(i\) for time \(t+1\) was distributed normally about the vector of
expected sizes, \(\boldsymbol{\mu_{gi}}\) with standard deviation
\(\sigma_g\), such that \begin{align}
\boldsymbol{z^{\prime}_i} &\sim Norm(\boldsymbol{\mu_{gi}}, \sigma_g) \\
\boldsymbol{\mu_{gi}} &= \boldsymbol{z_i}\beta_{g} + \boldsymbol{X_i}\theta_{g}
\end{align} where \(\beta_{g}\) is a vector of covariates for size,
\(\boldsymbol{X_i}\) is a set of cell-level environmental covariates,
and \(\theta_{g}\) is a vector of responses to the environmental
covariates.

\subsection{Flowering}

Individual flowering, \(\boldsymbol{l_i}\), was modelled for each
individual in cell \(i\) as a binary outcome (0: no flowers; 1: flowers)
following a Bernoulli distribution with probability
\(\boldsymbol{\psi_{li}}\) such that: \begin{align}
\boldsymbol{l_i} &\sim Bern(\boldsymbol{\psi_{li}}) \\
logit(\boldsymbol{\psi_{li}}) &= \boldsymbol{z_i}\beta_{l} + \boldsymbol{X_i}\theta_{l}
\end{align} where \(\beta_{l}\) is a vector of covariates for size,
\(\boldsymbol{X_i}\) is a set of cell-level environmental covariates,
and \(\theta_{l}\) is a vector of responses to the environmental
covariates.

\subsection{Seeds}

The number of seeds produced, \(\boldsymbol{d_i}\) by each flowering
individual in cell \(i\) for time \(t\) was Poisson distributed about
the vector of expected seed counts, \(\boldsymbol{\mu_{di}}\), such that
\begin{align}
\boldsymbol{d_i} &\sim Poisson(\boldsymbol{\mu_{di}}) \\
log(\boldsymbol{\mu_{di}}) &= \boldsymbol{z_i}\beta_{d} + \boldsymbol{X_i}\theta_{d}
\end{align} where \(\beta_{d}\) is a vector of covariates for size,
\(\boldsymbol{X_i}\) is a set of cell-level environmental covariates,
and \(\theta_{d}\) is a vector of responses to the environmental
covariates. In each cell \(i\), the seeds produced stay in the cell,
emigrate through short distance dispersal, or perish. Seeds that survive
may either enter the seed bank or recruit directly.

\subsection{Dispersal}

The total number of immigrant seeds arriving in a cell, \(D_i\) is
calculated as the sum of the seeds from nearby cells that disperse from
cell \(j\) to cell \(i\), such that: \begin{equation}
D_i = \sum\limits_{j=1}^J \sum\limits_{n=1}^{N_j} \boldsymbol{d_j}p_{emig}p_{ji}
\end{equation} where \(J\) is the number of cells dispersing into \(i\),
\(N_j\) is the number of individuals in cell \(j\), \(d_j\) is the
vector of seed numbers produced by individuals in cell \(j\),
\(p_{emig}\) is the probability that seeds produced in cell \(j\)
emigrate, and \(p_{ji}\) is the probability that a seed emigrating from
cell \(j\) is dispersed to cell \(i\).

\subsection{Recruits}

The number of recruits, \(n_{rcr,i}\) in cell \(i\) in time \(t\), may
originate either from the seed bank, from seeds produced in cell \(i\)
in year \(t\), or from immigrant seeds arriving in year \(t\), such
that: \begin{align}
n_{rcr,i} &= B_i p_{rcrB}p_{est} + \sum\limits_{n=1}^{N_i} \boldsymbol{d_i}(1-p_{emig})p_{rcrDirect}p_{est} + D_i p_{rcrDirect}p_{est}\\
\boldsymbol{z^{\prime}_{rcr,i}} &\sim Norm(\mu_{rcr.z}, \sigma_{rcr.z})
\end{align} where \(B_i\) is the number of seeds in the seed bank,
\(p_{rcrB}\) is the probability that a seed germinates from the seed
bank, \(p_{est}\) is the probability that a new seedling establishes,
\(p_rcrDirect\) is the probability that a seed germinates in the year it
was produced, \(z^{\prime}_{rcr,i}\) is the size distribution of new
recruits in year \(t+1\), \(\mu_{rcr.z}\) is the mean recruit size, and
\(\sigma_{rcr.z}\) is the standard deviation of recruit size.

\subsection{Seed Bank}

Finally, the seed bank for year \(t+1\) is calculated as the sum of the
seeds remaining in the seed bank, the seeds produced in cell \(i\) and
entering into the seed bank, and the immigrant seeds entering into the
seed bank, such that: \begin{equation}
B^{\prime}_i = B_i(1-p_{rcrB})s_B + \sum\limits_{n=1}^{N_i} \boldsymbol{d_i}(1-p_{emig})(1-p_{rcrDirect})s_B + D_i(1-p_{rcrDirect})s_B
\end{equation} where \(s_B\) is the probability that a seed survives in
the seed bank to the next year.

\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}

\newpage
\section{Species Distribution Model details}

This section of the appendix contains additional information regarding
the structure of the species distribution models. The full R code is
available from \url{https://github.com/Sz-Tim/sdmMethodComp}. Include
more information on the CA\textsubscript{p} model and some on the
CA\textsubscript{i} model.

\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}

\newpage
\section{Scenario details}

This section of the appendix contains additional information regarding
the data and modelling scenarios. The full R code is available from
\url{https://github.com/Sz-Tim/sdmMethodComp}. Include the table showing
the exact implementation of each scenario.

\end{document}

### 2019-12-13 ms/supp/App_2.Rmd
returned to latex for equations

---
title: 'Appendix 2: Supplemental Results'
subtitle: 'The performance of presence-based and process-based species distribution models'
author: 'Tim M. Szewczyk, Marek Petrik, Jenica M. Allen'
output:
  pdf_document: 
    number_sections: yes
    toc: yes
    toc_depth: 1
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 1
editor_options:
  chunk_output_type: inline
header-includes:
  - \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{A.\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{A.\arabic{figure}}}  \usepackage{longtable}  \usepackage{caption}
---

\beginsupplement

---

This appendix contains supplemental tables and figures to provide additional detail for the results, including maps of $\lambda$ and $N$ for each species (true, predicted, difference), as well as figures and a table for sensitivity and specificity with the same structures as those for TSS in the main manuscript.

\begin{figure}
	\centering\includegraphics[width=\linewidth]{../../figs/Supp_SensvIdeal.jpg}
	\caption{\label{fig:SensitivityvIdeal} Effect of scenario on median sensitivity relative to the '\emph{ideal}' scenario. Sensitivity represents the proportion of true presences that were correctly predicted as presences.}
\end{figure}

\begin{figure}
	\centering\includegraphics[width=\linewidth]{../../figs/Supp_SpecvIdeal.jpg}
	\caption{\label{fig:SpecificityvIdeal} Effect of scenario on median specificity relative to the '\emph{ideal}' scenario. Specificity represents the proportion of true absences that were correctly predicted as absences.}
\end{figure}

### 2019-12-13 ms/supp/App_2.tex
returned to latex for equations

--- 2019-12-13 ms/supp/App_1.tex
@@
 \hypersetup{unicode=true,
-            pdftitle={Appendix 1: Supplemental Methods},
+            pdftitle={Appendix 2: Supplemental Results},
             pdfauthor={Tim M. Szewczyk, Marek Petrik, Jenica M. Allen},
@@
 
-  \title{Appendix 1: Supplemental Methods}
+  \title{Appendix 2: Supplemental Results}
     \pretitle{\vspace{\droptitle}\centering\huge}
@@
 
-This appendix contains supplemental methods pertaining to the virtual
-species generation, species distribution model specifics, and scenario
-particulars.
+This appendix contains supplemental tables and figures to provide
+additional detail for the results, including maps of \(\lambda\) and
+\(N\) for each species (true, predicted, difference), as well as figures
+and a table for sensitivity and specificity with the same structures as
+those for TSS in the main manuscript.
 
-\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
+\begin{figure}
+    \centering\includegraphics[width=\linewidth]{../../figs/Supp_SensvIdeal.jpg}
+    \caption{\label{fig:SensitivityvIdeal} Effect of scenario on median sensitivity relative to the '\emph{ideal}' scenario.
+Sensitivity represents the proportion of true presences that were correctly predicted as presences.}
+\end{figure}
 
-\section{General model structure}
-\subsection{Integral Projection Model overview}
-
-To generate fully-known true distributions for the virtual species, we
-used the general structure of an Integral Projection Model (IPM) to
-calculate the intrinsic growth rate \(\lambda\) in each cell of the
-gridded landscape, and adapted the regression-based structure of an IPM
-into an individual-level, simulation-based cellular automata (CA) model
-to produce spatiotemporally dynamic abundance distributions. The
-equations below apply to both models.
-
-IPMs use the size distribution, \(z\), of individuals at time \(t\),
-along with a kernel \(K(z^{\prime}, z)\), to predict the size
-distribution, \(z^{\prime}\), at time \(t+1\): \begin{equation}
-n_{t+1}(z^{\prime}) = \int_{\Omega} K(z^{\prime}, z) n_t(z) dz
-\end{equation}
-
-Here, \(\Omega\) represents the range of possible sizes for the species
-or population. The kernel \(K(z^{\prime}, z)\) is composed of a growth
-and survival component, \(P(z, z^{\prime})\), representing the fate of
-individuals from time \(t\) to \(t+1\), and a fecundity component,
-\(F(z, z^{\prime})\), representing new individuals added between time
-\(t\) and \(t+1\). In practice, the integral is approximated using a
-discretized transition matrix, and the intrinsic growth rate \(\lambda\)
-is calculated as the first eigenvalue of the transition matrix.
-
-The \(P\) and \(F\) kernels are decomposed further into more mechanistic
-conditional probabilities and parameters, many of which are functions of
-the size distribution \(z\). For example: \begin{equation}
-\begin{split}
-K(z^{\prime}, z) & = P(z^{\prime}, z) + F(z^{\prime}, z) \\
-    & = s(z) g(z^{\prime}|z) + p_{flower}(z) f_{seeds}(z) p_{estab} f_{rcrSize}(z^{\prime})
-\end{split}
-\end{equation}
-
-where \(s(z)\) is the survival probability of individuals based on size,
-\(g(z^{\prime}|z)\) is the probability density of size \(z^{\prime}\)
-for an individual of size \(z\), \(p_{flower}(z)\) is the probability
-that an individual of size \(z\) produces flowers, \(f_{seeds}(z)\) is
-the expected number of seeds produced by an individual of size \(z\)
-given that they flower, \(p_{estab}\) is the probability that a seed
-germinates and establishes as a new recruit, and
-\(f_{rcrSize}(z^{\prime})\) is the expected size distribution of new
-recruits at time \(t+1\). On a gridded landscape, the population in each
-cell could be modelled independently using the above structure, with
-environmental effects incorporated by allowing the environment to
-influence parameter values (e.g., \(f_{seeds}\)).
-
-\subsection{Adding a seed bank}
-
-This basic IPM structure is very flexible, allowing for discrete life
-stages, reproduction-dependent mortality, and environmental covariates.
-To incorporate a seed bank where seeds that do not germinate between
-time \(t\) and \(t+1\) may survive to \(t+2\) or beyond, the fecundity
-kernel is altered, with seed bank \(B\), such that: \begin{align}
-n_{t+1}(z^{\prime}) &= B_t s_{rcrB} p_{estab} f_{rcrSize}(z^{\prime}) + \int_{\Omega}[P(z^{\prime}, z)+F(z^{\prime}, z)] n_t(z) dz \\
-F(z^{\prime}, z) &= p_{flower}(z)  f_{seeds}(z) s_{rcrDirect} p_{estab} f_{rcrSize}(z^{\prime}) \\
-B_{t+1} &= B_t s_{survB} (1-s_{rcrB})  + p_{flower}(z)  f_{seeds}(z) (1-s_{rcrDirect}) s_{survB} n_t(z) dz
-\end{align}
-
-where \(B_t\) is the number of seeds in the seed bank at time \(t\),
-\(s_{rcrB}\) is the probability a seed recruits from the seed bank,
-\(s_{survB}\) is the probability a seed survives in the seed bank from
-time \(t\) to \(t+1\), and \(s_{rcrDirect}\) is the probability a seed
-produced in year \(t\) germinates between year \(t\) and \(t+1\). Two
-notes about the structure and definitions: 1) a seed added to the seed
-bank must fail to recruit in time \(t\) and must also survive from \(t\)
-to \(t+1\), and 2) the probability of recruiting is best interpreted as
-the probability of germinating and is therefore separate from the
-probability of establishing. Both of these could be defined differently
-to combine each set of processes, though keeping them separate allows
-for seeds to perish.
-
-\subsection{Adding dispersal}
-
-The above equations assume isolated populations in each cell. However,
-for a typical plant species, dispersal occurs when seeds move from the
-cell where they are produced to a different cell. In an IPM with a seed
-bank, this will affect the fecundity kernel \(F(z^{\prime}, z)\) and the
-seed bank \(B\). Specifically, the number of seeds in the seed bank in
-cell \(i\) at time \(t+1\) will be the number of seeds surviving in the
-seed bank \(B_i\) from time \(t\) to \(t+1\), plus the number of seeds
-produced in cell \(i\) at time \(t\) that remain in cell \(i\) and do
-not recruit directly, plus the number of seeds entering cell \(i\) from
-cells \(j = 1, ..., J\) as immigrants in time \(t\) that then fail to
-recruit directly: \begin{equation}
-    \begin{aligned}
-        B_{i,t+1} = \enspace
-        & B_{i,t} s_{survSB} (1-s_{rcrSB}) \enspace+  \\
-        & \int_{\Omega} p_{flower,i}(z)  f_{seeds,i}(z) (1-p_{emig}) (1-s_{rcrDirect}) s_{survB} n_{i,t}(z) dz \enspace+  \\
-        & \sum_{j=1}^{J} \int_{\Omega} [p_{flower,j}(z)  f_{seeds,j}(z) p_{emig} n_{t,j}(z) dz] p_{SDD,ji} (1-s_{rcrDirect}) s_{survB}\\
-    \end{aligned}
-\end{equation} \begin{equation}
-\begin{aligned}
-n_{i,t+1}(z^{\prime}) = \enspace
-& B_{i,t} s_{rcrSB} p_{estab_i} f_{rcrSize}(z^{\prime}) \enspace+   \\
-& \int_{\Omega} [s_i(z) g_i(z^{\prime}|z) + p_{flower_i}(z)  f_{seeds_i}(z) (1-p_{emig}) s_{rcrDirect} p_{estab_i} f_{rcrSize}(z^{\prime})] n_{i,t}(z) dz \enspace +  \\
-& \sum_{j=1}^{J} \int_{\Omega} [p_{flower,j}(z)  f_{seeds,j}(z) p_{emig} n_{t, j}(z) dz] p_{SDD,ji} s_{rcrDirect} p_{estab,i} f_{rcrSize}(z^{\prime})\\
-\end{aligned}
-\end{equation} for each cell \(i\) which is a target cell of each cell
-\(j\) of \(J\) cells, where the integral describes the seed production
-in each cell \(j\) and \(p_{SDD,ji}\) is the probability that a seed
-dispersed from \(j\) lands in \(i\). Note that, as above, seeds added to
-the seed bank must survive overwinter as well as fail to recruit
-directly.
-
-\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
-
-\newpage
-\section{Regression equations for the virtual species}
-
-We used the above IPM structure, including a seed bank and dispersal, as
-the basis for the virtual species. Thus, the population in each cell
-\(i\) has size distribution \(z\), where \(z^{\prime}\) is the size
-distribution the next year, and \(\boldsymbol{z_i}\) is a matrix with
-columns for: 1, \(z\), \(z^2\), and \(z^3\).
-
-\subsection{Survival}
-
-Annual survival, \(\boldsymbol{s_i}\), was modelled for each individual
-in cell \(i\) as a binary outcome (0: mortality; 1: survival) following
-a Bernoulli distribution with probability \(\boldsymbol{\psi_{si}}\)
-such that: \begin{align}
-\boldsymbol{s_i} & \sim Bern(\boldsymbol{\psi_{si}}) \\
-logit(\boldsymbol{\psi_{si}}) & = \boldsymbol{z_i}\beta_{s} + \boldsymbol{X_i}\theta_{s}
-\end{align} where \(\beta_{s}\) is a vector of covariates for size,
-\(\boldsymbol{X_i}\) is a set of cell-level environmental covariates,
-and \(\theta_{s}\) is a vector of responses to the environmental
-covariates.
-
-\subsection{Growth}
-
-The size distribution, \(\boldsymbol{z^{\prime}_i}\) of individuals in
-cell \(i\) for time \(t+1\) was distributed normally about the vector of
-expected sizes, \(\boldsymbol{\mu_{gi}}\) with standard deviation
-\(\sigma_g\), such that \begin{align}
-\boldsymbol{z^{\prime}_i} &\sim Norm(\boldsymbol{\mu_{gi}}, \sigma_g) \\
-\boldsymbol{\mu_{gi}} &= \boldsymbol{z_i}\beta_{g} + \boldsymbol{X_i}\theta_{g}
-\end{align} where \(\beta_{g}\) is a vector of covariates for size,
-\(\boldsymbol{X_i}\) is a set of cell-level environmental covariates,
-and \(\theta_{g}\) is a vector of responses to the environmental
-covariates.
-
-\subsection{Flowering}
-
-Individual flowering, \(\boldsymbol{l_i}\), was modelled for each
-individual in cell \(i\) as a binary outcome (0: no flowers; 1: flowers)
-following a Bernoulli distribution with probability
-\(\boldsymbol{\psi_{li}}\) such that: \begin{align}
-\boldsymbol{l_i} &\sim Bern(\boldsymbol{\psi_{li}}) \\
-logit(\boldsymbol{\psi_{li}}) &= \boldsymbol{z_i}\beta_{l} + \boldsymbol{X_i}\theta_{l}
-\end{align} where \(\beta_{l}\) is a vector of covariates for size,
-\(\boldsymbol{X_i}\) is a set of cell-level environmental covariates,
-and \(\theta_{l}\) is a vector of responses to the environmental
-covariates.
-
-\subsection{Seeds}
-
-The number of seeds produced, \(\boldsymbol{d_i}\) by each flowering
-individual in cell \(i\) for time \(t\) was Poisson distributed about
-the vector of expected seed counts, \(\boldsymbol{\mu_{di}}\), such that
-\begin{align}
-\boldsymbol{d_i} &\sim Poisson(\boldsymbol{\mu_{di}}) \\
-log(\boldsymbol{\mu_{di}}) &= \boldsymbol{z_i}\beta_{d} + \boldsymbol{X_i}\theta_{d}
-\end{align} where \(\beta_{d}\) is a vector of covariates for size,
-\(\boldsymbol{X_i}\) is a set of cell-level environmental covariates,
-and \(\theta_{d}\) is a vector of responses to the environmental
-covariates. In each cell \(i\), the seeds produced stay in the cell,
-emigrate through short distance dispersal, or perish. Seeds that survive
-may either enter the seed bank or recruit directly.
-
-\subsection{Dispersal}
-
-The total number of immigrant seeds arriving in a cell, \(D_i\) is
-calculated as the sum of the seeds from nearby cells that disperse from
-cell \(j\) to cell \(i\), such that: \begin{equation}
-D_i = \sum\limits_{j=1}^J \sum\limits_{n=1}^{N_j} \boldsymbol{d_j}p_{emig}p_{ji}
-\end{equation} where \(J\) is the number of cells dispersing into \(i\),
-\(N_j\) is the number of individuals in cell \(j\), \(d_j\) is the
-vector of seed numbers produced by individuals in cell \(j\),
-\(p_{emig}\) is the probability that seeds produced in cell \(j\)
-emigrate, and \(p_{ji}\) is the probability that a seed emigrating from
-cell \(j\) is dispersed to cell \(i\).
-
-\subsection{Recruits}
-
-The number of recruits, \(n_{rcr,i}\) in cell \(i\) in time \(t\), may
-originate either from the seed bank, from seeds produced in cell \(i\)
-in year \(t\), or from immigrant seeds arriving in year \(t\), such
-that: \begin{align}
-n_{rcr,i} &= B_i p_{rcrB}p_{est} + \sum\limits_{n=1}^{N_i} \boldsymbol{d_i}(1-p_{emig})p_{rcrDirect}p_{est} + D_i p_{rcrDirect}p_{est}\\
-\boldsymbol{z^{\prime}_{rcr,i}} &\sim Norm(\mu_{rcr.z}, \sigma_{rcr.z})
-\end{align} where \(B_i\) is the number of seeds in the seed bank,
-\(p_{rcrB}\) is the probability that a seed germinates from the seed
-bank, \(p_{est}\) is the probability that a new seedling establishes,
-\(p_rcrDirect\) is the probability that a seed germinates in the year it
-was produced, \(z^{\prime}_{rcr,i}\) is the size distribution of new
-recruits in year \(t+1\), \(\mu_{rcr.z}\) is the mean recruit size, and
-\(\sigma_{rcr.z}\) is the standard deviation of recruit size.
-
-\subsection{Seed Bank}
-
-Finally, the seed bank for year \(t+1\) is calculated as the sum of the
-seeds remaining in the seed bank, the seeds produced in cell \(i\) and
-entering into the seed bank, and the immigrant seeds entering into the
-seed bank, such that: \begin{equation}
-B^{\prime}_i = B_i(1-p_{rcrB})s_B + \sum\limits_{n=1}^{N_i} \boldsymbol{d_i}(1-p_{emig})(1-p_{rcrDirect})s_B + D_i(1-p_{rcrDirect})s_B
-\end{equation} where \(s_B\) is the probability that a seed survives in
-the seed bank to the next year.
-
-\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
-
-\newpage
-\section{Species Distribution Model details}
-
-This section of the appendix contains additional information regarding
-the structure of the species distribution models. The full R code is
-available from \url{https://github.com/Sz-Tim/sdmMethodComp}. Include
-more information on the CA\textsubscript{p} model and some on the
-CA\textsubscript{i} model.
-
-\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
-
-\newpage
-\section{Scenario details}
-
-This section of the appendix contains additional information regarding
-the data and modelling scenarios. The full R code is available from
-\url{https://github.com/Sz-Tim/sdmMethodComp}. Include the table showing
-the exact implementation of each scenario.
+\begin{figure}
+    \centering\includegraphics[width=\linewidth]{../../figs/Supp_SpecvIdeal.jpg}
+    \caption{\label{fig:SpecificityvIdeal} Effect of scenario on median specificity relative to the '\emph{ideal}' scenario.
+Specificity represents the proportion of true absences that were correctly predicted as absences.}
+\end{figure}
 

### 2020-02-05 code/6_figures_ms.R
updated figures

@@ -9,11 +9,11 @@
 ########
 
 rmarkdown::render("ms/supp/App_1.Rmd", 
-                  output_file="App_1.pdf",
+                  output_file="App_A.pdf",
                   knit_root_dir="../")
 
 rmarkdown::render("ms/supp/App_2.Rmd", 
-                  output_file="App_2.pdf",
+                  output_file="App_B.pdf",
                   knit_root_dir="../")
 
 
@@ -24,7 +24,7 @@ rmarkdown::render("ms/supp/App_2.Rmd",
 
 # load workspace
 pkgs <- c("tidyverse", "magrittr", "stringr", "here", "viridis",
-          "grid", "gtable", "ggnewscale")
+          "grid", "gtable", "ggnewscale", "gridExtra")
 suppressMessages(invisible(lapply(pkgs, library, character.only=T)))
 walk(dir("code", "fn", full.names=T), source)
 theme_set(theme_bw())
@@ -67,12 +67,14 @@ TSS.ci <- TSS.df %>% group_by(sp, SDM, issue, Boundary) %>%
 senspe.ci <- TSS.df %>% group_by(sp, SDM, issue, Boundary) %>% 
   summarise(med.sen=median(sensitivity, na.rm=T), 
             mn.sen=mean(sensitivity, na.rm=T),
+            sd.sen=sd(sensitivity, na.rm=T),
             q025.sen=quantile(sensitivity, probs=0.025, na.rm=T), 
             q25.sen=quantile(sensitivity, probs=0.25, na.rm=T), 
             q75.sen=quantile(sensitivity, probs=0.75, na.rm=T), 
             q975.sen=quantile(sensitivity, probs=0.975, na.rm=T),
             med.spe=median(specificity, na.rm=T), 
             mn.spe=mean(specificity, na.rm=T),
+            sd.spe=sd(specificity, na.rm=T),
             q025.spe=quantile(specificity, probs=0.025, na.rm=T), 
             q25.spe=quantile(specificity, probs=0.25, na.rm=T), 
             q75.spe=quantile(specificity, probs=0.75, na.rm=T), 
@@ -335,6 +337,23 @@ ggplot(TSS.ci, aes(x=issue, y=mn, colour=SDM, group=SDM)) +
[…]

### 2020-02-05 code/fn_aux.R
added function for plotting slope estimates

@@ -515,6 +515,42 @@ plot_sdm_reg <- function(p, sdm.diag, pars, par.v, X, X.mx, cols,
 
 
 
+
+##-- plot distribution of predicted slopes
+#' Compare vital rate regression slopes
+#' @param p True parameters
+#' @param sdm.diag List of diagnostic outputs
+#' @param pars Parameter to plot
+#' @param par.v Indexes of parameters
+#' @param cols Color for each element in \code{ipm.diag}
+#' @param xlab x-axis label
+#' @param ylab y-axis label
+#' @param xlim x-axis limits
+#' @param ylim y-axis limits
+plot_sdm_slopes <- function(p, sdm.diag, pars) {
+  library(tidyverse); theme_set(theme_bw())
+  slope.df <- imap(sdm.diag[1:4], ~map(.x, ~pluck(., pars)) %>% 
+                     do.call(rbind, .) %>% 
+                     as.data.frame %>% mutate(scenario=.y)) %>%
+    do.call(rbind, .) %>%
+    pivot_longer(cols=1:(ncol(.)-1), names_to="variable")
+  true.df <- data.frame(value=p[[pars]], 
+                        variable=names(sdm.diag[[1]][[1]][[pars]]))
+  p_dens <- ggplot(slope.df, aes(x=value, colour=scenario)) + geom_density() + 
+    geom_vline(data=true.df, aes(xintercept=value)) + 
+    facet_wrap(~variable, scales="free")
+  p_box <- ggplot(slope.df, aes(x=scenario, y=value, colour=scenario)) + 
+    geom_boxplot() + geom_hline(data=true.df, aes(yintercept=value)) + 
+    facet_wrap(~variable, scales="free")
+  return(list(p_dens=p_dens, p_box=p_box, slope.df=slope.df, true.df=true.df))
+}
+
+
+
+
+
+
+
 #' @title Layout panels in a grid with nested strips
 #'
 #' @description \code{facet_nest()} form a matrix of panels defined by row and

### 2020-02-05 code/fn_sim.R
comment

@@ -228,7 +228,7 @@ simulate_data <- function(n.cell, lo, hi, p, X, n_z, sdd.ji, p.ji, N_init, sp,
       p_est.i[,yr_k] <- p_est1
     }
     
-    # debug
+    ## debug
     # k <- k+1
     # length(occupied)
     # summary(N)

### 2020-02-05 ms/supp/App_1.Rmd
rename App A/B, added text and figures

@@
 ---
-title: 'Appendix 1: Supplemental Methods'
-subtitle: 'The performance of presence-based and process-based species distribution models'
+title: 'Appendix A: Supplemental Methods'
+subtitle: 'The performance of presence-based and process-based species distribution models under realistic conditions'
 author: 'Tim M. Szewczyk, Marek Petrik, Jenica M. Allen'
@@
 
-This appendix contains supplemental methods pertaining to the virtual species generation, species distribution model specifics, and scenario particulars.
+This appendix contains supplemental methods pertaining to the virtual species, species distribution models, and scenarios.
 
@@
 We used the above IPM structure, including a seed bank and dispersal, as the basis for the virtual species.
-Thus, the population in each cell $i$ has size distribution $z$, where $z^{\prime}$ is the size distribution the next year, and $\boldsymbol{z_i}$ is a matrix with columns for: 1, $z$, $z^2$, and $z^3$.
+Thus, the population in each cell $i$ has size distribution $z$, where $z^{\prime}$ is the size distribution the next year, and $\boldsymbol{z_i}$ is a matrix with columns of 1 (intercept) and $z$.
 
@@
 \section{Species Distribution Model details}
-This section of the appendix contains additional information regarding the structure of the species distribution models.
-The full R code is available from https://github.com/Sz-Tim/sdmMethodComp.
-Include more information on the CA\textsubscript{p} model and some on the CA\textsubscript{i} model.
+This section of the appendix contains additional information regarding the structure of the species distribution models. The full R code is available from https://github.com/Sz-Tim/sdmMethodComp.
+
+\subsection{IPM}
+The Integral Projection Models for each species followed the structure described in the previous section.
+For each dataset (100 datasets x 2 species x 4 data scenarios), we fit regressions for each vital rate regression using the R package \emph{MuMIn} to compare all combinations of climatic variables.
+The full models were:
+```{r, eval=F}
+# pr(survival)
+glm(surv ~ size + temp + temp_sq + precip + precip_sq, family="binomial")
+# growth | survival
+lm(size_next ~ size + temp + temp_sq + precip + precip_sq)
+# pr(flowering | survival)
+glm(flower ~ size + temp + temp_sq + precip + precip_sq, family="binomial")
+# number of seeds | flowering
+glm(n_seed ~ size + temp + temp_sq + precip + precip_sq, family="poisson")
+# pr(germination)
+glm(germ ~ temp + temp_sq + precip + precip_sq, family="binomial")
+```
+The size distribution for recruits was taken as the sample mean and standard deviation of recruits in each cell.
+Parameters relating to seed bank dynamics and dispersal were assumed to be known, and the effects of this assumption were evaluated in the modelling mis-specification scenarios.
+
+\subsection{CA\textsubscript{i}}
+The individual-level cellular automaton model (CA\textsubscript{i}) simulates individuals in each cell of the landscape using the structure of the IPM.
+Effectively, the fitted regressions are used to generate expected values for each individual, accounting for size and environment, and random values are then drawn from the corresponding distributions to simulate outcomes.
+For example, to assign survival, a value is drawn from a binomial distribution with the probability calculated from the fitted survival GLM.
+Thus, we simulated local processes within each cell.
+Regional processes were represented as the dispersal of seeds both within short distance dispersal neighborhoods and to a pre-determined number of random cells in the landscape.
+The CA\textsubscript{i} model is thus a simulation-based adaptation of the IPM.
+However, this requires imposing density dependence to both maintain realism and to work within the confines of modern computing capabilities.
+We implemented density dependent seedling establishment such that $p_{est,i,t} = min(p_{est,i}, N_{rcr}/(p_{germ}(D_i+B_i+d_i(1-p_{emig}))))$, following the naming convention described above, effectively placing a cap on the number of new recruits (Ellner & Rees 2006).
+
+\subsection{CA\textsubscript{p}}
+The population-level cellular automaton model (CA\textsubscript{p}) simulates stage-structured populations in each cell of the landscape using population-level values for demographic rates.
+Thus, rather than simulating, for example, the survival of each individual, the CA\textsubscript{p} model simulates the total expected number of surviving individuals in each cell.
+We follow the model structure and sequence from a previously published model focused on the invasive shrub glossy buckthorn (\emph{Frangula alnus}) in the northeastern United States (Szewczyk et al.
+2019).
+This includes the same life cycle processes as in the IPM and CA\textsubscript{i}: survival, growth (=aging), flowering, seed production, short and long distance dispersal, seed bank dynamics, and germination.
+However, individuals were not tracked or predicted, but rather averages or totals within age categories (shrub: 4 age categories; biennial: 2 age categories).
+We implemented density dependence as a ceiling-type carrying capacity, where the carrying capacity $K$ was predicted for each cell.
+In the samples used to parameterize the regressions (consisting of 3 years of data), cells were assumed to be approximately at carrying capacity if the calculated $lambda_t = N_t/N_{t-1}$ changed less than 5\% between consecutive years.
+Thus, we parameterized the following full models, including year as a random effect:
+```{r, eval=F}
+# carrying capacity
+glmer(K ~ temp + temp_sq + precip + precip_sq + (1|yr), family="poisson")
+# adult survival (shrub only)
+glmer(surv_adult ~ temp + temp_sq + precip + precip_sq + (1|yr), family="binomial")
+# juvenile survival
+glmer(surv_juv ~ temp + temp_sq + precip + precip_sq + (1|yr), family="binomial")
+# pr(flowering | survival)
+glmer(flower ~ size + temp + temp_sq + precip + precip_sq + (1|yr), family="binomial")
+# number of seeds | flowering
+glmer(n_seed ~ size + temp + temp_sq + precip + precip_sq + (1|yr), family="poisson")
+# pr(germination)
+glm(germ ~ temp + temp_sq + precip + precip_sq, family="binomial")
+```
+As in the IPM and CA\textsubscript{i}, dispersal included long distance and short distance dispersal of seeds.
+Identical assumptions and implementations were used for the CA\textsubscript{p} model as described above.
 
@@
 Include the table showing the exact implementation of each scenario.
+
+\subsection{Data scenarios}
+\subsubsection{Ideal}
+In the \emph{ideal} scenario, each data set consisted of cells sampled from the equilibrium distribution at the end of the virtual species simulation (tmax = 300).
+Cells with N > 5 were selected with uniform probability.
+For MaxEnt, 300 occupied cells were drawn for each data set.
+For the process-based SDMs, 25 cells were drawn.
+Within each of the 25 cells, up to 100 (CA\textsubscript{i}, IPM) or 1000 (CA\textsubscript{p}) individuals were randomly selected to be used in the vital rate regressions, with stratification among small, medium, and large individuals (CA\textsubscript{i}, IPM) or juveniles and adults (CA\textsubscript{p}).
+For cells with fewer individuals than the maximum, all individuals were included regardless of size.
+The final three years of the simulation were used to parameterize the CA\textsubscript{p} model, while only the final two years were used for the CA\textsubscript{i} model and IPM.
+The data available included abundance, survival, size, flowering status, fruit production, and germination rates.
+For estimating carrying capacity in the CA\textsubscript{p} model, we assumed cells where N changed less than 5\% between years were approximately at carrying capacity, and used the average abundance across years for those cells.
+
+\subsubsection{Non-equilibrium}
+The sampling regime under the \emph{non-equilibrium} scenario was identical to that of the \emph{ideal} scenario, except that samples were drawn from the distribution 100 years after introduction rather than the equilibrium distribution of 300 years.
+
+\subsubsection{Geographic bias}
+The \emph{geographic bias} sampling scenario represents the bias in sampling locations common in ecological studies, where data are more likely to be acquired in locations that are easier to reach.
+To establish realistic patterns of data collection, we used previously published data on invasive plant occurrences in the United States (Allen & Bradley 2016) along with human population density (Manson et al.
+2017) and the primary road network (US Census Bureau 2017).
+In each cell of the landscape, we binarized invasive plant observation (0: no observations, 1: at least 1 observation), the total human population density, and the total length of primary roads.
+We fit a logistic regression including quadratic terms for each covariate.
+Finally, we used the predicted probability of observation from the regression as the probability that each cell was sampled from the virtual species distributions.
+
+\subsubsection{Measurement error}
+The \emph{measurement error} sampling scenario added noise to the observations used to fit each SDM.
+Because the type of data required differs drastically between SDM methods, we adjusted the mechanism for adding error accordingly.
+In each case, datasets were first generated as described in the \emph{ideal} scenario.
+To mimic errors in geolocation or identification for MaxEnt, we randomly selected 3\% of the cells in each dataset, replacing them with cells chosen randomly from the landscape regardless of occupancy status.
+To mimic error in field measurements for the population-level datasets (i.e., CA\textsubscript{p}), we added noise to the observations of abundance ($N_{obs} = Norm(N_{true}, N_{true}*0.02)$) and seed production ($seed_{obs} = Norm(seed_{true}, seed_{true}*0.05)$).
+To mimic error in field measurements for the individual-level datasets (i.e., CA\textsubscript{i}, IPM), we added noise to the observations of growth ($g_{obs} = Norm(g_{true}, 0.1)$) and seed production ($seed_{obs} = Norm(seed_{true}, seed_{true}*0.05)$).
+Thus, counting error for seeds and individuals increased proportionally to the true values, while measurement error for size was constant.
+
+\subsection{Modelling scenarios}
+All modelling scenarios were performed using the \emph{ideal} data sets.
+Rather than addressing common issues with the data used to inform the SDMs, the focus was on the structure and implementation of the SDMs themselves.
+
+\subsubsection{Incorrect covariates}
+In the \emph{incorrect covariates} modelling scenario, we fit all SDMs with correlated but more general climatic variables than were used in the generative process for the virtual species.
+We used Mean Annual Temperature in place of the Minimum Temperature of the Coldest Month (r: 0.978; Fig.
+\ref{fig:CovScatter}a), and Annual Precipitation in place of the Precipitation in May (r: 0.624; Fig.
+\ref{fig:CovScatter}b).
+Though correlated, the geographic distribution of the difference is highly structured for both temperature and precipitation (Fig.
+\ref{fig:CovMap}).
+\begin{figure}
+	\centering
+	\captionsetup{width=\linewidth}
+	\includegraphics[width=\linewidth]{../../figs/Supp_CovScat.jpg}
+	\caption{\label{fig:CovScatter} Scatter plots of z-transformed environmental covariates.
+(a) Mean Annual Temperature and Minimum Temperature of the Coldest Month, and (b) Annual Precipitation and Precipitation in May.
+The more general variables are highly correlated with the more specific variables across the eastern United States at a 5x5km resolution.
+The true demographic rates were related to the more specific variables, while the more general were used in the \emph{incorrect covariates} modelling scenario.}
+\end{figure}
+\begin{figure}
+	\centering
+	\captionsetup{width=\linewidth}
+	\includegraphics[width=\linewidth]{../../figs/Supp_CovMap.jpg}
+	\caption{\label{fig:CovMap} Map of the difference between z-transformed environmental covariates.
+(a) Mean Annual Temperature and Minimum Temperature of the Coldest Month and (b) Annual Precipitation and Precipitation in May.
+Though highly correlated, the residuals show strong spatial patterning in each case.}
+\end{figure}
+
+\subsubsection{No seed bank}
+This scenario applied only to the process-based SDMs.
+Seed bank dynamics are difficult to quantify, and are poorly known in general relative to other aspects of species' life cycles and ecology.
+A simple solution to this problem is to assume negligible effects of the seed bank on species' distributions or that any important effects are compensated for by being implicitly included in estimates of related processes such as direct germination.
+In this modelling scenario, we assume in each process-based SDM that seed bank mortality is complete such that all seedlings in year $t$ originate from seeds produced in year $t-1$.
+
+\subsubsection{Under dispersal}
+This scenario applied only to the process-based SDMs.
+Dispersal occurred via both short distance and long distance processes, and both were reduced in the \emph{under dispersal} scenario.
+Because two virtual species had different dispersal strategies, where the shrub had more effective short distance dispersal while the biennial relied more heavily on long distance dispersal, we modified dispersal proportionally.
+For long distance dispersal, the number of annual events was decreased to 20\% of the true value.
+For short distance dispersal, the rate of the exponential kernel was decreased by 50\%.
+Consequently, the total number of seeds produced was unaffected, but chance dispersal events to new regions were less likely, and seeds were more likely to be dispersed to nearby cells within the short distance dispersal neighborhood.
+
+\subsubsection{Over dispersal}
+This scenario applied only to the process-based SDMs.
+For the \emph{over dispersal} scenario, the model mis-specification was reversed compared to the \emph{under dispersal} scenario.
+Thus, for long distance dispersal, the number of annual events was increased to 5 times the true value.
+For short distance dispersal, the rate of the exponential kernel was increased to 2 times the true value.
+Consequently, the total number of seeds produced was unaffected, but chance dispersal events to new regions were more likely, and cells toward the periphery of the short distance dispersal neighborhood were more likely to receive seeds.
+
+---
+
+\section{References}
+Allen, J.
+M., & Bradley, B.
+A.
+(2016).
+Out of the weeds?
+Reduced plant invasion risk with climate change in the continental United States.
+Biological Conservation, 203(November), 306–312.
+https://doi.org/10.1016/j.biocon.2016.09.015
+
+Ellner, S. P., & Rees, M. (2006). Integral Projection Models for Species with Complex Demography. The American Naturalist, 167(3), 410–428. https://doi.org/10.1086/499438
+
+Manson, S., Schroeder, J., Van Riper, D., & Ruggles, S. (2017). IPUMS National Historical Geographic Information System: Version 12.0 [Database]. https://doi.org/http://doi.org/10.18128/D050.V12.0
+
+Szewczyk, T.
+M., Lee, T., Ducey, M.
+J., Aiello-Lammens, M.
+E., Bibaud, H., & Allen, J.
+M.
+(2019).
+Local management in a regional context : Simulations with process-based species distribution models.
+Ecological Modelling, 413(August), 108827.
+https://doi.org/10.1016/j.ecolmodel.2019.108827
+
+U.S. Census Bureau. (2017). 2017 TIGER/Line Shapefiles Technical Documentation. Retrieved from https://www.census.gov/geo/maps-data/data/pdfs/tiger/tgrshp2012/TGRSHP2012_TechDoc.pdf

### 2020-02-05 ms/supp/App_2.Rmd
rename App A/B, added text and figures

@@
 ---
-title: 'Appendix 2: Supplemental Results'
-subtitle: 'The performance of presence-based and process-based species distribution models'
+title: 'Appendix B: Supplemental Results'
+subtitle: 'The performance of presence-based and process-based species distribution models under realistic conditions'
 author: 'Tim M. Szewczyk, Marek Petrik, Jenica M. Allen'
@@
 header-includes:
-  - \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{A.\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{A.\arabic{figure}}}  \usepackage{longtable}  \usepackage{caption}
+  - \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{B.\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{B.\arabic{figure}}}  \usepackage{longtable}  \usepackage{caption}
 ---
@@
 
-This appendix contains supplemental tables and figures to provide additional detail for the results, including maps of $\lambda$ and $N$ for each species (true, predicted, difference), as well as figures and a table for sensitivity and specificity with the same structures as those for TSS in the main manuscript.
+This appendix contains supplemental figures to provide additional detail for the results, including figures for sensitivity and specificity, and additional detail on predicted slopes for the IPM and CA\textsubscript{i} models.
+
+---
+
+\newpage
+\section{Sensitivity and specificity}
+
+\subsection{Overall averages}
+The True Skill Statistic (TSS) evaluates the combined ability of a model to predict the presences and the absences.
+While TSS is useful for comparing overall performance, the individual components can provide insight as well.
+Here, we show summary plots for the effect of each scenario on the sensitivity (the proportion of true presences correctly predicted as presences) and specificity (the proportion of true absences correctly predicted as absences).
 
 \begin{figure}
-	\centering\includegraphics[width=\linewidth]{../../figs/Supp_SensvIdeal.jpg}
-	\caption{\label{fig:SensitivityvIdeal} Effect of scenario on median sensitivity relative to the '\emph{ideal}' scenario.
+	\centering\includegraphics[width=.75\linewidth]{../../figs/Supp_Sens_mn+CI.jpg}
+	\caption{\label{fig:SensitivityMed} Sensitivity mean and 95\% confidence intervals across 100 sampled data sets for each SDM and scenario, compared to true distributions defined by $\lambda > 1$ and $N > 0$.
+Scenarios include: no sampling or modeling issues (ideal), sampling issues (measurement error, sampling bias, non-equilibrium), and modeling issues (incorrect covariates, no seed bank, under dispersal, over dispersal).
 Sensitivity represents the proportion of true presences that were correctly predicted as presences.}
@@
 \begin{figure}
-	\centering\includegraphics[width=\linewidth]{../../figs/Supp_SpecvIdeal.jpg}
-	\caption{\label{fig:SpecificityvIdeal} Effect of scenario on median specificity relative to the '\emph{ideal}' scenario.
+	\centering\includegraphics[width=.75\linewidth]{../../figs/Supp_Spec_mn+CI.jpg}
+	\caption{\label{fig:SpecificityMed} Specificity mean and 95\% confidence intervals across 100 sampled data sets for each SDM and scenario, compared to true distributions defined by $\lambda > 1$ and $N > 0$.
+Scenarios include: no sampling or modeling issues (ideal), sampling issues (measurement error, sampling bias, non-equilibrium), and modeling issues (incorrect covariates, no seed bank, under dispersal, over dispersal).
 Specificity represents the proportion of true absences that were correctly predicted as absences.}
 \end{figure}
+
+\newpage
+\subsection{Scenario effects}
+Some scenarios, such as \emph{non-equilibrium} for MaxEnt or \emph{under-dispersal} for the process-based models, result in consistent under-prediction, as seen by a decline in sensitivity and an increase in specificity.
+In contrast, \emph{incorrect covariates} has more complex effects, leading to a decline in both sensitivity and specificity in the process-based models.
+In this case, not only are presences more poorly predicted, but absences are more poorly predicted as well.
+
+\begin{figure}
+	\centering\includegraphics[width=.9\linewidth]{../../figs/Supp_SensvIdeal.jpg}
+	\caption{\label{fig:SensitivityvIdeal} Effect of scenario on median sensitivity relative to the \emph{ideal} scenario.
+Sensitivity represents the proportion of true presences that were correctly predicted as presences.}
+\end{figure}
+
+\begin{figure}
+	\centering\includegraphics[width=.9\linewidth]{../../figs/Supp_SpecvIdeal.jpg}
+	\caption{\label{fig:SpecificityvIdeal} Effect of scenario on median specificity relative to the \emph{ideal} scenario.
+Specificity represents the proportion of true absences that were correctly predicted as absences.}
+\end{figure}
+
+---
+
+\newpage
+\section{Estimated slopes}
+The IPM and CA\textsubscript{i} models fit regressions of the same form as the generative models, and so error in the estimated slopes can be directly compared.
+The pattern of accuracy across scenarios reflects the observed trends in TSS.
+The effect of data scenario is minimal, with the exception of non-equilibrium for the shrub.
+The slopes clearly show the improvement from sampling newer populations with a broader age distribution, relative to all other scenarios where populations tended to be older.
+This effect is absent from the biennial regressions, as individuals are not long-lived.
+\begin{figure}
+	\centering\includegraphics[width=.8\linewidth]{../../figs/diag/slope_box_shrub.pdf}
+	\caption{\label{fig:slopesShrub} Effect of data scenario on slope estimates for the shrub.
+Regressions are parameterized more accurately when using \emph{non-equilibrium} populations which contain a more even size distribution.}
+\end{figure}
+
+\begin{figure}
+	\centering\includegraphics[width=.8\linewidth]{../../figs/diag/slope_box_biennial.pdf}
+	\caption{\label{fig:slopesBiennial} Effect of data scenario on slope estimates for the biennial Regressions are parameterized consistently with minimal differences among data scenarios.}
+\end{figure}

### 2020-02-05 ms/supp/App_A.tex
rename App A/B, added text and figures

--- 2019-12-13 ms/supp/App_1.tex
@@
 \hypersetup{unicode=true,
-            pdftitle={Appendix 1: Supplemental Methods},
+            pdftitle={Appendix A: Supplemental Methods},
             pdfauthor={Tim M. Szewczyk, Marek Petrik, Jenica M. Allen},
@@
 \urlstyle{same}  % don't use monospace font for urls
+\usepackage{color}
+\usepackage{fancyvrb}
+\newcommand{\VerbBar}{|}
+\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
+\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
+% Add ',fontsize=\small' for more characters per line
+\usepackage{framed}
+\definecolor{shadecolor}{RGB}{248,248,248}
+\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
+\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
+\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
+\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\BuiltInTok}[1]{#1}
+\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
+\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
+\newcommand{\ExtensionTok}[1]{#1}
+\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ImportTok}[1]{#1}
+\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\NormalTok}[1]{#1}
+\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
+\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
+\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\RegionMarkerTok}[1]{#1}
+\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
 \usepackage{graphicx,grffile}
@@
 
-  \title{Appendix 1: Supplemental Methods}
+  \title{Appendix A: Supplemental Methods}
     \pretitle{\vspace{\droptitle}\centering\huge}
@@
   \subtitle{The performance of presence-based and process-based species distribution
-models}
+models under realistic conditions}
   \author{Tim M. Szewczyk, Marek Petrik, Jenica M. Allen}
@@
 This appendix contains supplemental methods pertaining to the virtual
-species generation, species distribution model specifics, and scenario
-particulars.
+species, species distribution models, and scenarios.
 
@@
 distribution the next year, and \(\boldsymbol{z_i}\) is a matrix with
-columns for: 1, \(z\), \(z^2\), and \(z^3\).
+columns of 1 (intercept) and \(z\).
 
@@
 the structure of the species distribution models. The full R code is
-available from \url{https://github.com/Sz-Tim/sdmMethodComp}. Include
-more information on the CA\textsubscript{p} model and some on the
-CA\textsubscript{i} model.
+available from \url{https://github.com/Sz-Tim/sdmMethodComp}.
+
+\subsection{IPM}
+
+The Integral Projection Models for each species followed the structure
+described in the previous section. For each dataset (100 datasets x 2
+species x 4 data scenarios), we fit regressions for each vital rate
+regression using the R package \emph{MuMIn} to compare all combinations
+of climatic variables. The full models were:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{# pr(survival)}
+\KeywordTok{glm}\NormalTok{(surv }\OperatorTok{~}\StringTok{ }\NormalTok{size }\OperatorTok{+}\StringTok{ }\NormalTok{temp }\OperatorTok{+}\StringTok{ }\NormalTok{temp_sq }\OperatorTok{+}\StringTok{ }\NormalTok{precip }\OperatorTok{+}\StringTok{ }\NormalTok{precip_sq, }\DataTypeTok{family=}\StringTok{"binomial"}\NormalTok{)}
+\CommentTok{# growth | survival}
+\KeywordTok{lm}\NormalTok{(size_next }\OperatorTok{~}\StringTok{ }\NormalTok{size }\OperatorTok{+}\StringTok{ }\NormalTok{temp }\OperatorTok{+}\StringTok{ }\NormalTok{temp_sq }\OperatorTok{+}\StringTok{ }\NormalTok{precip }\OperatorTok{+}\StringTok{ }\NormalTok{precip_sq)}
+\CommentTok{# pr(flowering | survival)}
+\KeywordTok{glm}\NormalTok{(flower }\OperatorTok{~}\StringTok{ }\NormalTok{size }\OperatorTok{+}\StringTok{ }\NormalTok{temp }\OperatorTok{+}\StringTok{ }\NormalTok{temp_sq }\OperatorTok{+}\StringTok{ }\NormalTok{precip }\OperatorTok{+}\StringTok{ }\NormalTok{precip_sq, }\DataTypeTok{family=}\StringTok{"binomial"}\NormalTok{)}
+\CommentTok{# number of seeds | flowering}
+\KeywordTok{glm}\NormalTok{(n_seed }\OperatorTok{~}\StringTok{ }\NormalTok{size }\OperatorTok{+}\StringTok{ }\NormalTok{temp }\OperatorTok{+}\StringTok{ }\NormalTok{temp_sq }\OperatorTok{+}\StringTok{ }\NormalTok{precip }\OperatorTok{+}\StringTok{ }\NormalTok{precip_sq, }\DataTypeTok{family=}\StringTok{"poisson"}\NormalTok{)}
+\CommentTok{# pr(germination)}
+\KeywordTok{glm}\NormalTok{(germ }\OperatorTok{~}\StringTok{ }\NormalTok{temp }\OperatorTok{+}\StringTok{ }\NormalTok{temp_sq }\OperatorTok{+}\StringTok{ }\NormalTok{precip }\OperatorTok{+}\StringTok{ }\NormalTok{precip_sq, }\DataTypeTok{family=}\StringTok{"binomial"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+The size distribution for recruits was taken as the sample mean and
+standard deviation of recruits in each cell. Parameters relating to seed
+bank dynamics and dispersal were assumed to be known, and the effects of
+this assumption were evaluated in the modelling mis-specification
+scenarios.
+
+\subsection{CA\textsubscript{i}}
+
+The individual-level cellular automaton model (CA\textsubscript{i})
+simulates individuals in each cell of the landscape using the structure
+of the IPM. Effectively, the fitted regressions are used to generate
+expected values for each individual, accounting for size and
+environment, and random values are then drawn from the corresponding
+distributions to simulate outcomes. For example, to assign survival, a
+value is drawn from a binomial distribution with the probability
+calculated from the fitted survival GLM. Thus, we simulated local
+processes within each cell. Regional processes were represented as the
+dispersal of seeds both within short distance dispersal neighborhoods
+and to a pre-determined number of random cells in the landscape. The
+CA\textsubscript{i} model is thus a simulation-based adaptation of the
+IPM. However, this requires imposing density dependence to both maintain
+realism and to work within the confines of modern computing
+capabilities. We implemented density dependent seedling establishment
+such that
+\(p_{est,i,t} = min(p_{est,i}, N_{rcr}/(p_{germ}(D_i+B_i+d_i(1-p_{emig}))))\),
+following the naming convention described above, effectively placing a
+cap on the number of new recruits (Ellner \& Rees 2006).
+
+\subsection{CA\textsubscript{p}}
+
+The population-level cellular automaton model (CA\textsubscript{p})
+simulates stage-structured populations in each cell of the landscape
+using population-level values for demographic rates. Thus, rather than
+simulating, for example, the survival of each individual, the
+CA\textsubscript{p} model simulates the total expected number of
+surviving individuals in each cell. We follow the model structure and
+sequence from a previously published model focused on the invasive shrub
+glossy buckthorn (\emph{Frangula alnus}) in the northeastern United
+States (Szewczyk et al.~2019). This includes the same life cycle
+processes as in the IPM and CA\textsubscript{i}: survival, growth
+(=aging), flowering, seed production, short and long distance dispersal,
+seed bank dynamics, and germination. However, individuals were not
+tracked or predicted, but rather averages or totals within age
+categories (shrub: 4 age categories; biennial: 2 age categories). We
+implemented density dependence as a ceiling-type carrying capacity,
+where the carrying capacity \(K\) was predicted for each cell. In the
+samples used to parameterize the regressions (consisting of 3 years of
+data), cells were assumed to be approximately at carrying capacity if
+the calculated \(lambda_t = N_t/N_{t-1}\) changed less than 5\% between
+consecutive years. Thus, we parameterized the following full models,
+including year as a random effect:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{# carrying capacity}
+\KeywordTok{glmer}\NormalTok{(K }\OperatorTok{~}\StringTok{ }\NormalTok{temp }\OperatorTok{+}\StringTok{ }\NormalTok{temp_sq }\OperatorTok{+}\StringTok{ }\NormalTok{precip }\OperatorTok{+}\StringTok{ }\NormalTok{precip_sq }\OperatorTok{+}\StringTok{ }\NormalTok{(}\DecValTok{1}\OperatorTok{|}\NormalTok{yr), }\DataTypeTok{family=}\StringTok{"poisson"}\NormalTok{)}
+\CommentTok{# adult survival (shrub only)}
+\KeywordTok{glmer}\NormalTok{(surv_adult }\OperatorTok{~}\StringTok{ }\NormalTok{temp }\OperatorTok{+}\StringTok{ }\NormalTok{temp_sq }\OperatorTok{+}\StringTok{ }\NormalTok{precip }\OperatorTok{+}\StringTok{ }\NormalTok{precip_sq }\OperatorTok{+}\StringTok{ }\NormalTok{(}\DecValTok{1}\OperatorTok{|}\NormalTok{yr), }\DataTypeTok{family=}\StringTok{"binomial"}\NormalTok{)}
+\CommentTok{# juvenile survival}
+\KeywordTok{glmer}\NormalTok{(surv_juv }\OperatorTok{~}\StringTok{ }\NormalTok{temp }\OperatorTok{+}\StringTok{ }\NormalTok{temp_sq }\OperatorTok{+}\StringTok{ }\NormalTok{precip }\OperatorTok{+}\StringTok{ }\NormalTok{precip_sq }\OperatorTok{+}\StringTok{ }\NormalTok{(}\DecValTok{1}\OperatorTok{|}\NormalTok{yr), }\DataTypeTok{family=}\StringTok{"binomial"}\NormalTok{)}
+\CommentTok{# pr(flowering | survival)}
+\KeywordTok{glmer}\NormalTok{(flower }\OperatorTok{~}\StringTok{ }\NormalTok{size }\OperatorTok{+}\StringTok{ }\NormalTok{temp }\OperatorTok{+}\StringTok{ }\NormalTok{temp_sq }\OperatorTok{+}\StringTok{ }\NormalTok{precip }\OperatorTok{+}\StringTok{ }\NormalTok{precip_sq }\OperatorTok{+}\StringTok{ }\NormalTok{(}\DecValTok{1}\OperatorTok{|}\NormalTok{yr), }\DataTypeTok{family=}\StringTok{"binomial"}\NormalTok{)}
+\CommentTok{# number of seeds | flowering}
+\KeywordTok{glmer}\NormalTok{(n_seed }\OperatorTok{~}\StringTok{ }\NormalTok{size }\OperatorTok{+}\StringTok{ }\NormalTok{temp }\OperatorTok{+}\StringTok{ }\NormalTok{temp_sq }\OperatorTok{+}\StringTok{ }\NormalTok{precip }\OperatorTok{+}\StringTok{ }\NormalTok{precip_sq }\OperatorTok{+}\StringTok{ }\NormalTok{(}\DecValTok{1}\OperatorTok{|}\NormalTok{yr), }\DataTypeTok{family=}\StringTok{"poisson"}\NormalTok{)}
+\CommentTok{# pr(germination)}
+\KeywordTok{glm}\NormalTok{(germ }\OperatorTok{~}\StringTok{ }\NormalTok{temp }\OperatorTok{+}\StringTok{ }\NormalTok{temp_sq }\OperatorTok{+}\StringTok{ }\NormalTok{precip }\OperatorTok{+}\StringTok{ }\NormalTok{precip_sq, }\DataTypeTok{family=}\StringTok{"binomial"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+As in the IPM and CA\textsubscript{i}, dispersal included long distance
+and short distance dispersal of seeds. Identical assumptions and
+implementations were used for the CA\textsubscript{p} model as described
+above.
 
@@
 
+\subsection{Data scenarios}
+\subsubsection{Ideal}
+
+In the \emph{ideal} scenario, each data set consisted of cells sampled
+from the equilibrium distribution at the end of the virtual species
+simulation (tmax = 300). Cells with N \textgreater{} 5 were selected
+with uniform probability. For MaxEnt, 300 occupied cells were drawn for
+each data set. For the process-based SDMs, 25 cells were drawn. Within
+each of the 25 cells, up to 100 (CA\textsubscript{i}, IPM) or 1000
+(CA\textsubscript{p}) individuals were randomly selected to be used in
+the vital rate regressions, with stratification among small, medium, and
+large individuals (CA\textsubscript{i}, IPM) or juveniles and adults
+(CA\textsubscript{p}). For cells with fewer individuals than the
+maximum, all individuals were included regardless of size. The final
+three years of the simulation were used to parameterize the
+CA\textsubscript{p} model, while only the final two years were used for
+the CA\textsubscript{i} model and IPM. The data available included
+abundance, survival, size, flowering status, fruit production, and
+germination rates. For estimating carrying capacity in the
+CA\textsubscript{p} model, we assumed cells where N changed less than
+5\% between years were approximately at carrying capacity, and used the
+average abundance across years for those cells.
+
+\subsubsection{Non-equilibrium}
+
+The sampling regime under the \emph{non-equilibrium} scenario was
+identical to that of the \emph{ideal} scenario, except that samples were
+drawn from the distribution 100 years after introduction rather than the
+equilibrium distribution of 300 years.
+
+\subsubsection{Geographic bias}
+
+The \emph{geographic bias} sampling scenario represents the bias in
+sampling locations common in ecological studies, where data are more
+likely to be acquired in locations that are easier to reach. To
+establish realistic patterns of data collection, we used previously
+published data on invasive plant occurrences in the United States (Allen
+\& Bradley 2016) along with human population density (Manson et
+al.~2017) and the primary road network (US Census Bureau 2017). In each
+cell of the landscape, we binarized invasive plant observation (0: no
+observations, 1: at least 1 observation), the total human population
+density, and the total length of primary roads. We fit a logistic
+regression including quadratic terms for each covariate. Finally, we
+used the predicted probability of observation from the regression as the
+probability that each cell was sampled from the virtual species
+distributions.
+
+\subsubsection{Measurement error}
+
+The \emph{measurement error} sampling scenario added noise to the
+observations used to fit each SDM. Because the type of data required
+differs drastically between SDM methods, we adjusted the mechanism for
+adding error accordingly. In each case, datasets were first generated as
+described in the \emph{ideal} scenario. To mimic errors in geolocation
+or identification for MaxEnt, we randomly selected 3\% of the cells in
+each dataset, replacing them with cells chosen randomly from the
+landscape regardless of occupancy status. To mimic error in field
+measurements for the population-level datasets (i.e.,
+CA\textsubscript{p}), we added noise to the observations of abundance
+(\(N_{obs} = Norm(N_{true}, N_{true}*0.02)\)) and seed production
+(\(seed_{obs} = Norm(seed_{true}, seed_{true}*0.05)\)). To mimic error
+in field measurements for the individual-level datasets (i.e.,
+CA\textsubscript{i}, IPM), we added noise to the observations of growth
+(\(g_{obs} = Norm(g_{true}, 0.1)\)) and seed production
+(\(seed_{obs} = Norm(seed_{true}, seed_{true}*0.05)\)). Thus, counting
+error for seeds and individuals increased proportionally to the true
+values, while measurement error for size was constant.
+
+\subsection{Modelling scenarios}
+
+All modelling scenarios were performed using the \emph{ideal} data sets.
+Rather than addressing common issues with the data used to inform the
+SDMs, the focus was on the structure and implementation of the SDMs
+themselves.
+
+\subsubsection{Incorrect covariates}
+
+In the \emph{incorrect covariates} modelling scenario, we fit all SDMs
+with correlated but more general climatic variables than were used in
+the generative process for the virtual species. We used Mean Annual
+Temperature in place of the Minimum Temperature of the Coldest Month (r:
+0.978; Fig. \ref{fig:CovScatter}a), and Annual Precipitation in place of
+the Precipitation in May (r: 0.624; Fig. \ref{fig:CovScatter}b). Though
+correlated, the geographic distribution of the difference is highly
+structured for both temperature and precipitation (Fig.
+\ref{fig:CovMap}).
+
+\begin{figure}
+    \centering
+    \captionsetup{width=\linewidth}
+    \includegraphics[width=\linewidth]{../../figs/Supp_CovScat.jpg}
+    \caption{\label{fig:CovScatter} Scatter plots of z-transformed environmental covariates.
+(a) Mean Annual Temperature and Minimum Temperature of the Coldest Month, and (b) Annual Precipitation and Precipitation in May.
+The more general variables are highly correlated with the more specific variables across the eastern United States at a 5x5km resolution.
+The true demographic rates were related to the more specific variables, while the more general were used in the \emph{incorrect covariates} modelling scenario.}
+\end{figure}
+\begin{figure}
+    \centering
+    \captionsetup{width=\linewidth}
+    \includegraphics[width=\linewidth]{../../figs/Supp_CovMap.jpg}
+    \caption{\label{fig:CovMap} Map of the difference between z-transformed environmental covariates.
+(a) Mean Annual Temperature and Minimum Temperature of the Coldest Month and (b) Annual Precipitation and Precipitation in May.
+Though highly correlated, the residuals show strong spatial patterning in each case.}
+\end{figure}
+
+\subsubsection{No seed bank}
+
+This scenario applied only to the process-based SDMs. Seed bank dynamics
+are difficult to quantify, and are poorly known in general relative to
+other aspects of species' life cycles and ecology. A simple solution to
+this problem is to assume negligible effects of the seed bank on
+species' distributions or that any important effects are compensated for
+by being implicitly included in estimates of related processes such as
+direct germination. In this modelling scenario, we assume in each
+process-based SDM that seed bank mortality is complete such that all
+seedlings in year \(t\) originate from seeds produced in year \(t-1\).
+
+\subsubsection{Under dispersal}
+
+This scenario applied only to the process-based SDMs. Dispersal occurred
+via both short distance and long distance processes, and both were
+reduced in the \emph{under dispersal} scenario. Because two virtual
+species had different dispersal strategies, where the shrub had more
+effective short distance dispersal while the biennial relied more
+heavily on long distance dispersal, we modified dispersal
+proportionally. For long distance dispersal, the number of annual events
+was decreased to 20\% of the true value. For short distance dispersal,
+the rate of the exponential kernel was decreased by 50\%. Consequently,
+the total number of seeds produced was unaffected, but chance dispersal
+events to new regions were less likely, and seeds were more likely to be
+dispersed to nearby cells within the short distance dispersal
+neighborhood.
+
+\subsubsection{Over dispersal}
+
+This scenario applied only to the process-based SDMs. For the
+\emph{over dispersal} scenario, the model mis-specification was reversed
+compared to the \emph{under dispersal} scenario. Thus, for long distance
+dispersal, the number of annual events was increased to 5 times the true
+value. For short distance dispersal, the rate of the exponential kernel
+was increased to 2 times the true value. Consequently, the total number
+of seeds produced was unaffected, but chance dispersal events to new
+regions were more likely, and cells toward the periphery of the short
+distance dispersal neighborhood were more likely to receive seeds.
+
+\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
+
+\section{References}
+
+Allen, J. M., \& Bradley, B. A. (2016). Out of the weeds? Reduced plant
+invasion risk with climate change in the continental United States.
+Biological Conservation, 203(November), 306--312.
+\url{https://doi.org/10.1016/j.biocon.2016.09.015}
+
+Ellner, S. P., \& Rees, M. (2006). Integral Projection Models for
+Species with Complex Demography. The American Naturalist, 167(3),
+410--428. \url{https://doi.org/10.1086/499438}
+
+Manson, S., Schroeder, J., Van Riper, D., \& Ruggles, S. (2017). IPUMS
+National Historical Geographic Information System: Version 12.0
+{[}Database{]}. \url{https://doi.org/http://doi.org/10.18128/D050.V12.0}
+
+Szewczyk, T. M., Lee, T., Ducey, M. J., Aiello-Lammens, M. E., Bibaud,
+H., \& Allen, J. M. (2019). Local management in a regional context :
+Simulations with process-based species distribution models. Ecological
+Modelling, 413(August), 108827.
+\url{https://doi.org/10.1016/j.ecolmodel.2019.108827}
+
+U.S. Census Bureau. (2017). 2017 TIGER/Line Shapefiles Technical
+Documentation. Retrieved from
+\url{https://www.census.gov/geo/maps-data/data/pdfs/tiger/tgrshp2012/TGRSHP2012_TechDoc.pdf}
+
 \end{document}

### 2020-02-05 ms/supp/App_B.tex
rename App A/B, added text and figures

--- 2020-02-05 ms/supp/App_2.Rmd
@@
----
-title: 'Appendix B: Supplemental Results'
-subtitle: 'The performance of presence-based and process-based species distribution models under realistic conditions'
-author: 'Tim M. Szewczyk, Marek Petrik, Jenica M. Allen'
-output:
-  pdf_document: 
-    number_sections: yes
-    toc: yes
-    toc_depth: 1
-  html_document:
-    number_sections: yes
-    toc: yes
-    toc_depth: 1
-editor_options:
-  chunk_output_type: inline
-header-includes:
-  - \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{B.\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{B.\arabic{figure}}}  \usepackage{longtable}  \usepackage{caption}
----
+\documentclass[]{article}
+\usepackage{lmodern}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+\else % if luatex or xelatex
+  \ifxetex
+    \usepackage{mathspec}
+  \else
+    \usepackage{fontspec}
+  \fi
+  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
+\fi
+% use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+% use microtype if available
+\IfFileExists{microtype.sty}{%
+\usepackage{microtype}
+\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\usepackage[margin=1in]{geometry}
+\usepackage{hyperref}
+\hypersetup{unicode=true,
+            pdftitle={Appendix B: Supplemental Results},
+            pdfauthor={Tim M. Szewczyk, Marek Petrik, Jenica M. Allen},
+            pdfborder={0 0 0},
+            breaklinks=true}
+\urlstyle{same}  % don't use monospace font for urls
+\usepackage{graphicx,grffile}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\IfFileExists{parskip.sty}{%
+\usepackage{parskip}
+}{% else
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
+}
+\setlength{\emergencystretch}{3em}  % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{5}
+% Redefines (sub)paragraphs to behave more like sections
+\ifx\paragraph\undefined\else
+\let\oldparagraph\paragraph
+\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+\let\oldsubparagraph\subparagraph
+\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
 
-\beginsupplement
+%%% Use protect on footnotes to avoid problems with footnotes in titles
+\let\rmarkdownfootnote\footnote%
+\def\footnote{\protect\rmarkdownfootnote}
 
----
+%%% Change title format to be more compact
+\usepackage{titling}
 
-This appendix contains supplemental figures to provide additional detail for the results, including figures for sensitivity and specificity, and additional detail on predicted slopes for the IPM and CA\textsubscript{i} models.
+% Create subtitle command for use in maketitle
+\providecommand{\subtitle}[1]{
+  \posttitle{
+    \begin{center}\large#1\end{center}
+    }
+}
 
----
+\setlength{\droptitle}{-2em}
+
+  \title{Appendix B: Supplemental Results}
+    \pretitle{\vspace{\droptitle}\centering\huge}
+  \posttitle{\par}
+  \subtitle{The performance of presence-based and process-based species distribution
+models under realistic conditions}
+  \author{Tim M. Szewczyk, Marek Petrik, Jenica M. Allen}
+    \preauthor{\centering\large\emph}
+  \postauthor{\par}
+    \date{}
+    \predate{}\postdate{}
+
+\newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{B.\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{B.\arabic{figure}}}  \usepackage{longtable}  \usepackage{caption}
+
+\begin{document}
+\maketitle
+
+{
+\setcounter{tocdepth}{1}
+\tableofcontents
+}
+\setcounter{table}{0}  \renewcommand{\thetable}{B.\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{B.\arabic{figure}}
+
+\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
+
+This appendix contains supplemental tables and figures to provide
+additional detail for the results, including maps of \(\lambda\) and
+\(N\) for each species (true, predicted, difference), figures and tables
+for sensitivity and specificity, and additional detail on predicted
+slopes for the IPM and CA\textsubscript{i} models.
+
+\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
+
+\section{Prediction maps}
+
+\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
 
@@
 \subsection{Overall averages}
-The True Skill Statistic (TSS) evaluates the combined ability of a model to predict the presences and the absences.
-While TSS is useful for comparing overall performance, the individual components can provide insight as well.
-Here, we show summary plots for the effect of each scenario on the sensitivity (the proportion of true presences correctly predicted as presences) and specificity (the proportion of true absences correctly predicted as absences).
 
+The True Skill Statistic (TSS) evaluates the combined ability of a model
+to predict the presences and the absences. While TSS is useful for
+comparing overall performance, the individual components can provide
+insight as well. Here, we show summary plots for the effect of each
+scenario on the sensitivity (the proportion of true presences correctly
+predicted as presences) and specificity (the proportion of true absences
+correctly predicted as absences).
+
+\newpage
 \begin{figure}
@@
 
+\newpage
 \begin{figure}
@@
 \subsection{Scenario effects}
-Some scenarios, such as \emph{non-equilibrium} for MaxEnt or \emph{under-dispersal} for the process-based models, result in consistent under-prediction, as seen by a decline in sensitivity and an increase in specificity.
-In contrast, \emph{incorrect covariates} has more complex effects, leading to a decline in both sensitivity and specificity in the process-based models.
-In this case, not only are presences more poorly predicted, but absences are more poorly predicted as well.
+
+Some scenarios, such as \emph{non-equilibrium} for MaxEnt or
+\emph{under-dispersal} for the process-based models, result in
+consistent under-prediction, as seen by a decline in sensitivity and an
+increase in specificity. In contrast, \emph{incorrect covariates} has
+more complex effects, leading to a decline in both sensitivity and
+specificity in the process-based models. In this case, not only are
+presences more poorly predicted, but absences are more poorly predicted
+as well.
 
@@
 
----
+\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
 
@@
 \section{Estimated slopes}
-The IPM and CA\textsubscript{i} models fit regressions of the same form as the generative models, and so error in the estimated slopes can be directly compared.
-The pattern of accuracy across scenarios reflects the observed trends in TSS.
-The effect of data scenario is minimal, with the exception of non-equilibrium for the shrub.
-The slopes clearly show the improvement from sampling newer populations with a broader age distribution, relative to all other scenarios where populations tended to be older.
-This effect is absent from the biennial regressions, as individuals are not long-lived.
+
+The IPM and CA\textsubscript{i} models fit regressions of the same form
+as the generative models, and so error in the estimated slopes can be
+directly compared. The pattern of accuracy across scenarios reflects the
+observed trends in TSS. The effect of data scenario is minimal, with the
+exception of non-equilibrium for the shrub. The slopes clearly show the
+improvement from sampling newer populations with a broader age
+distribution, relative to all other scenarios where populations tended
+to be older. This effect is absent from the biennial regressions, as
+individuals are not long-lived.
+
 \begin{figure}
@@
 \end{figure}
+
+\end{document}

### 2020-02-07 ms/supp/App_1.Rmd
slope comparisons updated

@@
 \section{General model structure}
-\subsection{Integral Projection Model overview}
+\subsection{Integral Projection Models (IPMs)}
 To generate fully-known true distributions for the virtual species, we used the general structure of an Integral Projection Model (IPM) to calculate the intrinsic growth rate $\lambda$ in each cell of the gridded landscape, and adapted the regression-based structure of an IPM into an individual-level, simulation-based cellular automata (CA) model to produce spatiotemporally dynamic abundance distributions.
-The equations below apply to both models.
 
-IPMs use the size distribution, $z$, of individuals at time $t$, along with a kernel $K(z^{\prime}, z)$, to predict the size distribution, $z^{\prime}$, at time $t+1$:
+IPMs (Easterling 2000; Merow et al.
+2014, 2017) use the size distribution, $z$, of individuals at time $t$, along with a kernel $K(z^{\prime}, z)$, to predict the size distribution, $z^{\prime}$, at time $t+1$:
 \begin{equation}
@@
 Thus, the population in each cell $i$ has size distribution $z$, where $z^{\prime}$ is the size distribution the next year, and $\boldsymbol{z_i}$ is a matrix with columns of 1 (intercept) and $z$.
+For the IPMs, we used the following regression equations to populate discretized transition matrices for each cell to then calculate the deterministic intrinsic growth rate $\lambda$.
+For the CA\textsubscript{i} models, we used the same regression equations to generate expected values for each individual, and then drew simulated outcomes from the specified distribtutions.
 
@@
 
+Easterling, M.
+R., Ellner, S.
+P., & Dixon, P.
+M.
+(2000).
+Size-specific sensitivity: Applying a new structured population model.
+Ecology, 81(3), 694–708.
+https://doi.org/10.1890/0012-9658(2000)081[0694:SSSAAN]2.0.CO;2
+
 Ellner, S. P., & Rees, M. (2006). Integral Projection Models for Species with Complex Demography. The American Naturalist, 167(3), 410–428. https://doi.org/10.1086/499438
@@
 Manson, S., Schroeder, J., Van Riper, D., & Ruggles, S. (2017). IPUMS National Historical Geographic Information System: Version 12.0 [Database]. https://doi.org/http://doi.org/10.18128/D050.V12.0
+
+Merow, C., Dahlgren, J.
+P., Metcalf, C.
+J.
+E., Childs, D.
+Z., Evans, M.
+E.
+K., Jongejans, E., … Mcmahon, S.
+M.
+(2014).
+Advancing population ecology with integral projection models: A practical guide.
+Methods in Ecology and Evolution, 5(2), 99–110.
+https://doi.org/10.1111/2041-210X.12146
+
+Merow, C., Bois, S.
+T., Allen, J.
+M., Xie, Y., & Silander, J.
+A.
+(2017).
+Climate change both facilitates and inhibits invasive plant ranges in New England.
+Proceedings of the National Academy of Sciences, 114(16), E3276–E3284.
+https://doi.org/10.1073/pnas.1609633114
 

### 2020-02-07 ms/supp/App_2.Rmd
slope comparisons updated

@@
 \begin{figure}
-	\centering\includegraphics[width=.8\linewidth]{../../figs/diag/slope_box_shrub.pdf}
-	\caption{\label{fig:slopesShrub} Effect of data scenario on slope estimates for the shrub.
-Regressions are parameterized more accurately when using \emph{non-equilibrium} populations which contain a more even size distribution.}
+	\centering\includegraphics[width=\linewidth]{../../figs/diag/slope_dens_shrub.pdf}
+	\caption{\label{fig:slopesShrub} Effect of data scenario on slope estimates for the vital rate regressions for the shrub.
+Each density curve shows the distribution of fitted slope values across 100 data sets used in the IPM and CA\textsubscript{i} models.
+The true value is indicated by the vertical dashed line.
+For (a) survival probability, the slopes are estimated more accurately when using \emph{non-equilibrium} populations which contain a more even size distribution.
+In contrast, (b) growth, (c) flowering probability, and (e) germination probability show little variation among scenarios, while for (d) seed production, added \emph{measurement error} increased variability.
+Note that the axes vary among panels.}
 \end{figure}
@@
 \begin{figure}
-	\centering\includegraphics[width=.8\linewidth]{../../figs/diag/slope_box_biennial.pdf}
-	\caption{\label{fig:slopesBiennial} Effect of data scenario on slope estimates for the biennial Regressions are parameterized consistently with minimal differences among data scenarios.}
+	\centering\includegraphics[width=\linewidth]{../../figs/diag/slope_dens_biennial.pdf}
+	\caption{\label{fig:slopesBiennial} Effect of data scenario on slope estimates for the vital rate regressions for the biennial.
+Each density curve shows the distribution of fitted slope values across 100 data sets used in the IPM and CA\textsubscript{i} models.
+The true value is indicated by the vertical dashed line.
+The fitted slopes were consistent with minimal differences among data scenarios for (a) survival probability, (b) growth, (c) flowering probability, (d) seed production, and (e) germination probability.
+Peaks at 0 represent data sets where the optimal model did not include that term.
+Note that the axes vary among panels.}
 \end{figure}

### 2020-02-07 ms/supp/App_A.tex
slope comparisons updated

@@
 into an individual-level, simulation-based cellular automata (CA) model
-to produce spatiotemporally dynamic abundance distributions. The
-equations below apply to both models.
+to produce spatiotemporally dynamic abundance distributions.
 
-IPMs use the size distribution, \(z\), of individuals at time \(t\),
-along with a kernel \(K(z^{\prime}, z)\), to predict the size
-distribution, \(z^{\prime}\), at time \(t+1\): \begin{equation}
+IPMs (Easterling 2000; Merow et al.~2014, 2017) use the size
+distribution, \(z\), of individuals at time \(t\), along with a kernel
+\(K(z^{\prime}, z)\), to predict the size distribution, \(z^{\prime}\),
+at time \(t+1\): \begin{equation}
 n_{t+1}(z^{\prime}) = \int_{\Omega} K(z^{\prime}, z) n_t(z) dz
@@
 distribution the next year, and \(\boldsymbol{z_i}\) is a matrix with
-columns of 1 (intercept) and \(z\).
+columns of 1 (intercept) and \(z\). For the IPMs, we used the following
+regression equations to populate discretized transition matrices for
+each cell to then calculate the deterministic intrinsic growth rate
+\(\lambda\). For the CA\textsubscript{i} models, we used the same
+regression equations to generate expected values for each individual,
+and then drew simulated outcomes from the specified distribtutions.
 
@@
 
+Easterling, M. R., Ellner, S. P., \& Dixon, P. M. (2000). Size-specific
+sensitivity: Applying a new structured population model. Ecology, 81(3),
+694--708.
+\url{https://doi.org/10.1890/0012-9658(2000)081\%5B0694:SSSAAN\%5D2.0.CO;2}
+
 Ellner, S. P., \& Rees, M. (2006). Integral Projection Models for
@@
 {[}Database{]}. \url{https://doi.org/http://doi.org/10.18128/D050.V12.0}
+
+Merow, C., Dahlgren, J. P., Metcalf, C. J. E., Childs, D. Z., Evans, M.
+E. K., Jongejans, E., \ldots{} Mcmahon, S. M. (2014). Advancing
+population ecology with integral projection models: A practical guide.
+Methods in Ecology and Evolution, 5(2), 99--110.
+\url{https://doi.org/10.1111/2041-210X.12146}
+
+Merow, C., Bois, S. T., Allen, J. M., Xie, Y., \& Silander, J. A.
+(2017). Climate change both facilitates and inhibits invasive plant
+ranges in New England. Proceedings of the National Academy of Sciences,
+114(16), E3276--E3284. \url{https://doi.org/10.1073/pnas.1609633114}
 

### 2020-02-07 ms/supp/App_B.tex
slope comparisons updated

@@
 
-This appendix contains supplemental tables and figures to provide
-additional detail for the results, including maps of \(\lambda\) and
-\(N\) for each species (true, predicted, difference), figures and tables
-for sensitivity and specificity, and additional detail on predicted
-slopes for the IPM and CA\textsubscript{i} models.
-
-\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
-
-\section{Prediction maps}
+This appendix contains supplemental figures to provide additional detail
+for the results, including figures for sensitivity and specificity, and
+additional detail on predicted slopes for the IPM and
+CA\textsubscript{i} models.
 
@@
 
-\newpage
 \begin{figure}
@@
 \begin{figure}
-    \centering\includegraphics[width=.8\linewidth]{../../figs/diag/slope_box_shrub.pdf}
+    \centering\includegraphics[width=\linewidth]{../../figs/diag/slope_dens_shrub.pdf}
     \caption{\label{fig:slopesShrub} Effect of data scenario on slope estimates for the shrub.
-Regressions are parameterized more accurately when using \emph{non-equilibrium} populations which contain a more even size distribution.}
+Survival regressions for the shrub are parameterized more accurately when using \emph{non-equilibrium} populations which contain a more even size distribution.
+Note that the axes vary among panels.}
 \end{figure}
@@
 \begin{figure}
-    \centering\includegraphics[width=.8\linewidth]{../../figs/diag/slope_box_biennial.pdf}
-    \caption{\label{fig:slopesBiennial} Effect of data scenario on slope estimates for the biennial Regressions are parameterized consistently with minimal differences among data scenarios.}
+    \centering\includegraphics[width=\linewidth]{../../figs/diag/slope_dens_biennial.pdf}
+    \caption{\label{fig:slopesBiennial} Effect of data scenario on slope estimates for the biennial Regressions are parameterized consistently with minimal differences among data scenarios.
+Peaks at 0 represent data sets where the optimal model did not include that term.
+Note that the axes vary among panels.}
 \end{figure}

### 2020-02-07 code/check_diagnostics.R

@@ -1,8 +1,9 @@
 
 sp <- c("barberry", "garlic_mustard")[2]
+spType <- c("shrub", "biennial")[(sp=="garlic_mustard")+1]
 write.plots <- TRUE
 by.issue <- FALSE
-pkgs <- c("tidyverse", "magrittr", "stringr", "here", "viridis")
+pkgs <- c("tidyverse", "magrittr", "stringr", "here", "viridis", "gridExtra")
 suppressMessages(invisible(lapply(pkgs, library, character.only=T)))
 walk(dir("code", "fn", full.names=T), source)
 sp_i <- read.csv("data/species_3km.csv") %>% filter(Name==sp)
@@ -39,6 +40,54 @@ x2.mx <- cbind(x2, x2^2)
 
 
 ########-------------------------------
+## IPM: Slope comparisons
+########
+
+param.ls <- list(c("s_z", "s_x"), c("g_z", "g_x"), 
+                 c("fl_z", "fl_x"), c("seed_z", "seed_x"), 
+                 c("germ_x"))
+ipm.slopes.ls <- ipm.true.ls <- gg.p <- vector("list", length(param.ls))
+reg_vars <- c("(Intercept)"="(Intercept)", "size"="size",
+              "minTemp"="bio10_6", "minTemp^2"="bio10_6_sq", 
+              "precMay"="bio10_prMay", "precMay^2"="bio10_prMay_sq")
+paramTypes <- c("(a) survival", "(b) growth", "(c) flowering", 
+                "(d) seeds", "(e) germination")
+for(i in seq_along(param.ls)) {
+  params <- param.ls[[i]]
+  params <- setNames(params, str_split_fixed(params, "_", 2)[,2])
+  paramType <- paramTypes[i]
+  ipm.slopes <- map_dfr(params, ~plot_sdm_slopes(p, ipm.diag, .)$slope.df) %>%
+    mutate(variable=factor(variable, levels=reg_vars, labels=names(reg_vars)),
+           param=paramType,
+           scenario=factor(scenario, 
+                           levels=c("none", "noise", "sampBias", "nonEq"),
+                           labels=c("Ideal", "Measurement error",
+                                    "Sampling bias", "Non-equilibrium")))
[…]

### 2020-02-13 ms/supp/App_1.Rmd
updated appendixes

@@
 
-\section{General model structure}
-\subsection{Integral Projection Models (IPMs)}
+\section{Integral Projection Model structure}
+\subsection{Overview}
 To generate fully-known true distributions for the virtual species, we used the general structure of an Integral Projection Model (IPM) to calculate the intrinsic growth rate $\lambda$ in each cell of the gridded landscape, and adapted the regression-based structure of an IPM into an individual-level, simulation-based cellular automata (CA) model to produce spatiotemporally dynamic abundance distributions.
@@
 \newpage
-\section{Regression equations for the virtual species}
+\section{Virtual species simulation}
 We used the above IPM structure, including a seed bank and dispersal, as the basis for the virtual species.

### 2020-02-13 ms/supp/App_2.Rmd
updated appendixes

@@
 
-\newpage
 \section{Sensitivity and specificity}
@@
 Here, we show summary plots for the effect of each scenario on the sensitivity (the proportion of true presences correctly predicted as presences) and specificity (the proportion of true absences correctly predicted as absences).
+\clearpage
 
@@
 
-\newpage
+\clearpage
 \subsection{Scenario effects}
@@
 
+\clearpage
 \begin{figure}
@@
 
----
-
-\newpage
+\clearpage
 \section{Estimated slopes}
@@
 The pattern of accuracy across scenarios reflects the observed trends in TSS.
-The effect of data scenario is minimal, with the exception of non-equilibrium for the shrub.
+The effect of data scenario is minimal, with the exception of \emph{non-equilibrium} for the shrub survival probability regressions.
 The slopes clearly show the improvement from sampling newer populations with a broader age distribution, relative to all other scenarios where populations tended to be older.
 This effect is absent from the biennial regressions, as individuals are not long-lived.
+Additionally, slopes for the growth regression were poorly estimated for the shrub, resulting from the imposition of a size maximum in the virtual species.
+
 \begin{figure}

### 2020-02-13 ms/supp/App_B.tex
updated appendixes

@@
 predicted as presences) and specificity (the proportion of true absences
-correctly predicted as absences).
+correctly predicted as absences). \clearpage
 
@@
 
-\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
-
-\newpage
+\clearpage
 \section{Estimated slopes}
@@
 observed trends in TSS. The effect of data scenario is minimal, with the
-exception of non-equilibrium for the shrub. The slopes clearly show the
-improvement from sampling newer populations with a broader age
-distribution, relative to all other scenarios where populations tended
-to be older. This effect is absent from the biennial regressions, as
-individuals are not long-lived.
+exception of \emph{non-equilibrium} for the shrub survival probability
+regressions. The slopes clearly show the improvement from sampling newer
+populations with a broader age distribution, relative to all other
+scenarios where populations tended to be older. This effect is absent
+from the biennial regressions, as individuals are not long-lived.
+Additionally, slopes for the growth regression were poorly estimated for
+the shrub, resulting from the imposition of a size maximum in the
+virtual species.
 
@@
     \centering\includegraphics[width=\linewidth]{../../figs/diag/slope_dens_shrub.pdf}
-    \caption{\label{fig:slopesShrub} Effect of data scenario on slope estimates for the shrub.
-Survival regressions for the shrub are parameterized more accurately when using \emph{non-equilibrium} populations which contain a more even size distribution.
+    \caption{\label{fig:slopesShrub} Effect of data scenario on slope estimates for the vital rate regressions for the shrub.
+Each density curve shows the distribution of fitted slope values across 100 data sets used in the IPM and CA\textsubscript{i} models.
+The true value is indicated by the vertical dashed line.
+For (a) survival probability, the slopes are estimated more accurately when using \emph{non-equilibrium} populations which contain a more even size distribution.
+In contrast, (b) growth, (c) flowering probability, and (e) germination probability show little variation among scenarios, while for (d) seed production, added \emph{measurement error} increased variability.
 Note that the axes vary among panels.}
@@
     \centering\includegraphics[width=\linewidth]{../../figs/diag/slope_dens_biennial.pdf}
-    \caption{\label{fig:slopesBiennial} Effect of data scenario on slope estimates for the biennial Regressions are parameterized consistently with minimal differences among data scenarios.
+    \caption{\label{fig:slopesBiennial} Effect of data scenario on slope estimates for the vital rate regressions for the biennial.
+Each density curve shows the distribution of fitted slope values across 100 data sets used in the IPM and CA\textsubscript{i} models.
+The true value is indicated by the vertical dashed line.
+The fitted slopes were consistent with minimal differences among data scenarios for (a) survival probability, (b) growth, (c) flowering probability, (d) seed production, and (e) germination probability.
 Peaks at 0 represent data sets where the optimal model did not include that term.

### 2020-02-13 code/4_analysis.R
calculating values for both species

@@ -8,7 +8,7 @@
 ## Setup
 ########
 # file specifications
-sp <- "sp2"
+sp <- "sp1"
 overwrite <- TRUE
 
 # load workspace
@@ -33,6 +33,13 @@ out_P <- map2(f_P, i_P, ~readRDS(.x) %>% mutate(SDM=.y[1], issue=.y[2])) %>%
                             lambda<1 & prP>=0.5 ~ "S:0 P:1",
                             lambda>=1 & prP<0.5 ~ "S:1 P:0",
                             lambda<1 & prP<0.5 ~ "S:0 P:0"))
+out_P$B.f[out_P$issue=="noSB"] <- NA
+
+# Compare issues to none
+out_P <- full_join(out_P, 
+                   filter(out_P, issue=="none") %>% select(-c(1:79, 92)), 
+                   by=c("id", "SDM"), suffix=c("", ".none")) %>%
+  mutate(sp=sp)
 
 # Calculate log likelihood
 out_LL <- out_P %>% 
@@ -43,7 +50,8 @@ out_LL <- out_P %>%
   group_by(SDM, issue) %>%
   summarise(LL.N=sum(logP.N),
             LL.lam=sum(logP.lam)) %>%
-  gather(Boundary, LogLik, 3:4)
+  gather(Boundary, LogLik, 3:4) %>%
+  mutate(sp=sp)
 out_LL$Boundary <- factor(out_LL$Boundary, labels=c("lambda > 1", "N > 0"))
 
 # Aggregate TSS
@@ -62,7 +70,8 @@ out_TSS <- rbind(map(f_TSS, ~readRDS(.)[[1]]) %>%
                                     sensitivity=map_dbl(.x, ~.$sensitivity), 
                                     specificity=map_dbl(.x, ~.$specificity), 
                                     Boundary="lambda > 1", 
-                                    SDM=.y[1], issue=.y[2])))
+                                    SDM=.y[1], issue=.y[2]))) %>%
+  mutate(sp=sp)
 
 # Calculate MaxEnt diagnostics
 f_MxE <- list.files(here("out", sp), pattern="Diag_MxE", full.names=T)
@@ -72,7 +81,8 @@ AUC_MxE <- map2_dfr(diagRaw_MxE, i_MxE,
                      ~tibble(AUC=c(map_dbl(.x[[1]], ~.@auc), 
                                    map_dbl(.x[[2]], ~.@auc)),
[…]

### 2020-02-13 ms/ms_latex/ms_sdmComp.tex

@@
 
-\title{The performance of presence-based and process-based species distribution models}
+\title{The performance of presence-based and process-based species distribution models under realistic conditions}
 
@@
 \begin{keyword}
-keywords
+integral projection model; cellular automaton; virtual species; demography; niche model; mis-specification
 \end{keyword}
@@
 \label{S:1}
-% SDMs and the prevalence of occurrenc-based methods (And)
-1. General introduction about species distribution models and the historical prevalence of occurrence-based SDMs.
+% SDMs and the prevalence of occurrence-based methods (And)
+Species distribution models (SDMs) have been used extensively to predict where species are expected to occur, both presently and in the future \citep{Guisan2005,Phillips2008,Elith2009,Allen2016,Araujo2019}.
+The majority of SDMs found in the literature are occurrence-based, using approaches such as Maxent to predict a species' distribution based on the relationship of a set of geo-referenced detections with environmental variables \citep{Elith2009}.
+A wide variety of techniques have been developed \citep{Elith2009,Fitzpatrick2013,Norberg2019}, with much attention in the literature to the relative strengths and weaknesses of different approaches \citep{Elith2009,Merow2014b,Fernandes2018,Norberg2019}, and resulting strategies for avoiding common biases or errors in available datasets \citep{Merow2013,Boria2014,Merow2016}.
+In some cases, local absence data are available as well to help inform these models \citep{Koshkina2017}.
+Under many conditions, these methods are highly capable of predicting current distributions CITE, and several large online databases provide easy access to adequate and ever-growing datasets CITE.
 
 % Process-based SDMs as an alternative, including a specific example or two (But)
-2. The rise of process-based SDMs (or at least the rise in advocacy for process-based SDMs), including some specific examples of their use, since that was the main question raised at ISEM.
+However, occurrence-based methods require several assumptions which may limit their use in key conservation situations.
+In particular, occurrence-based SDMs assume that populations are at equilibrium with the environment, such that the present distribution reflects constraints due to broad environmental conditions \citep{Elith2009}.
+Extrapolating to novel environmental conditions is frequently unreliable, as evidenced by the mismatches seen between the observed and predicted ranges of species in non-native continents CITE.
+In a world with a changing climate, the assumption of distributions at equilibrium is already invalid.
+Distributional shifts consistent with a warming climate have been observed in many taxa, ranging from plants \citep{Lenoir2010,Engler2011} to mammals \citep{Moritz2008a,Rowe2015} to birds \citep{Tingley2009a,Tingley2012} to moths \citep{Chen2009b}.
+The disparity between past occurrences and current climate will continue to increase.
+Process-based SDMs have been proposed as an alternative approach that may be less sensitive to these key limitations of occurrence-based SDMs \citep{Merow2011a,Evans2016,Merow2017,Cabral2017}.
+Rather than occurrences, process-based SDMs rely on geographically varying demographic rates, such as survival or flowering probability, to ultimately predict the distribution as an emergent property of environmentally dependent population dynamics \citep{Evans2016,Cabral2017}.
 
 % Expected benefits of process-based SDMs (And)
-3. Expected benefits of process-based SDMs
+Process-based SDMs are a set of highly flexible of modelling approaches which encompass individual-based and population-based methods, as well as analytical and simulation-based techniques.
+For example, Integral Projection Models (IPMs) incorporate effects of individual size in addition to climatic conditions to predict the intrinsic growth rate, $\lambda$, in each pixel of the landscape via demographic rates \citep{Easterling2000,Merow2017}.
+In contrast, cellular automata (CA) models simulate the movement of individuals across a gridded landscape, incorporating regional dispersal dynamics with local population growth based on individual-level or population-level demography \citep{Merow2011a,Szewczyk2019}.
+Process-based SDMs thus have several expected advantages over occurrence-based SDMs.
+First, process-based SDMs predict the rates of biological processes across a geographic region using regressions with environmental conditions.
+Consequently, there are no assumptions about equilibrium between the climate and the current distribution, but rather that the data used to fit the regressions are representative of the environmental gradient.
+Second, process-based SDMs predict easily interpreted and biologically meaningful quantities, such as intrinsic growth rates or abundances, as well as the underlying demographic rates and, potentially, spatially varying sensitivities.
+These additional parameters can help to inform applied management decisions \citep{Matzek2015,Szewczyk2019}.
+Third, the models are perhaps conceptually nearer to reality, representing the distribution as an emergent property of local and regional processes \citep{Merow2014b,Evans2016}.
 
 % Benefits of occurrence-based, and the unknowns of process-based performance (But) 
-4. Potential downsides of process-based SDMs
+There are, however, limitations and challenges with successfully implementing process-based SDMs.
+In stark contrast to occurrence-based SDMs, process-based SDMs have heavy data requirements \citep{Merow2014a,Evans2016}.
+Each demographic rate must be estimated either as a global parameter or as dependent on the environment and possibly size \citep{Merow2014a,Szewczyk2019}.
+Relevant rates will depend on the species, but may include annual survival probability, mean annual growth, variability in annual growth, flowering probability, mean per capita seed production, seedbank survival rates, and germination probabilities, in addition to dispersal parameters if movement between cells is included \citep{Merow2011a,Szewczyk2019}.
+Process-based SDMs are also structurally complex.
+Missing or mis-specified processes may have cascading impacts that are difficult to predict \citep{Merow2014c}.
+Similarly, problems and solutions resulting from common imperfections in datasets have been investigated for occurrence-based SDMs \citep{Wisz2008,Boria2014,Merow2016,Fernandes2018}.
+The individualized nature of process-based SDMs, with model structures based on species' life cycles \citep{Merow2011a,Merow2017,Szewczyk2019}, may render the effects and solutions of data-based problems more difficult to generalize.
+Thus, while process-based SDMs are intuitively appealing, it would be prudent to verify their expected benefits and appraise their resilience to errors before widespread adoption.
 
 % Mission statement (Therefore)
-We evaluate the performance of an occurrence-based SDM and three process-based SDMs for two virtual species.
-We investigate the ability of each SDM  to predict the true distributions under ideal conditions, as well as under various realistic data and modelling scenarios.
-Specifically, we ask: 1) Do process-based models outperform MaxEnt overall?
-2) are process-based or occurrence-based SDMs more susceptible to particular data deficiencies?
-And 3) do common modelling errors negate any improved performance observed in the process-based SDMs?
+Here, we directly evaluate the performance of the occurrence-based SDM Maxent and three process-based SDMs using two virtual species and multiple virtual sampling regimes.
+We investigate the ability of each SDM to predict the true distributions under ideal conditions, as well as under various data and modelling scenarios to represent a range of realistic challenges.
+Specifically, we ask: 1) Do process-based models outperform Maxent overall?
+2) Are process-based or occurrence-based SDMs more susceptible to particular data deficiencies?
+3) Do common modelling mis-specifications negate any improved performance otherwise observed in the process-based SDMs?
 
 \section{Methods}
-% General outline of the process (?)
-% Range boundary definitions
-% Virtual species simulation
-% Virtual ecologist simulation
-% SDM methods we compared
-
 \label{S:2}
 \begin{figure}
-	\centering\includegraphics[width=5in]{../../figs/process_outline.png}
-	\caption{\label{fig:outline} Simulation, sampling, and prediction process. Distributions were simulated for virtual species based on environmental effects on vital rates and demographic processes.}
+	\centering\includegraphics[width=5in]{figs/process_outline.png}
+	\caption{\label{fig:outline} Simulation, sampling, and prediction process.
+Distributions were simulated for virtual species based on environmental effects on size-dependent vital rates and demographic processes.
+From the simulated distribution for each virtual species, we drew samples to parameterize four SDMs (occurrence-based: Maxent; process-based analytic: IPM; process-based stochastic: CA\textsubscript{p}, CA\textsubscript{i}).
+We evaluated the performance of each SDM under four sampling scenarios and four modelling scenarios.}
 \end{figure}
+
 % General outline
 We evaluated the performance of four species distribution models (SDMs) in predicting the ranges of two simulated species under a variety of data quality and modelling scenarios (Fig.
-1), including one occurrence-based method (MaxEnt) and three process-based methods (Integral Projection Model: IPM; individual-level cellular automata: CA\textsubscript{i}; and population-level cellular automata: CA\textsubscript{p}).
-IPMs predict a deterministic intrinsic growth rate $\lambda$ in each cell of the landscape based on regressions of vital rates with environmental variables and individual sizes.
-In contrast, CA models are simulation-based and explicitly model spatio-temporal dynamics among cells.
-The CA\textsubscript{i} model simulates the growth, survival, and fecundity of individuals, while the CA\textsubscript{p} model uses population averages to summarize similar processes.
-See additional details for the process-based models in Appendix 1.
+\ref{fig:outline}) on a gridded landscape.
+We compared one occurrence-based SDM (Maxent) and three process-based SDMs (Integral Projection Model: IPM; individual-level cellular automata: CA\textsubscript{i}; and population-level cellular automata: CA\textsubscript{p}).
+For each species under each scenario, we parameterized the SDMs using 100 samples from the true distributions.
+We evaluated ability of each SDM to predict the true distribution, and the relative effects of each scenario on predictive accuracy.
+Additional details for the virtual species, process-based models, and scenarios are included in Appendix A.
+All code is available online at \textless https://github.com/Sz-Tim/sdmMethodComp\textgreater.
+
+\subsection{Species Distribution Models}
+Maxent employs the machine learning technique of maximum entropy to relate a set of presences with environmental conditions, optimizing a probability surface against a background based on pseudo-absences to predict the relative suitability across the landscape for the species \citep{Phillips2008}.
+Maxent models were fit in the \emph{dismo} R package with a threshold of 0.95 using the logistic output with 10 replicates for each dataset.
+
+IPMs predict a deterministic intrinsic growth rate $\lambda$ in each cell of the landscape by integrating across a kernel based on regressions of vital rates with environmental variables and individual sizes \citep{Easterling2000,Ellner2006,Merow2014a,Merow2017}, such that:
+\begin{equation}
+    n_{t+1}(z^{\prime}) = \int_{\Omega} K(z^{\prime}, z) n_t(z) dz 
+\end{equation}
+\begin{equation}
+    K(z^{\prime}, z) = P(z^{\prime}, z) + F(z^{\prime}, z)
+\end{equation}
+where $z$ represents size as a continuous variable, $n$ is the distribution across sizes, $\Omega$ is the range of possible sizes, $K(z^{\prime}, z)$ is the kernel representing the probability of transitions among sizes, $P(z^{\prime}, z)$ is the growth and survival component, and $F(z^{\prime}, z)$ is the fecundity component.
+To compose $P(z^{\prime}, z)$, we included regressions for growth and survival.
+To compose $F(z^{\prime}, z)$, we included regressions for flowering probability, seed production, and germination.
+Additionally, we incorporated a seedbank as a discrete stage (Appendix A: Section 1.2).
+Short distance dispersal was implemented by adjusting $F(z^{\prime}, z)$ for cells within each short distance dispersal neighborhood to account for the increase in the propagule pressure due to seed immigrants from nearby cells as well as the decrease from seed emigrants out of each cell (Appendix A: Section 1.3).
+
+In contrast to analytical IPMs, CA models are simulation-based and explicitly model spatio-temporal dynamics among cells.
+The individual-level cellular automaton model (CA\textsubscript{i}) simulates individuals in each cell of the landscape by building on the structure of the IPM.
+We simulated local processes within each cell, including survival, growth, flowering, seed production, germination, and seedbank dynamics.
+We used identical vital rate regressions as for the IPM, but then simulated individuals and outcomes (e.g., annual survival, growth, etc.) in each cell by generating random numbers from the corresponding distributions using the expected values from the regressions.
+Regional processes were represented as the dispersal of seeds both within short distance dispersal neighborhoods and to a set number of random cells in the landscape.
+
+The population-level cellular automaton model (CA\textsubscript{p}) uses population averages to summarize similar processes, and models approximately age-structured populations which include pre-reproductive juvenile stages and a single adult stage \citep{Merow2011a,Szewczyk2019}.
+Seeds are similarly dispersed via short distance dispersal among nearby cells and via stochastic long distance dispersal to any cell on the landscape.
+All three process-based SDMs thus model local populations within each cell driven by similar demographic processes, and dispersal of seeds among cells at a more regional scale.
 
@@
 We simulated two virtual species in the eastern United States, based approximately on Japanese barberry (\emph{Berberis thunbergii}), an invasive shrub with bird-dispersed seeds, and garlic mustard (\emph{Alliaria petiolata}), an invasive biennial with seeds dispersed by non-volant animals and water.
-Both species have previously been the focus of process-based SDMs (CITE).
-To simulate each species, we defined relationships between environmental variables, individual size, and vital rates (Fig.
-\ref{fig:outline}, Appendix 1) on a gridded landscape spanning the Eastern Temperate Forest and Northern Forest ecozones within the United States ($\sim$5x5km resolution: 115,105 cells, CITE).
+Both species have previously been the focus of process-based SDMs \citep{Merow2017}.
+For each species, we defined relationships where vital rates were dependent on environmental variables and individual size (Fig.
+\ref{fig:outline}, Appendix A) on a gridded landscape spanning the Eastern Temperate Forest and Northern Forest ecozones within the United States ($\sim$5x5km resolution: 115,105 cells \citep{Omernik2014}).
 Vital rates included annual individual growth, annual survival, flowering probability, seed production, and germination probability.
-For environmental covariates, we used the minimum temperature of the coldest month and the precipitation in May, following previous work (CITE PNAS, Chelsa).
-Based on these regressions, we calculated the intrinsic growth rate $\lambda$ in each cell following the methods used for IPMs, where each cell was treated as a population.
-Cells with $\lambda > 1$ were considered to be capable of containing persistent populations, constituting the first definition of the true distribution for each species.
+For individuals of a given size, each vital rate was determined by the response to the minimum temperature of the coldest month and the precipitation in May, following previous work \citep{Merow2017}.
+We calculated a mean value for each climatic variable in each cell using the CHELSA climate dataset \citep{Karger2017}.
+We defined the true range for each species using two methods: cells where the intrinsic growth rate $\lambda>1$ (i.e., analytic), and cells containing individuals at the end of the simulation (i.e., dynamic).
+
+We calculated the intrinsic growth rate $\lambda$ in each cell following the methods used for IPMs, where each cell was treated as a population.
+Vital rate regressions and parameters were initially informed based on previous work on each species in New England, USA \citep{Merow2017}, but modified \emph{ad hoc} to generate equilibrium distributions approximating the current distribution of each species in the eastern United States.
+For the shrub, dispersal parameters were based on dispersal rates of similarly bird-dispersed invasive shrubs \citep{Merow2011a,Szewczyk2019}.
+For the biennial, short distance dispersal was decreased and long distance dispersal increased to qualitatively reflect the mode of dispersal.
+For the shrub, $\lambda$ was calculated in each cell as the first eigenvalue of the transition matrix constructed from the vital rate regressions (Appendix A).
+The two-year life cycle of the biennial required calculating $\lambda$ by iterating populations through the transition matrix until the distribution had stabilized, at which point $\lambda$ is calculated as $N_t/N_{t-1}$.
+Cells with $\lambda > 1$ were considered to be capable of supporting persistent populations, constituting the true $\lambda$-based definition for each species.
 
 To simulate populations, we initialized 10 random cells with 10 individuals, with individual sizes drawn from a uniform distribution constrained by the allowable range for each species.
-Then, we calculated the expected survival probability, growth rate, flowering probability, and seed production for each individual according to the vital rate regressions (Appendix 1).
-Then, we drew random values from the appropriate distributions for each vital rate for each individual, implemented short- and long-distance dispersal of seeds according to the dispersal mode of each species, and generated new individuals in each cell from germinating seeds.
-Populations in each cell were subject to density dependence through reduced seedling establishment (CITE) after the abundance exceeded a predetermined threshold.
+Then, we calculated the expected survival probability, growth rate, flowering probability, and seed production for each individual according to the same vital rate regressions used in the calculation of $\lambda$ (Appendix A).
+We drew random values from the appropriate distributions for each vital rate for each individual, implemented short- and long-distance dispersal of seeds according to the dispersal mode of each species, and generated new individuals in each cell from germinating seeds.
+Populations in each cell were subject to density dependence through reduced seedling establishment \citep{Ellner2006,Rebarber2012,Merow2014a} after the abundance exceeded a predetermined threshold.
 The populations were simulated through 300 years, at which point the distribution appeared stable, indicating that equilibrium had been reached.
-The cells with $N > 0$ in the final year thus constituted the second definition of the true distribution for each species.
+Cells with $N > 0$ in the final year were considered to be the equilibrium population, constituting the true $N$-based definition for each species.
 
 \subsection{Sampling and Modelling Scenarios}
-We evaluated four sampling scenarios: \emph{ideal} (year 300; uniform sampling probability among occupied cells), \emph{non-equilibrium} (year 100; uniform sampling probability among occupied cells), \emph{geographic bias} (year 300; sampling probability $\propto$ human and road density in occupied cells), and \emph{measurement error} (year 300; MaxEnt: 3\% of observations drawn from full landscape regardless of occupancy; process-based: error added to growth, seed production, and abundance observations).
-For each species, we generated 100 sets of sampled cells for each scenario.
-For MaxEnt, each sample consisted of 300 cells (CITE), while for the process-based SDMs, each sample consisted of 25 cells (CITE).
-For process-based models, a maximum of 1000 (CA\textsubscript{p}) or 100 (CA\textsubscript{i}, IPM) individuals were sampled from within each selected grid cell to reflect real-world logistical limitations (Appendix 1, https://github.com/Sz-Tim/sdmMethodComp).
+To generate datasets for fitting the SDMs, we drew samples from the simulated (i.e., $N$-based) distribution for each species (Fig.
+\ref{fig:outline}: \emph{Observed Samples}).
+We evaluated four sampling scenarios: \emph{ideal}, \emph{non-equilibrium}, \emph{geographic bias}, and \emph{measurement error}.
+For each species, we generated 100 data sets under each sampling scenario.
+For Maxent, each data set consisted of presences from 300 cells \citep{Wisz2008,Allen2016,Fernandes2018}.
+For the process-based SDMs, each data set consisted of 25 cells \citep{Merow2017}, from which a maximum of 100 (CA\textsubscript{i}, IPM) or 1000 (CA\textsubscript{p}) individuals were sampled per cell to reflect real-world logistical limitations in measuring demographic and vital rates (Appendix A).
+The \emph{ideal} data sets were drawn from the true virtual species distributions at year 300 with uniform sampling probability among cells occupied by at least 5 individuals.
+The \emph{non-equilibrium} datasets were instead drawn from the virtual species distributions at year 100.
+The \emph{geographic bias} data sets were drawn from year 300, but with the sampling probability for each cell proportional to the human population density and road density.
+For the final data set, we imposed \emph{measurement error}.
+For Maxent, we replaced 3\% of the observations in each data set with random cells drawn from full landscape regardless of occupancy.
+For the process-based SDMs, we added error to the measurements of annual growth: $g_{obs} = Norm(g_{true}, 0.1)$; seed production: $seed_{obs} = Norm(seed_{true}, seed_{true}*0.05)$; and population abundance: $N_{obs} = Norm(N_{true}, N_{true}*0.02)$.
+Thus, the \emph{measurement error} scenario reflects imprecision in the data sets rather than biased measurement.
 
-In addition, we assessed the effect of four scenarios of modeling misspecification: \emph{incorrect covariates}, \emph{no seed bank} (process-based only), \emph{over-estimated dispersal} (process-based only), \emph{under-estimated dispersal} (process-based only).
-Each of these misspecifications was implemented using the \emph{ideal} datasets.
-For \emph{incorrect covariates}, the SDMs were fit with correlated but more general climate variables (mean annual temperature, annual precipitation) rather than the more specific covariates used in the generative process and for all other scenarios (minimum temperature of the coldest month, precipitation in May).
+We also evaluated four scenarios of modelling mis-specification: \emph{incorrect covariates}, \emph{no seedbank} (process-based only), \emph{over-estimated dispersal} (process-based only), \emph{under-estimated dispersal} (process-based only).
+Each of these mis-specifications was implemented using the \emph{ideal} datasets.
+For \emph{incorrect covariates}, the SDMs were fit with correlated (Appendix A) but more general climate variables (mean annual temperature, annual precipitation) rather than the more specific covariates used in the generative process and for all other scenarios (minimum temperature of the coldest month, precipitation in May).
+For \emph{no seedbank}, the predictive SDMs assumed that all seedlings in the sampled datasets originated from seeds produced in the same year, with complete over-winter mortality of non-germinating seeds.
+For \emph{over-estimated dispersal}, the maximum short distance dispersal distance was increased by 2 cells, the mean short distance dispersal distance was doubled, and the number of annual long distance dispersal events multiplied by five.
+For \emph{under-estimated dispersal}, the maximum short distance dispersal distance was decreased by 2 cells, the mean short distance dispersal distance was halved, and the number of annual long distance dispersal events divided by five.
 
-For each scenario (Fig \ref{fig:outline}), the SDMs were parameterized using the corresponding sampled datasets.
-To select environmental covariates for each vital rate regression for each process-based SDMs, we compared all combinations of the climatic variables and their squares with the R package \emph{MuMIn}.
-For the CA\textsubscript{p} and CA\textsubscript{i} models, simulations were run for 300 years to generate predicted distributions.
-Using each definition of the true simulated ranges (i.e., $\lambda > 1$ and $N > 0$), we calculated the True Skill Statistic (TSS) for each of the 100 predicted distributions per scenario per SDM.
-We calculated TSS as $sensitivity + specificity - 1$, where $sensitivity$ is the proportion of true presences that are predicted correctly, and $specificity$ is the proportion of true absences that are predicted correctly, such that TSS ranges from $-1$ (all cells predicted incorrectly) to $1$ (all cells predicted correctly).
+For each scenario (Fig.
+\ref{fig:outline}), the SDMs were parameterized using the corresponding sampled datasets.
+As explanatory variables, we used two climatic variables (minimum temperature of the coldest month, precipitation in May) and their squares, with the exception of the \emph{incorrect covariates} scenario in which we instead used the more general variables.
+To select environmental covariates for each process-based SDMs, we compared all combinations of the climatic variables and their squares with the R package \emph{MuMIn} for each vital rate regression.
+For the CA\textsubscript{i} models and IPMs, this included regressions for annual survival, annual growth, flowering probability, seed production for flowering individuals, and germination probability.
+For the CA\textsubscript{p} models, the demographic rate regressions included carrying capacity, juvenile survival, adult survival (shrub only), flowering probability, seed production for flowering individuals, and germination probability.
+With the fitted CA\textsubscript{p} and CA\textsubscript{i} models, simulations were run for 300 years to generate predicted distributions.
+For the fitted IPMs, we calculated $\lambda$ in each cell.
+
+To evaluate the performance of each SDM under each scenario, we calculated the True Skill Statistic (TSS) \citep{Allouche2006} for each of the 100 predicted distributions per scenario per SDM (Maxent: 100 datasets x 5 scenarios x 2 species = 1000 predicted distributions; CA, IPM: 100 datasets x 8 scenarios x 2 species = 1600 predicted distributions).
+We compared each predicted distribution to the true simulated range based on population viability (true $\lambda > 1$) and based on population presence (true $N > 0$).
+TSS is calculated as \emph{sensitivity} + \emph{specificity} $-$ 1, where \emph{sensitivity} is the proportion of true presences that are predicted correctly, and \emph{specificity} is the proportion of true absences that are predicted correctly, such that TSS ranges from $-1$ (all cells predicted incorrectly) to $1$ (all cells predicted correctly).
+
+To compare TSS among SDM types overall, we used the R package \emph{dunn.test} to perform Dunn's Test of Multiple Comparisons Using Rank Sums, a non-parametric post-hoc test for pairwise comparisons following a Kruskal-Wallis test.
+We compared SDM performance for each definition of range boundary, using the median TSS of each SDM for each species under each scenario, restricting the analyses to scenarios applicable to all four SDMs.
 
@@
 % Talk results
-% 1. Across all scenarios, IPM is best for lambda, CAi/p are best for N, with MaxEnt in the middle for both
-% 2. Under ideal conditions, process-based outperform MaxEnt
-% 3. For non-equilibrium distributions, process-based outperform MaxEnt
+% 1. Across all scenarios, IPM is best for lambda, CAi/p are best for N, with Maxent in the middle for both
+% 2. Under ideal conditions, process-based outperform Maxent
+% 3. For non-equilibrium distributions, process-based outperform Maxent
 % 4. Process-based are much, much more sensitive to covariate choice -- implies sensitivity to regression accuracy
@@
 % 6. Structural errors seem to have erratic effects
-% 0. Process-based are better IF you are confident of the process, but MaxEnt is more robust to error and uncertainty
+% 0. Process-based are better IF you are confident of the process, but Maxent is more robust to error and uncertainty
 
@@
 %	N = CA, MxE, IPM (avgs)
-%	MaxEnt somewhat better for lambda
+%	Maxent somewhat better for lambda
 % Effect of data issues
@@
 \ref{fig:TSSmn}).
-Across scenarios applicable to all four SDMs (Table \ref{table:ranks}), the IPM best predicted the $\lambda$-based distributions (mean rank: 1.5) and the CA models best predicted the $N$-based distributions (mean rank: CA\textsubscript{i}=1.6, CA\textsubscript{p}=1.9).
-MaxEnt was intermediate for both range boundaries on average (mean rank: $\lambda$-based=2.7, $N$-based=2.9)
+Across scenarios applicable to all four SDMs (Table \ref{table:ranks}), the IPM best predicted the $\lambda$-based distributions (\emph{P\textsubscript{IPM-CA\textsubscript{i}}}: 0.02, \emph{P\textsubscript{IPM-CA\textsubscript{p}}}: 0.01, \emph{P\textsubscript{IPM-Maxent}}: 0.01), with no significant difference between the CA models and Maxent.
+Similarly, the CA models best predicted the $N$-based distributions (\emph{P\textsubscript{IPM-CA\textsubscript{i}}}: \textless0.01, \emph{P\textsubscript{IPM-CA\textsubscript{p}}}: \textless0.01, \emph{P\textsubscript{Maxent-CA\textsubscript{i}}}: \textless0.01, \emph{P\textsubscript{Maxent-CA\textsubscript{p}}}: 0.02) with no significant difference between the CA models.
+Maxent was intermediate for both range boundaries on average (mean rank: $\lambda$-based=2.7, $N$-based=2.9), though its rank was not significantly different from the process-based model(s) that performed most poorly for each range boundary.
 
@@
 		IPM & 1 & 1.5 & 1 & 4\\
-		MaxEnt & 1 & 2.7 & 2.5 & 4\\
+		Maxent & 1 & 2.7 & 2.5 & 4\\
 		[.5ex]\hline \\[-2.5ex]
@@
 		IPM & 3 & 3.6 & 4 & 4\\
-		MaxEnt & 1 & 2.9 & 3 & 4\\
+		Maxent & 1 & 2.9 & 3 & 4\\
 		[.5ex]\hline \\[-2ex]
@@
 % Effect of data issues
-For MaxEnt, TSS decreased when fit using samples from non-equilibrium populations, driven by smaller predicted ranges as indicated by a decrease in sensitivity (Fig.
-SXX:Sensitivity-Ideal) and a smaller increase in specificity (Fig.
-SXX:Specificity-Ideal).
-In contrast, non-equilibrium samples had minimal effect on TSS for the CA\textsubscript{p} model, with a modest improvement for the long-lived shrub in the IPM and CA\textsubscript{i} models driven by increases in sensitivity (Fig.
-SXX:Sensitivity-Ideal).
-Measurement error and sampling bias had negligible effects on the process-based models, and small to moderate effects on the performance of MaxEnt (Fig.
+For Maxent, TSS decreased when fit using samples from \emph{non-equilibrium} populations, driven by smaller predicted ranges as indicated by a decrease in sensitivity (Fig.
+B.1) and a smaller increase in specificity (Fig.
+B.2).
+In contrast, \emph{non-equilibrium} samples had minimal effect on TSS for the CA\textsubscript{p} model, with a modest improvement for the long-lived shrub in the IPM and CA\textsubscript{i} models driven by increases in sensitivity (Fig.
+B.1).
+\emph{Measurement error} and \emph{sampling bias} had negligible effects on the process-based models, and small to moderate effects on the performance of Maxent (Fig.
 \ref{fig:TSSmn}, Fig.
-SXX:Sensitivity-Ideal, Fig.
-SXX:Specificity-Ideal).
+B.1, Fig.
+B.2).
 
 \begin{figure}
-	\centering\includegraphics[width=5in]{../../figs/fig_TSS_mn+CI.jpg}
-	\caption{\label{fig:TSSmn} True skill statistic (TSS) mean and 95\% confidence intervals across 100 sampled datasets for each SDM and scenario, compared to true distributions defined by $\lambda > 1$ and $N > 0$.
-Scenarios include: no sampling or modeling issues (ideal), sampling issues (measurement error, sampling bias, non-equilibrium), and modeling issues (incorrect covariates, no seed bank, under dispersal, over dispersal).}
+	\centering\includegraphics[width=5in]{figs/fig_TSS_mn+CI.jpg}
+	\caption{\label{fig:TSSmn} True skill statistic (TSS) mean and 95\% confidence intervals across 100 sampled data sets for each SDM and scenario, compared to true distributions defined by $\lambda > 1$ and $N > 0$.
+Scenarios include: no sampling or modelling issues (ideal), sampling issues (measurement error, sampling bias, non-equilibrium), and modelling issues (incorrect covariates, no seedbank, under dispersal, over dispersal).}
 \end{figure}
@@
 \begin{figure}
-	\centering\includegraphics[width=\linewidth]{../../figs/fig_TSSvIdeal.jpg}
+	\centering\includegraphics[width=\linewidth]{figs/fig_TSSvIdeal.jpg}
 	\caption{\label{fig:TSSvIdeal} Effect of scenario on median true skill statistic (TSS) relative to the '\emph{ideal}' scenario.
@@
 % Effect of modelling issues
-Modeling with incorrect covariates was universally problematic for model performance, though it had a greater impact on TSS in the shrub than the biennial species (Fig.
+modelling with \emph{incorrect covariates} was universally problematic for model performance, though it had a greater impact on TSS in the shrub than the biennial species (Fig.
 \ref{fig:TSSvIdeal}).
-Incorrect covariates did not lead to systematic over-prediction or under-prediction, but rather a decrease in both sensitivity and specificity, particularly for the process-based models (Fig.
-SXX:Sensitivity-Ideal, Fig.
-SXX:Specificity-Ideal).
-Excluding the seed bank in the process-based models had minimal effect on TSS for the shrub, but variable effects on TSS for the biennial.
-Relative to ‘ideal’, all process-based models showed reduced sensitivity and increased specificity for the biennial, indicating a smaller predicted range.
-The IPM was generally more resistant to mischaracterized dispersal compared to the simulation-based CA\textsubscript{i} and CA\textsubscript{p} models.
+\emph{Incorrect covariates} did not lead to systematic over-prediction or under-prediction, but rather a decrease in both sensitivity and specificity, particularly for the process-based models (Fig.
+B.3, Fig.
+B.4).
+Excluding the seedbank in the process-based models had minimal effect on TSS for the shrub, but variable effects on TSS for the biennial.
+Relative to \emph{ideal}, all process-based models showed reduced sensitivity and increased specificity for the biennial, indicating a smaller predicted range.
+The IPM was generally more resistant to mis-characterized dispersal compared to the simulation-based CA\textsubscript{i} and CA\textsubscript{p} models.
 Underestimation of dispersal parameters led to decreased predicted ranges, and overestimation to increased predicted ranges relative to ‘ideal’.
@@
 For instance, the IPM predicts $\lambda$ in each cell of the landscape.
-Discrepancies in predicted $\lambda$ values were biased downward in the northern portion of the study region (FIGURE 4), more strongly so for the shrub.
+Discrepancies in predicted $\lambda$ values were biased downward in the northern portion of the study region (Fig.
+\ref{fig:N_Lambda}), more strongly so for the shrub.
+Mirroring the effect on predicted presence, residual error was most variable and most extensive when \emph{incorrect covariates} were used.
+
 Similarly, the CA models predict the abundance $N$ for each species.
-Discrepancies in predicted abundance from the CA\textsubscript{p} model showed a spatial banding of upward bias at the range margins for both species (FIGURE 4).
-Immediately interior to the range margin, the CA\textsubscript{p} and CA\textsubscript{i} model showed upward bias for the shrub (FIGURE 4).
+Discrepancies in predicted abundance from the CA\textsubscript{p} model showed a spatial banding of upward bias at the range margins for both species (Fig.
+\ref{fig:N_Lambda}), indicating a tendency to over-predict at range edges.
+Immediately interior to the range margin, the CA\textsubscript{p} and CA\textsubscript{i} model showed upward bias for the shrub (Fig.
+\ref{fig:N_Lambda}).
+The CA\textsubscript{p} model dramatically underestimated abundance of the long-lived shrub with \emph{non-equilibrium} datasets, despite minimal impact on predicted presence or absence.
+As with the IPM, using \emph{incorrect covariates} tended to result in larger error in abundance predictions for both CA\textsubscript{p} and CA\textsubscript{i}.
+
+\begin{figure}
+	\centering\includegraphics[width=\linewidth]{figs/map_N_Lambda.jpg}
+	\caption{\label{fig:N_Lambda} Residual error in predicted $\lambda$ and $N$ under each scenario. Darker colors indicate larger residuals (under-prediction: purple, red; over-prediction: green, blue)}
+\end{figure}
 
@@
 % Talk Discussion
-% 1. Across all scenarios, IPM is best for lambda, CAi/p are best for N, with MaxEnt in the middle for both
-% 2. Under ideal conditions, process-based outperform MaxEnt
-% 3. For non-equilibrium distributions, process-based outperform MaxEnt
+% 1. Across all scenarios, IPM is best for lambda, CAi/p are best for N, with Maxent in the middle for both
+% 2. Under ideal conditions, process-based outperform Maxent
+% 3. For non-equilibrium distributions, process-based outperform Maxent
 % 4. Process-based are much, much more sensitive to covariate choice -- implies sensitivity to regression accuracy
-% 5. Dispersal is critical for CA models
+% 5. Dispersal was critical for CA models, but had essentially no effect for IPMs
 % 6. Structural errors seem to have erratic effects
-% 0. Process-based are better IF you are confident of the process, but MaxEnt is more robust to error and uncertainty
+% 0. Process-based are better IF you are confident of the process, but Maxent is more robust to error and uncertainty
 
 % Overview
-This is the first paragraph summarizing some general stuff.
-The process-based SDMs provide a lot more information, and they do, in fact, perform a bit better on average.
-They are less impacted by samples from non-equilibrium populations, so that is in line with expectations.
-However, they are more susceptible to mismatches in the covariates than MaxEnt, and the structural errors in the models don't necessarily have predictable effects.
+Overall, the virtual species distributions were predicted best by the process-based models when the range boundary definitions were appropriately matched.
+The $\lambda$-based ranges were best predicted by the IPM, and the $N$-based ranges were best predicted by the CA models, with Maxent typically intermediate in each case.
+Maxent showed highest sensitivity to \emph{non-equilibrium} samples.
+In contrast, the process-based models were relatively unaffected by \emph{non-equilibrium} samples, with moderate improvements for the longer-lived shrub.
+However, the process-based models were much more sensitive to the use of \emph{incorrect covariates} than Maxent.
+The simulation-based CA models were influenced by modelling mis-specifications, with the magnitude of the effect varying across species.
+While process-based SDMs indeed have many benefits over simpler occurrence-based SDMs, our results highlight the importance of accurately describing the environmental drivers impacting each vital rate.
 
-% Process-based > MaxEnt under ideal conditions, and especially non-equilibrium distributions
-Expand on the summary, bringing in past work on non-equilibrium distributions.
-Since process-based models aim to describe the biological processes that ultimately produce an emergent distribution, it really doesn't matter whether the samples are from a non-equilibrium species as long as the populations contain a sufficiently broad span of ages or sizes.
-This points to the fact that in parameterizing process-based models, care should be taken to select populations with a mix of ages or sizes rather from a range of environmental conditions than focusing on the geographic distribution per se, in essence capturing more of the parameter space.
-Thus, for non-equilibrium species, a well-described process-based SDM should do better than an occurrence-based model.
-.
-% However, process-based are highly sensitive
-%	- covariate choice: even highly correlated covariates lead to dramatically decreased ability
-%	- dispersal: critical for CA models
-% 	- structural errors seem to have erratic effects
-However, process-based models appear to be sensitive to several parts of the modelling process.
-First, they are much more sensitive to covariate choice, and by generalization, to the alignment of the modelled relationship between vital rates and the environment with reality.
-The flexible structure of MaxEnt, perhaps along with the top-down perspective, makes it much less influenced by errors in the covariates.
-Even with covariates that are highly correlated with the true covariates, there is still a large decrease in predictive ability for the process-based models.
+% Range boundaries
+IPMs predict $\lambda$ in each cell of the landscape, and consequently directly predict a $\lambda$-based range boundary.
+Similarly, the simulation-based CA models predict abundance, and thus directly predict a $N$-based range boundary.
+For Maxent, the range boundary definition is less explicit.
+The geo-located presences used as input imply an abundance-based range, but predicted distributions are frequently constrained by a data-based threshold (e.g., relative suitability values that encapsulate 95\% of the observed data) to exclude both erroneous data and sink populations, which implies an effort to predict a $\lambda$-based range.
+There may or may not be a big difference between the two range definitions, but it will depend on the species.
+For example, the magnitude of the difference should vary with the number or distribution of marginal cells (lots of marginal cells mean large areas with middling occupancy probabilities), the typical annual stochasticity in the population abundances (more stochasticity means more discrepancies are likely), and dispersal limitation (high dispersal or a long introduction history means fewer areas that have not been reached, but also possibly a larger proportion of sink populations).
+
+% Process-based > Maxent under ideal conditions, and less susceptible to sampling effects
+Under ideal conditions, the process-based SDMs, properly matched to the range boundary definition, outperformed Maxent (Fig.
+\ref{fig:TSSmn}).
+This generally held true across the sampling scenarios, with minimal impact of \emph{measurement error} or \emph{sampling bias} on the process-based SDMs.
+
+% Climate change and alien species
+Critically, there was little or no decrease in performance in the process-based models with \emph{non-equilibrium} samples (Fig.
+\ref{fig:TSSvIdeal}, B.3, B.4).
+In fact, the individual-based models (IPM, CA\textsubscript{i}) showed somewhat improved performance for the shrub, as the younger populations had a more even age distribution, allowing for more accurate estimation of the size-dependent vital rate regressions (Fig.
+B.5a).
+Because process-based SDMs aim to describe the biological processes that ultimately result in an emergent geographic distribution, care should be taken to select sample populations with an adequately broad range of sizes or ages.
+In contrast, the performance of Maxent declined the most in the \emph{non-equilibrium} scenario, consistent with past work (CITE).
+These results thus support the proposal that process-based SDMs are advantageous in situations where populations are not at equilibrium \citep{Merow2011a,Evans2016,Merow2017,Cabral2017}, such as for distributions of alien species or for future distributions as the climate continues to change.
+
+% But caveat –– labor intensive, so only worth it for important species
+However, there are undeniable logistical limitations to constructing process-based SDMs.
+Ideally, a process-based SDM for a plant would include rates across a size or age range and environmental gradient for survival, growth, flowering, seed production, and a seedbank \citep{Ellner2006,Merow2014a}, in addition to short distance and long distance dispersal dynamics.
+Each process must be informed with data collected over the course of multiple years, and the sampled environmental space should adequately represent the environmental space used for prediction.
+For many species, some data may be available in the literature or borrowed from ecologically similar species \citep{Merow2011a,Szewczyk2019}, and some values can be estimated using parameter-oriented parameterization \citep{Grimm2005,Merow2011a,Szewczyk2019}.
+Further, sensitivity analyses can determine effects of parameter uncertainty to identify which parameters are most critical \citep{Prowse2016,Aiello-Lammens2017}.
+Nevertheless, this process is realistically too laborious to perform in a single study on hundreds of species, as is possible with occurrence-based SDMs \citep{Allen2016}.
+Process-based SDMs therefore may be better suited for species of particular concern, including invasive species or threatened and endangered species.
+
+% Dangers of process-based SDMs: covariate choice
+In our simulations, the process-based models were sensitive to several of the modelling scenarios.
+Importantly, they were more sensitive in general to the choice of covariates than was Maxent (Fig.
+\ref{fig:TSSvIdeal}, B.3, B.4).
+In the \emph{incorrect covariates} scenario, we fit the predictive models using mean annual temperature and annual precipitation, while the virtual species were generated based on relationships with the minimum temperature of the coldest month and the precipitation in May.
+The variables were highly correlated (Fig.
+A.1), but the differences are spatially structured (Fig.
+A.2).
+The quadratic relationships we allowed in the vital rate regressions were unable to adequately compensate for this spatial discrepancy in the covariates.
+In contrast, the flexible nature of the algorithms which underlie Maxent appear diminish these effects by allowing for more complex and idiosyncratic relationships between occurrences and the environment.
+Process-based SDMs directly aim to model the processes that determine species distributions, and so their predictions are more sensitive to mismatches between the modelled processes and the truth.
+For example, if seedling mortality is driven primarily by the minimum temperature in the spring, a process-based SDM may predict the distribution quite poorly if either mean, maximum, or minimum annual temperature are used instead, despite their high correlation.
+Using more flexible relationships in the process-based SDMs, such as generalized additive models, may reduce the impact of covariate choice, but with the added risks associated with more complexity.
+In the construction of a process-based SDM, it is critical to understand each aspect of the biology of the focal species well in order to correctly specify the environmental drivers affecting each vital rate.
+
+% Dangers of process-based SDMs: seedbank 
+Seedbank dynamics are extremely difficult to quantify (CITE).
+Even for many well-studied plant species, the seedbank is poorly understood (CITE), and process-based SDMs often either assume that the impact is negligible to the overall distribution \citep{Merow2017} or assume simple survival rates based on literature \citep{Szewczyk2019}.
+In the \emph{no seedbank} scenario, we assumed that the seedbank had negligible impact; the support for such an assumption was dependant on the species and the SDM (Fig.
+\ref{fig:TSSvIdeal}, B.3, B.4).
+The sensitivity, specificity, and TSS of all three process-based SDMs were unchanged from ignoring the seedbank for the shrub.
+For the biennial, however, excluding the seedbank modestly decreased sensitivity and modestly increased specificity, reflecting a smaller predicted distribution.
+The overall impact on TSS was consequently variable across SDMs, depending on the magnitude of the effects on sensitivity and specificity.
+Our results suggest that seedbank dynamics are likely to be a more important component in process-based SDMs for short-lived species than for long-lived species, due to differences in the reproductive strategies associated with the different life histories, as well as the buffering effect on inter-annual abundance resulting from long-lived individuals \citep{Morris2003}.
+
+% Dangers of process-based SDMs: dispersal
+The accuracy of the CA model predictions were sensitive to dispersal.
+Drastically underestimating dispersal resulted in smaller predicted ranges, while overpredicting dispersal resulted in larger predicted ranges.
+The predictions for the biennial tended to be a bit more sensitive, particularly in terms of specificity (Fig.
+B.4).
+
+% Other advantages and uses of process-based models
+There are a bunch of reasons you might want to build a process-based model aside from increased accuracy. I could draw on some examples from the buckthorn paper.
+
+% Conclusions
+Process-based models provide a great deal more information that can make SDMs more directly useful to management decisions.
+In our simulations, the well-specified process-based SDMs predicted the true distributions better.
+Our results also support the assertion in the literature that process-based SDMs are less susceptible to non-equilibrium, suggesting that they're a good option for alien species and climate change scenarios.
+However, easy-to-make modelling errors, assumptions, and mis-specifications may have a big impact on their accuracy, so caution should be taken to ensure that the environmental drivers in the model match the true biology of the species as closely as possible.
+It's also a good idea to run global sensitivity analyses to identify the parameters and processes that are most likely to shape the distribution \citep{Prowse2016,Aiello-Lammens2017}.
+For species of particular concern, process-based SDMs are tools that can provide more accurate and more applicable predictions.
 
 % So: Why or when to use each?
-%	- MaxEnt is probably best for most species: fairly robust, already know about non-equilibrium, etc
+%	- Maxent is probably best for most species: fairly robust, already know about non-equilibrium, etc
 %		- But what about climate change? Isn't everything at disequilibrium now? Shit...
@@
 %		- redo each with just the lifespan changed? But that will probably just make it more complicated....
+% Dangers of process-based SDMs
+%	- covariate choice: even highly correlated covariates lead to dramatically decreased ability -- EXTRAPOLATION IS MORE DANGEROUS??
+%	- dispersal: critical for CA models
+% 	- structural errors seem to have erratic effects
 % Range boundaries: What is each SDM predicting?
 
-% Range boundaries
-Range boundaries can be defined in several ways.
-For the true distribution of each species, we use two definitions: abundance-based (cells with $N > 0$) and $\lambda$-based (cells with $\lambda > 1$).
-The process-based SDMs compared here each predict a specific type of range boundary.
-IPMs predict $\lambda$ in each cell of the landscape, while the simulation-based CA models predict abundance.
-For MaxEnt, the range boundary definition is less explicit.
-The geo-located presences used as input imply an abundance-based range, but predicted distributions are frequently constrained by a data-based threshold (e.g., relative suitability values that encapsulate 95\% of the observed data) to exclude both erroneous data and sink populations, which implies an effort to predict a $\lambda$-based range.
-Generally, process-based SDMs explicitly predict one type of range or another, while occurrence-based SDMs may be rather ambiguous.
-There may or may not be a big difference between the two range definitions, but it will depend on the species.
-For example, the magnitude of the difference should vary with the number or distribution of marginal cells (lots of marginal cells mean large areas with middling occupancy probabilities), the typical annual stochasticity in the population abundances (more stochasticity means more discrepancies are likely), and dispersal limitation (high dispersal or a long introduction history means fewer areas that have not been reached, but also possibly a larger proportion of sink populations).
-
 \section{Acknowledgments}
-This project was funding through National Science Foundation award IIS-1717368.
+This project was funded through National Science Foundation award IIS-1717368.
 
 \section{Bibliography}
+\bibliography{sdmMethodComp}
 

### 2020-02-13 code/6_figures_ms.R
updated names

@@ -29,7 +29,7 @@ suppressMessages(invisible(lapply(pkgs, library, character.only=T)))
 walk(dir("code", "fn", full.names=T), source)
 theme_set(theme_bw())
 sp.names <- c("shrub", "biennial")
-SDM_col <- c(MaxEnt="#fd8d3c", IPM="#08306b", 
+SDM_col <- c(Maxent="#fd8d3c", IPM="#08306b", 
              'CA[i]'="#2171b5", 'CA[p]'="#6baed6")
 
 # TSS
@@ -48,7 +48,7 @@ TSS.df$issue <- factor(TSS.df$issue,
                                 "Over dispersal"))
 TSS.df$sp <- factor(TSS.df$sp, levels=c("sp1", "sp2"), labels=sp.names)
 TSS.df$Boundary <- paste0("True~range:~", TSS.df$Boundary)
-TSS.df$SDM <- lvls_revalue(TSS.df$SDM, c("CA[p]", "CA[i]", "IPM", "MaxEnt"))
+TSS.df$SDM <- lvls_revalue(TSS.df$SDM, c("CA[p]", "CA[i]", "IPM", "Maxent"))
 TSS.ci <- TSS.df %>% group_by(sp, SDM, issue, Boundary) %>% 
   summarise(med=median(TSS, na.rm=T), 
             mn=mean(TSS, na.rm=T),
@@ -98,7 +98,7 @@ LL.df$issue <- factor(LL.df$issue,
                                 "Over dispersal"))
 LL.df$sp <- factor(LL.df$sp, levels=c("sp1", "sp2"), labels=sp.names)
 LL.df$Boundary <- paste0("True~range:~", LL.df$Boundary)
-LL.df$SDM <- lvls_revalue(LL.df$SDM, c("CA[p]", "CA[i]", "IPM", "MaxEnt"))
+LL.df$SDM <- lvls_revalue(LL.df$SDM, c("CA[p]", "CA[i]", "IPM", "Maxent"))
 
 # Predictions
 P.df <- rbind(read.csv("out/sp1/out_P.csv"),
@@ -122,14 +122,14 @@ P.df$issue <- factor(P.df$issue,
                               "No Seedbank", "Under dispersal",
                               "Over dispersal"))
 P.df$sp <- factor(P.df$sp, levels=c("sp1", "sp2"), labels=sp.names)
[…]

### 2020-02-13 ms/ms_latex/ms_sdmComp.tex
Updates from Overleaf

@@
 \label{S:3}
-% Talk results
-% 1. Across all scenarios, IPM is best for lambda, CAi/p are best for N, with Maxent in the middle for both
-% 2. Under ideal conditions, process-based outperform Maxent
-% 3. For non-equilibrium distributions, process-based outperform Maxent
-% 4. Process-based are much, much more sensitive to covariate choice -- implies sensitivity to regression accuracy
-% 5. Dispersal is critical for CA models
-% 6. Structural errors seem to have erratic effects
-% 0. Process-based are better IF you are confident of the process, but Maxent is more robust to error and uncertainty
-
-% Possible paragraphs:
-% Rankings:
-%	Ideal & mean(all scenarios): corresponding process-based is best on average
-%	lambda = IPM, MxE, CA (avgs)
-%	N = CA, MxE, IPM (avgs)
-%	Maxent somewhat better for lambda
-% Effect of data issues
-%	Measurement error
-% Effect of modelling issues
-% Differences in output
 
@@
 % Effect of modelling issues
-modelling with \emph{incorrect covariates} was universally problematic for model performance, though it had a greater impact on TSS in the shrub than the biennial species (Fig.
+Modelling with \emph{incorrect covariates} decreased performance for all SDMs, though the decline tended to be larger on the process-based SDMs.
+It generally had a greater impact on TSS in the shrub than the biennial species (Fig.
 \ref{fig:TSSvIdeal}).
@@
 The IPM was generally more resistant to mis-characterized dispersal compared to the simulation-based CA\textsubscript{i} and CA\textsubscript{p} models.
-Underestimation of dispersal parameters led to decreased predicted ranges, and overestimation to increased predicted ranges relative to ‘ideal’.
+Underestimation of dispersal parameters led to decreased predicted ranges, and overestimation to increased predicted ranges relative to \emph{ideal}.
 The effects were more extreme for the biennial than for the shrub.
@@
 \ref{fig:N_Lambda}), indicating a tendency to over-predict at range edges.
-Immediately interior to the range margin, the CA\textsubscript{p} and CA\textsubscript{i} model showed upward bias for the shrub (Fig.
+Immediately interior to the range margin, the CA\textsubscript{p} and CA\textsubscript{i} models showed upward bias for the shrub (Fig.
 \ref{fig:N_Lambda}).
@@
 \label{S:4}
-
-% Talk Discussion
-% 1. Across all scenarios, IPM is best for lambda, CAi/p are best for N, with Maxent in the middle for both
-% 2. Under ideal conditions, process-based outperform Maxent
-% 3. For non-equilibrium distributions, process-based outperform Maxent
-% 4. Process-based are much, much more sensitive to covariate choice -- implies sensitivity to regression accuracy
-% 5. Dispersal was critical for CA models, but had essentially no effect for IPMs
-% 6. Structural errors seem to have erratic effects
-% 0. Process-based are better IF you are confident of the process, but Maxent is more robust to error and uncertainty
 
@@
 % Dangers of process-based SDMs: dispersal
-The accuracy of the CA model predictions were sensitive to dispersal.
-Drastically underestimating dispersal resulted in smaller predicted ranges, while overpredicting dispersal resulted in larger predicted ranges.
+The accuracy of the CA model predictions were sensitive to dispersal (Fig.
+\ref{fig:TSSvIdeal}).
+As expected, underestimating dispersal resulted in smaller predicted ranges, while overestimating dispersal resulted in larger predicted ranges.
 The predictions for the biennial tended to be a bit more sensitive, particularly in terms of specificity (Fig.
 B.4).
+In contrast, the IPM predictions were less affected by errors in the dispersal parameters (Fig.
+\ref{fig:TSSvIdeal}).
 
@@
 
+\section{Acknowledgments}
+This project was funded through National Science Foundation award IIS-1717368.
+
+\section{Bibliography}
+\bibliography{ms/ms_latex/sdmMethodComp}
+
+\end{document}
+
+%% RESULTS
+% Talk results
+% 1. Across all scenarios, IPM is best for lambda, CAi/p are best for N, with Maxent in the middle for both
+% 2. Under ideal conditions, process-based outperform Maxent
+% 3. For non-equilibrium distributions, process-based outperform Maxent
+% 4. Process-based are much, much more sensitive to covariate choice -- implies sensitivity to regression accuracy
+% 5. Dispersal is critical for CA models
+% 6. Structural errors seem to have erratic effects
+% 0. Process-based are better IF you are confident of the process, but Maxent is more robust to error and uncertainty
+
+% Possible paragraphs:
+% Rankings:
+%	Ideal & mean(all scenarios): corresponding process-based is best on average
+%	lambda = IPM, MxE, CA (avgs)
+%	N = CA, MxE, IPM (avgs)
+%	Maxent somewhat better for lambda
+% Effect of data issues
+%	Measurement error
+% Effect of modelling issues
+% Differences in output
+
+%% DISCUSSION
+% Talk Discussion
+% 1. Across all scenarios, IPM is best for lambda, CAi/p are best for N, with Maxent in the middle for both
+% 2. Under ideal conditions, process-based outperform Maxent
+% 3. For non-equilibrium distributions, process-based outperform Maxent
+% 4. Process-based are much, much more sensitive to covariate choice -- implies sensitivity to regression accuracy
+% 5. Dispersal was critical for CA models, but had essentially no effect for IPMs
+% 6. Structural errors seem to have erratic effects
+% 0. Process-based are better IF you are confident of the process, but Maxent is more robust to error and uncertainty
+
 % So: Why or when to use each?
@@
 % Range boundaries: What is each SDM predicting?
-
-\section{Acknowledgments}
-This project was funded through National Science Foundation award IIS-1717368.
-
-\section{Bibliography}
-\bibliography{sdmMethodComp}
-
-\end{document}

### 2020-02-13 ms/ms_latex/ms_sdmComp.tex
discussion paragraphs

@@
 % Process-based predictions (lambda, N, etc)
-The process-based SDMs generate predictions for a number of biological quantities of interest.
-For instance, the IPM predicts $\lambda$ in each cell of the landscape.
 Discrepancies in predicted $\lambda$ values were biased downward in the northern portion of the study region (Fig.
@@
 Mirroring the effect on predicted presence, residual error was most variable and most extensive when \emph{incorrect covariates} were used.
-
-Similarly, the CA models predict the abundance $N$ for each species.
 Discrepancies in predicted abundance from the CA\textsubscript{p} model showed a spatial banding of upward bias at the range margins for both species (Fig.
@@
 
-% Process-based > Maxent under ideal conditions, and less susceptible to sampling effects
-Under ideal conditions, the process-based SDMs, properly matched to the range boundary definition, outperformed Maxent (Fig.
+% Process-based > Maxent under ideal conditions, and data scenarios – esp. non-equilibrium
+Under ideal conditions, the process-based SDMs, when properly matched to their respective range boundary definitions, outperformed Maxent (Fig.
 \ref{fig:TSSmn}).
-This generally held true across the sampling scenarios, with minimal impact of \emph{measurement error} or \emph{sampling bias} on the process-based SDMs.
-
-% Climate change and alien species
-Critically, there was little or no decrease in performance in the process-based models with \emph{non-equilibrium} samples (Fig.
+The SDMs were also generally unaffected by either the addition of \emph{measurement error} or to geographic \emph{sampling bias}, though the predictions of Maxent were affected somewhat more under both scenarios.
+Similarly, there was little or no decrease in performance in the process-based SDMs with \emph{non-equilibrium} samples (Fig.
 \ref{fig:TSSvIdeal}, B.3, B.4).
-In fact, the individual-based models (IPM, CA\textsubscript{i}) showed somewhat improved performance for the shrub, as the younger populations had a more even age distribution, allowing for more accurate estimation of the size-dependent vital rate regressions (Fig.
+In fact, the individual-based models (IPM, CA\textsubscript{i}) showed marginally improved performance for the shrub, as the younger populations had a more even age distribution, allowing for more accurate estimation of the size-dependent vital rate regressions (Fig.
 B.5a).
@@
 B.4).
-In contrast, the IPM predictions were less affected by errors in the dispersal parameters (Fig.
+In contrast, errors in the dispersal parameters had limited impact on the predictions of the IPM (Fig.
 \ref{fig:TSSvIdeal}).
+The movement of seeds between cells was insufficient to dramatically extend the boundary of sustainable populations, predicted by $\lambda$ in the IPM.
+By using both the IPM and CA\textsubscript{i} models, which are constructed using the same environmental and size relationships, sink populations in marginal cells can be distinguished, where individuals exist, but populations are unable to persist.
 
 % Other advantages and uses of process-based models
-There are a bunch of reasons you might want to build a process-based model aside from increased accuracy. I could draw on some examples from the buckthorn paper.
+Apart from potentially increased accuracy, process-based SDMs have other important benefits.
+While occurrence-based models often predict relative habitat suitability, IPMs predict the growth rate $\lambda$ and CA models predict relative abundance.
+However, the biological quantities provided by a process-based SDM are much more numerous.
+For instance, the models used here each predict survival rates, flowering rates, seed production, and germination rates in each cell, either at the level of the individual or as population averages.
+Further, process-based SDMs are naturally able to incorporate management actions \citep{Szewczyk2019} or a dynamic environment (CITE).
+Uncertainty in the covariates, for example in the form of climate change scenarios, and uncertainty in the parameters can be incorporated through multiple iterations of the models to better describe the range of plausible outcomes.
+The flexible nature of process-based SDMs thus provides an important opportunity to evaluate potential actions and plan for an inherently uncertain future.
 

### 2020-02-13 ms/ms_latex/ms_sdmComp.tex
revised

@@
 % Range boundaries
-IPMs predict $\lambda$ in each cell of the landscape, and consequently directly predict a $\lambda$-based range boundary.
-Similarly, the simulation-based CA models predict abundance, and thus directly predict a $N$-based range boundary.
+IPMs predict $\lambda$ in each cell of the landscape, and consequently directly predict the $\lambda$-based range boundary.
+Similarly, the simulation-based CA models predict abundance, thus directly predicting the $N$-based range boundary.
 For Maxent, the range boundary definition is less explicit.
-The geo-located presences used as input imply an abundance-based range, but predicted distributions are frequently constrained by a data-based threshold (e.g., relative suitability values that encapsulate 95\% of the observed data) to exclude both erroneous data and sink populations, which implies an effort to predict a $\lambda$-based range.
-There may or may not be a big difference between the two range definitions, but it will depend on the species.
-For example, the magnitude of the difference should vary with the number or distribution of marginal cells (lots of marginal cells mean large areas with middling occupancy probabilities), the typical annual stochasticity in the population abundances (more stochasticity means more discrepancies are likely), and dispersal limitation (high dispersal or a long introduction history means fewer areas that have not been reached, but also possibly a larger proportion of sink populations).
+The geo-located presences used as input imply an abundance-based range, but predicted distributions are frequently constrained by a data-based threshold (e.g., relative suitability values that encapsulate 95\% of the observed data) to exclude both erroneous data and sink populations \citep{Elith2009a,Allen2016}, which implies an intent to predict a $\lambda$-based range.
+The practical difference between the range definitions will depend on the biology of the species.
+For example, the magnitude of the difference should vary with the number or distribution of marginal cells, where large areas with marginal cells may result in different patterns of occupancy.
+Similarly, species with high annual stochasticity in population abundances are more likely to exhibit patchier distributions due to chance local extinctions \citep{Morris2003}.
+Alternatively, species with high dispersal, or invasive species with a longer introduction history, are less likely to be dispersal limited, but may instead have a larger proportion of non-viable sink populations.
+Selection of process-based SDMs requires clearly specifying the range boundary definition in a way that occurrence-based SDMs do not.
 
